### 2023-05-29 src/tex/ms.tex
Commit the content

% Define document class
\documentclass[reprint,superscriptaddress,nobibnotes,amsmath,amssymb,aps,prx,hidelinks]{revtex4-2}

% Import relevant packages
\usepackage{showyourwork}
\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage{bm}
\usepackage{multirow}
\usepackage[separate-uncertainty=true]{siunitx}
\usepackage{amsmath}
\usepackage[shortlabels]{enumitem}
\usepackage[page]{appendix}
\usepackage{xfrac}
\usepackage{printlen}
\usepackage[version=4]{mhchem}
\usepackage{blindtext}
% \usepackage[hidelinks]{hyperref}
\sisetup{
    list-units      = brackets,
    range-units     = brackets,
    range-phrase    = -,
    list-pair-separator= {, },
    list-separator  = {, },
    list-final-separator = {, }
    }
\renewcommand\appendixname{}
\newcommand{\papertitle}{Accurate Estimation of Diffusion Coefficients and their Uncertainties from Computer Simulation}
% \newcommand{\oMSD}{\ensuremath{\overline{\bm{s}^2}}}
% \newcommand{\oMSDs}[1]{\ensuremath{\overline{s^2}}(#1)}
% \newcommand{\oMSDi}{\ensuremath{\overline{s^2}_i}}
% \newcommand{\oMSDj}{\ensuremath{\overline{s^2}_j}}

\newcommand{\oMSD}{\ensuremath{\bm{x}}}
\newcommand{\oMSDs}[1]{\ensuremath{x}(#1)}
\newcommand{\oMSDi}{\ensuremath{x_i}}
\newcommand{\oMSDj}{\ensuremath{x_j}}
\newcommand{\oMSDn}{\ensuremath{x_n}}
\newcommand{\oMSDm}{\ensuremath{x_{n + m}}}
\newcommand{\moMSDn}{\ensuremath{\left<\oMSDn\right>}}
\newcommand{\moMSDm}{\ensuremath{\left<\oMSDm\right>}}
\newcommand{\modelmatrix}{\mathbf{A}}
\newcommand{\model}{\bm{m}}
\newcommand{\prob}[1]{\ensuremath{p(#1)}}
\newcommand{\probsq}[1]{\ensuremath{p[#1]}}
\newcommand{\nind}[1]{\ensuremath{N^\prime_{#1}}}
\newcommand{\code}[1]{#1}

\newcommand{\MSD}[1]{\big\langle\Delta\mathbf{r}{(#1)}^2\big\rangle}
\newcommand{\Dest}{\ensuremath{\widehat{D}^*}}
\newcommand{\D}{\ensuremath{D^*}}
\newcommand{\var}[1]{\ensuremath{\sigma^2[#1]}}
\newcommand{\varest}[1]{\ensuremath{\widehat{\sigma}^2[#1]}}
\DeclareMathOperator{\cov}{cov}
\DeclareMathOperator*{\argmax}{argmax}

% Redefine \maketitle so that it can be used twice (for supplementary)
\makeatletter
\def\maketitle{
\@author@finish
\title@column\titleblock@produce
\suppressfloats[t]}
\makeatother

% Begin!
\begin{document}

\let\oldaddcontentsline\addcontentsline% Store \addcontentsline
\renewcommand{\addcontentsline}[3]{}

% Title
\title{\papertitle}

% Author list
\author{Andrew R. McCluskey}
  \email{andrew.mccluskey@ess.eu}
  \affiliation{European Spallation Source ERIC, Ole Maaløes vej 3, 2200 København N, DK}
\author{Samuel W. Coles}
  \affiliation{Department of Chemistry, University of Bath, Claverton Down, Bath, BA2 7AY, UK}
  \affiliation{The Faraday Institution, Quad One, Harwell Science and Innovation Campus, Didcot, OX11 0RA, UK}
\author{Benjamin J. Morgan}
  \email{b.j.morgan@bath.ac.uk}
  \affiliation{Department of Chemistry, University of Bath, Claverton Down, Bath, BA2 7AY, UK}
  \affiliation{The Faraday Institution, Quad One, Harwell Science and Innovation Campus, Didcot, OX11 0RA, UK}

% Abstract
\begin{abstract}
    Self-diffusion coefficients, $\D$, are routinely estimated from molecular dynamics simulations by fitting a linear model to the observed mean-squared displacements (MSDs) of mobile species.
    MSDs derived from simulation suffer from statistical noise, which introduces uncertainty in the resulting estimate of $\D$.
    An optimal scheme for estimating $\D$ will minimise this uncertainty, i.e., will have high statistical efficiency, and will give an accurate estimate of the uncertainty itself.
    We present a scheme for estimating $\D$ from a single simulation trajectory with high statistical efficiency and accurately estimating the uncertainty in the predicted value.
    The statistical distribution of MSDs observable from a given simulation is modelled as a multivariate normal distribution using an analytical covariance matrix for an equivalent system of freely diffusing particles, which we parameterise from the available simulation data. 
    We then perform Bayesian regression to sample the distribution of linear models that are compatible with this model multivariate normal distribution, to obtain a statistically efficient estimate of $D^*$ and an accurate estimate of the associated statistical uncertainty.
\end{abstract}

\maketitle

\section{Introduction}

Mass transport is a fundamental physical process that is central to our understanding of fluids~\cite{sendner_interfacial_2009,shimizu_structural_2015,ghoufi_ultrafast_2016} and plays a critical role in biochemical systems~\cite{maccmmon_dynamics_1977,robertson_diffusion_2006}, and in solid-state devices such as batteries, fuel cells, and chemical sensors~\cite{eames_ionic_2015,morgan_understanding_2011,walsh_taking_2018}.
Molecular dynamics simulations are widely used to study microscopic transport processes, as they give direct insight into atomic-scale transport mechanisms and can be used to calculate macroscopic transport coefficients~\cite{morgan_relationships_2014,morgan_mechanistic_2021,poletayev_defect_2022,klepis_long_2009,wang_application_2011,zelovich_hydroxide_2019}.
These transport coefficients are formally defined in terms of ensemble averages.
Dynamical simulations, however, sample the full ensemble space stochastically, and parameters derived from simulation data are therefore only estimates of the true parameter of interest.
The statistical uncertainty associated with such estimates depends on the details of the simulation---e.g., size and timescale---and on the choice of estimation method.
An optimal estimation method will minimise the uncertainty in the computed quantity, i.e., it will have high statistical efficiency, and will also allow this uncertainty to be accurately estimated.

One commonly used parameter for quantifying atomic-scale mass transport is the self-diffusion coefficient, $\D$, which describes diffusion in the absence of a chemical potential gradient.
$\D$ is related to the ensemble-average mean squared displacement (MSD), $\MSD{t}$, via the Einstein relation~\cite{einstein_uber_1905,helfand_transport_1960},
%
\begin{equation}
    \D = \lim_{t\to\infty}\frac{\MSD{t}}{6t},
    \label{equ:einstein}
\end{equation}
%
where $t$ is elapsed time.
Because numerical simulations are finite in both time and space, MSDs obtained from simulation data always deviate from the true ensemble average MSD.
One can, however, compute an estimate of the self-diffusion coefficient, $\Dest$, by fitting a linear model to the observed MSD and using the gradient of this fitted model in place of $\MSD{t} / t$ in Eqn.~\ref{equ:einstein}.

The simplest approach to fitting a linear model to MSD data from simulation is ordinary least squares regression (OLS).
OLS gives analytical expressions for the ``best fit'' regression coefficients (the slope and intercept) and their respective uncertainties, making it easy to implement and quick to perform.
This procedure, however, is appropriate only for data that are both statistically independent and identically distributed.
Neither of these conditions hold for MSD data obtained from simulation, which instead are serially correlated and usually have unequal variances.
As a consequence, OLS is statistically inefficient, giving a relatively large statistical uncertainty in $\Dest$.
Furthermore, using the textbook OLS expression for the uncertainty in $\Dest$ significantly underestimates the true uncertainty in this estimate.
This underestimated uncertainty may cause overconfidence in the accuracy of values of $\D$ estimated using OLS, and using these data in downstream analyses may result in faulty inferences.
While the uncertainty associated with OLS estimates of $\D$ can, in principle, be accurately estimated by directly sampling over multiple repeated simulations, this approach greatly increases the total computational cost and therefore is often not practical.

Here, we describe an approximate Bayesian regression method for estimating $\D$ with near-maximal statistical efficiency while also accurately estimating the corresponding statistical uncertainty, using data from a single simulation.
We model the statistical population of simulation MSDs as a multivariate normal distribution, using an analytical covariance matrix derived for an equivalent system of freely diffusing particles, with this covariance matrix parameterised from the observed simulation data.
We then use Markov-chain Monte Carlo to sample the posterior distribution of linear models compatible with this multivariate normal model.
The resulting posterior distribution provides an efficient estimate for $\D$ and allows the associated statistical uncertainty in $\Dest$ to be accurately quantified.
This method is implemented in the open-source Python package \textsc{kinisi}~\cite{mccluskey_kinisi_2022}.

\section{Background}

For a simulation of equivalent particles, the observed mean squared displacement as a function of time, $\oMSDs{t}$, can be computed as an average over equivalent particles and time origins:
%
\begin{equation}
  \oMSDs{t} = \frac{1}{N(t)}\sum^{N(t)}_{j=1}{\left[\Delta\mathbf{r}_j(t)\right]}^2,
  \label{equ:observed_msd}
\end{equation}
%
where $N(t)$ is the total number of observed squared-displacements at time $t$.
The resulting observed MSD is a vector, $\oMSD$, with individual elements $\oMSDi$.
Each element of this vector differs from the true ensemble-average MSD for that time by some unknown amount.
Fitting a linear model to $\oMSD$ gives an estimated self-diffusion coefficient, $\Dest$, which again differs from the true self-diffusion coefficient, $\D$, by some unknown amount.

Performing repeated simulations starting from different random seeds or with different histories will produce a set of replica trajectories, where each trajectory gives a different, statistically equivalent, observed MSD.
The set of all possible replica trajectories defines a population of hypothetical observed MSDs, and the MSD obtained from any one trajectory can be considered a random sample, $\bm{X}$, drawn from the multivariate probability distribution that describes this population, i.e, $\bm{X} \sim \prob{\oMSD}$.
Each potential MSD sample could, in principle, be fitted to a linear model to obtain a corresponding estimate for the self-diffusion coefficient; $\bm{X} \mapsto \Dest$.
The population of all such estimates therefore defines a probability distribution $\prob{\Dest}$.
The estimated diffusion coefficient obtained from a single simulation corresponds to a random sample drawn from this distribution, while the uncertainty in $\Dest$ is described by the shape of the full distribution $\prob{\Dest}$.

The statistical properties of $p(\Dest)$ depend on both the input MSD data and the choice of regression scheme used to obtain a ``best fit'' linear model.
An optimal estimation scheme for $\D$ should be unbiased, i.e., the expected value, $\mathbb{E}(\Dest)$, should equal the true self-diffusion coefficient $\D$, and should be maximally statistically efficient, i.e., the spread of $p(\Dest)$ around $\D$ should be minimised.
An estimation scheme should also provide an accurate estimate of the uncertainty in $\Dest$, to allow this estimated parameter to be used in subsequent inferential analysis.

For data that are both statistically independent and identically normally distributed, ordinary least squares regression (OLS) is unbiased and statistically efficient, and gives accurate estimates of the uncertainties in the resulting regression coefficients.
MSD data obtained from simulation, however, are neither statistically independent nor identically distributed.
The variances, $\var{\oMSDi}$, are correlated, since the displacement of each particle at time $t+\Delta t$ is necessarily similar to its displacement at time $t$, and hence, $\oMSDs{t}$ is similar to $\oMSDs{t+\Delta t}$.
These variances are also typically unequal---the data are heteroscedastic;
for example, for a random walk, the MSD variance increases quadratically with time \cite{smith_random_1996}. 
Because the key assumptions of the OLS method are not valid for MSD data, OLS gives statistically inefficient estimates of $\D$, while the estimated regression uncertainties obtained from the standard OLS statistical formulae significantly underestimate the true uncertainty in $\prob{\Dest_\mathrm{OLS}}$ (Fig.~\ref{fig:glswlsols}a).

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/glswlsols.pdf}
    \caption{
        Example distributions of estimated self-diffusion coefficients, $\Dest$, calculated using (a) ordinary least squares (OLS), (b) weighted least squares (WLS), and (c) generalised least squares (GLS),
        from MSD data from \num{4096} individual simulations of \num{128} particles undergoing a \SI{128}{step} 3D lattice random walk, with a step size chosen so that the true diffusion coefficient $\D = 1$.
        In each panel, the grey curve shows the best-fit normal distribution for the simulation data, the upper horizontal bar shows the standard deviation of this distribution, and the lower horizontal bar shows the average estimated standard distribution given by the analytical expression for $\sigma[\prob{\Dest}]$ for each regression method.}
    \label{fig:glswlsols}
    \script{glswlsols.py}
\end{figure}

Some improvement can be made by using weighted least squares (WLS) (Fig.~\ref{fig:glswlsols}b), where the residual for each observed MSD value is weighted by the reciprocal of its variance, $1/(\var{\oMSDi})$.
Like OLS, WLS is an unbiased estimator, and for heteroscedastic data it has higher statistical efficiency than OLS.
WLS still disregards correlations in $\oMSD$, however, and is therefore statistically inefficient, while the WLS estimated uncertainties for the regression coefficients still underestimate the true uncertainty in $\prob{\Dest_\mathrm{WLS}}$.

To optimally estimate the true ensemble-average MSD, and hence $\D$, from simulation data, it is necessary to account for both the changing variance and correlation structure of $\oMSD$.
Within the framework of linear regression, this can be achieved using generalised least squares (GLS).
GLS gives estimated regression coefficients, $\widehat{\beta}$, via
%
\begin{equation}
    \widehat{\beta} = \left(\modelmatrix^{\top}\mathbf{\Sigma}^{-1}\modelmatrix\right)^{-1}\modelmatrix^{\top}\mathbf{\Sigma}^{-1}\oMSD,
    \label{equ:gls}
\end{equation} 
%
where $\modelmatrix$ is the model matrix $\begin{bmatrix}\mathbf{1} & \bm{t}\end{bmatrix}$, with $\bm{t}$ the vector of observed times, and $\mathbf{\Sigma}$ is the covariance matrix for the observed MSD values.
For correlated heteroscedastic data, such as MSD data, GLS offers the theoretical maximum statistical efficiency---it achieves the Cram\'er--Rao bound~\cite{cramer_mathematical_1946,rao_information_1945,rao_selected_1994,darmois_sur_1945,aitken_on_1942}---and provides accurate analytical estimates of the uncertainty in the predicted regression coefficients (Fig.~\ref{fig:glswlsols}c).

A second method for estimating the ensemble-average MSD, and thus $\Dest$, from simulation data is Bayesian regression.
Like GLS, Bayesian regression can take into account both the changing variance and the correlation structure inherent in the data.
Rather than providing a singular ``best-fit'' estimate like GLS, Bayesian regression produces a posterior probability distribution for the regression coefficients.
The mean of this distribution serves as the point estimate of the coefficients and, in the absence of additional prior information, is equivalent to the GLS estimate, while the spread of the distribution quantifies the uncertainty in these estimates.
For data that is both heteroscedastic and correlated, such as MSD data from simulations, Bayesian regression, like GLS, is formally fully statistically efficient.

To estimate $\D$ from some observed MSD data, $\oMSD$, using Bayesian regression, we compute the posterior probability distribution $\prob{\model|\oMSD}$ for a linear model $\model = 6\D \bm{t} + c$, where $\D$ and $c$ are parameters to be estimated.
This posterior distribution is described by Bayes' theorem,
%
\begin{equation}
    \prob{\model|\oMSD} = \frac{\prob{{\oMSD|\model}}\prob{\model}}{\prob{\oMSD}},
    \label{equ:bayes}
\end{equation}
%
where $\prob{\oMSD|\model}$ is the probability of observing data $\oMSD$ given model $\model$, often described as the ``likelihood'', and $\prob{\oMSD}$ is the marginal probability of the observed data $\oMSD$.
Integrating over $\prob{\model|\oMSD}$ with respect to $c$ yields the marginal posterior distribution $\prob{\D|\oMSD}$, from which the best point-estimate $\Dest$ and distribution variance $\varest{\Dest}$ can be computed.

Given a sufficiently large number of observed squared displacements at each time $t$, the central limit theorem applies, and $\oMSD$ can be considered a sample from a multivariate normal distribution with log-likelihood
%
\begin{equation}
    \begin{aligned}{}
      \ln \prob{\oMSD|\model} = -\frac{1}{2}\big[ & \ln(\left|\mathbf{\Sigma}\right|) + {(\oMSD - \model)}^{\!\top}\mathbf{\Sigma}^{-1}(\oMSD - \model) \\ 
      & + k \ln(2\pi)\big],
    \end{aligned}
    \label{equ:loglike}
\end{equation}
%
where $\mathbf{\Sigma}$ is the observed MSD covariance matrix and $k$ is the length of the vector $\oMSD$, i.e., the number of time intervals for which we have observed MSD data.
Providing this likelihood function can be calculated, we can compute the posterior distribution $\prob{\model | \oMSD}$, which gives an optimally efficient point-estimate for $\D$ and a complete description of the associated uncertainty in $\Dest$.

\section{Approximating $\bm{\Sigma}$ from simulation data}

For Bayesian regression and GLS, we require the covariance matrix for the observed MSD, $\mathbf{\Sigma}$, which is generally unknown.
To proceed, we replace $\mathbf{\Sigma}$ with a model covariance matrix, $\mathbf{\Sigma^\prime}$, with a known analytical form, that we parameterise from the available simulation data.
Providing the correlation structure of $\mathbf{\Sigma^\prime}$ is similar to that of $\mathbf{\Sigma}$, this model correlation matrix can be used in approximate Bayesian or GLS schemes to estimate the ensemble-average MSD, and hence $\D$, with high efficiency and accurate estimated uncertainties.

We model the covariance matrix for the observed MSD from a given simulation using the covariance matrix for the MSD of an equivalent system of freely diffusing particles, $\mathbf{\Sigma^\prime}$.
For observed MSDs computed by averaging over non-overlapping time windows, the covariance matrix $\mathbf{\Sigma^\prime}$, in the long time limit, has elements (see SI)
%
\begin{equation}
  \Sigma^\prime\left[\oMSDi, \oMSDj\right]= \Sigma^\prime\left[\oMSDj, \oMSDi\right] =
  \var{\oMSDi} \frac{\nind{i}}{\nind{j}},\hspace{1em} \forall\,i \leq j,
  \label{equ:cvv}
\end{equation} 
%
where $\var{\oMSDi}$ are the time-dependent variances of the observed MSD, and $\nind{i}$ is the total number of non-overlapping observed squared-displacements for time-interval $i$.
We estimate the variances $\var{\oMSDi}$ using the standard result that the variance of the mean of a sample scales inversely with the number of independent constituent observations.
Specifically, we compute an estimated variance $\varest{\oMSDi}$ by rescaling the observed variance of the squared displacement for time interval $i$ by the number of non-overlapping contributing time windows, $\nind{i}$;
%
\begin{equation}
  \varest{\oMSDi} = \frac{1}{\nind{i}}\var{\Delta r_i^2}.
  \label{equ:varestMSD}
\end{equation}
%

The estimated variance $\varest{\oMSD}$ can be calculated from a single simulation trajectory, and provides an accurate estimate of the true variance $\var{\oMSD}$.
To demonstrate this, we performed \num{4096} independent simulations of \num{128} particles undergoing a three-dimensional cubic-lattice random walk of \num{128} steps per particle.
Using data from all \num{4096} simulations, we first compute the true simulation MSD and its variance (Fig.~\ref{fig:msd}a).
We also compute the MSD and estimated variance using data from a single simulation trajectory (Fig.~\ref{fig:msd}b), using the scheme described above.
A quantitative comparison between the true MSD variance and the single-trajectory estimated MSD variance is made in Fig.~\ref{fig:msd}c: the close numerical agreement confirms that Eqn~\ref{equ:varestMSD} can be used to estimate $\var{\oMSD}$, which can then be used to parameterise the model covariance matrix $\mathbf{\Sigma^\prime}$ via Eqn.~\ref{equ:cvv}.
%
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/msd.pdf}
    \caption{
        Comparison of the numerical variance in observed MSD from multiple replica simulations and the estimated variance in observed MSD given by rescaling the variance in observed squared displacements (Eqn.~\ref{equ:varestMSD}).
        Panel (a) shows the mean observed MSD from \num{4096} simulations of \num{128} particles undergoing a 3D lattice random walk of \num{128} steps per particle, with error bars of $\pm2\sigma[\oMSDi]$.
        Panel (b) shows the MSD from just one simulation, with error bars of $\pm2\widehat{\sigma}[\oMSDi]$, obtained via Eqn.~\ref{equ:varestMSD}.
        Panel (c) plots the numerical variance against the estimated variance from a single simulation as a function of timestep $i$.
        %Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences.
    }
    \label{fig:msd}
    \script{msd.py}
\end{figure}
%

The practical implementation of both GLS and Bayesian regression requires that the covariance matrix $\mathbf{\Sigma^\prime}$ is invertible (positive definite); see Eqns.~\ref{equ:gls} and \ref{equ:loglike}.
The estimated MSD variances derived from simulation data via Eqn.~\ref{equ:varestMSD} are statistically noisy and using these to directly parameterise $\mathbf{\Sigma^\prime}$ can yield non-invertible singular matrices.
To make our scheme numerically tractable, we therefore fit our estimated MSD variances to the analytical variance for an analogous system of particles undergoing random walks~\cite{smith_random_1996};
%
\begin{equation}
    \var{\oMSDi} = a\frac{t^2_i}{\nind{i}},
    \label{equ:variance}
\end{equation}
%
where $a$ is a scaling parameter determined by fitting Eqn.~\ref{equ:variance} to the directly estimated MSD variances.
This smoothing of $\var{\oMSDi}$ guarantees that the resulting model covariance matrix $\mathbf{\Sigma^\prime}$ is invertible and thus suitable for GLS or Bayesian regression.

To illustrate the complete numerical procedure for deriving the model covariance matrix, $\bm{\Sigma^\prime}$, we present in Fig.~\ref{fig:covariances} the MSD covariance matrix for \num{4096} random-walk simulations, as described above, at three differing levels of approximation:
the numerically converged covariance matrix, $\bm{\Sigma}$, computed using the data from all \num{4096} simulations (Fig.~\ref{fig:covariances}a);
the corresponding analytical model covariance matrix, $\mathbf{\Sigma^\prime}$, defined by Eqn.~\ref{equ:cvv} and parametrised using analytical variances $\var{\oMSDi}$ (Fig.~\ref{fig:covariances}b); and the average model covariance matrix obtained by parametrising Eqn.~\ref{equ:cvv} using smoothed variances estimated from individual simulation trajectories, and averaging over the resulting set of all \num{4096} matrices (Fig.~\ref{fig:covariances}c).
%
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/covariances.pdf}
    \caption{
        (a) The numerical MSD covariance matrix $\bm{\Sigma}$ calculated using MSD data from \num{4096} simulations of \num{128} particles undergoing a 3D lattice random walk of \num{128} steps per particle.
        (b) The analytical MSD covariance matrix $\bm{\Sigma^\prime}$ (Eqn.~\ref{equ:cvv}), parametrised using analytical random-walk variances $\var{\oMSDi}$.
        (c) The MSD covariance matrix obtained applying the numerical scheme described in the main text to each individual random walk simulation, averaged over all \num{4096} such simulations.
        Colour bars in (a--c) show the covariance, $\Sigma\left[\oMSDi, \oMSDj\right]$.
        The off-diagonal panels show difference plots, computed as per-element ratios between pairs of covariance matrices (a--c).
        %Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences.
    }
    \label{fig:covariances}
    \script{covariances.py}
\end{figure}
%

While the analytical and average estimated covariance matrices show some systematic deviation from the numerically converged covariance matrix, the general correlation structure is preserved.
The discrepancy between the model and numerical covariance matrices largely stems from the approximation made in deriving the analytical form that $t$ is large, which leads to an overestimation of the variance at low $t$.
Despite this, the average estimated covariance matrix reproduces well the correlation structure of the true numerical covariance matrix, indicating that the component covariance matrices estimated from individual simulation trajectories may be used within approximte GLS or Bayesian regression schemes to estimate $\D$ and $\var{\Dest}$.

\section{Validation}

To demonstrate the complete approximate Bayesian regression scheme, as described above, we present two distinct examples.
First, we consider a simple 3D-lattice random walk, where the true self-diffusion coefficient $\D$ is specified by the simulation parameters, and a well-converged numerical covariance matrix can be obtained with relatively low computational cost, which allows us to directly compare the estimates produced by our method to ``best case'' estimates from a hypothetical method with access to the true covariance matrix.
Second, we consider an example real-world system---the lithium-ion solid electrolyte \ce{Li7La3Zr2O12} (LLZO)---which represents a practical application of our method to a well-studied material of practical interest for solid-state lithium-ion batteries \cite{MuruganEtAl_AngewChemIntEd2007,burbano_sparse_2016,morgan_lattice_2017,SquiresEtAl_PhysRevMater2022}.

\begin{figure*}[ht!]
    \centering
    \includegraphics[width=\textwidth]{figures/random_walk.pdf}
    \caption{
        (a) Observed MSD from a single simulation of 128 particles undergoing a 3D-lattice random walk of 128 steps per particle (dark line).
        The green shading shows the corresponding posterior distribution $\prob{\model|\oMSD}$ of linear models compatible with the observed MSD data $\oMSD$, calculated using the scheme described in the main text.
        The variegated shading indicates compatibility intervals of \SIlist[list-final-separator = {, and }]{1;2;3}{\sigma}$[\prob{\model|\oMSD}]$.
        (b) The marginal posterior distribution $\prob{\Dest|\oMSD}$ obtained from the posterior distribution of linear models in (a).
        The mean of this distribution gives the point estimate $\Dest$ for this simulation input data.
        The blue horizontal bar shows an interval of one standard deviation in $\prob{\Dest|\oMSD}$.
        (c) Probability distribution of point-estimates $\prob{\Dest}$ obtained from \num{4096} individual random-walk simulations.
        Each simulation has been analysed as in (a) and (b) to yield a single corresponding point estimate $\Dest$.
        The grey line shows the distribution of point estimates, $\prob{\Dest_\mathrm{num}}$, obtained using Bayesian regression with a mean vector and numerical covariance matrix derived from the complete dataset of all \num{4096} simulations.
        The pink horizontal bar shows an interval of one standard deviation in $\prob{\Dest}$.
        (d) Probability distribution of estimated variances, $\varest{\Dest}$, for individual random-walk simulations, compared to the true sample variance (pink vertical line) $\var{\Dest}$.
    } 
    \label{fig:random_walk}
    \script{random_walk.py}
\end{figure*}

Fig.~\ref{fig:random_walk}a shows the observed MSD from a single 3D-lattice random-walk simulation, along with the estimated posterior distribution of linear models compatible with the observed MSD data, $\prob{\model|\oMSD}$, calculated via Eqns.~\ref{equ:bayes} and \ref{equ:loglike}.
The corresponding marginal posterior distribution of estimated diffusion coefficients $\prob{\D|\oMSD}$ is shown in Fig.~\ref{fig:random_walk}b; this distribution is approximately Gaussian and is centred on the true self-diffusion coefficient $\D = \num{1}$, demonstrating that for this example trajectory we obtain a good point-estimate of $\D$.

To evaluate the overall performance of our method, we repeat our analysis on the full set of \num{4096} random-walk simulations.
Fig.~\ref{fig:random_walk}c presents a histogram of the resulting point estimates of $\D$, with each estimate derived as the mean of the posterior distribution $\prob{\D|\oMSD}$ using input data from an individual simulation.
We also show the probability distribution of estimated diffusion coefficients obtained using Bayesian regression with a mean vector and covariance matrix derived numerically from all \num{4096} simulations (solid line).
This latter distribution represents the distribution of ``best possible'' estimates of $\D$ and exhibits the minimum possible theoretical variance.
The close agreement between these two distributions demonstrates that our approximate Bayesian regression scheme yields nearly optimal estimates of $\D$ using data from individual simulations.
The distribution of estimated diffusion coefficients from single simulations is slightly broader than the exact numerical results.
This minor deviation is a consequence of the overestimation of $\varest{\oMSDi}$ at short times, noted above, which results from our use of the long-time limit in the derivation of the analytical model covariance matrix.

We next consider the degree to which our method can quantify the uncertainty in $\Dest$ when using input data from a single simulation.
Fig.~\ref{fig:random_walk}d shows the distribution of estimated variances $\varest{\Dest}$, with each sample calculated from an individual simulation trajectory.
We also show the true variance of individual point estimates, $\var{\Dest}$, which characterises the spread of the histogram in Fig.~\ref{fig:random_walk}c.
The distribution of estimated variances is biased relative to the true variance and skewed, due to numerical differences between the true covariance matrix $\mathbf{\Sigma}$ and the model covariance matrix $\mathbf{\Sigma^\prime}$ (see the SI for details).
In general, however, the distribution of the estimated variance shows good agreement with the true sample variance.
Notably, the precision of this estimate is significantly greater than obtained using OLS or WLS and their corresponding textbook statistical formulae.

We next benchmark our method using data from simulations of the lithium-ion solid electrolyte cubic \ce{Li7La3Zr2O12} (c-LLZO).
We performed a single simulation of \num{1536} atoms (\num{448} Li ions) at \SI{1000}{K} for \SI{1.6}{\nano\second} (full simulation details are provided in the SI).
To generate multiple statistically equivalent trajectories, the resulting simulation data was partitioned into \num{512} effective trajectories, each approximately $\sim\SI{25}{\ps}$ in length, and containing data for \num{56} lithium ions.
We then perform the same approximate Bayesian regression analysis as above on each effective trajectory, excluding the first \SI{10}{ps} of MSD data in each case to remove short-time data corresponding to ballistic and sub-diffusive regimes \cite{he_statistical_2018}.

The resulting distribution of the point estimates, $\Dest$, from analysis of all \num{512} effective trajectories is shown in Fig.~\ref{fig:diffusion}a.
Again, the corresponding distribution of $\Dest$ estimates derived using Bayesian regression and a well-converged numerical covariance matrix calculated from the full LLZO dataset is also shown for comparison.
The distribution $\prob{\Dest}$ obtained using the model covariance matrix and parametrised separately for each individual effective simulation is highly similar to that obtained using the aggregate numerical covariance matrix calculated from the complete simulation dataset.
This close agreement mirrors the results for our random walk simulations, and confirms that our method yields accurate and statistically efficient estimates for $\D$, even for real-world simulation data.

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/diffusion.pdf}
    \caption{
      (a) Probability distribution of point estimates $\prob{\Dest}$ for \num{512} effective simulations of LLZO (orange histogram).
      The grey line shows the distribution $\prob{\Dest_\mathrm{num}}$ obtained using Bayesian regression with the complete LLZO dataset as input.
      The pink bar shows an interval of one standard deviation $\sigma[\prob{\Dest}]$.
      (b) Probability distribution of estimated variances, $\varest{\Dest}$, for individual LLZO effective simulations, compared to the true sample variance (pink vertical line) $\var{\Dest}$.
      %Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences. 
    }
    \label{fig:diffusion}
    \script{diffusion.py}
\end{figure}

We also consider the probability distribution of estimates of the variance in $\Dest$ calculated for each effective trajectory (Fig.~\ref{fig:diffusion}b), which we compare to the true variance in $\Dest$ for our method; i.e., the variance of the histogram in Fig.~\ref{fig:diffusion}a.
While the estimated variances deviate somewhat from the true distribution $\prob{\var{\Dest}}$, the agreement is reasonable and mirrors our results for the random walk simulations.
Hence, our method provides reasonably accurate estimates of the uncertainty in $\Dest$ for our c-LLZO dataset, even when applied to single effective trajectories with limited displacement data (only 56 mobile ions, and \SI{25}{ps} simulation length).

% To further benchmark our method, we performed a scaling analysis using sets \num{512} random-walk simulations with varying total number of particles or total simulation length (number of steps).
% We then analyse the resulting simulation data using four estimation methods to generate distributions of compatible linear models:
% \begin{enumerate}
%   \item Bayesian regression using a numerical covariance matrix derived from the complete set of \num{512} simulations.
%   This analysis is equivalent to GLS, and gives the smallest theoretical variance in the resulting point estimates $\Dest$.
%   \item Approximate Bayesian regression as described above, with an estimated model covarianec matrix computed for each individual trajectory using the corresponding simulation data.
%   \item Approximate Bayesian regression with a diagonal covariance matrix, $\bm{\Sigma}=\bm{\sigma}\mathbf{I}$, with the diagonal terms estimated from the simulation data.
%   This analysis is equivalent to WLS.
%   \item Bayesian regression with a covariance matrix $\bm{\Sigma}=\mathbf{I}$, the identity matrix.
%   This analysis is equivalent to OLS.
% \end{enumerate}
% Our analysis is shown in Fig.~[TODO] across two panels. The first panel illustrates the scaling performance of each method as we vary the number of atoms. The second panel demonstrates the scaling performance of each method as we adjust the total simulation time.

\section{Summary and Discussion}

We have introduced and evaluated an approximate Bayesian regression method for estimating the self-diffusion coefficient, $\D$, from molecular dynamics simulation data.
We consider the observed mean-squared displacement data from a single simulation as a random sample, $\bm{X}$, from a population of potential MSDs generated by equivalent replica simulations, $\bm{X}\sim p(\oMSD)$.
We model this population using a multivariate normal distribution, $p(\oMSD) = \mathcal{N}(\model, \mathbf{\Sigma})$, with mean vector $\model = 6\D\bm{t} + c$, where $\D$ and $c$ are model parameters to be determined.

To model the covariance matrix, we use an analytical solution derived for an equivalent system of freely diffusing particles.
To parameterise this model covariance matrix, we rescale the variance of the observed squared displacements from the input simulation trajectory, followed by a smoothing step to ensure a positive-definite matrix.
The resulting model covariance matrix preserves the correlation structure of the true simulation MSD covariance matrix, and gives a multivariate normal model for the population of observable simulation MSDs that depends solely on the model parameters, $\D$ and $c$.

We use Markov-Chain Monte Carlo to sample the posterior distribution of linear models compatible with the observed MSD data.
This approach yields a marginal posterior distribution, $\prob{\D | \oMSD}$, that gives a statistically efficient point estimate for $\D$ and allows the associated statistical uncertainty, $\var{\Dest}$, to be quantified.

We have benchmarked our approach using simulation data for an ideal 3D lattice random walk and for the lithium-ion solid electrolyte \ce{Li7La3Zr2O12} (LLZO).
In both cases, we obtain a distribution of estimates for $\D$ that closely matches the theoretically optimal distribution obtained using a numerical covariance matrix derived from a large number of replica simulation trajectories.

We obtain estimates for $\D$ that are unbiased, with near-optimal statistical efficiency, using input data from single simulation trajectories.
The approximate Bayesian regression scheme therefore provides more accurate single-point estimates of the self-diffusion coefficient than the commonly used OLS or WLS methods, when applied to the same input simulation data.
The improved statistical efficiency of this method, when compared to OLS or WLS, enables the estimation of $\D$ with equivalent accuracy from considerably smaller simulations---either in terms of timescale or system size. 
This reduces the overall computational cost when compared to studies that use OLS or WLS for estimating a linear fit to simulation MSD data.
Alternatively, this approach provides the possibility to estimate $\D$ with greater precision, given simulation trajectories of equal size.

Our method also provides reasonable estimates of the statistical uncertainty in the estimated value $\Dest$, in contrast to OLS and WLS which systematically significantly underestimate the uncertainty in regression coefficients when applied to simulated MSD data. 
While these estimated statistical uncertainties can still differ from the true (but unknown) uncertainty in $\Dest$, particularly when using short-timescale simulation data, they allow for scientifically meaningful comparisons to be made between estimated diffusion coefficients across different materials or under varying conditions, such as changes in temperature, or between computational findings and experimental results.
Furthermore, these uncertainties allow for quantitative downstream analysis, such as the application of Arrhenius (on non-Arrhenius) type models to describe the temperature dependence of self-diffusion.

The approximate Bayesian regression scheme presented here provides a statistically efficient means of estimating the self-diffusion coefficient, $\D$, from molecular dynamics simulation data.
It improves upon textbook approaches by providing accurate point estimates of $\D$ with near-optimal statistical efficiency, while also providing a reasonable description of the uncertainty in these estimates.
The high statistical efficiency of our method allows for the use of smaller simulations, which can significantly reduce computational costs.
Overall, our method offers significant advantages over more conventional methods of estimating self-diffusion coefficients from atomistic simulations.
We have implemented this procedure in the open-source package \textsc{kinisi} \cite{mccluskey_kinisi_2022}, which we hope will support its use within the broader simulation community across a range of materials science contexts.
\section*{Data availability}

Electronic Supplementary Information (ESI) available: A complete set of analysis/plotting scripts allowing for a fully reproducible and automated analysis workflow, using \textsc{showyourwork}~\cite{luger_showyourwork_2021}, for this work is available at Ref.~\cite{mccluskey_github_2022} under an MIT license.
All raw simulation files are available on Zenodo shared under CC BY-SA 4.0 licences, the random walk simulations and other analysis-linked data~\cite{mccluskey_zenodo_2022} and the LLZO raw simulation trajectories~\cite{coles_llzo_zenodo_2022}.
The method outlined in this work is implemented in the open-source Python package \textsc{kinisi}~\cite{mccluskey_kinisi_2022}, which is available under an MIT license.

\section*{CR\lowercase{e}d\lowercase{i}T author statement}

A.R.M.: Conceptualization, Formal Analysis, Investigation, Methodology, Software, Visualisation, Writing---original draft.
S.W.C.: Methodology, Resources, Writing---review and editing.
B.J.M.: Conceptualization, Methodology, Software, Writing---review and editing.

\section*{Acknowledgements}

The authors thank Jacob M.\ Dean and Tim Rogers for their valuable input in checking the mathematical derivations that make up the appendices, the beta-testers for the \textsc{kinisi} package, and Rodrigo Luger and Daniel Foreman-Mackey for their help using \textsc{showyourwork}~\cite{luger_showyourwork_2021}.
This work used the Isambard 2 UK National Tier-2 HPC Service (http://gw4.ac.uk/isambard/) operated by GW4 and the UK Met Office, and funded by EPSRC (EP/T022078/1).
The authors gratefully acknowledge the University of Bath's Research Computing Group~\cite{bath_research_2018} for their support in this work, in the LLZO molecular dynamics simulations.
Other simulations and analyses were carried out using the Data Management and Software Centre computing cluster at the European Spallation Source ERIC.
S.W.C. and B.J.M. acknowledge the support of the Faraday Institution through the CATMAT project (grant number FIRG016). 
B.J.M. acknowledges support from the Royal Society (UF130329 and URF\textbackslash R\textbackslash 191006).

\bibliographystyle{naturemag}
\bibliography{bib}
\let\addcontentsline\oldaddcontentsline

\onecolumngrid
\clearpage 
\twocolumngrid

\appendix
\renewcommand\thesection{SI.\Roman{section}}
\counterwithout{figure}{section}
\renewcommand\thefigure{SI.\arabic{figure}}
\setcounter{figure}{0}
\counterwithout{equation}{section}
\renewcommand\theequation{SI.\arabic{equation}}
\setcounter{equation}{0}
\counterwithout{table}{section}
\renewcommand\thetable{SI.\arabic{table}}
\setcounter{table}{0}
\pagenumbering{arabic} 
\renewcommand\thepage{SI.\arabic{page}}
\addtocontents{toc}{\protect\setcounter{tocdepth}{0}}

\title{Supplemental Material for ``\papertitle''}
\maketitle
\onecolumngrid

This document presents supplementary material for the manuscript ``\papertitle''.
It contains the following sections:
\begin{enumerate}
    % \item Evidence for the effect of the number of independent observations on the distribution of mean-squared displacements.
    \item The derivation of the covariance matrix in the long-time limit for freely diffusion particles.
    \item Quantification of the scaling behaviour in the posterior distribution for the diffusion coefficient with system size.
    \item Discussion of the skewed distribution of the estimated variance in the diffusion coefficient.
    \item Details of the LLZO molecular dynamics simulations.
    \item Discussion of the importance of only using non-diffusive simulation in the analysis.
    \item Details of the implementation of the model covariance method in the \textsc{kinisi} package.
\end{enumerate}
A repository containing the analysis and plotting code used to generate all results and figures in the main manuscript and this supplemental material document is available at \url{www.github.com/arm61/msd-errors}~\cite{mccluskey_github_2022}, under MIT (code) and CC BY-SA 4.0 (figures and text) licenses.
This repository includes a fully reproducible \code{showyourwork} workflow, which allows complete reproduction of the analysis, plotting of figures and compilation of the manuscripts.
The corresponding input datasets are openly available under the CC BY-SA 4.0 licence \cite{mccluskey_zenodo_2022,coles_llzo_zenodo_2022}.

\vspace{\columnsep}
\twocolumngrid

% \section{Effect of number of independent observations on the distribution of mean-squared displacement}
% \label{sec:chi2}
% %
% In the main text, we note that for a large enough number of observations, the observed MSD is a multivariate normal distribution. 
% This is made clear on inspection of the observed MSD at a single time interval (which should follow a normal distribution), where squared particle displacements, in the diffusive regime, are $\chi^2_k$-distributed, where $k$ is the number of independent observations of a given displacement. 
% Fig.~\ref{fig:freedom} shows that when there are sufficiently many independent observations of ${\left[\Delta\mathbf{r}_i(t)_\alpha\right]}^2$, each population of observable MSD values $p(\bar{s}^2_i)$ is normally distributed, i.e.\ as the number of observations increases the $\chi^2_k$-distribution tends towards a normal distribution.
% A single observation of \SI{16}{atoms} (i.e.\ where $t>0.5t_{\max}$), the distribution is not yet normally distributed.
% %
% \begin{figure*}
%     \centering
%     \includegraphics[width=\textwidth]{figures/freedom.pdf}
%     \caption{
%         A comparison between the observed mean-squared displacement (blue histograms) and the relevant $\chi^2_k$-distribution.
%         The simulations were run for \SI{128}{\second}, therefore there is only a single observation per atom per dimension, such that $k=N_{\mathrm{atoms}}d$; (a) $k = 2 \times 3$, (b) $k = 4 \times 3$, (c) $k = 8 \times 3$, (d) $k = 128 \times 3$.
%         % Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences. 
%     }
%     \label{fig:freedom}
%     \script{freedom.py}
% \end{figure*}
% %

\section{Derivation of the long-time limit covariance matrix for a system of freely diffusing particles.}
\label{sec:ran}
%
The main manuscript asserts that the long-time limit covariance matrix for a simulation of freely diffusing particles can be expressed as
%
\begin{equation}
  \Sigma^\prime\left[\oMSDi, \oMSDj\right]= \Sigma^\prime\left[\oMSDj, \oMSDi\right] =
  \var{\oMSDi} \frac{\nind{i}}{\nind{j}},\hspace{1em} \forall\,i \leq j,
  \label{equ:cvv_SI}
\end{equation}
%
where $\oMSDi$ is the observed mean-squared displacement (MSD) for time interval $i$ and $\nind{i}$ is the number of statistically independent observed squared displacements averaged over to compute the mean value (see Eqn.~2 in the main manuscript).

We must first consider the expected variance for the MSD at timestep $i$, $\var{\oMSDi}$, following the equivalent derivation by Smith and Gillan~\cite{smith_random_1996}, and then extend this to the covariance $\Sigma^\prime\left[\oMSDi, \oMSDj\right]$ to obtain the result given above.

For a single particle undergoing a one-dimensional random walk with step size $\kappa$, each step gives a displacement $h = \pm \kappa$.
After $n$ steps, the MSD, $\oMSDn$, is
%
\begin{equation}
  \begin{aligned}
    \oMSDn &= \left[\sum_i^n h_i\right]^2\\
           &= \sum_i^n\sum_j^n h_i h_j \\
           &= \sum_i^n h_i^2 + \sum_i^n\sum_{j\neq i}^n h_ih_j.
  \end{aligned}
\end{equation}
%
The expected MSD in the long-time limit, $\mathbb{E}(\oMSDn) = \moMSDn$, is obtained by averaging over all permutations of $h_i$ and $h_j$:
%
\begin{equation}
  \begin{aligned}
    \moMSDn &= \sum_i^n \left<h_i^2\right> + \sum_i^n\sum_{j\neq i}^n\left<h_i h_j\right>
  \end{aligned}
\end{equation}
%
For a random walk, the second term averages to zero for all $h_i$ and $h_j$, and
%
\begin{equation}
  \begin{aligned}
    \moMSDn &= \sum_i^n \left<h_i^2\right> \\
            &= n\kappa^2.
  \end{aligned}
  \label{equ:msd_SI}
\end{equation}
%
Hence the expected value for the mean-squared displacement increases linearly with the number of steps taken.
% First, we will outline a brief note on nomenclature, we will use $\MSD{n}$ to refer to the mean-squared displacement in a single simulation, i.e., this is the mean over all particles and observations of the given number of hops, $n$, $\Big\langle\MSD{n}\Big\rangle$ will be used for the average of the mean-squared displacements from many simulations. 
% We can think of these as a sample and population mean respectively, where $\MSD{n} \to \Big\langle\MSD{n}\Big\rangle$ with increasing numbers of simulations.

The variance in the observed MSD, $\var{\oMSDn}$ is given by the standard statistical formula
%
\begin{equation}
    \var{\oMSDn} = \left<{\left[\oMSDn - \moMSDn\right]}^2\right>,
\end{equation}
%
which can be expanded as
%
\begin{equation}
    \begin{aligned}
        \var{\oMSDn} &= \left<\oMSDn^2\right> - 2\moMSDn\left<\oMSDn\right> + \moMSDn^2,\\
                     &= \left<\oMSDn^2\right> - \moMSDn^{\,2}.
    \end{aligned}
\end{equation}
%

The first term can be expanded in terms of displacements $h$ as
%
\begin{equation}
  \left<\oMSDn^2\right> = \left<\sum_i^n\sum_j^n\sum_k^n\sum_l^n h_i h_j h_k h_l\right>
  \label{equ:big_av}
\end{equation}
%
which can be simplified by noting that $h_i$, $h_j$, $h_k$, and $h_l$ are uncorrelated when $i \neq j \neq k \neq l$, and the only terms that contribute to the average are those with:
%
\begin{itemize}
    \item[(a)] $i = j = k = l$;
    \item[(b)] $(i = j) \neq (k = l)$;
    \item[(c)] $(i = k) \neq (j = l)$;
    \item[(d)] $(i = l) \neq (j = k)$. 
\end{itemize}
%
Each of these conditions guarantees a positive product $h_ih_jh_kh_l$.

From (a) we obtain
%
\begin{equation}
    \left<\sum_i^nh_i^4\right> = n\kappa^4,
\end{equation}
%
and from (b), (c), and (d), which are equivalent, we obtain
%
\begin{equation}
    \left<\sum_i^n\sum_j^nh_i^2h_j^2\right> = (n\kappa^2)^2 = n^2\kappa^4.
\end{equation}
%
This gives
%
\begin{equation}
    \left<\oMSDn^2\right> = (3n^2 + n)\kappa^4,
\end{equation}
%
which, in the limit $n \to \infty$, approaches
%
\begin{equation}
    \left<\oMSDn^2\right> = 3n^2\kappa^4.
    \label{equ:infty}
\end{equation}
%
Combining this with the square of Eqn.~\ref{equ:msd_SI} gives the variance in the mean-squared displacement as
%
\begin{equation}
    \var{\oMSDn} = 3n^2\kappa^4 - n^2\kappa^4 = 2n^2\kappa^4,
\end{equation}
%
i.e., the variance increases quadratically with the number of steps taken.

The result above gives the variance in the mean squared displacement of a single particle if we consider a single time origin.
In practical simulations with more than one mobile particle we can obtain improved statistics by averaging over independent observed squared displacements; either averaging over equivalent mobile particles or averaging over statistically independent sections of the observed simulation trajectory, which we consider to be the case for mutually non-overlapping time intervals.
Now the variance in MSD is scaled by the usual number of independent contributing observations to give
%
\begin{equation}
    \var{\oMSDn} = \frac{2n^2\kappa^4}{\nind{n}},
\label{equ:der_var}
\end{equation}
%
where $\nind{n}$ is the total number of statistically independent squared displacements that contribute to $\oMSDi$.

The results in Eqns.~\ref{equ:msd_SI} \&~\ref{equ:der_var} can be extended to a $d$-dimensional lattice random walk, where for each step is along one of the $d$-dimensions, by combining the results for one dimension. 
This means that the MSD becomes 
%
\begin{equation}
    \moMSDn_{d} = \sum^d{\frac{n\kappa^2}{d}} = n\kappa^2,
\end{equation}
%
and the variance in the MSD becomes
%
\begin{equation}
    \var{\oMSDn}_d = \sum^d{\frac{2n^2\kappa^4}{d^2\nind{n}}} = \frac{2n^2\kappa^4}{d\nind{n}},
\end{equation}
%
where the $n$ in both Eqn.~\ref{equ:msd_SI} \&~\ref{equ:der_var} have been replaced with $n/d$ as the steps are spread (equally) over all dimensions.

This can be extended to consider the covariance between two different numbers of steps, $n$ and $n+m$, in the random walk where the expected MSDs will be
%
\begin{equation}
    \begin{aligned}
        \moMSDn &= n\kappa^2; \\
        \moMSDm &= (n+m)\kappa^2.
    \end{aligned}
\end{equation}
%
The covariance between these is defined as
%
\begin{equation}
    \Sigma \left[\oMSDn, \oMSDm \right] = \left<{\left[{\oMSDn - \moMSDn}\right] \left[{\oMSDm - \moMSDm}\right]}\right>, 
\end{equation}
%
which can be expanded,
%
\begin{equation}
    \begin{aligned}
        \Sigma \left[\oMSDn, \oMSDm \right] = \big\langle & \oMSDn\oMSDm - \oMSDn\moMSDm \\
        & - \moMSDn\oMSDm + \moMSDn\moMSDm\big\rangle,
    \end{aligned}
\end{equation}
%
and then reformulated to give
%
\begin{equation}
    \Sigma \left[\oMSDn, \oMSDm \right] = \left<\oMSDn\oMSDm\right> - \moMSDn\moMSDm,
    \label{equ:pair}
\end{equation}
%
where 
%
\begin{equation}
    \begin{aligned}
        \moMSDn\moMSDm &= \oMSDn\oMSDm \\
                       &= nd^2(n+m)\kappa^2 \\
                       &= n(n+m)\kappa^4
    \end{aligned}
\end{equation}
%
and, by analogy to Eqn.~\ref{equ:big_av},
%
\begin{equation}
    \left<\oMSDn\oMSDm\right> = \left<\sum_i^n\sum_j^n\sum_k^{n+m}\sum_l^{n+m} h_i h_j h_k h_l\right>,
\end{equation}
%
which we can rewrite as
%
\begin{equation}
    \begin{aligned}
        \left<\oMSDn\oMSDm\right> = \Bigg\langle & \sum_{i=1}^n\sum_{j=1}^n\sum_{k=1}^{n}\sum_{l=1}^{n} h_i h_j h_k h_l \\
        & + \sum_{i=1}^n\sum_{j=1}^n\sum_{k=1}^{n}\sum_{l=n+1}^{n+m} h_i h_j h_k h_l \\
        & + \sum_{i=1}^n\sum_{j=1}^n\sum_{k=n+1}^{n+m}\sum_{l=1}^{n} h_i h_j h_k h_l \\
        & + \sum_{i=1}^n\sum_{j=1}^n\sum_{k=n+1}^{n+m}\sum_{l=n+1}^{n+m} h_i h_j h_k h_l \Bigg\rangle.
    \end{aligned}
    \label{equ:vbig}
\end{equation}
%
The second and third terms in Eqn.~\ref{equ:vbig} tend to \num{0} as there is an equal probability of positive and negative displacements. 
This reduces Eqn.~\ref{equ:vbig} to 
%
\begin{equation}
    \begin{aligned}
        \left<\oMSDn\oMSDm\right> = & \Bigg\langle \sum_{i=1}^n\sum_{j=1}^n\sum_{k=1}^{n}\sum_{l=1}^{n} h_i h_j h_k h_l \Bigg\rangle \\
        & + \Bigg\langle \sum_{i=1}^n\sum_{j=1}^n\sum_{k=n+1}^{n+m}\sum_{l=n+1}^{n+m} h_i h_j h_k h_l \Bigg\rangle, 
    \end{aligned}
\end{equation}
%
and using Eqn.~\ref{equ:infty} gives
%
\begin{equation}
    \left<\oMSDn\oMSDm\right> = 3n^2\kappa^4 + \Bigg\langle \sum_{i=1}^n\sum_{j=1}^n\sum_{k=n+1}^{n+m}\sum_{l=n+1}^{n+m} h_i h_j h_k h_l \Bigg\rangle.
\end{equation}
%
We can rewrite this as
%
\begin{equation}
    \left<\oMSDn\oMSDm\right> = 3n^2\kappa^4 + \Bigg\langle \sum_{i=1}^n\sum_{j=1}^n h_i h_j \Bigg\rangle \Bigg\langle \sum_{k=n+1}^{n+m}\sum_{l=n+1}^{n+m} h_k h_l \Bigg\rangle,
\end{equation}
%
where the following holds,
%
\begin{equation}
    \begin{aligned}
        \left<\oMSDn\oMSDm\right> &= 3n^2\kappa^4 + n\kappa^2 m\kappa^2 \\
                                  &= 3n\kappa^4 + nm\kappa^4.
    \end{aligned}
\end{equation}
%
Putting this result into Eqn.~\ref{equ:pair} allows the covariance to be written as
%
\begin{equation}
    \begin{aligned}
    \Sigma^\prime \left[\oMSDn, \oMSDm \right] &= 3n^2\kappa^4 + nm\kappa^4 - n(n+m)\kappa^4 \\
                                        &= 3n^2\kappa^4 - n^2\kappa^4 = 2n^2\kappa^4,
    \end{aligned}
    \label{equ:cov_der}
\end{equation}
where we use the $\Sigma^\prime$ notation to identify that this is in the long-time limit.

In this case, the covariance depends only on the number of overlapping points, $n$, between the two time intervals. 
We can easily rationalise this by accepting that for a random walk any non-overlapping point will be completely uncorrelated and therefore have a covariance of \num{0}. 
Similar to the case for the variance, the covariance derived in Eqn.~\ref{equ:cov_der} is that for a single particle at a single time origin. 
The number of independent observed squared displacements for a given covariance should be the minimum number of shared independent observed squared displacements between the two time intervals, which is $\nind{n+m}$. 
Therefore, scaled by the number of contributing independent observations, the covariance, in the long-time limit, is
%
\begin{equation}
    \Sigma^\prime \left[\oMSDn, \oMSDm \right] = \frac{2n^2\kappa^4}{\nind{n+m}}.
\end{equation}
%
Similar to the MSD and the variance, the covariance can be written for $d$-dimensions as 
%
\begin{equation}
    \Sigma^\prime \left[\oMSDn, \oMSDm \right] = \frac{2n^2\kappa^4}{d\nind{n+m}}.
\end{equation}
%
The covariance can be calculated directly from the variance by recognising that both depend on the number of overlapping points, $n$, as follows
%
\begin{equation}
    \Sigma^\prime \left[\oMSDn, \oMSDm \right] = \var{\oMSDn}\frac{\nind{n}}{\nind{n+m}}.
\end{equation}
%
This is then rewritten in terms of $i$ and $j$ to give, Eqn.~\ref{equ:cvv_SI}.

Using the equivalence of $2d\D t \equiv n\kappa^2$~\cite{howard_reports_1964}, Eqns.~\ref{equ:msd_SI} \&~\ref{equ:cvv_SI} can be rewritten in terms of $t$ (or $t_1$ and $t_2$) and the diffusion coefficient, for any dimensionality of lattice random walk,
%
\begin{equation}
    \oMSDs{t} = 2d \D t,
    \label{equ:der_msd_fick}
\end{equation}
%
and 
%
\begin{equation}
    \Sigma^\prime \left[\oMSDs{t_1}, \oMSDs{t_2} \right] = 8d{(\D)}^2 t_1^2\frac{N^\prime(t_2)}{N^\prime(t_1)},\hspace{1em} \forall\,t_1 \leq t_2.
\end{equation}
%

\section{Scaling of the estimation of $\prob{\Dest}$}
\label{sec:scaling}
%
The true sample variance from \num{512} diffusion coefficient estimates, $\var{\Dest}$, using either the model covariance method, outlined in the main text, the independent method (using estimated variances but ignoring the need for the full covariance matrix, analogous to WLS), the independent \& identical method (analogous to OLS), and using the full numerical covariance matrix as a function of simulation length and number of atoms are compared in Fig.~\ref{fig:stat_eff}.

All estimation methods have the same power law exponent, $a\approx-1$, when the number of atoms is varied, due to the number of observations linearly scaling with number of atoms. 
This is the observation of the denominator ($n - 1$, where $n$ is the sample size) in the standard variance formula. 
The numerical and model covariance methods show the same relationship when the maximum simulation length is varied.
However, the independent and independent \& identical approaches have values for $a$ of \numrange{0.1}{0.3}.
Doubling the simulation length doubles the number of observations in the first half of the time range, while, the second half contains only new samples.
The full off-diagonal values of the covariance matrix are necessary to correctly account for these new samples (as the off-diagonal elements link between the different time intervals), leading to the poor scaling of the WLS and OLS analogues that is observed. 
%
% \begin{table*}
%     \centering
%     \caption{Parameters for the scaling of the variance of the diffusion coefficient for truly random walks, from Figs.~\ref{fig:stat_eff}a \& b.}
%     \label{tab:ipl_true}
%     \begin{tabular}{c  c | rclcl | c | c | c}
%          & & numerical & This & Independent & Independent \& Identical \\
%         \hline
%         \multirow{2}{*}{$1/N_{\mathrm{atoms}}$} & $a$ & \num{-9.81 \pm 0.06e-1} & \num{-9.03 \pm 0.18e-1} & \num{-9.95 \pm 0.12e-1} & \num{-9.78 \pm 0.14e-1} \\
%           & $b$/atoms$^{-a}$ & \num{1.42 \pm 0.03e-3} & \num{1.36 \pm 0.08e-3} & \num{7.27 \pm 0.26e-3} & \num{2.15 \pm 0.09e-2} \\
%         \multirow{2}{*}{$1/\Delta t_{\max}$} & $a$ & \num{-1.20 \pm 0.01e-0} & \num{-1.05 \pm 0.02e-0} & \num{-5.69 \pm 0.17e-1} & \num{-8.33 \pm 1.80e-2} \\ 
%         & $b$ & \num{3.78 \pm 0.15e-3} & \num{2.74 \pm 0.14e-3} & \num{8.99 \pm 0.51e-4} & \num{2.81 \pm 0.24e-4} \\ 
%     \end{tabular}
% \end{table*}
%
%
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/stat_eff.pdf}
    \caption{
        Scaling of statistical efficiency of different approaches;
        the decrease in $\sigma^2[\hat{\mu}(\Dest_\alpha)]$ as a function of number of atoms (a \& c) and maximum simulation length (b \& d).
        The solid lines indicate the fitted power law relationship for each of the numerical results (blue), the model covariance approach (green), where the variances are found by rescaling but the data considered independent (orange), and where the variances are considered to be identical and independent (pink). 
        For all except the numerical result, the variance was found from \num{512} determinations of the mean diffusion coefficient from equivalent simulations of walks of \SI{128}{atoms} for \num{128} steps and a step size of \num{1}.
        % Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences. 
        }
    \label{fig:stat_eff}
    \script{stat_eff.py}
\end{figure}
%

The scaling in the variance that is estimated from each individual simulation using the model covariance method also shows different behaviour between changing the number of atoms and changing the maximum simulation length (Fig.~\ref{fig:efficiency}).
Initially, we compare the scaling behaviour of the estimated variance with the numerically determined variance (Figs.~\ref{fig:efficiency}a \&~\ref{fig:efficiency}b), where the power law is similar regardless of how the system changes in size. 
The width of the distribution of $\prob{\varest{\Dest}}$ (the error bars in Figs.~\ref{fig:efficiency}a \&~\ref{fig:efficiency}b) also show power law behaviour (Figs~\ref{fig:efficiency}c \&~\ref{fig:efficiency}d). 
In this case, the exponent is different if the number of atoms is changed ($a \approx -2$) compared to the maximum simulation length ($a \approx -1$). 
The precision of the variance estimate will improve more if the number of atoms is doubled rather than if the simulation is run for twice as long. 
When the number of atoms is doubled, the number of contributing independent observations also doubles for all time intervals but when the simulation is doubled in length, the number of contributing independent observations is only doubled up to the original simulation length.
%
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/efficiency.pdf}
    \caption{
        Scaling of estimated variance (a \& b), estimated variance precision (c \& d), and estimated variance accuracy (e \& f) from the method outlined in the main text;
        Green points indicate mean values for the estimated variance, where error bars are present they indicate a \SI{95}{\percent} confidence interval. 
        Blue points indicate the variance found from numerical simulations.
        The precision is quantified with the difference between the upper and lower bounds of the confidence interval and the accuracy from the bias compared to the numerical result.
        The variance was found from \num{512} determinations of the diffusion coefficient from equivalent simulations of walks. 
        When $N_{\mathrm{atoms}}$ was varied $t_{\max}$ was \SI{128} steps and when $t_{\max}$ was varied $N_{\mathrm{atoms}}$ \SI{128}{atoms}; for all simulations the step size was \num{1}.
        % Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences. 
        }
    \label{fig:efficiency}
    \script{efficiency.py}
\end{figure}
%

\section{Skewness of distribution of $\varest{\Dest}$}
\label{sec:skew}
%
The skew observed in the estimated variance for the diffusion coefficient when using the model covariance matrix (Fig.~\ref{fig:random_walk}d) arises from differences between the numerical and model covariance matrices
If the numerical covariance matrix was used in place of the model covariance matrix, the resulting distribution of $\varest{\Dest}$ is not be skewed (Fig.~\ref{fig:true_cov}).
Additionally, the distribution of $\Dest$ is no longer broader than the numerical distribution, as is the case in Fig~\ref{fig:random_walk}c, which arises from the model covariance matrix (and analytical covariance matrix) overestimating the variance at low $t$ (Fig.~\ref{fig:covariances}), due to the long-time limit assumption made in its derivation.
%
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/true_cov.pdf}
    \caption{
        The cause of the skewed variance estimation when the model covariance matrix is used; 
        distributions of $\mu(\Dest)$ (a) and $\sigma^2(\Dest)$ (b) determined using the observed mean-squared displacements and the numerical covariance matrix (from \num{1024} equivalent simulations), showing the unskewed distribution of variances, the numerical distribution of $\mu(\Dest)$ is shown with the blue solid line and the sample variance in (a) is shown with the vertical green line in (b).
        Values found using random walks of \SI{128}{atoms} for \num{128} steps and a step size of \num{1}. 
        % Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences. 
        }
    \label{fig:true_cov}
    \script{true_cov.py}
\end{figure}
%

\section{LLZO simulations}
%
Classical molecular dynamics were run using the \textsc{metalwalls} code~\cite{marin_metalwalls_2020}. 
The DIPPIM polarisable ion force field parameterised by Burbano \emph{et al.}~\cite{burbano_sparse_2016} was used due to its ability to account accurately for the effect of ion polarisability on diffusion~\cite{wilson_polarization_1993,burbano_sparse_2016}. 
Simulations were run for the cubic phase of LLZO in NVT ensemble at \SI{1000}{\kelvin} and were extended to a length of \SI{2}{\nano\second} with a \SI{2}{\femto\second} timestep. 
Temperature was maintained by means of Nos\'{e}-Hoover thermostat with a relaxation time of \SI{121}{\femto\second} (5000 $\hbar / E_{h}$)~\cite{nose_unified_1984,hoover_canonical_1985,martyna_nose_1992}. 
As in the prior study of Burbano \emph{et al.}~\cite{burbano_sparse_2016}, simulations were performed using $2 \times 2 \times 2$ supercells with \SI{1536}{atoms} in a manner similar to Ref.~\citenum{burbano_sparse_2016} with the system being in the perfect crystalline form at the beginning.

\section{Effect of non-diffusive simulation}
\label{sec:nondiff}
%
In the modelling of the diffusive process, it is necessary to exclude simulation not in the diffusive regime, where the atoms are ``caged'' by their neighbours. 
The model covariance matrix has been derived assuming the diffusive regime (the Einstein relation implicitly assumes diffusion), therefore only diffusion should be included in our analysis. 
The non-diffusive regime typically dominates on short-time scales, giving way to diffusion at larger time scales. 
Fig.~\ref{fig:true_llzo} investigates the effect of using starting the Bayesian linear model at increasing initial values of $t$ (i.e.\ the value of $t$ from which the linear model was applied), in essence reducing the percentage of simulation from non-diffusive particles in the analysis. 
Using simulation from non-diffusive particles leads to a bias in the estimation of the variance of $\prob{\Dest}$, which is reduced when less of the non-diffusive regime is used. 
%
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/true_llzo.pdf}
    \caption{
        A comparison of the effect of different starting times for the diffusive regime \SIlist[list-final-separator = {, and }]{0;2;4;6;8;10;15;20}{\pico\second} of the LLZO systems on $\varest{\Dest}$.
        % Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences. 
        }
    \label{fig:true_llzo}
    \script{true_llzo.py}
\end{figure*}
%

To avoid the spurious effect of non-diffusive data, this regime must be excluded. 
It is useful, in the first instance, to overestimate excluded regime and assess if the resulting estimated variance is small enough to draw conclusions about the data or perform hypothesis testing.  
A balance can be struck between the start of the diffusive regime (having this as high as possible to ensure minimal non-diffusive data) and the size of the resulting variance.

\section{Numerical implementation in \textsc{kinisi}}
\label{sec:implementation}
%
We have implemented the method outlined in the main text in the open-source Python package \textsc{kinisi}~\cite{mccluskey_kinisi_2022}. 
Here, we will discuss both \textsc{kinisi}'s broader implementation of the method and its application in the validation examples.

First, to find the maximum number of observations of squared displacements at each time interval, \textsc{kinisi} uses overlapping sliding window sampling. 
The maximum number of observations for a given time interval, $t$, is $N_{\mathrm{atoms}} \times (N_{t} - i)$ displacements, where $N_{\mathrm{atoms}}$ is the number of atoms, $N_{t}$ is the total number of timesteps, and $i$ is the index of the timestep (starting with \num{1} for the shortest timestep). 
The independent variance is estimated by rescaling the observed overlapping variance in the squared displacements by the number of non-overlapping squared displacements, $\nind{i} = N_{\mathrm{atoms}} \times N_{t} / i$, as shown in Eqn.~\ref{equ:varestMSD}. 
The population mean for the squared displacements and the independent variances are then stored.

Eqn.~\ref{equ:cvv_SI} defines the parameterisation of the covariance matrix from the variances and the number of independent observations. 
The covariance matrix is only constructed for values of $t$ where the particles are considered to be diffusing, which can be controlled manually by the user. 
In this work, the particles were assumed to be diffusing for $t>\SI{4}{\second}$ for the random walks and \SI{10}{\pico\second} for the LLZO system in Fig.~\ref{fig:diffusion}. 
The LLZO system was investigated at a range of starting times in Fig.~\ref{fig:true_llzo}.

To obtain an initial guess for the gradient and intercept of the linear model that describes the observed MSD, \textsc{kinisi} uses ordinary least squares. 
This initial guess is then used as a starting point for the minimisation of the negative maximum a posteriori (the maximum of the posterior distribution in Eqn.~\ref{equ:bayes}), with the prior probability that the diffusion coefficient is greater than or equal to zero~\cite{broyden_convergence_1970,fletcher_new_1970,goldfarb_family_1970,shanno_conditioning_1970}.
The log-likelihood calculation (Eqn.~\ref{equ:loglike}) makes use of the Moore-Penrose generalisation of the inverse of a Hermitian matrix~\cite{moore_on_1920,bjerhammar_application_1951,penrose_generlized_1955}.

To sample the joint posterior probability distribution of the linear model, \textsc{kinisi} uses the \textsc{emcee} package~\cite{foremanmackey_emcee_2019} was used, which implements Goodman and Weare's affine invariant Markov chain Monte Carlo ensemble sampler~\cite{goodman_ensemble_2010}. 
A physically-informed prior probability that $\D > 0$ is imposed on all analysis.
The sampling is performed with \num{32} walkers, for \num{1500} steps, the first \num{500} steps are removed as a burn-in period, finally, the sampled chains are thinned, such that only every \num{10}th value is used. 
This gives \num{3200} points sampled in the posterior distribution, these points can be histogrammed (as in Fig~\ref{fig:random_walk}) and summary statistics taken.

\end{document}

### 2023-05-29 src/code/llzo/many_runs.py

@@ -1,0 +1,98 @@
+import MDAnalysis as mda
+import numpy as np
+from kinisi.parser import MDAnalysisParser
+from kinisi.diffusion import MSDBootstrap
+from tqdm import tqdm
+
+n = snakemake.params['n']
+start_diff = snakemake.params['start_diff']
+
+def universe_slice(u: mda.Universe, data_file: str, slice_start: int, slice_end: int) -> mda.Universe:
+    """
+    Concatenate a list of Universes.
+
+    :param u_list: List of Universe objects
+    :param data_file: File path for data
+    :param slice_start: Step to start the slice from
+    :return: A single flattened Universe
+    """
+    j = 0
+    trajectory = np.zeros((len(u.trajectory[slice_start:slice_end]), len(u.atoms), 3))
+    for t in tqdm(u.trajectory[slice_start:slice_end]):
+        trajectory[j] = t.positions
+        j += 1
+    shorter_universe = mda.Universe(data_file, trajectory)
+    for t in tqdm(shorter_universe.trajectory):
+        for i in shorter_universe.atoms:
+            if i.type == 'LI':
+                i.type = 'Li'
+            if i.type == 'Li1':
+                i.type = 'Li'
+            if i.type == 'Li2':
+                i.type = 'Li'
+            if i.type == 'L':
+                i.type = 'La'
+            if i.type == 'Z':
+                i.type = 'Zr'
+        t.dimensions = np.array([26.175106662244392197, 26.175106662244392197, 26.175106662244392197, 90, 90, 90])
+    return shorter_universe
+
+file = f'src/data/llzo/traj{n}.xyz'
+
+
+u = mda.Universe(file, file)
+uu = universe_slice(u, file, 0, 500)
+step_skip = 100  # sampling rate
+timestep = 5.079648e-4  # ps
+da_params = {'specie': 'Li', 'time_step': timestep, 'step_skip': step_skip}
+p = MDAnalysisParser(uu, **da_params)
+
+ngp = np.zeros((8, 4, p.delta_t.size))
+no = np.zeros((8, 4, p.delta_t.size))
+dt = np.zeros((8, 4, p.delta_t.size))
+msd = np.zeros((8, 4, p.delta_t.size))
+msd_true = np.zeros((8, 4, p.delta_t.size))
+msd_std = np.zeros((8, 4, p.delta_t.size))
[…]

### 2023-05-29 src/code/llzo/true_ls.py

@@ -1,0 +1,62 @@
+# Copyright (c) Andrew R. McCluskey
+# Distributed under the terms of the MIT License
+# author: Andrew R. McCluskey (arm61)
+
+import numpy as np
+from scipy.stats import linregress
+from scipy.linalg import pinvh
+from emcee import EnsembleSampler
+from scipy.optimize import minimize
+from kinisi.diffusion import _straight_line
+
+np.random.seed(1)
+
+start_diff = snakemake.params['start_diff']
+
+timestep = np.load(f'src/data/llzo/diffusion_0_{start_diff}.npz')['dt'][0, 0]
+no = np.load(f'src/data/llzo/diffusion_0_{start_diff}.npz')['no'][0, 0]
+d = np.zeros((16, 8, 4, timestep.size))
+for i in range(0, 16, 1):
+    d[i] = np.load(f'src/data/llzo/diffusion_{i}_{start_diff}.npz')['msd_true']
+
+true_mean = d.reshape(-1, d.shape[-1]).mean(0)
+max_ngp = np.argwhere(timestep > start_diff)[0][0]
+rcov = np.cov(d.reshape(-1, d.shape[-1]).T)[max_ngp:, max_ngp:]
+
+_, logdet = np.linalg.slogdet(rcov)
+logdet += np.log(2 * np.pi) * d.reshape(-1, d.shape[-1]).mean(0)[max_ngp:].size
+inv = pinvh(rcov)
+
+def log_likelihood(theta: np.ndarray) -> float:
+    """
+    Get the log likelihood for multivariate normal distribution.
+    :param theta: Value of the gradient and intercept of the straight line.
+    :return: Log-likelihood value.
+    """
+    if theta[0] < 0:
+        return -np.inf
+    model = _straight_line(timestep[max_ngp:], *theta)
+    diff = (model - d.reshape(-1, d.shape[-1]).mean(0)[max_ngp:])
+    logl = -0.5 * (logdet + np.matmul(diff.T, np.matmul(inv, diff)))
+    return logl
+
+ols = linregress(timestep[max_ngp:], d.reshape(-1, d.shape[-1]).mean(0)[max_ngp:])
+
+
+def nll(*args):
+    """
+    General purpose negative log-likelihood.
+    Returns:
+        (:py:attr:`float`): Negative log-likelihood.
+    """
+    return -log_likelihood(*args)
+
+
[…]

### 2023-05-29 src/code/random_walks/efficiency.py

@@ -1,0 +1,110 @@
+import numpy as np
+from scipy.optimize import curve_fit
+from uravu.distribution import Distribution
+from uravu.relationship import Relationship
+
+jump = 1
+
+xaxis = np.array([16, 32, 64, 128, 256, 512, 1024])
+
+kinisi_atoms = np.zeros((xaxis.size, 512, 3200))
+weighted_atoms = np.zeros((xaxis.size, 512, 3200))
+ordinary_atoms = np.zeros((xaxis.size, 512, 3200))
+kinisi_length = np.zeros((xaxis.size, 512, 3200))
+weighted_length = np.zeros((xaxis.size, 512, 3200))
+ordinary_length = np.zeros((xaxis.size, 512, 3200))
+true_d_atoms = np.zeros((xaxis.size, 3200))
+true_d_length = np.zeros((xaxis.size, 3200))
+for i, ii in enumerate(xaxis):
+    length = 128
+    true_d_atoms[i] = np.load(f"src/data/random_walks/numerical/D_{jump}_{ii}_{length}.npz")["diff_c"]
+    file_open = np.load(f"src/data/random_walks/kinisi/rw_{jump}_{ii}_{length}_s512.npz")
+    kinisi_atoms[i] = file_open['diff_c']
+    file_open = np.load(f"src/data/random_walks/weighted/rw_{jump}_{ii}_{length}_s512.npz")
+    weighted_atoms[i] = file_open['diff_c']
+    file_open = np.load(f"src/data/random_walks/ordinary/rw_{jump}_{ii}_{length}_s512.npz")
+    ordinary_atoms[i] = file_open['diff_c']
+for i, ii in enumerate(xaxis):
+    atoms = 128
+    true_d_length[i] = np.load(f"src/data/random_walks/numerical/D_{jump}_{atoms}_{ii}.npz")["diff_c"]
+    file_open = np.load(f"src/data/random_walks/kinisi/rw_{jump}_{atoms}_{ii}_s512.npz")
+    kinisi_length[i] = file_open['diff_c']
+    file_open = np.load(f"src/data/random_walks/weighted/rw_{jump}_{atoms}_{ii}_s512.npz")
+    weighted_length[i] = file_open['diff_c']
+    file_open = np.load(f"src/data/random_walks/ordinary/rw_{jump}_{atoms}_{ii}_s512.npz")
+    ordinary_length[i] = file_open['diff_c']
+
[…]

### 2023-05-29 src/code/random_walks/glswlsols.py

@@ -1,0 +1,136 @@
+# Copyright (c) Andrew R. McCluskey
+# Distributed under the terms of the MIT License
+# author: Andrew R. McCluskey (arm61)
+
+import numpy as np
+from scipy.stats import linregress
+from scipy.linalg import pinvh
+from emcee import EnsembleSampler
+from scipy.optimize import minimize
+from kinisi.matrix import find_nearest_positive_definite
+from kinisi.diffusion import _straight_line
+
+np.random.seed(1)
+
+
+def nll(*args):
+    """
+    General purpose negative log-likelihood.
+    Returns:
+        (:py:attr:`float`): Negative log-likelihood.
+    """
+    return -log_likelihood(*args)
+
+
+atoms = int(snakemake.params['atoms'])
+length = int(snakemake.params['length'])
+jump = int(snakemake.params['jump'])
+
+true_msd = np.load(f'src/data/random_walks/numerical/rw_1_{atoms}_{length}_s4096.npz')['data'][:, 4:]
+true_cov = np.cov(true_msd.T)
+timestep = np.arange(4, length, 1)
+
+y = true_msd.T
+A = np.array([timestep, np.ones(timestep.size)]).T
+W = true_cov
+V = np.diag(true_cov.diagonal())
+
+g1 = np.matmul(np.linalg.inv(np.matmul(A.T, np.matmul(np.linalg.pinv(W), A))),
+               np.matmul(A.T, np.matmul(np.linalg.pinv(W), y)))[0] / 6
+g2 = np.matmul(np.linalg.inv(np.matmul(A.T, np.matmul(np.linalg.pinv(V), A))),
+               np.matmul(A.T, np.matmul(np.linalg.pinv(V), y)))[0] / 6
+g3 = np.matmul(np.linalg.inv(np.matmul(A.T, A)), np.matmul(A.T, y))[0] / 6
+
+W = find_nearest_positive_definite(W)
+
+_, logdet = np.linalg.slogdet(W)
+logdet += np.log(2 * np.pi) * true_msd[0].size
+inv = pinvh(W)
+
+
+def log_likelihood(theta: np.ndarray) -> float:
+    """
+    Get the log likelihood for multivariate normal distribution.
+    :param theta: Value of the gradient and intercept of the straight line.
+    :return: Log-likelihood value.
+    """
+    if theta[0] < 0:
+        return -np.inf
+    model = _straight_line(timestep, *theta)
[…]

### 2023-05-29 src/code/random_walks/kinisi_rw.py

@@ -1,0 +1,46 @@
+# Copyright (c) Andrew R. McCluskey
+# Distributed under the terms of the MIT License
+# author: Andrew R. McCluskey (arm61)
+
+import numpy as np
+from kinisi.diffusion import MSDBootstrap
+from tqdm import tqdm
+from random_walk import get_disp3d, walk
+
+jump = snakemake.params['jump']
+atoms = snakemake.params['atoms']
+length = int(snakemake.params['length'])
+size = int(snakemake.params['n'])
+
+timestep = np.arange(1, length + 1, 1, dtype=int)
+data = np.zeros((size, timestep.size, 4))
+covariance = np.zeros((size, timestep.size - 4, timestep.size - 4))
+npd_covariance = np.zeros((size, timestep.size - 4, timestep.size - 4))
+n_o = np.zeros((size, timestep.size))
+diff_c = np.zeros((size, 3200))
+intercept = np.zeros((size, 3200))
+
+for seed in tqdm(range(size)):
+    rng = np.random.RandomState(seed)
+    np.random.seed(seed)
+    disp_3d, n_samples = get_disp3d(walk, length, atoms, jump_size=jump, seed=rng)
+
+    diff = MSDBootstrap(timestep, disp_3d, n_samples, random_state=rng, progress=False)
+    diff.diffusion(use_ngp=False, random_state=rng, dt_skip=4, progress=False)
+    diff_c[seed] = diff.gradient.samples / 6
+    intercept[seed] = diff.intercept.samples
+    data[seed, :, 0] = diff.dt
+    data[seed, :, 1] = diff.n
+    data[seed, :, 2] = diff.s
+    data[seed, 4:, 3] = diff._model_v
+    npd_covariance[seed] = diff._npd_covariance_matrix
+    covariance[seed] = diff.covariance_matrix
+    n_o[seed] = diff._n_o
+
+np.savez(f'src/data/random_walks/kinisi/rw_1_{atoms}_{length}_s{size}.npz',
+         diff_c=diff_c,
+         intercept=intercept,
+         data=data,
+         covariance=covariance,
+         npd_covariance=npd_covariance,
+         n_o=n_o)

### 2023-05-29 src/code/random_walks/longtime.py

@@ -1,0 +1,56 @@
+# Copyright (c) Andrew R. McCluskey
+# Distributed under the terms of the MIT License
+# author: Andrew R. McCluskey (arm61)
+
+import numpy as np
+from scipy.stats import multivariate_normal, linregress
+from scipy.optimize import minimize
+from emcee import EnsembleSampler
+from kinisi.matrix import find_nearest_positive_definite
+from uravu.utils import straight_line
+
+atoms = 1024
+lengths = [128, 256, 52, 1024, 2048, 4096]
+jumps = [1, 8, 32]
+
+data = np.zeros((2, len(jumps), len(lengths), 3200))
+
+for j, jump in enumerate(jumps):
+    for k, length in enumerate(lengths):
+        simulation = np.load(f'../data/numerical_{jump}_{atoms}_{length}.npz')['data']
+        covariance = np.cov(simulation.T)
+        timestep = np.linspace(1, length, length, dtype=int)[1:-1]
+
+        covariance = find_nearest_positive_definite(covariance)
+        mv = multivariate_normal(simulation.mean(0), covariance, allow_singular=True)
+
+        def log_likelihood(theta):
+            """
+            Get the log likelihood for multivariate normal distribution.
+            Args:
+                theta (:py:attr:`array_like`): Value of the gradient and intercept of the straight line.
+            Returns:
+                (:py:attr:`float`): Log-likelihood value.
+            """
+            model = straight_line(timestep, *theta)
+            logl = mv.logpdf(model)
+            return logl
+
+        ols = linregress(timestep, simulation.mean(0))
+
+        def nll(*args):
+            """
+            General purpose negative log-likelihood.
+            Returns:
+                (:py:attr:`float`): Negative log-likelihood.
+            """
+            return -log_likelihood(*args)
+
+        max_likelihood = minimize(nll, np.array([ols.slope, ols.intercept])).x
+        pos = max_likelihood + max_likelihood * 1e-3 * np.random.randn(32, max_likelihood.size)
+        sampler = EnsembleSampler(*pos.shape, log_likelihood)
+        sampler.run_mcmc(pos, 1500, progress=True, progress_kwargs={'desc': "Likelihood Sampling"})
+        flatchain = sampler.get_chain(flat=True, discard=500)
+        data[0, j, k] = flatchain[:, 0] / 6
+        data[1, j, k] = flatchain[:, 1]
+np.save('src/data/longtime.npy', data)

### 2023-05-29 src/code/random_walks/nonindependent_sampling.py

@@ -1,0 +1,46 @@
+# Copyright (c) Andrew R. McCluskey
+# Distributed under the terms of the MIT License
+# author: Andrew R. McCluskey (arm61)
+
+import numpy as np
+from sklearn.utils import resample
+from tqdm import tqdm
+from uravu.distribution import Distribution
+from random_walk import walk, anticorrelated_walk, get_disp3d
+from kinisi.diffusion import MSDBootstrap
+
+jump = snakemake.params['jump']
+atoms = snakemake.params['atoms']
+length = int(snakemake.params['length'])
+correlated = snakemake.params['correlation']
+size = int(snakemake.params['n'])
+
+timestep = np.linspace(1, length, length, dtype=int)
+res = np.zeros((size, length - 2))
+res_e = np.zeros((size, length - 2))
+for seed in tqdm(range(size)):
+    data = np.zeros((length - 2, 3))
+
+    rng = np.random.RandomState(seed)
+    np.random.seed(seed)
+    if correlated == 'correlated':
+        disp_3d = get_disp3d(anticorrelated_walk, length, atoms, seed=rng)
+    else:
+        disp_3d = get_disp3d(walk, length, atoms, seed=rng)
+
+    diff = MSDBootstrap(timestep, disp_3d, random_state=rng, progress=False)
+
+    i = 0
+    for n in range(len(diff.dt))[1:]:
+        r2_sum = np.sum(diff._displacements[n]**2, axis=2)
+        resampled = [np.mean(resample(r2_sum.flatten())) for j in range(1000)]
+        distro = Distribution(resampled)
+        while (not distro.normal):
+            distro.add_samples([np.mean(resample(r2_sum.flatten())) for j in range(100)])
+        res[seed, i] = distro.n
+        res_e[seed, i] = np.var(distro.samples, ddof=1)
+        i += 1
+
+np.savez(f'src/data/random_walks/nonindependent/{correlated}/rw_{int(jump)}_{atoms}_{length}_s{size}.npz',
+         res_msd=res,
+         res_var=res_e)

### 2023-05-29 src/code/random_walks/numerical_rw.py

@@ -1,0 +1,28 @@
+# Copyright (c) Andrew R. McCluskey
+# Distributed under the terms of the MIT License
+# author: Andrew R. McCluskey (arm61)
+
+import numpy as np
+from tqdm import tqdm
+from random_walk import walk
+
+jump = snakemake.params['jump']
+atoms = snakemake.params['atoms']
+length = int(snakemake.params['length'])
+size = int(snakemake.params['n'])
+
+timestep = np.arange(1, length + 1, 1, dtype=int)
+k = np.zeros((size, timestep.size))
+for seed in tqdm(range(size)):
+    rng = np.random.RandomState(seed)
+
+    cum_steps = walk(atoms, timestep, seed=rng, jump_size=jump)
+
+    for i, n in enumerate(timestep):
+        disp = np.concatenate([cum_steps[:, np.newaxis, i],
+                               np.subtract(cum_steps[:, i + 1:], cum_steps[:, :-(i + 1)])],
+                              axis=1)
+        d2 = np.sum(disp**2, axis=-1)
+        k[seed, i] = np.mean(d2.flatten())
+
+np.savez(f'src/data/random_walks/numerical/rw_1_{atoms}_{length}_s{size}.npz', data=k)

### 2023-05-29 src/code/random_walks/ordinary_rw.py

@@ -1,0 +1,72 @@
+# Copyright (c) Andrew R. McCluskey
+# Distributed under the terms of the MIT License
+# author: Andrew R. McCluskey (arm61)
+
+import numpy as np
+from scipy.stats import linregress
+from scipy.linalg import pinvh
+from scipy.optimize import minimize
+from emcee import EnsembleSampler
+from kinisi.diffusion import _straight_line
+from tqdm import tqdm
+
+
+jump = snakemake.params['jump']
+atoms = snakemake.params['atoms']
+length = int(snakemake.params['length'])
+size = int(snakemake.params['n'])
+
+gradient = np.zeros((size, 3200))
+intercept = np.zeros((size, 3200))
+data = np.load(f'src/data/random_walks/kinisi//rw_1_{atoms}_{length}_s{size}.npz')['data']
+
+for seed in tqdm(range(size)):
+    rng = np.random.RandomState(seed)
+    np.random.seed(seed)
+
+    x = data[seed, 4:, 0]
+    y = data[seed, 4:, 1]
+    i_matrix = np.eye(y.size)
+    add = np.log(2 * np.pi) * y.size
+
+    def log_likelihood(theta: np.ndarray) -> float:
+        """
+        Get the log likelihood for multivariate normal distribution.
+        :param theta: Value of the gradient and intercept of the straight line.
+        :return: Log-likelihood value.
+        """
+        if theta[0] < 0:
+            return -np.inf
+        model = x * theta[0] + theta[1]
+        diff = (model - y)
+        var = diff.var()
+        logdet = np.log(var ** y.size)
+        logdet += np.log(2 * np.pi) * y.size
+        inv = i_matrix / var
+        logl = -0.5 * (logdet + np.matmul(diff.T, np.matmul(inv, diff)))
+        return logl
+
+    ols = linregress(x, y)
+    slope = ols.slope
+    c = 1e-20
+    if slope < 0:
+        slope = 1e-20
+
+    def nll(*args) -> float:
+        """
+        General purpose negative log-likelihood.
+        :return: Negative log-likelihood
+        """
+        return -log_likelihood(*args)
+
+    max_likelihood = minimize(nll, np.array([slope, c])).x
+    pos = max_likelihood + max_likelihood * 1e-3 * np.random.randn(32, max_likelihood.size)
+    sampler = EnsembleSampler(*pos.shape, log_likelihood)
[…]

### 2023-05-29 src/code/random_walks/random_walk.py

@@ -1,0 +1,62 @@
+# Copyright (c) Andrew R. McCluskey
+# Distributed under the terms of the MIT License
+# author: Andrew R. McCluskey (arm61)
+
+from typing import Union, List
+import numpy as np
+
+
+def walk(atoms: int,
+         timesteps: np.ndarray,
+         jump_size: int = 1,
+         seed: Union[np.random.mtrand.RandomState, None] = None) -> np.ndarray:
+    """
+    Perform a random walk.
+
+    :param atoms: number of atoms
+    :param timesteps: the timestep values
+    :param jump_size: size of jump
+    :param seed: random seed source
+    :return: cumulative sum of steps for walk
+    """
+    possible_moves = np.zeros((6, 3))
+    j = 0
+    for i in range(0, 6, 2):
+        possible_moves[i, j] = jump_size
+        possible_moves[i + 1, j] = -jump_size
+        j += 1
+    choices = seed.choice(len(range(len(possible_moves))), size=(atoms, len(timesteps)))
+    steps = np.zeros((atoms, len(timesteps), 3))
+    for i in range(steps.shape[0]):
+        for j in range(steps.shape[1]):
+            steps[i, j] = possible_moves[choices[i, j]]
+    cum_steps = np.cumsum(steps, axis=1)
+    return cum_steps
+
+
+def get_disp3d(which_walk: callable,
+               steps: int,
+               atoms: int,
+               jump_size: int = 1,
+               seed: Union[np.random.mtrand.RandomState, None] = None) -> List[np.ndarray]:
+    """
+    Return the three-dimensional displacements from a given random walk.
+
+    :param which_walk: the walk to perform.
+    :param steps: number of timesteps.
+    :param atoms: number of atoms.
+    :param jump_size: size of jump
+    :param seed: random seed source
+    :return: Three-dimensional displacements
+    """
+    dt = np.linspace(1, steps, steps, dtype=int)
+    cum_steps = which_walk(atoms, dt, jump_size=jump_size, seed=seed)
+    disp_3d = []
+    n_samples = np.array([])
+    for i, n in enumerate(dt):
+        disp = np.concatenate([cum_steps[:, np.newaxis, i],
+                               np.subtract(cum_steps[:, i + 1:], cum_steps[:, :-(i + 1)])],
+                              axis=1)
+        disp_3d.append(disp)
+        n_samples = np.append(n_samples, dt[-1] / n * atoms)
+    return disp_3d, n_samples

### 2023-05-29 src/code/random_walks/stat_eff.py

@@ -1,0 +1,101 @@
+import numpy as np
+from scipy.optimize import curve_fit
+
+jump = 1
+
+xaxis = np.array([16, 32, 64, 128, 256, 512, 1024])
+
+kinisi_atoms = np.zeros((xaxis.size))
+weighted_atoms = np.zeros((xaxis.size))
+ordinary_atoms = np.zeros((xaxis.size))
+numerical_atoms = np.zeros((xaxis.size))
+kinisi_length = np.zeros((xaxis.size))
+weighted_length = np.zeros((xaxis.size))
+ordinary_length = np.zeros((xaxis.size))
+numerical_length = np.zeros((xaxis.size))
+for i, ii in enumerate(xaxis):
+    length = 128
+    file_open = np.load(f"src/data/random_walks/kinisi/rw_{jump}_{ii}_{length}_s512.npz")
+    kinisi_atoms[i] = file_open['diff_c'].mean(-1).var(-1, ddof=1)
+    file_open = np.load(f"src/data/random_walks/weighted/rw_{jump}_{ii}_{length}_s512.npz")
+    weighted_atoms[i] = file_open['diff_c'].mean(-1).var(-1, ddof=1)
+    file_open = np.load(f"src/data/random_walks/ordinary/rw_{jump}_{ii}_{length}_s512.npz")
+    ordinary_atoms[i] = file_open['diff_c'].mean(-1).var(-1, ddof=1)
+    file_open = np.load(f"src/data/random_walks/numerical/D_{jump}_{ii}_{length}.npz")
+    numerical_atoms[i] = file_open['diff_c'].var(ddof=1)
+for i, ii in enumerate(xaxis):
+    atoms = 128
+    file_open = np.load(f"src/data/random_walks/kinisi/rw_{jump}_{atoms}_{ii}_s512.npz")
+    kinisi_length[i] = file_open['diff_c'].mean(-1).var(-1, ddof=1)
+    file_open = np.load(f"src/data/random_walks/weighted/rw_{jump}_{atoms}_{ii}_s512.npz")
+    weighted_length[i] = file_open['diff_c'].mean(-1).var(-1, ddof=1)
+    file_open = np.load(f"src/data/random_walks/ordinary/rw_{jump}_{atoms}_{ii}_s512.npz")
+    ordinary_length[i] = file_open['diff_c'].mean(-1).var(-1, ddof=1)
[…]

### 2023-05-29 src/code/random_walks/true_ls.py

@@ -1,0 +1,65 @@
+# Copyright (c) Andrew R. McCluskey
+# Distributed under the terms of the MIT License
+# author: Andrew R. McCluskey (arm61)
+
+import numpy as np
+from scipy.stats import linregress
+from scipy.linalg import pinvh
+from emcee import EnsembleSampler
+from scipy.optimize import minimize
+from kinisi.matrix import find_nearest_positive_definite
+from kinisi.diffusion import _straight_line
+
+np.random.seed(1)
+
+atoms = int(snakemake.params['atoms'])
+length = int(snakemake.params['length'])
+jump = int(snakemake.params['jump'])
+
+true_msd = np.load(f'src/data/random_walks/numerical/rw_1_{atoms}_{length}_s4096.npz')['data']
+timestep = np.arange(1, length + 1, 1, dtype=int)
+true_mean = true_msd.mean(0)
+rcov = np.cov(true_msd.T)
+
+max_ngp = np.argwhere(timestep > 4)[0][0]
+rcov = find_nearest_positive_definite(rcov[max_ngp:, max_ngp:])
+
+_, logdet = np.linalg.slogdet(rcov)
+logdet += np.log(2 * np.pi) * true_msd.mean(0)[max_ngp:].size
+inv = pinvh(rcov)
+
+
+def log_likelihood(theta: np.ndarray) -> float:
+    """
+    Get the log likelihood for multivariate normal distribution.
+    :param theta: Value of the gradient and intercept of the straight line.
+    :return: Log-likelihood value.
+    """
+    if theta[0] < 0:
+        return -np.inf
+    model = _straight_line(timestep[max_ngp:], *theta)
+    diff = (model - true_msd.mean(0)[max_ngp:])
+    logl = -0.5 * (logdet + np.matmul(diff.T, np.matmul(inv, diff)))
+    return logl
+
+
+ols = linregress(timestep[max_ngp:], true_msd.mean(0)[max_ngp:])
+
+
+def nll(*args):
+    """
+    General purpose negative log-likelihood.
+    Returns:
+        (:py:attr:`float`): Negative log-likelihood.
+    """
+    return -log_likelihood(*args)
+
+
+max_likelihood = minimize(nll, np.array([ols.slope, ols.intercept])).x
+pos = max_likelihood + max_likelihood * 1e-3 * np.random.randn(32, max_likelihood.size)
[…]

### 2023-05-29 src/code/random_walks/truecov_rw.py

@@ -1,0 +1,69 @@
+# Copyright (c) Andrew R. McCluskey
+# Distributed under the terms of the MIT License
+# author: Andrew R. McCluskey (arm61)
+
+import numpy as np
+from scipy.stats import linregress
+from scipy.linalg import pinvh
+from scipy.optimize import minimize
+from emcee import EnsembleSampler
+from kinisi.diffusion import _straight_line
+from tqdm import tqdm
+
+jump = snakemake.params['jump']
+atoms = snakemake.params['atoms']
+length = int(snakemake.params['length'])
+size = int(snakemake.params['n'])
+
+rcov = np.cov(
+    np.load(f'src/data/random_walks/numerical/rw_1_{atoms}_{length}_s4096.npz')['data'][:, 4:].T)
+kinisi = np.load(f'src/data/random_walks/kinisi/rw_1_{atoms}_{length}_s{size}.npz')['data']
+
+gradient = np.zeros((size, 3200))
+intercept = np.zeros((size, 3200))
+
+for seed in tqdm(range(size)):
+    x = kinisi[seed, 4:, 0]
+    y = kinisi[seed, 4:, 1]
+
+    _, logdet = np.linalg.slogdet(rcov)
+    logdet += np.log(2 * np.pi) * y.size
+    inv = pinvh(rcov)
+
+    def log_likelihood(theta: np.ndarray) -> float:
+        """
+        Get the log likelihood for multivariate normal distribution.
+        :param theta: Value of the gradient and intercept of the straight line.
+        :return: Log-likelihood value.
+        """
+        if theta[0] < 0:
+            return -np.inf
+        model = _straight_line(x, *theta)
+        diff = (model - y)
+        logl = -0.5 * (logdet + np.matmul(diff.T, np.matmul(inv, diff)))
+        return logl
+
+    ols = linregress(x, y)
+    slope = ols.slope
+    c = 1e-20
+    if slope < 0:
+        slope = 1e-20
+
+    def nll(*args) -> float:
+        """
+        General purpose negative log-likelihood.
+        :return: Negative log-likelihood
+        """
+        return -log_likelihood(*args)
+
+    max_likelihood = minimize(nll, np.array([slope, c])).x
+    pos = max_likelihood + max_likelihood * 1e-3 * np.random.randn(32, max_likelihood.size)
+    sampler = EnsembleSampler(*pos.shape, log_likelihood)
[…]

### 2023-05-29 src/code/random_walks/weighted_rw.py

@@ -1,0 +1,72 @@
+# Copyright (c) Andrew R. McCluskey
+# Distributed under the terms of the MIT License
+# author: Andrew R. McCluskey (arm61)
+
+import numpy as np
+from scipy.stats import linregress
+from scipy.linalg import pinvh
+from scipy.optimize import minimize
+from emcee import EnsembleSampler
+from kinisi.diffusion import _straight_line
+from tqdm import tqdm
+
+jump = snakemake.params['jump']
+atoms = snakemake.params['atoms']
+length = int(snakemake.params['length'])
+size = int(snakemake.params['n'])
+
+gradient = np.zeros((size, 3200))
+intercept = np.zeros((size, 3200))
+data = np.load(f'src/data/random_walks/kinisi/rw_1_{atoms}_{length}_s{size}.npz')
+
+for seed in tqdm(range(size)):
+    rng = np.random.RandomState(seed)
+    np.random.seed(seed)
+
+    x = data['data'][seed, 4:, 0]
+    y = data['data'][seed, 4:, 1]
+    yerr = data['data'][seed, 4:, 2]  
+
+    rcov = np.diag(yerr ** 2)
+
+    _, logdet = np.linalg.slogdet(rcov)
+    logdet += np.log(2 * np.pi) * y.size
+    inv = pinvh(rcov)
+
+    def log_likelihood(theta: np.ndarray) -> float:
+        """
+        Get the log likelihood for multivariate normal distribution.
+        :param theta: Value of the gradient and intercept of the straight line.
+        :return: Log-likelihood value.
+        """
+        if theta[0] < 0:
+            return -np.inf
+        model = _straight_line(x, *theta)
+        diff = (model - y)
+        logl = -0.5 * (logdet + np.matmul(diff.T, np.matmul(inv, diff)))
+        return logl
+
+    ols = linregress(x, y)
+    slope = ols.slope
+    c = 1e-20
+    if slope < 0:
+        slope = 1e-20
+
+    def nll(*args) -> float:
+        """
+        General purpose negative log-likelihood.
+        :return: Negative log-likelihood
+        """
+        return -log_likelihood(*args)
+
+    max_likelihood = minimize(nll, np.array([slope, c])).x
+    pos = max_likelihood + max_likelihood * 1e-3 * np.random.randn(32, max_likelihood.size)
+    sampler = EnsembleSampler(*pos.shape, log_likelihood)
[…]

### 2023-05-29 src/scripts/bmvn_msd.py

@@ -1,0 +1,63 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+from scipy.stats import norm
+
+import utils._fig_params as fp
+from utils.plotting_helper import CREDIBLE_INTERVALS
+import paths
+
+jump = 1
+atoms = 128
+length = 128
+
+kinisi = np.load(paths.data / f"random_walks/kinisi/rw_{jump}_{atoms}_{length}_s4096.npz")
+dt = kinisi['data'][0, :, 0]
+distribution = kinisi['diff_c'][0] * 6 * dt[:, np.newaxis] + kinisi['intercept'][0]
+
+true = 1 / 6
+
+figsize = (3.4, 3.4)
+fig = plt.figure(figsize=figsize)
+gs = gridspec.GridSpec(1, 5, figure=fig, wspace=0.9, hspace=0.7)
+
+axes = []
+titles = []
+
+axes.append(fig.add_subplot(gs[0, :3]))
+titles.append(r"$\mathbf{a}$ - $x(t)$")
+for ci in CREDIBLE_INTERVALS:
+    axes[-1].fill_between(dt,
+                          *np.percentile(distribution, ci[:-1], axis=1) * 1e-1,
+                          alpha=ci[-1],
+                          color=fp.colors[2],
+                          lw=0)
+axes[-1].plot(dt, kinisi['data'][0, :, 1] * 1e-1, color=fp.colors[1], label=r'$x(t)$')
+axes[-1].set_xlabel(r"$t$ / s")
+axes[-1].set_ylabel(r"$x(t)$ / $10^1$m$^2$")
+axes[-1].set_xlim(0, None)
+axes[-1].set_ylim(0, None)
+axes[-1].set_xticks([0, 64, 128])
+axes[-1].set_yticks([0, 6, 12])
+
+axes.append(fig.add_subplot(gs[0, 3:]))
+titles.append(r"$\mathbf{b}$ - $p(\hat{D}^* | x(t))$")
+y, x = np.histogram(kinisi['diff_c'], bins=fp.NBINS, density=True)
+axes[-1].stairs(y * 1e-1, x, fill=True, color=fp.colors[2], label=r'Estimated posterior')
+axes[-1].axvline(true, color=fp.colors[0], label='True $D^*$')
+axes[-1].set_xlabel(r"$D^*$ / m$^2$s$^{-1}$")
[…]

### 2023-05-29 src/scripts/covariances.py

@@ -1,0 +1,157 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+from matplotlib.lines import Line2D
+import seaborn as sns
+
+import utils._fig_params as fp
+import paths
+
+colors = ["#0173B2", "#029E73", "#D55E00", "#CC78BC", "#ECE133", "#56B4E9"]
+
+jump = 1
+dimensionality = 3
+D = 6 / (dimensionality * 2)
+length = 128
+atoms = 128
+label_offset = (0, 1.1)
+rng = np.random.RandomState(0)
+np.random.seed(0)
+
+newcmp1 = sns.light_palette(colors[0], as_cmap=True)
+newcmp2 = sns.light_palette(colors[1], as_cmap=True)
+newcmp3 = sns.light_palette(colors[2], as_cmap=True)
+newcmp4 = sns.diverging_palette(311.4, 161.2, as_cmap=True)
+
+true_msd = np.load(paths.data / f"random_walks/numerical/rw_{jump}_{atoms}_{length}_s4096.npz")["data"]
+true_cov = np.cov(true_msd[:, 4:length].T)
+
+kinisi_data = np.load(paths.data / f"random_walks/kinisi/rw_{jump}_{atoms}_{length}_s4096.npz")
+
+kinisi_cov = kinisi_data["covariance"].mean(0)
+no = kinisi_data["n_o"][0]
+timestep = np.arange(1, length + 1, 1)
+
+ts_mesh = np.meshgrid(timestep[4:], timestep[4:])
+
+anal_cov = np.zeros((length, length))
+for i in range(0, anal_cov.shape[0]):
+    for j in range(i, anal_cov.shape[1]):
+        value = 8 * dimensionality**2 * D**2 * timestep[i]**2 / (dimensionality * no[j])
+        anal_cov[i, j] = value
+        anal_cov[j, i] = np.copy(anal_cov[i, j])
+anal_cov = anal_cov[4:, 4:]
+
+max_cov = np.ceil(np.max([anal_cov.max(), true_cov.max(), kinisi_cov.max()]))
+min_cov = np.floor(np.min([anal_cov.min(), true_cov.min(), kinisi_cov.min()]))
+levels = np.linspace(min_cov, max_cov, 15)
+
[…]

### 2023-05-29 src/scripts/diffusion.py

@@ -1,0 +1,73 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+from scipy.stats import norm
+
+import utils._fig_params as fp
+from utils.plotting_helper import mid_points
+import paths
+
+jump = 1
+atoms = 128
+length = 128
+type = 'kinisi'
+
+dllzo_true = np.load(paths.data / "llzo/true_10.npz")['diff_c']
+d = np.zeros((16, 8, 4, 3200))
+for i in range(0, 16, 1):
+    d[i] = np.load(paths.data / f'llzo/diffusion_{i}_10.npz')['d']
+d = d.reshape(-1, 3200)
+llzo_x = np.linspace(dllzo_true.min(), dllzo_true.max(), 1000)
+
+figsize = (3.2, 1.42)
+fig = plt.figure(figsize=figsize)
+gs = gridspec.GridSpec(1, 2, figure=fig, wspace=1)
+
+axes = []
+titles = []
+
+axes.append(fig.add_subplot(gs[0, 0]))
+y, x = np.histogram(d.mean(-1), bins=int(fp.NBINS * 0.5), density=True)
+axes[-1].stairs(y * 1e-4, x * 1e5, fill=True, color=fp.colors[2], label='$p(\hat{D}^*_{\mathrm{Li}^+})$')
+axes[-1].plot(llzo_x * 1e5,
+              norm.pdf(llzo_x, dllzo_true.mean(), dllzo_true.std()) * 1e-4,
+              color=fp.colors[4],
+              label='$p(\hat{D}^*_{\mathrm{Li}^+} | \mathbfit{x})$')
+axes[-1].plot(np.array([d.mean(-1).mean() - d.mean(-1).std(),
+                        d.mean(-1).mean() + d.mean(-1).std()]) * 1e5,
+              [axes[-1].get_ylim()[1] * 1.05, axes[-1].get_ylim()[1] * 1.05],
+              color=fp.colors[3],
+              label='$\sigma(\hat{D}^*_{\mathrm{Li}^+})$',
+              marker="|")
+axes[-1].set_yticks([0, 6, 12])
[…]

### 2023-05-29 src/scripts/efficiency.py

@@ -1,0 +1,103 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+from scipy.stats import norm, skew
+from uravu.distribution import Distribution
+from uravu.axis import Axis
+
+import utils._fig_params as fp
+from utils.plotting_helper import to_string, mid_points
+import paths
+
+jump = 1
+
+xaxis = np.array([16, 32, 64, 128, 256, 512, 1024])
+
+figsize = (3.64, 4.5)
+fig = plt.figure(figsize=figsize)
+gs = gridspec.GridSpec(2, 2, figure=fig, wspace=0.6, hspace=0.4)
+axes = []
+titles = []
+letters = ['a', 'b', 'c', 'd', 'e', 'f', 'g', 'h', 'i', 'j', 'k', 'l']
+
+data = np.load(paths.data / f"random_walks/efficiency.npz")
+
+axes.append(fig.add_subplot(gs[0, 0]))
+axes[-1].plot(xaxis, data['ty'], marker='.', ls="", color=fp.colors[0])
+a = Axis([Distribution(i.var(-1, ddof=1)) for i in data['ky']])
+n = [np.mean(i.var(-1, ddof=1)) for i in data['ky']]
+axes[-1].errorbar(xaxis, n, a.ci(), marker='.', ls="", color=fp.colors[2])
+axes[-1].set_xscale('log', base=2)
+axes[-1].set_yscale('log', base=2)
+axes[-1].set_xlabel(r'$N_{{\mathrm{{atoms}}}}$ / atoms')
+axes[-1].set_ylabel(r"$\hat{\sigma}^2 (\hat{D}^*)$")
+axes[-1].set_xticks([32, 128, 512])
+axes[-1].set_xticklabels([32, 128, 512])
+axes[-1].minorticks_off()
+titles.append(f'{letters[0]}')
+axes[-1].set_title(r"$\hat{\sigma}^2(\hat{D}^*)$ scaling")
+
+axes.append(fig.add_subplot(gs[1, 0]))
+axes[-1].plot(xaxis, (a.ci()[1] - a.ci()[0]) * 1e5, '.', c=fp.colors[2])
+axes[-1].set_xscale('log', base=2)
+axes[-1].set_yscale('log', base=2)
+axes[-1].set_xlabel(r'$N_{{\mathrm{{atoms}}}}$ / atoms')
[…]

### 2023-05-29 src/scripts/glswlsols.py

@@ -1,0 +1,90 @@
+import numpy as np
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+import matplotlib.transforms as transforms
+from scipy.stats import norm
+from uravu.distribution import Distribution
+
+import utils._fig_params as fp
+import paths
+
+data = np.load('src/data/random_walks/numerical/glswlsols_1_128_128.npz')
+
+figsize = (3.744, 3.4)
+fig = plt.figure(figsize=figsize)
+gs = gridspec.GridSpec(3, 1, figure=fig, hspace=0.8)
+
+x = np.linspace(0.75, 1.25, 1000)
+
+axes = []
+titles = []
+axes.append(fig.add_subplot(gs[0, 0]))
+bins = np.linspace((data['ols_pop']).min(), (data['ols_pop']).max(), fp.NBINS * 2)
+axes[-1].hist(data['ols_pop'], bins=bins, density=True, color=fp.colors[5])
+axes[-1].plot(x, norm.pdf(x, *norm.fit(data['ols_pop'])), color=fp.colors[5])
+axes[-1].plot([1 - data['ols_pop'].std(), 1 + data['ols_pop'].std()],
+              [axes[-1].get_ylim()[1] * 1.1, axes[-1].get_ylim()[1] * 1.1],
+              color='#1E5B84',
+              marker='|')
+axes[-1].plot([1 - data['ols_est'][:, 0].std(), 1 + data['ols_est'][:, 0].std()],
+              [axes[-1].get_ylim()[1] * 1.1, axes[-1].get_ylim()[1] * 1.1],
+              color=fp.colors[3],
+              marker='|')
+axes[-1].set_xticks([0.8, 0.9, 1, 1.1, 1.2])
+axes[-1].set_xlim(0.75, 1.25)
+axes[-1].set_yticks([0, 3, 6])
+axes[-1].set_ylabel(r'$p[\hat{D}^* (D^*)^{-1}]$')
+titles.append("a")
+axes[-1].set_title("OLS")
+
+axes.append(fig.add_subplot(gs[1, 0], sharex=axes[0]))
+axes[-1].hist(data['wls_pop'], bins=bins, density=True, color=fp.colors[5])
+axes[-1].plot(x, norm.pdf(x, *norm.fit(data['wls_pop'])), color=fp.colors[5])
[…]

### 2023-05-29 src/scripts/longtime.py

@@ -1,0 +1,81 @@
+import numpy as np
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+from scipy.stats import norm
+from uncertainties import ufloat
+from tueplots import figsizes
+
+import utils._fig_params as fp
+import paths
+
+lengths = [16, 32, 64, 128, 256, 512, 1024]
+
+figsize = figsizes.icml2022_half(nrows=1, ncols=1, height_to_width_ratio=0.5)['figure.figsize']
+fig = plt.figure(figsize=figsize)
+gs = gridspec.GridSpec(1, 1, figure=fig)
+
+axes = []
+titles = []
+cap = ['True']
+mpl.rcParams["axes.spines.right"] = True
+
+i = 0
+data = np.zeros((len(lengths), 2, 3200))
+for ll, length in enumerate(lengths):
+    d = np.load(paths.data / f"random_walks/numerical/D_1_128_{length}.npz")
+    data[ll, 0] = d["diff_c"]
+    data[ll, 1] = d["intercept"]
+
+axes.append(fig.add_subplot(gs[0, 0]))
+axes[-1].errorbar(
+    lengths,
+    data[:, 0].mean(-1),
+    data[:, 0].std(-1) * 2,
+    marker=".",
+    #   ls="-",
+    lw=0.5,
+    color=fp.colors[0])
+axes[-1].set_xscale("log", base=2)
+axes[-1].set_xlabel("Simulation Length / s")
+axes[-1].set_ylabel("$\hat{D}^*$ / m$^{2}$s$^{-1}$")
+axes[-1].yaxis.label.set_color(fp.colors[0])
+axes[-1].tick_params(axis='y', colors=fp.colors[0])
+axes[-1].set_xticks([32, 128, 512])
+axes[-1].set_xticklabels([32, 128, 512])
+axes[-1].minorticks_off()
+
+axes.append(axes[-1].twinx())
+axes[-1].errorbar(
+    lengths,
+    data[:, 1].mean(-1),
+    data[:, 1].std(-1) * 2,
+    marker=".",
+    #   ls="--",
+    lw=0.5,
+    color=fp.colors[1])
+axes[-1].set_xscale("log", base=2)
+axes[-1].set_xlabel("Simulation Length / s")
+axes[-1].set_ylabel(r"$\hat{D}_{\mathrm{offset}}$ / m$^{2}$")
[…]

### 2023-05-29 src/scripts/msd.py

@@ -1,0 +1,82 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+from scipy.stats import norm
+
+import utils._fig_params as fp
+from utils.plotting_helper import mid_points
+import paths
+
+jump = 1
+atoms = 128
+length = 128
+
+kinisi = np.load(paths.data / f"random_walks/kinisi/rw_{jump}_{atoms}_{length}_s4096.npz")['data']
+numerical = np.load(paths.data / f"random_walks/numerical/rw_{jump}_{atoms}_{length}_s4096.npz")['data']
+dt = kinisi[0, :, 0]
+
+figsize = (3.73, 0.85)
+fig = plt.figure(figsize=figsize)
+gs = gridspec.GridSpec(1, 3, figure=fig, wspace=0.7, hspace=0.7)
+
+axes = []
+titles = []
+
+axes.append(fig.add_subplot(gs[0, 0]))
+titles.append(r"a")
+axes[-1].set_title("Numerical")
+m, c, b = axes[-1].errorbar(dt, numerical.mean(0), numerical.std(0), color=fp.colors[0], elinewidth=0.2)
+[bar.set_alpha(0.5) for bar in b]
+[cap.set_alpha(0.5) for cap in c]
+axes[-1].set_xlabel(r"$t^{}$")
+axes[-1].set_ylabel(r"$x(t)$")
+axes[-1].set_xlim(0, None)
+axes[-1].set_ylim(0, None)
+axes[-1].set_xticks([0, 64, 128])
+axes[-1].set_yticks([0, 400, 800])
+
+axes.append(fig.add_subplot(gs[0, 1], sharey=axes[-1]))
+titles.append(r"b")
+axes[-1].set_title("Estimated")
+m, c, b = axes[-1].errorbar(dt, kinisi[0, :, 1], kinisi[0, :, 2], color=fp.colors[2], elinewidth=0.2)
+[bar.set_alpha(0.5) for bar in b]
+[cap.set_alpha(0.5) for cap in c]
+axes[-1].set_xlabel(r"$t^{}$")
+axes[-1].set_ylabel(r"$x(t)$")
+axes[-1].set_xlim(0, None)
+axes[-1].set_ylim(0, None)
+axes[-1].set_xticks([0, 64, 128])
+axes[-1].set_yticks([0, 400, 800])
+
+axes.append(fig.add_subplot(gs[0, 2]))
+titles.append(r"c")
[…]

### 2023-05-29 src/scripts/paths.py

@@ -1,0 +1,28 @@
+"""
+Exposes common paths useful for manipulating datasets and generating figures.
+"""
+from pathlib import Path
+
+# Absolute path to the top level of the repository
+root = Path(__file__).resolve().parents[2].absolute()
+
+# Absolute path to the `src` folder
+src = root / "src"
+
+# Absolute path to the `src/data` folder (contains datasets)
+data = src / "data"
+
+# Absolute path to the `src/static` folder (contains static images)
+static = src / "static"
+
+# Absolute path to the `src/scripts` folder (contains figure/pipeline scripts)
+scripts = src / "scripts"
+
+# Absolute path to the `src/tex` folder (contains the manuscript)
+tex = src / "tex"
+
+# Absolute path to the `src/tex/figures` folder (contains figure output)
+figures = tex / "figures"
+
+# Absolute path to the `src/tex/output` folder (contains other outputs)
+output = tex / "output"

### 2023-05-29 src/scripts/random_walk.py

@@ -1,0 +1,121 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+from scipy.stats import norm
+
+import utils._fig_params as fp
+from utils.plotting_helper import mid_points, CREDIBLE_INTERVALS
+import paths
+
+jump = 1
+atoms = 128
+length = 128
+type = 'kinisi'
+
+true = np.load(paths.data / f"random_walks/{type}/rw_{jump}_{atoms}_{length}_s4096.npz")['diff_c']
+dinfty_true = np.load(paths.data / f"random_walks/numerical/D_1_128_128.npz")['diff_c']
+
+rw_x = np.linspace(true.mean(-1).min(), true.mean(-1).max(), 1000)
+
+kinisi = np.load(paths.data / f"random_walks/kinisi/rw_{jump}_{atoms}_{length}_s4096.npz")
+dt = kinisi['data'][0, :, 0]
+distribution = kinisi['diff_c'][0] * 6 * dt[:, np.newaxis] + kinisi['intercept'][0]
+
+figsize = (8.06, 1.3)
+print(figsize)
+fig = plt.figure(figsize=figsize)
+gs = gridspec.GridSpec(1, 4, figure=fig, wspace=0.8, width_ratios=[87, 63, 63, 63])
+
+axes = []
+titles = []
+
+axes.append(fig.add_subplot(gs[0, 0]))
+titles.append("a")
+axes[-1].set_title("Single simulation")
+axes[-1].fill_between(dt,
+                      *np.percentile(distribution, CREDIBLE_INTERVALS[0][:-1], axis=1),
+                      alpha=CREDIBLE_INTERVALS[0][-1],
+                      color=fp.colors[1],
+                      lw=0,
+                      label=r'$p(m | \mathbfit{x})$')
+for ci in CREDIBLE_INTERVALS[1:]:
+    axes[-1].fill_between(dt, *np.percentile(distribution, ci[:-1], axis=1), alpha=ci[-1], color=fp.colors[1], lw=0)
+axes[-1].plot(dt, kinisi['data'][0, :, 1], color='#1E5B84', label=r"$\mathbfit{x}$")
+axes[-1].set_xlabel(r"$t$")
+axes[-1].set_ylabel(r"$x(t)$")
[…]

### 2023-05-29 src/scripts/stat_eff.py

@@ -1,0 +1,140 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+from scipy.stats import norm
+from uravu.distribution import Distribution
+from uravu.axis import Axis
+
+import utils._fig_params as fp
+from utils.plotting_helper import mid_points
+import paths
+
+correlation = 'true'
+jump = 1
+
+
+def find_exp(number) -> int:
+    base10 = np.log10(abs(number))
+    return int(abs(np.floor(base10)))
+
+
+def f(x, a, b):
+    return b * x**a
+
+
+xaxis = np.array([16, 32, 64, 128, 256, 512, 1024])
+
+figsize = (3.66, 2)
+fig = plt.figure(figsize=figsize)
+gs = gridspec.GridSpec(1, 2, figure=fig, wspace=0.6, hspace=0.4)
+
+x = np.linspace(xaxis.min(), xaxis.max(), 1000)
+
+axes = []
+titles = []
+letters = ['a', 'b']
+
+data = np.load(paths.data / f"random_walks/stat_eff.npz")
+
+write_out = open(paths.output / f"stat_eff_table.txt", 'w')
+
+write_out.write(r"\multirow{2}{*}{" + r"$1/N_{\mathrm{atoms}}$" + r"} ")
+write_out.write(r"& $a$ & \num{" + f"{data['na'] * 10 ** find_exp(data['na']):.2f}" + " \pm " +
+                f"{data['dna'] * 10 ** find_exp(data['na']):.2f}e-{find_exp(data['na'])}" + r"} ")
+write_out.write(r"& \num{" + f"{data['ka'] * 10 ** find_exp(data['ka']):.2f}" + " \pm " +
+                f"{data['dka'] * 10 ** find_exp(data['ka']):.2f}e-{find_exp(data['ka'])}" + r"} ")
+write_out.write(r"& \num{" + f"{data['wa'] * 10 ** find_exp(data['wa']):.2f}" + " \pm " +
+                f"{data['dwa'] * 10 ** find_exp(data['wa']):.2f}e-{find_exp(data['wa'])}" + r"} ")
+write_out.write(r"& \num{" + f"{data['pa'] * 10 ** find_exp(data['pa']):.2f}" + " \pm " +
[…]

### 2023-05-29 src/scripts/true_cov.py

@@ -1,0 +1,67 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+import matplotlib.transforms as transforms
+from matplotlib.ticker import FuncFormatter
+from scipy.stats import norm
+from uravu.distribution import Distribution
+from uravu.axis import Axis
+
+import utils._fig_params as fp
+from utils.plotting_helper import to_string, CREDIBLE_INTERVALS
+import paths
+
+jump = 1
+atoms = 128
+length = 128
+type = 'true_cov'
+
+true = np.load(paths.data / f"random_walks/{type}/rw_{jump}_{atoms}_{length}_s4096.npz")['diff_c']
+dinfty_true = np.load(paths.data / "random_walks/numerical/D_1_128_128.npz")['diff_c']
+
+figsize = (3.4, 1.4)
+fig = plt.figure(figsize=figsize)
+gs = gridspec.GridSpec(1, 2, figure=fig, wspace=1)
+
+axes = []
+titles = []
+
+x_ = np.linspace(true.mean(-1).min(), true.mean(-1).max(), 1000)
+
+axes.append(fig.add_subplot(gs[0, 0]))
+y, x = np.histogram(true.mean(-1), density=True, bins=fp.NBINS)
+axes[-1].stairs(y * 1e-1, x, fill=True, color=fp.colors[2], label='$p(\hat{D}^*)$')
+axes[-1].plot(x_,
+              norm.pdf(x_, dinfty_true.mean(), dinfty_true.std()) * 1e-1,
+              color=fp.colors[4],
+              label=r'$p(\hat{D}^* | \mathbfit{x})$')
+axes[-1].plot([dinfty_true.mean() - dinfty_true.std(),
+               dinfty_true.mean() + dinfty_true.std()], [axes[-1].get_ylim()[1] * 1.05, axes[-1].get_ylim()[1] * 1.05],
+              color=fp.colors[3],
+              label='$\sigma(\hat{D}^*)$')
+axes[-1].set_xlabel(r"$\hat{D}^*$")
+axes[-1].set_ylabel(r"$p(\hat{D}^*)$")
+titles.append(r"a")
+axes[-1].set_title("$D^*$ distribution")
+axes[-1].legend(loc='upper left', bbox_to_anchor=(0.65, 1))
+
[…]

### 2023-05-29 src/scripts/true_llzo.py

@@ -1,0 +1,109 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+from matplotlib.ticker import FuncFormatter
+from scipy.stats import norm
+from uravu.distribution import Distribution
+from uravu.axis import Axis
+
+import utils._fig_params as fp
+from utils.plotting_helper import mid_points
+import paths
+
+jump = 1
+atoms = 128
+length = 128
+correlation = 'true'
+type = 'kinisi'
+
+start_diffs = [0, 2, 4, 6, 8, 10, 15, 20]
+d = np.zeros((len(start_diffs), 16, 8, 4, 3200))
+timestep = np.load(paths.data / f"llzo/diffusion_0_10.npz")['dt'][0, 0]
+dinfty_true = np.zeros((len(start_diffs), 3200))
+cinfty_true = np.zeros((len(start_diffs), 3200))
+for j, start in enumerate(start_diffs):
+    dinfty_true[j] = np.load(paths.data / f"llzo/true_{start}.npz")['diff_c']
+    cinfty_true[j] = np.load(paths.data / f"llzo/true_{start}.npz")['intercept']
+    for i in range(0, 16, 1):
+        d[j, i] = np.load(paths.data / f'llzo/diffusion_{i}_{start}.npz')['d']
+d = d.reshape(len(start_diffs), -1, 3200)
+
+figsize = (8.455, 3)
+fig = plt.figure(figsize=figsize)
+gs = gridspec.GridSpec(2, 4, figure=fig, wspace=0.6, hspace=1)
+letters = [['a', 'b', 'c', 'd'], ['e', 'f', 'g', 'h']]
+times = [[0, 2, 4, 6], [8, 10, 15, 20]]
+
+axes = []
+titles = []
+
+j = 0
+for i in [0, 4]:
+
+    axes.append(fig.add_subplot(gs[j, 0]))
+    y, x = np.histogram(d[i].var(-1, ddof=1), density=True, bins=int(fp.NBINS * 0.5))
+    axes[-1].stairs(y, x, fill=True, color=fp.colors[0])
+    axes[-1].axvline(d[i].mean(-1).var(ddof=1), c=fp.colors[3], ls='--')
[…]

### 2023-05-29 src/scripts/utils/_fig_params.py

@@ -1,0 +1,92 @@
+"""
+Figure parameters for kinisi to help make nice plots.
+
+Copyright (c) Andrew R. McCluskey and Benjamin J. Morgan
+
+Distributed under the terms of the MIT License
+
+@author: Andrew R. McCluskey
+"""
+
+from matplotlib import rcParams
+
+# blue, orange, green, pink, dark green, grey
+# colors = ["#0099c8", "#D55E00", "#029E73", "#CC78BC", "#006646", "#949494"]
+# colors = ["#5D5779", "#A46355", "#56C1A3", "#FFAABB", "#99DDFF", "#BBBBBB"]    
+colors=["#80b8d8", "#80ceb8", "#e9ae7f", "#cca7c6", "#b8b8b8", "#aad9f4"]
+FONTSIZE = 7
+NEARLY_BLACK = "#161616"
+LIGHT_GREY = "#F5F5F5"
+WHITE = "#ffffff"
+NBINS = 25
+
+MASTER_FORMATTING = {
+    "axes.formatter.limits": (-3, 3),
+    "xtick.major.pad": 1,
+    "ytick.major.pad": 1,
+    "ytick.color": NEARLY_BLACK,
+    "xtick.color": NEARLY_BLACK,
+    "axes.labelcolor": NEARLY_BLACK,
+    "axes.spines.bottom": True,
+    "axes.spines.left": True,
+    "axes.spines.right": False,
+    "axes.spines.top": False,
+    "axes.axisbelow": True,
+    "legend.frameon": False,
+    'axes.edgecolor': NEARLY_BLACK,
+    "pdf.fonttype": 42,
+    "ps.fonttype": 42,
+    "font.size": FONTSIZE,
+    "font.family": "sans-serif",
+    "savefig.bbox": "tight",
+    "axes.facecolor": WHITE,
+    "axes.labelpad": 2.0,
+    "axes.labelsize": FONTSIZE,
+    "axes.titlepad": 8,
+    "axes.titlesize": FONTSIZE,
+    "axes.grid": False,
+    "grid.color": WHITE,
+    "lines.markersize": 3.0,
+    "xtick.major.size": 2.0,
+    "ytick.major.size": 2.0,
+    "lines.scale_dashes": False,
+    "xtick.labelsize": FONTSIZE,
+    "ytick.labelsize": FONTSIZE,
+    "legend.fontsize": FONTSIZE,
[…]

### 2023-05-29 src/scripts/utils/plotting_helper.py

@@ -1,0 +1,59 @@
+from uncertainties import ufloat
+from uravu.distribution import Distribution
+
+ALPHABET = 'abcdefghijklmnopqrstuvwxyz'
+CREDIBLE_INTERVALS = [[15.86555, 84.13445, 1], [2.27500, 97.725, 0.6], [0.135, 99.865, 0.2]]
+
+
+def to_string(d: Distribution, units: str = None, exponent: int = 0, normal_override: bool = False) -> str:
+    """
+    Turns Distribution objects to LaTeX strings
+
+    :param d: Distribution to LaTeXify.
+    :param units: Units to include.
+    """
+    if normal_override or d.normal:
+        starter = 'num'
+        if units:
+            starter = 'SI'
+
+        if exponent != 0:
+            e = Distribution(d.samples / (10**exponent))
+            v = ufloat(e.samples.mean(), e.samples.std(ddof=1) * 1.96)
+            s = "$\\" + f"{starter}" + r"{" + f"{v:.2fL}E{exponent}" + r"}"
+        else:
+            v = ufloat(d.samples.mean(), d.samples.std(ddof=1) * 1.96)
+            s = "$\\" + f"{starter}" + r"{" + f"{v:.2fL}" + r"}"
+        if units is not None:
+            s += r"{" + f"{units}" + r"}$"
+        else:
+            s += '$'
+    else:
+        if exponent != 0:
+            s = r"$\big" + f"[({d.dist_max / (10 ** exponent):.2f}^" \
+                + r"{+" + f"{(d.con_int()[1] - d.dist_max) / (10 ** exponent):.2f}" \
+                + r"}_{-" + f"{(d.dist_max - d.con_int()[0]) / (10 ** exponent):.2f}" \
+                + r"})\times 10^{" + f"{exponent}" + r"}\big]"
+        else:
+            s = r"$" + f"({d.dist_max:.2f}^" + r"{+" \
+                + f"{(d.con_int()[1] - d.dist_max):.2f}" + r"}_{-" \
+                + f"{(d.dist_max - d.con_int()[0]):.2f}" + r"})"
+        if units is not None:
+            s += r"\,\si{" + f"{units}" + r"}$"
+        else:
+            s += '$'
+    return s
+
+
+def mid_points(axis):
+    """
+    Find the mid point in an axis, in axes space.
+    """
[…]

### 2023-05-29 src/code/random_walks/longtime.py
Removing unused files

@@ -1,56 +1,0 @@
-# Copyright (c) Andrew R. McCluskey
-# Distributed under the terms of the MIT License
-# author: Andrew R. McCluskey (arm61)
-
-import numpy as np
-from scipy.stats import multivariate_normal, linregress
-from scipy.optimize import minimize
-from emcee import EnsembleSampler
-from kinisi.matrix import find_nearest_positive_definite
-from uravu.utils import straight_line
-
-atoms = 1024
-lengths = [128, 256, 52, 1024, 2048, 4096]
-jumps = [1, 8, 32]
-
-data = np.zeros((2, len(jumps), len(lengths), 3200))
-
-for j, jump in enumerate(jumps):
-    for k, length in enumerate(lengths):
-        simulation = np.load(f'../data/numerical_{jump}_{atoms}_{length}.npz')['data']
-        covariance = np.cov(simulation.T)
-        timestep = np.linspace(1, length, length, dtype=int)[1:-1]
-
-        covariance = find_nearest_positive_definite(covariance)
-        mv = multivariate_normal(simulation.mean(0), covariance, allow_singular=True)
-
-        def log_likelihood(theta):
-            """
-            Get the log likelihood for multivariate normal distribution.
-            Args:
-                theta (:py:attr:`array_like`): Value of the gradient and intercept of the straight line.
-            Returns:
-                (:py:attr:`float`): Log-likelihood value.
-            """
-            model = straight_line(timestep, *theta)
-            logl = mv.logpdf(model)
-            return logl
-
-        ols = linregress(timestep, simulation.mean(0))
-
-        def nll(*args):
-            """
-            General purpose negative log-likelihood.
-            Returns:
-                (:py:attr:`float`): Negative log-likelihood.
-            """
-            return -log_likelihood(*args)
-
-        max_likelihood = minimize(nll, np.array([ols.slope, ols.intercept])).x
-        pos = max_likelihood + max_likelihood * 1e-3 * np.random.randn(32, max_likelihood.size)
-        sampler = EnsembleSampler(*pos.shape, log_likelihood)
-        sampler.run_mcmc(pos, 1500, progress=True, progress_kwargs={'desc': "Likelihood Sampling"})
-        flatchain = sampler.get_chain(flat=True, discard=500)
-        data[0, j, k] = flatchain[:, 0] / 6
-        data[1, j, k] = flatchain[:, 1]
-np.save('src/data/longtime.npy', data)

### 2023-05-29 src/code/random_walks/nonindependent_sampling.py

@@ -1,46 +1,0 @@
-# Copyright (c) Andrew R. McCluskey
-# Distributed under the terms of the MIT License
-# author: Andrew R. McCluskey (arm61)
-
-import numpy as np
-from sklearn.utils import resample
-from tqdm import tqdm
-from uravu.distribution import Distribution
-from random_walk import walk, anticorrelated_walk, get_disp3d
-from kinisi.diffusion import MSDBootstrap
-
-jump = snakemake.params['jump']
-atoms = snakemake.params['atoms']
-length = int(snakemake.params['length'])
-correlated = snakemake.params['correlation']
-size = int(snakemake.params['n'])
-
-timestep = np.linspace(1, length, length, dtype=int)
-res = np.zeros((size, length - 2))
-res_e = np.zeros((size, length - 2))
-for seed in tqdm(range(size)):
-    data = np.zeros((length - 2, 3))
-
-    rng = np.random.RandomState(seed)
-    np.random.seed(seed)
-    if correlated == 'correlated':
-        disp_3d = get_disp3d(anticorrelated_walk, length, atoms, seed=rng)
-    else:
-        disp_3d = get_disp3d(walk, length, atoms, seed=rng)
-
-    diff = MSDBootstrap(timestep, disp_3d, random_state=rng, progress=False)
-
-    i = 0
-    for n in range(len(diff.dt))[1:]:
-        r2_sum = np.sum(diff._displacements[n]**2, axis=2)
-        resampled = [np.mean(resample(r2_sum.flatten())) for j in range(1000)]
-        distro = Distribution(resampled)
-        while (not distro.normal):
-            distro.add_samples([np.mean(resample(r2_sum.flatten())) for j in range(100)])
-        res[seed, i] = distro.n
-        res_e[seed, i] = np.var(distro.samples, ddof=1)
-        i += 1
-
-np.savez(f'src/data/random_walks/nonindependent/{correlated}/rw_{int(jump)}_{atoms}_{length}_s{size}.npz',
-         res_msd=res,
-         res_var=res_e)

### 2023-05-29 src/scripts/bmvn_msd.py

@@ -1,63 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import matplotlib.gridspec as gridspec
-from scipy.stats import norm
-
-import utils._fig_params as fp
-from utils.plotting_helper import CREDIBLE_INTERVALS
-import paths
-
-jump = 1
-atoms = 128
-length = 128
-
-kinisi = np.load(paths.data / f"random_walks/kinisi/rw_{jump}_{atoms}_{length}_s4096.npz")
-dt = kinisi['data'][0, :, 0]
-distribution = kinisi['diff_c'][0] * 6 * dt[:, np.newaxis] + kinisi['intercept'][0]
-
-true = 1 / 6
-
-figsize = (3.4, 3.4)
-fig = plt.figure(figsize=figsize)
-gs = gridspec.GridSpec(1, 5, figure=fig, wspace=0.9, hspace=0.7)
-
-axes = []
-titles = []
-
-axes.append(fig.add_subplot(gs[0, :3]))
-titles.append(r"$\mathbf{a}$ - $x(t)$")
-for ci in CREDIBLE_INTERVALS:
-    axes[-1].fill_between(dt,
-                          *np.percentile(distribution, ci[:-1], axis=1) * 1e-1,
-                          alpha=ci[-1],
-                          color=fp.colors[2],
-                          lw=0)
-axes[-1].plot(dt, kinisi['data'][0, :, 1] * 1e-1, color=fp.colors[1], label=r'$x(t)$')
-axes[-1].set_xlabel(r"$t$ / s")
-axes[-1].set_ylabel(r"$x(t)$ / $10^1$m$^2$")
-axes[-1].set_xlim(0, None)
-axes[-1].set_ylim(0, None)
-axes[-1].set_xticks([0, 64, 128])
-axes[-1].set_yticks([0, 6, 12])
-
-axes.append(fig.add_subplot(gs[0, 3:]))
-titles.append(r"$\mathbf{b}$ - $p(\hat{D}^* | x(t))$")
-y, x = np.histogram(kinisi['diff_c'], bins=fp.NBINS, density=True)
-axes[-1].stairs(y * 1e-1, x, fill=True, color=fp.colors[2], label=r'Estimated posterior')
-axes[-1].axvline(true, color=fp.colors[0], label='True $D^*$')
-axes[-1].set_xlabel(r"$D^*$ / m$^2$s$^{-1}$")
[…]

### 2023-05-29 src/scripts/longtime.py

@@ -1,81 +1,0 @@
-import numpy as np
-import matplotlib as mpl
-import matplotlib.pyplot as plt
-import matplotlib.gridspec as gridspec
-from scipy.stats import norm
-from uncertainties import ufloat
-from tueplots import figsizes
-
-import utils._fig_params as fp
-import paths
-
-lengths = [16, 32, 64, 128, 256, 512, 1024]
-
-figsize = figsizes.icml2022_half(nrows=1, ncols=1, height_to_width_ratio=0.5)['figure.figsize']
-fig = plt.figure(figsize=figsize)
-gs = gridspec.GridSpec(1, 1, figure=fig)
-
-axes = []
-titles = []
-cap = ['True']
-mpl.rcParams["axes.spines.right"] = True
-
-i = 0
-data = np.zeros((len(lengths), 2, 3200))
-for ll, length in enumerate(lengths):
-    d = np.load(paths.data / f"random_walks/numerical/D_1_128_{length}.npz")
-    data[ll, 0] = d["diff_c"]
-    data[ll, 1] = d["intercept"]
-
-axes.append(fig.add_subplot(gs[0, 0]))
-axes[-1].errorbar(
-    lengths,
-    data[:, 0].mean(-1),
-    data[:, 0].std(-1) * 2,
-    marker=".",
-    #   ls="-",
-    lw=0.5,
-    color=fp.colors[0])
-axes[-1].set_xscale("log", base=2)
-axes[-1].set_xlabel("Simulation Length / s")
-axes[-1].set_ylabel("$\hat{D}^*$ / m$^{2}$s$^{-1}$")
-axes[-1].yaxis.label.set_color(fp.colors[0])
-axes[-1].tick_params(axis='y', colors=fp.colors[0])
-axes[-1].set_xticks([32, 128, 512])
-axes[-1].set_xticklabels([32, 128, 512])
-axes[-1].minorticks_off()
-
-axes.append(axes[-1].twinx())
-axes[-1].errorbar(
-    lengths,
-    data[:, 1].mean(-1),
-    data[:, 1].std(-1) * 2,
-    marker=".",
-    #   ls="--",
-    lw=0.5,
-    color=fp.colors[1])
-axes[-1].set_xscale("log", base=2)
-axes[-1].set_xlabel("Simulation Length / s")
-axes[-1].set_ylabel(r"$\hat{D}_{\mathrm{offset}}$ / m$^{2}$")
[…]

### 2023-05-30 src/tex/ms.tex
arXiv text

@@
         The off-diagonal panels show difference plots, computed as per-element ratios between pairs of covariance matrices (a--c).
-        %Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences.
     }
@@
 We also show the true variance of individual point estimates, $\var{\Dest}$, which characterises the spread of the histogram in Fig.~\ref{fig:random_walk}c.
-The distribution of estimated variances is biased relative to the true variance and skewed, due to numerical differences between the true covariance matrix $\mathbf{\Sigma}$ and the model covariance matrix $\mathbf{\Sigma^\prime}$ (see the SI for details).
+The distribution of estimated variances is biased relative to the true variance and skewed, due to numerical differences between the true covariance matrix $\mathbf{\Sigma}$ and the model covariance matrix $\mathbf{\Sigma^\prime}$.
 In general, however, the distribution of the estimated variance shows good agreement with the true sample variance.
@@
       (b) Probability distribution of estimated variances, $\varest{\Dest}$, for individual LLZO effective simulations, compared to the true sample variance (pink vertical line) $\var{\Dest}$.
-      %Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences. 
     }
@@
 Hence, our method provides reasonably accurate estimates of the uncertainty in $\Dest$ for our c-LLZO dataset, even when applied to single effective trajectories with limited displacement data (only 56 mobile ions, and \SI{25}{ps} simulation length).
-
-% To further benchmark our method, we performed a scaling analysis using sets \num{512} random-walk simulations with varying total number of particles or total simulation length (number of steps).
-% We then analyse the resulting simulation data using four estimation methods to generate distributions of compatible linear models:
-% \begin{enumerate}
-%   \item Bayesian regression using a numerical covariance matrix derived from the complete set of \num{512} simulations.
-%   This analysis is equivalent to GLS, and gives the smallest theoretical variance in the resulting point estimates $\Dest$.
-%   \item Approximate Bayesian regression as described above, with an estimated model covarianec matrix computed for each individual trajectory using the corresponding simulation data.
-%   \item Approximate Bayesian regression with a diagonal covariance matrix, $\bm{\Sigma}=\bm{\sigma}\mathbf{I}$, with the diagonal terms estimated from the simulation data.
-%   This analysis is equivalent to WLS.
-%   \item Bayesian regression with a covariance matrix $\bm{\Sigma}=\mathbf{I}$, the identity matrix.
-%   This analysis is equivalent to OLS.
-% \end{enumerate}
-% Our analysis is shown in Fig.~[TODO] across two panels.
-The first panel illustrates the scaling performance of each method as we vary the number of atoms.
-The second panel demonstrates the scaling performance of each method as we adjust the total simulation time.
 
@@
 This work used the Isambard 2 UK National Tier-2 HPC Service (http://gw4.ac.uk/isambard/) operated by GW4 and the UK Met Office, and funded by EPSRC (EP/T022078/1).
-The authors gratefully acknowledge the University of Bath's Research Computing Group~\cite{bath_research_2018} for their support in this work, in the LLZO molecular dynamics simulations.
+The authors acknowledge the University of Bath's Research Computing Group for their support in running the LLZO molecular dynamics simulations.
 Other simulations and analyses were carried out using the Data Management and Software Centre computing cluster at the European Spallation Source ERIC.
@@
 \maketitle
-\onecolumngrid
+% \onecolumngrid
 
@@
     \item The derivation of the covariance matrix in the long-time limit for freely diffusion particles.
-    \item Quantification of the scaling behaviour in the posterior distribution for the diffusion coefficient with system size.
-    \item Discussion of the skewed distribution of the estimated variance in the diffusion coefficient.
+    \item Analysis of the relative efficiency of the approximate Bayesian regression method, presented here, compared to OLS, WLS, and GLS, and the efficiency scaling of each method with number of mobile particles or total simulation time..
+    % \item Discussion of the skewed distribution of the estimated variance in the diffusion coefficient.
     \item Details of the LLZO molecular dynamics simulations.
-    \item Discussion of the importance of only using non-diffusive simulation in the analysis.
+    % \item Discussion of the importance of only using non-diffusive simulation in the analysis.
     \item Details of the implementation of the model covariance method in the \textsc{kinisi} package.
@@
 
-\vspace{\columnsep}
+% \vspace{\columnsep}
 \twocolumngrid
-
-% \section{Effect of number of independent observations on the distribution of mean-squared displacement}
-% \label{sec:chi2}
-% %
-% In the main text, we note that for a large enough number of observations, the observed MSD is a multivariate normal distribution. 
-% This is made clear on inspection of the observed MSD at a single time interval (which should follow a normal distribution), where squared particle displacements, in the diffusive regime, are $\chi^2_k$-distributed, where $k$ is the number of independent observations of a given displacement.
-% Fig.~\ref{fig:freedom} shows that when there are sufficiently many independent observations of ${\left[\Delta\mathbf{r}_i(t)_\alpha\right]}^2$, each population of observable MSD values $p(\bar{s}^2_i)$ is normally distributed, i.e.\ as the number of observations increases the $\chi^2_k$-distribution tends towards a normal distribution.
-% A single observation of \SI{16}{atoms} (i.e.\ where $t>0.5t_{\max}$), the distribution is not yet normally distributed.
-% %
-% \begin{figure*}
-%     \centering
-%     \includegraphics[width=\textwidth]{figures/freedom.pdf}
-%     \caption{
-%         A comparison between the observed mean-squared displacement (blue histograms) and the relevant $\chi^2_k$-distribution.
-%         The simulations were run for \SI{128}{\second}, therefore there is only a single observation per atom per dimension, such that $k=N_{\mathrm{atoms}}d$; (a) $k = 2 \times 3$, (b) $k = 4 \times 3$, (c) $k = 8 \times 3$, (d) $k = 128 \times 3$.
-%         % Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences. 
-%     }
-%     \label{fig:freedom}
-%     \script{freedom.py}
-% \end{figure*}
-% %
 
@@
 %
-The main manuscript asserts that the long-time limit covariance matrix for a simulation of freely diffusing particles can be expressed as
+In the main manuscript we present the result that the covariance matrix for a system of freely diffusing particles, in the long-time limit, has the form
 %
@@
 %
-where $\oMSDi$ is the observed mean-squared displacement (MSD) for time interval $i$ and $\nind{i}$ is the number of statistically independent observed squared displacements averaged over to compute the mean value (see Eqn.~2 in the main manuscript).
+where $\oMSDi$ is the observed mean-squared displacement (MSD) for time interval $i$ and $\nind{i}$ is the number of statistically independent observed squared displacements averaged over to compute the mean value.
 
-We must first consider the expected variance for the MSD at timestep $i$, $\var{\oMSDi}$, following the equivalent derivation by Smith and Gillan~\cite{smith_random_1996}, and then extend this to the covariance $\Sigma^\prime\left[\oMSDi, \oMSDj\right]$ to obtain the result given above.
+To derive this result, we first present a derivation of the expected variance for the MSD at timestep $i$, $\var{\oMSD}$, following the approach of Smith and Gillan~\cite{smith_random_1996}.
+We then derive an expression for the covariance $\Sigma^\prime\left[\oMSDi, \oMSDj\right]$ to obtain the result above.
 
 For a single particle undergoing a one-dimensional random walk with step size $\kappa$, each step gives a displacement $h = \pm \kappa$.
-After $n$ steps, the MSD, $\oMSDn$, is
+After $n$ steps, the MSD, $\oMSDn$, is given by
 %
@@
   \begin{aligned}
-    \moMSDn &= \sum_i^n \left<h_i^2\right> + \sum_i^n\sum_{j\neq i}^n\left<h_i h_j\right>
+    \moMSDn &= \sum_i^n \left<h_i^2\right> + \sum_i^n\sum_{j\neq i}^n\left<h_i h_j\right>.
   \end{aligned}
@@
 Hence the expected value for the mean-squared displacement increases linearly with the number of steps taken.
-% First, we will outline a brief note on nomenclature, we will use $\MSD{n}$ to refer to the mean-squared displacement in a single simulation, i.e., this is the mean over all particles and observations of the given number of hops, $n$, $\Big\langle\MSD{n}\Big\rangle$ will be used for the average of the mean-squared displacements from many simulations.
-% We can think of these as a sample and population mean respectively, where $\MSD{n} \to \Big\langle\MSD{n}\Big\rangle$ with increasing numbers of simulations.
 
-The variance in the observed MSD, $\var{\oMSDn}$ is given by the standard statistical formula
+The variance in the observed MSD, $\var{\oMSDn}$, is given by the standard statistical formula
 %
@@
 %
-
 The first term can be expanded in terms of displacements $h$ as
@@
 \begin{equation}
-  \left<\oMSDn^2\right> = \left<\sum_i^n\sum_j^n\sum_k^n\sum_l^n h_i h_j h_k h_l\right>
+  \left<\oMSDn^2\right> = \left<\sum_i^n\sum_j^n\sum_k^n\sum_l^n h_i h_j h_k h_l\right>,
   \label{equ:big_av}
 \end{equation}
-%
-which can be simplified by noting that $h_i$, $h_j$, $h_k$, and $h_l$ are uncorrelated when $i \neq j \neq k \neq l$, and the only terms that contribute to the average are those with:
-%
+which can be simplified by noting that $h_i$, $h_j$, $h_k$, and $h_l$ are uncorrelated when $i \neq j \neq k \neq l$, and the only terms that contribute to the average are those where $h_ih_jh_kh_l$ is guaranteed to be non-zero:
 \begin{itemize}
@@
 \end{itemize}
-%
-Each of these conditions guarantees a positive product $h_ih_jh_kh_l$.
-
 From (a) we obtain
-%
 \begin{equation}
@@
 \end{equation}
-%
 and from (b), (c), and (d), which are equivalent, we obtain
-%
 \begin{equation}
@@
 \end{equation}
-%
 This gives
@@
 %
-Combining this with the square of Eqn.~\ref{equ:msd_SI} gives the variance in the mean-squared displacement as
+Combining this result with Eqn.~\ref{equ:msd_SI}, we can express the variance in the mean-squared displacement as
 %
@@
     \var{\oMSDn} = 3n^2\kappa^4 - n^2\kappa^4 = 2n^2\kappa^4,
+    \label{equ:varmsd}
 \end{equation}
 %
-i.e., the variance increases quadratically with the number of steps taken.
+i.e., $\var{\oMSDn}$ increases quadratically with the number of steps taken, or, equivalently, with time.
 
-The result above gives the variance in the mean squared displacement of a single particle if we consider a single time origin.
-In practical simulations with more than one mobile particle we can obtain improved statistics by averaging over independent observed squared displacements; either averaging over equivalent mobile particles or averaging over statistically independent sections of the observed simulation trajectory, which we consider to be the case for mutually non-overlapping time intervals.
-Now the variance in MSD is scaled by the usual number of independent contributing observations to give
+Eqn.~\ref{equ:varmsd} gives the variance of the mean squared displacement for a single particle considering a single time-origin.
+We can obtain improved statistics by averaging over statistically equivalent observed squared displacements (see Eqn.~2 in the main text), which can be acheived by averaging over mobile particles or by averaging over time origins.
+This averaging over equivalent observations reduces the variance in the observed MSD to
 %
@@
 where $\nind{n}$ is the total number of statistically independent squared displacements that contribute to $\oMSDi$.
+In the long-time limit, $\nind{n}$ is given by the product of the number of mobile particles and the number of non-overlapping time-windows of length $i$ in our simulation trajectory.
+Note that $\nind{n}$ considers non-overlapping time windows, since mutually overlapping time-windows give correlated squared displacements.
 
-The results in Eqns.~\ref{equ:msd_SI} \&~\ref{equ:der_var} can be extended to a $d$-dimensional lattice random walk, where for each step is along one of the $d$-dimensions, by combining the results for one dimension.
-This means that the MSD becomes 
+The results for a one-dimensional lattice above (Eqns.~\ref{equ:msd_SI} \&~\ref{equ:der_var}) can be extended to a $d$-dimensional lattice, to give
 %
@@
 %
-and the variance in the MSD becomes
+with variance
 %
@@
 \end{equation}
-%
-where the $n$ in both Eqn.~\ref{equ:msd_SI} \&~\ref{equ:der_var} have been replaced with $n/d$ as the steps are spread (equally) over all dimensions.
+%.
+Because each step is equally likely to move a partice along each of the $d$ dimensions, the term $n$ in Eqns.~\ref{equ:msd_SI} \&~\ref{equ:der_var} is replaced here with $n/d$.
 
-This can be extended to consider the covariance between two different numbers of steps, $n$ and $n+m$, in the random walk where the expected MSDs will be
+The analysis above can be extended to consider the covariance between two different numbers of steps, $n$ and $n+m$, in the random walk where the expected MSDs will be
 %
@@
 %
-which can be expanded,
+which can be expanded as
 %
@@
 
-\section{Scaling of the estimation of $\prob{\Dest}$}
+\section{Scaling of $\var{\Dest}$ with simulation size and choice of estimator}
 \label{sec:scaling}
 %
-The true sample variance from \num{512} diffusion coefficient estimates, $\var{\Dest}$, using either the model covariance method, outlined in the main text, the independent method (using estimated variances but ignoring the need for the full covariance matrix, analogous to WLS), the independent \& identical method (analogous to OLS), and using the full numerical covariance matrix as a function of simulation length and number of atoms are compared in Fig.~\ref{fig:stat_eff}.
+Fig.~\ref{fig:stat_eff} presents an analysis of how $\var{\Dest}$ varies as a function of changing the number of mobile particles (Fig.~\ref{fig:stat_eff}a)and the total simulation time (number of steps) (Fig.~\ref{fig:stat_eff}b).
+We compare the performance of four methods for estimating $\D$ from the observed MSD data: OLS, WLS, the approximate Bayesian regression method described here, and GLS.
+Each data point is calculated as the variance over point-estimates of $\D$ from \num{512} individual 3D-lattice random walk simulations, for each combination of $N_\mathrm{atoms}$ and $\Delta t_\mathrm{max}$.
+For WLS and GLS we compute the variances and covariance matrix, respectively, using the complete set of \num{512} simulations.
+For the GLS dataset, this gives an optimally efficient estimator for linear regression of observed MSD data, corresponding to the Cram\'er--Rao lower bound of $\var{\Dest}$.
 
-All estimation methods have the same power law exponent, $a\approx-1$, when the number of atoms is varied, due to the number of observations linearly scaling with number of atoms. 
-This is the observation of the denominator ($n - 1$, where $n$ is the sample size) in the standard variance formula. 
-The numerical and model covariance methods show the same relationship when the maximum simulation length is varied.
-However, the independent and independent \& identical approaches have values for $a$ of \numrange{0.1}{0.3}.
-Doubling the simulation length doubles the number of observations in the first half of the time range, while, the second half contains only new samples.
-The full off-diagonal values of the covariance matrix are necessary to correctly account for these new samples (as the off-diagonal elements link between the different time intervals), leading to the poor scaling of the WLS and OLS analogues that is observed.
-%
-% \begin{table*}
-%     \centering
-%     \caption{Parameters for the scaling of the variance of the diffusion coefficient for truly random walks, from Figs.~\ref{fig:stat_eff}a \& b.}
-%     \label{tab:ipl_true}
-%     \begin{tabular}{c  c | rclcl | c | c | c}
-%          & & numerical & This & Independent & Independent \& Identical \\
-%         \hline
-%         \multirow{2}{*}{$1/N_{\mathrm{atoms}}$} & $a$ & \num{-9.81 \pm 0.06e-1} & \num{-9.03 \pm 0.18e-1} & \num{-9.95 \pm 0.12e-1} & \num{-9.78 \pm 0.14e-1} \\
-%           & $b$/atoms$^{-a}$ & \num{1.42 \pm 0.03e-3} & \num{1.36 \pm 0.08e-3} & \num{7.27 \pm 0.26e-3} & \num{2.15 \pm 0.09e-2} \\
-%         \multirow{2}{*}{$1/\Delta t_{\max}$} & $a$ & \num{-1.20 \pm 0.01e-0} & \num{-1.05 \pm 0.02e-0} & \num{-5.69 \pm 0.17e-1} & \num{-8.33 \pm 1.80e-2} \\ 
-%         & $b$ & \num{3.78 \pm 0.15e-3} & \num{2.74 \pm 0.14e-3} & \num{8.99 \pm 0.51e-4} & \num{2.81 \pm 0.24e-4} \\ 
-%     \end{tabular}
-% \end{table*}
-%
-%
+The approximate Bayesian regression method described in the main manuscript performs similarly to GLS with a converged exact covariance matrix, and gives significantly smaller uncertainty in $\Dest$ than OLS or WLS at all simulation sizes and lengths considered.
+In addition, our method gives better scaling than OLS or WLS as the total simulation time is increased.
+
 \begin{figure}
@@
     \caption{
-        Scaling of statistical efficiency of different approaches;
-        the decrease in $\sigma^2[\hat{\mu}(\Dest_\alpha)]$ as a function of number of atoms (a \& c) and maximum simulation length (b \& d).
-        The solid lines indicate the fitted power law relationship for each of the numerical results (blue), the model covariance approach (green), where the variances are found by rescaling but the data considered independent (orange), and where the variances are considered to be identical and independent (pink).
-        For all except the numerical result, the variance was found from \num{512} determinations of the mean diffusion coefficient from equivalent simulations of walks of \SI{128}{atoms} for \num{128} steps and a step size of \num{1}.
+        Statistical efficiency, $\var{\Dest}$, of OLS (pink), WLS (orange), GLS (blue), and our approximate Bayesian regression method (green), as a function of varying;
+        the decrease in $\sigma^2[\hat{\mu}(\Dest_\alpha)]$ as a function of (a) number of atoms, and (b) total simulation time.
+        The solid lines show a fitted power law relationship for each method.
         % Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences. 
@@
     \script{stat_eff.py}
-\end{figure}
-%
-
-The scaling in the variance that is estimated from each individual simulation using the model covariance method also shows different behaviour between changing the number of atoms and changing the maximum simulation length (Fig.~\ref{fig:efficiency}).
-Initially, we compare the scaling behaviour of the estimated variance with the numerically determined variance (Figs.~\ref{fig:efficiency}a \&~\ref{fig:efficiency}b), where the power law is similar regardless of how the system changes in size.
-The width of the distribution of $\prob{\varest{\Dest}}$ (the error bars in Figs.~\ref{fig:efficiency}a \&~\ref{fig:efficiency}b) also show power law behaviour (Figs~\ref{fig:efficiency}c \&~\ref{fig:efficiency}d).
-In this case, the exponent is different if the number of atoms is changed ($a \approx -2$) compared to the maximum simulation length ($a \approx -1$). 
-The precision of the variance estimate will improve more if the number of atoms is doubled rather than if the simulation is run for twice as long. 
-When the number of atoms is doubled, the number of contributing independent observations also doubles for all time intervals but when the simulation is doubled in length, the number of contributing independent observations is only doubled up to the original simulation length.
-%
-\begin{figure}
-    \centering
-    \includegraphics[width=\columnwidth]{figures/efficiency.pdf}
-    \caption{
-        Scaling of estimated variance (a \& b), estimated variance precision (c \& d), and estimated variance accuracy (e \& f) from the method outlined in the main text;
-        Green points indicate mean values for the estimated variance, where error bars are present they indicate a \SI{95}{\percent} confidence interval. 
-        Blue points indicate the variance found from numerical simulations.
-        The precision is quantified with the difference between the upper and lower bounds of the confidence interval and the accuracy from the bias compared to the numerical result.
-        The variance was found from \num{512} determinations of the diffusion coefficient from equivalent simulations of walks. 
-        When $N_{\mathrm{atoms}}$ was varied $t_{\max}$ was \SI{128} steps and when $t_{\max}$ was varied $N_{\mathrm{atoms}}$ \SI{128}{atoms}; for all simulations the step size was \num{1}.
-        % Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences. 
-        }
-    \label{fig:efficiency}
-    \script{efficiency.py}
-\end{figure}
-%
-
-\section{Skewness of distribution of $\varest{\Dest}$}
-\label{sec:skew}
-%
-The skew observed in the estimated variance for the diffusion coefficient when using the model covariance matrix (Fig.~\ref{fig:random_walk}d) arises from differences between the numerical and model covariance matrices
-If the numerical covariance matrix was used in place of the model covariance matrix, the resulting distribution of $\varest{\Dest}$ is not be skewed (Fig.~\ref{fig:true_cov}).
-Additionally, the distribution of $\Dest$ is no longer broader than the numerical distribution, as is the case in Fig~\ref{fig:random_walk}c, which arises from the model covariance matrix (and analytical covariance matrix) overestimating the variance at low $t$ (Fig.~\ref{fig:covariances}), due to the long-time limit assumption made in its derivation.
-%
-\begin{figure}
-    \centering
-    \includegraphics[width=\columnwidth]{figures/true_cov.pdf}
-    \caption{
-        The cause of the skewed variance estimation when the model covariance matrix is used; 
-        distributions of $\mu(\Dest)$ (a) and $\sigma^2(\Dest)$ (b) determined using the observed mean-squared displacements and the numerical covariance matrix (from \num{1024} equivalent simulations), showing the unskewed distribution of variances, the numerical distribution of $\mu(\Dest)$ is shown with the blue solid line and the sample variance in (a) is shown with the vertical green line in (b).
-        Values found using random walks of \SI{128}{atoms} for \num{128} steps and a step size of \num{1}. 
-        % Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences. 
-        }
-    \label{fig:true_cov}
-    \script{true_cov.py}
 \end{figure}
@@
 Temperature was maintained by means of Nos\'{e}-Hoover thermostat with a relaxation time of \SI{121}{\femto\second} (5000 $\hbar / E_{h}$)~\cite{nose_unified_1984,hoover_canonical_1985,martyna_nose_1992}.
-As in the prior study of Burbano \emph{et al.}~\cite{burbano_sparse_2016}, simulations were performed using $2 \times 2 \times 2$ supercells with \SI{1536}{atoms} in a manner similar to Ref.~\citenum{burbano_sparse_2016} with the system being in the perfect crystalline form at the beginning.
-
-\section{Effect of non-diffusive simulation}
-\label{sec:nondiff}
-%
-In the modelling of the diffusive process, it is necessary to exclude simulation not in the diffusive regime, where the atoms are ``caged'' by their neighbours. 
-The model covariance matrix has been derived assuming the diffusive regime (the Einstein relation implicitly assumes diffusion), therefore only diffusion should be included in our analysis. 
-The non-diffusive regime typically dominates on short-time scales, giving way to diffusion at larger time scales. 
-Fig.~\ref{fig:true_llzo} investigates the effect of using starting the Bayesian linear model at increasing initial values of $t$ (i.e.\ the value of $t$ from which the linear model was applied), in essence reducing the percentage of simulation from non-diffusive particles in the analysis.
-Using simulation from non-diffusive particles leads to a bias in the estimation of the variance of $\prob{\Dest}$, which is reduced when less of the non-diffusive regime is used. 
-%
-\begin{figure*}
-    \centering
-    \includegraphics[width=\textwidth]{figures/true_llzo.pdf}
-    \caption{
-        A comparison of the effect of different starting times for the diffusive regime \SIlist[list-final-separator = {, and }]{0;2;4;6;8;10;15;20}{\pico\second} of the LLZO systems on $\varest{\Dest}$.
-        % Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences. 
-        }
-    \label{fig:true_llzo}
-    \script{true_llzo.py}
-\end{figure*}
-%
-
-To avoid the spurious effect of non-diffusive data, this regime must be excluded. 
-It is useful, in the first instance, to overestimate excluded regime and assess if the resulting estimated variance is small enough to draw conclusions about the data or perform hypothesis testing.  
-A balance can be struck between the start of the diffusive regime (having this as high as possible to ensure minimal non-diffusive data) and the size of the resulting variance.
+As in the prior study of Burbano \emph{et al.}~\cite{burbano_sparse_2016}, simulations were performed using $2 \times 2 \times 2$ supercells with \SI{1536}{atoms} following the same simulation protocal as in Ref.~\citenum{burbano_sparse_2016}.
 
@@
 In this work, the particles were assumed to be diffusing for $t>\SI{4}{\second}$ for the random walks and \SI{10}{\pico\second} for the LLZO system in Fig.~\ref{fig:diffusion}. 
-The LLZO system was investigated at a range of starting times in Fig.~\ref{fig:true_llzo}.
+%The LLZO system was investigated at a range of starting times in Fig.~\ref{fig:true_llzo}.
 

### 2023-06-09 src/tex/ms.tex
Matchign v3 on arXiv

@@
 \usepackage{blindtext}
-% \usepackage[hidelinks]{hyperref}
 \sisetup{
@@
 \newcommand{\papertitle}{Accurate Estimation of Diffusion Coefficients and their Uncertainties from Computer Simulation}
-% \newcommand{\oMSD}{\ensuremath{\overline{\bm{s}^2}}}
-% \newcommand{\oMSDs}[1]{\ensuremath{\overline{s^2}}(#1)}
-% \newcommand{\oMSDi}{\ensuremath{\overline{s^2}_i}}
-% \newcommand{\oMSDj}{\ensuremath{\overline{s^2}_j}}
 
@@
 
-\let\oldaddcontentsline\addcontentsline% Store \addcontentsline
+\let\oldaddcontentsline\addcontentsline
 \renewcommand{\addcontentsline}[3]{}
@@
 The variances, $\var{\oMSDi}$, are correlated, since the displacement of each particle at time $t+\Delta t$ is necessarily similar to its displacement at time $t$, and hence, $\oMSDs{t}$ is similar to $\oMSDs{t+\Delta t}$.
-These variances are also typically unequal---the data are heteroscedastic;
-for example, for a random walk, the MSD variance increases quadratically with time \cite{smith_random_1996}. 
+These variances are also typically unequal---the data are heteroscedastic \cite{smith_random_1996,he_statistical_2018}. 
 Because the key assumptions of the OLS method are not valid for MSD data, OLS gives statistically inefficient estimates of $\D$, while the estimated regression uncertainties obtained from the standard OLS statistical formulae significantly underestimate the true uncertainty in $\prob{\Dest_\mathrm{OLS}}$ (Fig.~\ref{fig:glswlsols}a).
@@
 
-A second method for estimating the ensemble-average MSD, and thus $\Dest$, from simulation data is Bayesian regression.
+An alternative method for estimating the ensemble-average MSD, and thus $\Dest$, from simulation data is Bayesian regression.
 Like GLS, Bayesian regression can take into account both the changing variance and the correlation structure inherent in the data.
@@
         Panel (c) plots the numerical variance against the estimated variance from a single simulation as a function of timestep $i$.
-        %Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences.
     }
@@
 The discrepancy between the model and numerical covariance matrices largely stems from the approximation made in deriving the analytical form that $t$ is large, which leads to an overestimation of the variance at low $t$.
-Despite this, the average estimated covariance matrix reproduces well the correlation structure of the true numerical covariance matrix, indicating that the component covariance matrices estimated from individual simulation trajectories may be used within approximte GLS or Bayesian regression schemes to estimate $\D$ and $\var{\Dest}$.
+Despite this, the average estimated covariance matrix reproduces well the correlation structure of the true numerical covariance matrix, indicating that the covariance matrices estimated from individual simulation trajectories may be used within approximate GLS or Bayesian regression schemes to estimate $\D$ and $\var{\Dest}$.
 
 \section{Validation}
-
-To demonstrate the complete approximate Bayesian regression scheme, as described above, we present two distinct examples.
-First, we consider a simple 3D-lattice random walk, where the true self-diffusion coefficient $\D$ is specified by the simulation parameters, and a well-converged numerical covariance matrix can be obtained with relatively low computational cost, which allows us to directly compare the estimates produced by our method to ``best case'' estimates from a hypothetical method with access to the true covariance matrix.
-Second, we consider an example real-world system---the lithium-ion solid electrolyte \ce{Li7La3Zr2O12} (LLZO)---which represents a practical application of our method to a well-studied material of practical interest for solid-state lithium-ion batteries \cite{MuruganEtAl_AngewChemIntEd2007,burbano_sparse_2016,morgan_lattice_2017,SquiresEtAl_PhysRevMater2022}.
 
@@
 
+To demonstrate the complete approximate Bayesian regression scheme, as described above, we present two distinct examples.
+First, we consider a simple 3D-lattice random walk, where the true self-diffusion coefficient $\D$ is specified by the simulation parameters, and a well-converged numerical covariance matrix can be obtained with relatively low computational cost, which allows us to directly compare the estimates produced by our method to ``best case'' estimates from a hypothetical method with access to the true covariance matrix.
+Second, we consider an example real-world system---the lithium-ion solid electrolyte \ce{Li7La3Zr2O12} (LLZO)---which represents an application of our method to a well-studied material of practical interest for solid-state lithium-ion batteries \cite{MuruganEtAl_AngewChemIntEd2007,burbano_sparse_2016,morgan_lattice_2017,SquiresEtAl_PhysRevMater2022}.
+
 Fig.~\ref{fig:random_walk}a shows the observed MSD from a single 3D-lattice random-walk simulation, along with the estimated posterior distribution of linear models compatible with the observed MSD data, $\prob{\model|\oMSD}$, calculated via Eqns.~\ref{equ:bayes} and \ref{equ:loglike}.
@@
 We also show the true variance of individual point estimates, $\var{\Dest}$, which characterises the spread of the histogram in Fig.~\ref{fig:random_walk}c.
-The distribution of estimated variances is biased relative to the true variance and skewed, due to numerical differences between the true covariance matrix $\mathbf{\Sigma}$ and the model covariance matrix $\mathbf{\Sigma^\prime}$.
+The distribution of estimated variances is biased relative to the true variance and skewed, due to numerical differences between the true covariance matrix $\mathbf{\Sigma}$ and the model covariance matrix $\mathbf{\Sigma^\prime}$ (further details are provided in the SI).
 In general, however, the distribution of the estimated variance shows good agreement with the true sample variance.
@@
 
-\begin{figure}
+\begin{figure}[tb]
     \centering
@@
 Hence, our method provides reasonably accurate estimates of the uncertainty in $\Dest$ for our c-LLZO dataset, even when applied to single effective trajectories with limited displacement data (only 56 mobile ions, and \SI{25}{ps} simulation length).
+
+\section{$\var{\Dest}$ scaling and comparison to OLS, WLS, and GLS}
+
+\begin{figure}[htb]
+    \centering
+    \includegraphics[width=\columnwidth]{figures/stateff.pdf}
+    \caption{
+      Scaling of $\var{\Dest}$ with simulation size for OLS (pink), WLS (blue), our approximate Bayesian regression method (green), and GLS (orange).
+      (a) Scaling versus number of mobile particles, $N_\mathrm{atoms}$.
+      (b) Scaling versus total simulation time, $t_\mathrm{max}$.
+      Solid lines show fitted power law relationships for each dataset.
+      The WLS and GLS data are obtained using numerically determined variances and covariance, respectively, from a set of \num{512} repeat simulations for each combination of $N_\mathrm{atoms}$ and $t_\mathrm{max}$.
+    }
+    \label{fig:stat_eff}
+    \script{stat_eff.py}
+\end{figure}
+
+Fig.~\ref{fig:stat_eff} presents an analysis of the variation in $\var{\Dest}$ as the number of mobile particles (Fig.~\ref{fig:stat_eff}a) and the total simulation time (number of steps) (Fig.~\ref{fig:stat_eff}b) are changed.
+We compare four methods for estimating $\D$ from the observed MSD data: OLS, WLS, the approximate Bayesian regression method described here, and GLS.
+When estimating $\D$ using WLS and GLS, we calculate the variances and the covariance matrix, respectively, numerically, using the complete set of \num{512} simulations.
+Each data point in Fig.~\ref{fig:stat_eff} represents the variance across point-estimates of $\D$ derived from \num{512} individual 3D-lattice random walk simulations, for each combination of $N_\mathrm{atoms}$ and $t_\mathrm{max}$.
+The GLS dataset corresponds to an optimally efficient estimator for linear regression of observed MSD data, equivalent to performing Bayesian regression with the numerical covariance matrix and an uninformative prior.
+
+Our approximate Bayesian regression method performs similarly to GLS, with a numerically converged covariance matrix, and gives significantly reduced uncertainty in $\Dest$ compared to OLS or WLS, for all simulation sizes and lengths considered.
+Moreover, our method scales better than OLS or WLS as the total simulation time is increased.
+The approximate Bayesian regression method in this manuscript therefore presents a significant improvement over more conventional methods such as OLS and WLS, by enabling more precise estimates of $\D$ across varied simulation sizes at equivalent computational cost.
 
@@
 \maketitle
-% \onecolumngrid
 
@@
 \begin{enumerate}
-    % \item Evidence for the effect of the number of independent observations on the distribution of mean-squared displacements.
     \item The derivation of the covariance matrix in the long-time limit for freely diffusion particles.
-    \item Analysis of the relative efficiency of the approximate Bayesian regression method, presented here, compared to OLS, WLS, and GLS, and the efficiency scaling of each method with number of mobile particles or total simulation time..
-    % \item Discussion of the skewed distribution of the estimated variance in the diffusion coefficient.
+    \item Discussion of the origin of bias and skew in the distribution of the estimated variance of the estimated diffusion coefficient, $\prob{\varest{\Dest}}$.
     \item Details of the LLZO molecular dynamics simulations.
-    % \item Discussion of the importance of only using non-diffusive simulation in the analysis.
     \item Details of the implementation of the model covariance method in the \textsc{kinisi} package.
@@
 
-% \vspace{\columnsep}
 \twocolumngrid
@@
 Eqn.~\ref{equ:varmsd} gives the variance of the mean squared displacement for a single particle considering a single time-origin.
-We can obtain improved statistics by averaging over statistically equivalent observed squared displacements (see Eqn.~2 in the main text), which can be acheived by averaging over mobile particles or by averaging over time origins.
+We can obtain improved statistics by averaging over statistically equivalent observed squared displacements (see Eqn.~2 in the main text), which can be achieved by averaging over mobile particles or by averaging over time origins.
 This averaging over equivalent observations reduces the variance in the observed MSD to
@@
 %.
-Because each step is equally likely to move a partice along each of the $d$ dimensions, the term $n$ in Eqns.~\ref{equ:msd_SI} \&~\ref{equ:der_var} is replaced here with $n/d$.
+Because each step is equally likely to move a particle along each of the $d$ dimensions, the term $n$ in Eqns.~\ref{equ:msd_SI} \&~\ref{equ:der_var} is replaced here with $n/d$.
 
@@
         \moMSDn\moMSDm &= \oMSDn\oMSDm \\
-                       &= nd^2(n+m)\kappa^2 \\
+                       &= n\kappa^2(n+m)\kappa^2 \\
                        &= n(n+m)\kappa^4
@@
 %
-The second and third terms in Eqn.~\ref{equ:vbig} tend to \num{0} as there is an equal probability of positive and negative displacements. 
+The second and third terms in Eqn.~\ref{equ:vbig} tend to zero as there is an equal probability of positive and negative displacements. 
 This reduces Eqn.~\ref{equ:vbig} to 
@@
 In this case, the covariance depends only on the number of overlapping points, $n$, between the two time intervals. 
-We can easily rationalise this by accepting that for a random walk any non-overlapping point will be completely uncorrelated and therefore have a covariance of \num{0}. 
+We can rationalise this by noting that for a random walk any non-overlapping points will be completely uncorrelated and therefore have a covariance of \num{0}. 
 Similar to the case for the variance, the covariance derived in Eqn.~\ref{equ:cov_der} is that for a single particle at a single time origin. 
 The number of independent observed squared displacements for a given covariance should be the minimum number of shared independent observed squared displacements between the two time intervals, which is $\nind{n+m}$.
-Therefore, scaled by the number of contributing independent observations, the covariance, in the long-time limit, is
+Therefore, the covariance, scaled by the number of contributing independent observations, in the long-time limit, is
 %
@@
 
-\section{Scaling of $\var{\Dest}$ with simulation size and choice of estimator}
-\label{sec:scaling}
+\section{Bias and skew in $\prob{\varest{\Dest}}$}
+\label{sec:skew}
 %
-Fig.~\ref{fig:stat_eff} presents an analysis of how $\var{\Dest}$ varies as a function of changing the number of mobile particles (Fig.~\ref{fig:stat_eff}a)and the total simulation time (number of steps) (Fig.~\ref{fig:stat_eff}b).
-We compare the performance of four methods for estimating $\D$ from the observed MSD data: OLS, WLS, the approximate Bayesian regression method described here, and GLS.
-Each data point is calculated as the variance over point-estimates of $\D$ from \num{512} individual 3D-lattice random walk simulations, for each combination of $N_\mathrm{atoms}$ and $\Delta t_\mathrm{max}$.
-For WLS and GLS we compute the variances and covariance matrix, respectively, using the complete set of \num{512} simulations.
-For the GLS dataset, this gives an optimally efficient estimator for linear regression of observed MSD data, corresponding to the Cram\'er--Rao lower bound of $\var{\Dest}$.
+In the main manuscript, we present results for a set of \num{4096} 3D-lattice random walk simulations, each consisting of \num{128} particles undergoing \num{128} steps (Fig.~\ref{fig:random_walk}).
+Our approximate Bayesian regression scheme allows us to estimate the variance in $\Dest$, denoted as $\varest{\Dest}$, that would be obtained over a large number of repeat simulations.
+This estimate is calculated from the variance of the marginal posterior distribution $\prob{\D|\model}$, which we derive from analysis of a single simulation trajectory.
+As shown in Fig.~\ref{fig:random_walk}d, our estimate for the population variance $\var{\Dest}$, obtained from a single simulation, aligns reasonably with the true value.
+When considering the distribution of estimated variance, $\prob{\varest{\Dest}}$, however, we observe a systematic overestimation (bias) of the true value, along with visible skewness.
 
-The approximate Bayesian regression method described in the main manuscript performs similarly to GLS with a converged exact covariance matrix, and gives significantly smaller uncertainty in $\Dest$ than OLS or WLS at all simulation sizes and lengths considered.
-In addition, our method gives better scaling than OLS or WLS as the total simulation time is increased.
-
+This bias and skew arise from our use of estimated variances $\varest{\oMSDi}$ when parametrising the model covariance matrix $\bm{\Sigma^\prime}$.
+Fig.~\ref{fig:true_cov} presents equivalent results for $\prob{\Dest}$ and $\prob{\varest{\Dest}}$ for the same \num{4096} individual simulations, but calculated using a numerical covariance matrix, $\bm{\Sigma}_\mathrm{num}$ derived from all \num{4096} observed MSDs.
+The resulting distribution $\prob{\varest{\Dest}}$ (Fig.~\ref{fig:true_cov}b) is no longer biased or skewed.
+Furthermore, the distribution $\prob{\Dest}$ agrees even more closely with the numerically converged distribution obtained when combining data from all \num{4096} simulations (Fig~\ref{fig:true_cov}a), contrasting with the results presented in Fig.~\ref{fig:random_walk}b, where our approximate Bayesian regression scheme yields a slightly broadened distribution due to the use of the long-time limit in the derivation of the analytical form for $\bm{\Sigma^\prime}$.
+%
 \begin{figure}
     \centering
-    \includegraphics[width=\columnwidth]{figures/stat_eff.pdf}
+    \includegraphics[width=\columnwidth]{figures/true_cov.pdf}
     \caption{
-        Statistical efficiency, $\var{\Dest}$, of OLS (pink), WLS (orange), GLS (blue), and our approximate Bayesian regression method (green), as a function of varying;
-        the decrease in $\sigma^2[\hat{\mu}(\Dest_\alpha)]$ as a function of (a) number of atoms, and (b) total simulation time.
-        The solid lines show a fitted power law relationship for each method.
-        % Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences. 
+        (a) Probability distribution of point-estimates $\prob{\Dest}$ obtained from \num{4096} individual random-walk simulations, using the numerical covariance matrix $\bm{\Sigma}_\mathrm{num}$.
+        Each simulation has been analysed as in Fig.~\ref{fig:msd}(a) and (b) to yield a single corresponding point estimate $\Dest$.
+        The grey line shows the distribution of point estimates, $\prob{\Dest_\mathrm{num}}$, obtained using Bayesian regression with a mean vector and numerical covariance matrix derived from the complete dataset of all \num{4096} simulations.
+        The pink horizontal bar shows an interval of one standard deviation in $\prob{\Dest}$.
+        (b) Probability distribution of estimated variances, $\varest{\Dest}$, for individual random-walk simulations, using the numerical covariance matrix $\bm{\Sigma}_\mathrm{num}$, compared to the true sample variance (pink vertical line) $\var{\Dest}$.
         }
-    \label{fig:stat_eff}
-    \script{stat_eff.py}
+    \label{fig:true_cov}
+    \script{true_cov.py}
 \end{figure}
@@
 Classical molecular dynamics were run using the \textsc{metalwalls} code~\cite{marin_metalwalls_2020}. 
-The DIPPIM polarisable ion force field parameterised by Burbano \emph{et al.}~\cite{burbano_sparse_2016} was used due to its ability to account accurately for the effect of ion polarisability on diffusion~\cite{wilson_polarization_1993,burbano_sparse_2016}.
-Simulations were run for the cubic phase of LLZO in NVT ensemble at \SI{1000}{\kelvin} and were extended to a length of \SI{2}{\nano\second} with a \SI{2}{\femto\second} timestep. 
-Temperature was maintained by means of Nos\'{e}-Hoover thermostat with a relaxation time of \SI{121}{\femto\second} (5000 $\hbar / E_{h}$)~\cite{nose_unified_1984,hoover_canonical_1985,martyna_nose_1992}.
-As in the prior study of Burbano \emph{et al.}~\cite{burbano_sparse_2016}, simulations were performed using $2 \times 2 \times 2$ supercells with \SI{1536}{atoms} following the same simulation protocal as in Ref.~\citenum{burbano_sparse_2016}.
+We used the DIPPIM polarisable ion force field, as parameterised by Burbano \emph{et al.}~\cite{burbano_sparse_2016}, due to its proven accuracy in accounting for the effect of ion polarisability on diffusion~\cite{wilson_polarization_1993,burbano_sparse_2016}.
+We simulated the cubic phase of LLZO in NVT ensemble at a temperature of \SI{1000}{\kelvin}.
+Simulations were run for \SI{2}{\nano\second} with a \SI{2}{\femto\second} timestep. 
+To control temperature, we used a Nos\'{e}-Hoover thermostat, with a relaxation time of \SI{121}{\femto\second} (5000 $\hbar / E_{h}$)~\cite{nose_unified_1984,hoover_canonical_1985,martyna_nose_1992}.
+Simulations were performed using $2 \times 2 \times 2$ supercells with \SI{1536}{atoms} following the same protocol as in Ref.~\citenum{burbano_sparse_2016}.
 
-\section{Numerical implementation in \textsc{kinisi}}
+\section{Numerical implementation in KINISI}
 \label{sec:implementation}
 %
-We have implemented the method outlined in the main text in the open-source Python package \textsc{kinisi}~\cite{mccluskey_kinisi_2022}. 
-Here, we will discuss both \textsc{kinisi}'s broader implementation of the method and its application in the validation examples.
+We have implemented the approximate Bayesian estimation method described in the main text in the open-source Python package \textsc{kinisi}~\cite{mccluskey_kinisi_2022}. 
+This section provides numerical details regarding the implementation of this method.
 
-First, to find the maximum number of observations of squared displacements at each time interval, \textsc{kinisi} uses overlapping sliding window sampling. 
-The maximum number of observations for a given time interval, $t$, is $N_{\mathrm{atoms}} \times (N_{t} - i)$ displacements, where $N_{\mathrm{atoms}}$ is the number of atoms, $N_{t}$ is the total number of timesteps, and $i$ is the index of the timestep (starting with \num{1} for the shortest timestep).
-The independent variance is estimated by rescaling the observed overlapping variance in the squared displacements by the number of non-overlapping squared displacements, $\nind{i} = N_{\mathrm{atoms}} \times N_{t} / i$, as shown in Eqn.~\ref{equ:varestMSD}.
-The population mean for the squared displacements and the independent variances are then stored.
+\textsc{kinisi} uses overlapping sliding window sampling when calculating the observed mean squared displacement at each time interval $t$ (see Eqn.~\ref{equ:observed_msd}).
+For a given time interval, $t$, the maximum number of observations is $N_{\mathrm{atoms}} \times (N_{t} - i)$ displacements, where $N_{\mathrm{atoms}}$ is the number of atoms, $N_{t}$ is the total number of timesteps, and $i$ is the index of the timestep (where \num{1} is the index for the shortest timestep).
+To estimate the variance of the observed MSD, we rescale the variance of observed squared displacements by the number of non-overlapping time intervals in the simulation, $\nind{i} = N_{\mathrm{atoms}} \times N_{t} / i$, as presented in Eqn.~\ref{equ:varestMSD}.
 
-Eqn.~\ref{equ:cvv_SI} defines the parameterisation of the covariance matrix from the variances and the number of independent observations. 
-The covariance matrix is only constructed for values of $t$ where the particles are considered to be diffusing, which can be controlled manually by the user. 
-In this work, the particles were assumed to be diffusing for $t>\SI{4}{\second}$ for the random walks and \SI{10}{\pico\second} for the LLZO system in Fig.~\ref{fig:diffusion}. 
-%The LLZO system was investigated at a range of starting times in Fig.~\ref{fig:true_llzo}.
+The parametrisation of the covariance matrix from the variances $\var{\oMSDi}$ and the number of independent observations $\nind{i}$ is defined by Eqn.~\ref{equ:cvv}. 
+The covariance matrix is only constructed for values of $t$ where the particle motion is considered to be in the long-time diffusive limit, with this threshold set by the user to a value appropriate for their system and simulation data.
+For the examples presented in the main manuscript, we consider particles to be in the diffusive regime from $t=\num{4}$ for the random walk trajectories and from $t=\SI{10}{\pico\second}$ for the LLZO simulations..
 
-To obtain an initial guess for the gradient and intercept of the linear model that describes the observed MSD, \textsc{kinisi} uses ordinary least squares. 
-This initial guess is then used as a starting point for the minimisation of the negative maximum a posteriori (the maximum of the posterior distribution in Eqn.~\ref{equ:bayes}), with the prior probability that the diffusion coefficient is greater than or equal to zero~\cite{broyden_convergence_1970,fletcher_new_1970,goldfarb_family_1970,shanno_conditioning_1970}.
-The log-likelihood calculation (Eqn.~\ref{equ:loglike}) makes use of the Moore-Penrose generalisation of the inverse of a Hermitian matrix~\cite{moore_on_1920,bjerhammar_application_1951,penrose_generlized_1955}.
+\textsc{kinisi} uses ordinary least squares to obtain an initial guess for the gradient and intercept of the linear model describing the observed MSD.
+This initial guess is then used as the starting point for minimising the negative maximum a posteriori (the peak of the posterior distribution as per Eqn.~\ref{equ:bayes}), with the improper prior that $\D \ge 0$~\cite{broyden_convergence_1970,fletcher_new_1970,goldfarb_family_1970,shanno_conditioning_1970}.
+The log-likelihood calculation (Eqn.~\ref{equ:loglike}) uses the Moore-Penrose generalisation of the inverse of a Hermitian matrix~\cite{moore_on_1920,bjerhammar_application_1951,penrose_generlized_1955}.
 
-To sample the joint posterior probability distribution of the linear model, \textsc{kinisi} uses the \textsc{emcee} package~\cite{foremanmackey_emcee_2019} was used, which implements Goodman and Weare's affine invariant Markov chain Monte Carlo ensemble sampler~\cite{goodman_ensemble_2010}.
-A physically-informed prior probability that $\D > 0$ is imposed on all analysis.
-The sampling is performed with \num{32} walkers, for \num{1500} steps, the first \num{500} steps are removed as a burn-in period, finally, the sampled chains are thinned, such that only every \num{10}th value is used.
-This gives \num{3200} points sampled in the posterior distribution, these points can be histogrammed (as in Fig~\ref{fig:random_walk}) and summary statistics taken.
+To sample the joint posterior probability distribution of the linear model, \textsc{kinisi} uses the \textsc{emcee} package~\cite{foremanmackey_emcee_2019}, which implements Goodman and Weare's affine invariant Markov chain Monte Carlo ensemble sampler~\cite{goodman_ensemble_2010}.
+When sampling $\prob{\D|\model}$ we again apply the improper prior $\D \ge 0$.
+The sampling process uses \num{32} walkers for \num{1500} steps, with the first \num{500} steps discarded as a burn-in period.
+The sampled chains are thinned such that only every 10th value is retained, yielding \num{3200} points sampled from the posterior distribution $\prob{\D|\model}$.
+These points can then be plotted as a histogram (as in Fig~\ref{fig:random_walk}), and summary statistics $\Dest$ and $\varest{\Dest}$ can be derived.
 

### 2023-06-09 src/code/random_walks/glswlsols.py

@@ -12,16 +12,6 @@ from kinisi.diffusion import _straight_line
 
 np.random.seed(1)
 
-
-def nll(*args):
-    """
-    General purpose negative log-likelihood.
-    Returns:
-        (:py:attr:`float`): Negative log-likelihood.
-    """
-    return -log_likelihood(*args)
-
-
 atoms = int(snakemake.params['atoms'])
 length = int(snakemake.params['length'])
 jump = int(snakemake.params['jump'])
@@ -37,100 +27,20 @@ V = np.diag(true_cov.diagonal())
 
 g1 = np.matmul(np.linalg.inv(np.matmul(A.T, np.matmul(np.linalg.pinv(W), A))),
                np.matmul(A.T, np.matmul(np.linalg.pinv(W), y)))[0] / 6
+c1 = np.sqrt(np.linalg.inv(np.matmul(A.T, np.matmul(np.linalg.pinv(W), A)))[0][0]) / 6
 g2 = np.matmul(np.linalg.inv(np.matmul(A.T, np.matmul(np.linalg.pinv(V), A))),
                np.matmul(A.T, np.matmul(np.linalg.pinv(V), y)))[0] / 6
+c2 = np.sqrt(np.linalg.inv(np.matmul(A.T, np.matmul(np.linalg.pinv(V), A)))[0][0]) / 6
 g3 = np.matmul(np.linalg.inv(np.matmul(A.T, A)), np.matmul(A.T, y))[0] / 6
 
-W = find_nearest_positive_definite(W)
-
-_, logdet = np.linalg.slogdet(W)
-logdet += np.log(2 * np.pi) * true_msd[0].size
-inv = pinvh(W)
-
-
-def log_likelihood(theta: np.ndarray) -> float:
-    """
-    Get the log likelihood for multivariate normal distribution.
-    :param theta: Value of the gradient and intercept of the straight line.
-    :return: Log-likelihood value.
-    """
-    if theta[0] < 0:
-        return -np.inf
-    model = _straight_line(timestep, *theta)
-    diff = (model - true_msd[0])
-    logl = -0.5 * (logdet + np.matmul(diff.T, np.matmul(inv, diff)))
-    return logl
-
-
-ols = linregress(timestep, true_msd[0])
-
-max_likelihood = minimize(nll, np.array([ols.slope, ols.intercept])).x
-pos = max_likelihood + max_likelihood * 1e-3 * np.random.randn(32, max_likelihood.size)
-sampler = EnsembleSampler(*pos.shape, log_likelihood)
[…]

### 2023-06-09 src/code/random_walks/ordinary_rw.py

@@ -16,8 +16,10 @@ atoms = snakemake.params['atoms']
 length = int(snakemake.params['length'])
 size = int(snakemake.params['n'])
 
-gradient = np.zeros((size, 3200))
-intercept = np.zeros((size, 3200))
+gradient = np.zeros((size))
+gradient_err = np.zeros((size))
+intercept = np.zeros((size))
+intercept_err = np.zeros((size))
 data = np.load(f'src/data/random_walks/kinisi//rw_1_{atoms}_{length}_s{size}.npz')['data']
 
 for seed in tqdm(range(size)):
@@ -26,47 +28,15 @@ for seed in tqdm(range(size)):
 
     x = data[seed, 4:, 0]
     y = data[seed, 4:, 1]
-    i_matrix = np.eye(y.size)
-    add = np.log(2 * np.pi) * y.size
-
-    def log_likelihood(theta: np.ndarray) -> float:
-        """
-        Get the log likelihood for multivariate normal distribution.
-        :param theta: Value of the gradient and intercept of the straight line.
-        :return: Log-likelihood value.
-        """
-        if theta[0] < 0:
-            return -np.inf
-        model = x * theta[0] + theta[1]
-        diff = (model - y)
-        var = diff.var()
-        logdet = np.log(var ** y.size)
-        logdet += np.log(2 * np.pi) * y.size
-        inv = i_matrix / var
-        logl = -0.5 * (logdet + np.matmul(diff.T, np.matmul(inv, diff)))
-        return logl
-
-    ols = linregress(x, y)
-    slope = ols.slope
-    c = 1e-20
-    if slope < 0:
-        slope = 1e-20
-
-    def nll(*args) -> float:
-        """
-        General purpose negative log-likelihood.
-        :return: Negative log-likelihood
-        """
-        return -log_likelihood(*args)
-
-    max_likelihood = minimize(nll, np.array([slope, c])).x
-    pos = max_likelihood + max_likelihood * 1e-3 * np.random.randn(32, max_likelihood.size)
-    sampler = EnsembleSampler(*pos.shape, log_likelihood)
-    sampler.run_mcmc(pos, 1000 + 500, progress=False)
-    flatchain = sampler.get_chain(flat=True, thin=10, discard=500)
-    gradient[seed] = flatchain[:, 0] / 6
-    intercept[seed] = flatchain[:, 1]
+    lin = linregress(x, y)
+    
+    gradient[seed] =  lin.slope / 6
[…]

### 2023-06-09 src/code/random_walks/stat_eff.py

@@ -18,22 +18,21 @@ for i, ii in enumerate(xaxis):
     file_open = np.load(f"src/data/random_walks/kinisi/rw_{jump}_{ii}_{length}_s512.npz")
     kinisi_atoms[i] = file_open['diff_c'].mean(-1).var(-1, ddof=1)
     file_open = np.load(f"src/data/random_walks/weighted/rw_{jump}_{ii}_{length}_s512.npz")
-    weighted_atoms[i] = file_open['diff_c'].mean(-1).var(-1, ddof=1)
+    weighted_atoms[i] = file_open['diff_c'].var()
     file_open = np.load(f"src/data/random_walks/ordinary/rw_{jump}_{ii}_{length}_s512.npz")
-    ordinary_atoms[i] = file_open['diff_c'].mean(-1).var(-1, ddof=1)
+    ordinary_atoms[i] = file_open['diff_c'].var()
     file_open = np.load(f"src/data/random_walks/numerical/D_{jump}_{ii}_{length}.npz")
-    numerical_atoms[i] = file_open['diff_c'].var(ddof=1)
+    numerical_atoms[i] = file_open['diff_c'].var()
 for i, ii in enumerate(xaxis):
     atoms = 128
     file_open = np.load(f"src/data/random_walks/kinisi/rw_{jump}_{atoms}_{ii}_s512.npz")
     kinisi_length[i] = file_open['diff_c'].mean(-1).var(-1, ddof=1)
     file_open = np.load(f"src/data/random_walks/weighted/rw_{jump}_{atoms}_{ii}_s512.npz")
-    weighted_length[i] = file_open['diff_c'].mean(-1).var(-1, ddof=1)
+    weighted_length[i] = file_open['diff_c'].var()
     file_open = np.load(f"src/data/random_walks/ordinary/rw_{jump}_{atoms}_{ii}_s512.npz")
-    ordinary_length[i] = file_open['diff_c'].mean(-1).var(-1, ddof=1)
+    ordinary_length[i] = file_open['diff_c'].var()
     file_open = np.load(f"src/data/random_walks/numerical/D_{jump}_{atoms}_{ii}.npz")
-    numerical_length[i] = file_open['diff_c'].var(ddof=1)
-
+    numerical_length[i] = file_open['diff_c'].var()
 
 def f(x, a, b):
     return b * x**a

### 2023-06-09 src/code/random_walks/true_ls.py

@@ -9,57 +9,42 @@ from emcee import EnsembleSampler
 from scipy.optimize import minimize
 from kinisi.matrix import find_nearest_positive_definite
 from kinisi.diffusion import _straight_line
+from tqdm import tqdm
 
 np.random.seed(1)
 
 atoms = int(snakemake.params['atoms'])
 length = int(snakemake.params['length'])
 jump = int(snakemake.params['jump'])
+size = 4096
 
-true_msd = np.load(f'src/data/random_walks/numerical/rw_1_{atoms}_{length}_s4096.npz')['data']
-timestep = np.arange(1, length + 1, 1, dtype=int)
+true_msd = np.load(f'src/data/random_walks/numerical/rw_1_{atoms}_{length}_s{size}.npz')['data'][:, 4:]
+timestep = np.arange(1, length + 1, 1, dtype=int)[4:]
 true_mean = true_msd.mean(0)
-rcov = np.cov(true_msd.T)
+W = np.cov(true_msd.T)
 
-max_ngp = np.argwhere(timestep > 4)[0][0]
-rcov = find_nearest_positive_definite(rcov[max_ngp:, max_ngp:])
+gradient = np.zeros((size))
+gradient_err = np.zeros((size))
+intercept = np.zeros((size))
+intercept_err = np.zeros((size))
 
-_, logdet = np.linalg.slogdet(rcov)
-logdet += np.log(2 * np.pi) * true_msd.mean(0)[max_ngp:].size
-inv = pinvh(rcov)
+for seed in tqdm(range(size)):
+    rng = np.random.RandomState(seed)
+    np.random.seed(seed)
 
+    A = np.array([timestep, np.ones(timestep.size)]).T
+    y = true_msd[seed].T
+    g1 = np.matmul(np.linalg.inv(np.matmul(A.T, np.matmul(np.linalg.pinv(W), A))),
+                np.matmul(A.T, np.matmul(np.linalg.pinv(W), y)))
+    c1 = np.sqrt(np.linalg.inv(np.matmul(A.T, np.matmul(np.linalg.pinv(W), A))))
 
-def log_likelihood(theta: np.ndarray) -> float:
-    """
-    Get the log likelihood for multivariate normal distribution.
-    :param theta: Value of the gradient and intercept of the straight line.
-    :return: Log-likelihood value.
-    """
-    if theta[0] < 0:
-        return -np.inf
-    model = _straight_line(timestep[max_ngp:], *theta)
[…]

### 2023-06-09 src/code/random_walks/weighted_rw.py

@@ -15,58 +15,32 @@ atoms = snakemake.params['atoms']
 length = int(snakemake.params['length'])
 size = int(snakemake.params['n'])
 
-gradient = np.zeros((size, 3200))
-intercept = np.zeros((size, 3200))
+gradient = np.zeros((size))
+gradient_err = np.zeros((size))
+intercept = np.zeros((size))
+intercept_err = np.zeros((size))
 data = np.load(f'src/data/random_walks/kinisi/rw_1_{atoms}_{length}_s{size}.npz')
+
+V = np.diag(data['data'][:, 4:, 1].var(0))
 
 for seed in tqdm(range(size)):
     rng = np.random.RandomState(seed)
     np.random.seed(seed)
 
-    x = data['data'][seed, 4:, 0]
-    y = data['data'][seed, 4:, 1]
-    yerr = data['data'][seed, 4:, 2]  
+    A = np.array([data['data'][seed, 4:, 0], np.ones(data['data'][seed, 4:, 0].size)]).T
+    y = data['data'][seed, 4:, 1].T
 
-    rcov = np.diag(yerr ** 2)
+    g2 = np.matmul(np.linalg.inv(np.matmul(A.T, np.matmul(np.linalg.pinv(V), A))),
+                np.matmul(A.T, np.matmul(np.linalg.pinv(V), y)))
+    c2 = np.sqrt(np.linalg.inv(np.matmul(A.T, np.matmul(np.linalg.pinv(V), A))))
 
-    _, logdet = np.linalg.slogdet(rcov)
-    logdet += np.log(2 * np.pi) * y.size
-    inv = pinvh(rcov)
-
-    def log_likelihood(theta: np.ndarray) -> float:
-        """
-        Get the log likelihood for multivariate normal distribution.
-        :param theta: Value of the gradient and intercept of the straight line.
-        :return: Log-likelihood value.
-        """
-        if theta[0] < 0:
-            return -np.inf
-        model = _straight_line(x, *theta)
-        diff = (model - y)
-        logl = -0.5 * (logdet + np.matmul(diff.T, np.matmul(inv, diff)))
-        return logl
-
-    ols = linregress(x, y)
-    slope = ols.slope
-    c = 1e-20
-    if slope < 0:
-        slope = 1e-20
-
-    def nll(*args) -> float:
-        """
-        General purpose negative log-likelihood.
-        :return: Negative log-likelihood
-        """
[…]

### 2023-06-09 src/scripts/glswlsols.py

@@ -27,7 +27,7 @@ axes[-1].plot([1 - data['ols_pop'].std(), 1 + data['ols_pop'].std()],
               [axes[-1].get_ylim()[1] * 1.1, axes[-1].get_ylim()[1] * 1.1],
               color=fp.colors[3],
               marker='|')
-axes[-1].plot([1 - data['ols_est'][:, 0].std(), 1 + data['ols_est'][:, 0].std()],
+axes[-1].plot([1 - data['ols_est'], 1 + data['ols_est']],
               [axes[-1].get_ylim()[1] * 1.1, axes[-1].get_ylim()[1] * 1.1],
               color='#1E5B84',
               marker='|')
@@ -45,7 +45,7 @@ axes[-1].plot([1 - data['wls_pop'].std(), 1 + data['wls_pop'].std()],
               [axes[-1].get_ylim()[1] * 1.1, axes[-1].get_ylim()[1] * 1.1],
               color=fp.colors[3],
               marker='|')
-axes[-1].plot([1 - data['wls_est'][:, 0].std(), 1 + data['wls_est'][:, 0].std()],
+axes[-1].plot([1 - data['wls_est'], 1 + data['wls_est']],
               [axes[-1].get_ylim()[1] * 1.1, axes[-1].get_ylim()[1] * 1.1],
               color='#1E5B84',
               marker='|')
@@ -61,7 +61,7 @@ axes[-1].plot([1 - data['gls_pop'].std(), 1 + data['gls_pop'].std()],
               [axes[-1].get_ylim()[1] * 1.1, axes[-1].get_ylim()[1] * 1.1],
               color=fp.colors[3],
               marker='|')
-axes[-1].plot([1 - data['gls_est'][:, 0].std(), 1 + data['gls_est'][:, 0].std()],
+axes[-1].plot([1 - data['gls_est'], 1 + data['gls_est']],
               [axes[-1].get_ylim()[1] * 1.1, axes[-1].get_ylim()[1] * 1.1],
               color='#1E5B84',
               marker='|')

### 2023-06-09 src/scripts/random_walk.py

@@ -92,17 +92,17 @@ axes[-1].legend(loc='upper left', bbox_to_anchor=(0.6, 1))
 
 axes.append(fig.add_subplot(gs[0, 3]))
 y, x = np.histogram(true.var(-1, ddof=1), bins=fp.NBINS, density=True)
-axes[-1].stairs(y * 1e-3, x * 1e3, fill=True, color=fp.colors[0], label='$p[\hat{\sigma}^2(D^*)]$')
-axes[-1].axvline(true.mean(-1).var(ddof=1) * 1e3,
+axes[-1].stairs(y, x, fill=True, color=fp.colors[0], label='$p[\hat{\sigma}^2(D^*)]$')
+axes[-1].axvline(true.mean(-1).var(ddof=1),
                  ymax=0.92,
                  c=fp.colors[3],
                  label='$\sigma^2(\hat{D}^*)$',
                  ls='--')
-axes[-1].set_xlabel(r"$\hat{\sigma}^2 (\hat{D}^*)$ / $10^{-3}$")
-axes[-1].set_ylabel(r"$p[\hat{\sigma}^2 (\hat{D}^*)]$ / $10^3$")
+axes[-1].set_xlabel(r"$\hat{\sigma}^2 (\hat{D}^*)$")
+axes[-1].set_ylabel(r"$p[\hat{\sigma}^2 (\hat{D}^*)]$")
 axes[-1].set_xlim([0, None])
 # axes[-1].set_xticks([0, 2, 4])
-axes[-1].set_yticks([0, 1, 2])
+# axes[-1].set_yticks([0, 1, 2])
 titles.append("d")
 axes[-1].set_title("Many simulations")
 axes[-1].legend(loc='upper left', bbox_to_anchor=(0.5, 1))

### 2023-06-09 src/scripts/stat_eff.py

@@ -24,7 +24,7 @@ def f(x, a, b):
 
 xaxis = np.array([16, 32, 64, 128, 256, 512, 1024])
 
-figsize = (3.66, 2)
+figsize = (3.66, 1.5)
 fig = plt.figure(figsize=figsize)
 gs = gridspec.GridSpec(1, 2, figure=fig, wspace=0.6, hspace=0.4)
 
@@ -35,46 +35,6 @@ titles = []
 letters = ['a', 'b']
 
 data = np.load(paths.data / f"random_walks/stat_eff.npz")
-
-write_out = open(paths.output / f"stat_eff_table.txt", 'w')
-
-write_out.write(r"\multirow{2}{*}{" + r"$1/N_{\mathrm{atoms}}$" + r"} ")
-write_out.write(r"& $a$ & \num{" + f"{data['na'] * 10 ** find_exp(data['na']):.2f}" + " \pm " +
-                f"{data['dna'] * 10 ** find_exp(data['na']):.2f}e-{find_exp(data['na'])}" + r"} ")
-write_out.write(r"& \num{" + f"{data['ka'] * 10 ** find_exp(data['ka']):.2f}" + " \pm " +
-                f"{data['dka'] * 10 ** find_exp(data['ka']):.2f}e-{find_exp(data['ka'])}" + r"} ")
-write_out.write(r"& \num{" + f"{data['wa'] * 10 ** find_exp(data['wa']):.2f}" + " \pm " +
-                f"{data['dwa'] * 10 ** find_exp(data['wa']):.2f}e-{find_exp(data['wa'])}" + r"} ")
-write_out.write(r"& \num{" + f"{data['pa'] * 10 ** find_exp(data['pa']):.2f}" + " \pm " +
-                f"{data['dpa'] * 10 ** find_exp(data['pa']):.2f}e-{find_exp(data['pa'])}" + r"} \\ ")
-write_out.write(r" & $b$/atoms$^{-a}$ & \num{" + f"{data['nb'] * 10 ** find_exp(data['nb']):.2f}" + " \pm " +
-                f"{data['dnb'] * 10 ** find_exp(data['nb']):.2f}e-{find_exp(data['nb'])}" + r"} ")
-write_out.write(r"& \num{" + f"{data['kb'] * 10 ** find_exp(data['kb']):.2f}" + " \pm " +
[…]

### 2023-06-09 src/scripts/true_cov.py

@@ -30,9 +30,9 @@ x_ = np.linspace(true.mean(-1).min(), true.mean(-1).max(), 1000)
 
 axes.append(fig.add_subplot(gs[0, 0]))
 y, x = np.histogram(true.mean(-1), density=True, bins=fp.NBINS)
-axes[-1].stairs(y * 1e-1, x, fill=True, color=fp.colors[2], label='$p(\hat{D}^*)$')
+axes[-1].stairs(y, x, fill=True, color=fp.colors[2], label='$p(\hat{D}^*)$')
 axes[-1].plot(x_,
-              norm.pdf(x_, dinfty_true.mean(), dinfty_true.std()) * 1e-1,
+              norm.pdf(x_, dinfty_true.mean(), dinfty_true.std()),
               color=fp.colors[4],
               label=r'$p(\hat{D}^* | \mathbfit{x})$')
 axes[-1].plot([dinfty_true.mean() - dinfty_true.std(),
@@ -42,13 +42,14 @@ axes[-1].plot([dinfty_true.mean() - dinfty_true.std(),
 axes[-1].set_xlabel(r"$\hat{D}^*$")
 axes[-1].set_ylabel(r"$p(\hat{D}^*)$")
 titles.append(r"a")
+axes[-1].set_xlim(0.9, 1.1)
 axes[-1].set_title("$D^*$ distribution")
 axes[-1].legend(loc='upper left', bbox_to_anchor=(0.65, 1))
 
 axes.append(fig.add_subplot(gs[0, 1]))
 y, x = np.histogram(true.var(-1, ddof=1), density=True, bins=fp.NBINS)
-axes[-1].stairs(y * 1e-5, x * 1e5, fill=True, color=fp.colors[0], label='$p[\hat{\sigma}^2(D^*)]$')
-axes[-1].axvline(true.mean(-1).var(ddof=1) * 1e5, c=fp.colors[3], label='$\sigma^2(\hat{D}^*)$', ls='--')
+axes[-1].stairs(y, x, fill=True, color=fp.colors[0], label='$p[\hat{\sigma}^2(D^*)]$')
+axes[-1].axvline(true.mean(-1).var(ddof=1), c=fp.colors[3], label='$\sigma^2(\hat{D}^*)$', ls='--')
 axes[-1].set_xlabel(r"$\hat{\sigma}^2 (\hat{D}^*)$")
[…]

### 2023-06-09 src/scripts/efficiency.py
Remove unused scripts

@@ -1,103 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import matplotlib.gridspec as gridspec
-from scipy.stats import norm, skew
-from uravu.distribution import Distribution
-from uravu.axis import Axis
-
-import utils._fig_params as fp
-from utils.plotting_helper import to_string, mid_points
-import paths
-
-jump = 1
-
-xaxis = np.array([16, 32, 64, 128, 256, 512, 1024])
-
-figsize = (3.64, 4.5)
-fig = plt.figure(figsize=figsize)
-gs = gridspec.GridSpec(2, 2, figure=fig, wspace=0.6, hspace=0.4)
-axes = []
-titles = []
-letters = ['a', 'b', 'c', 'd', 'e', 'f', 'g', 'h', 'i', 'j', 'k', 'l']
-
-data = np.load(paths.data / f"random_walks/efficiency.npz")
-
-axes.append(fig.add_subplot(gs[0, 0]))
-axes[-1].plot(xaxis, data['ty'], marker='.', ls="", color=fp.colors[0])
-a = Axis([Distribution(i.var(-1, ddof=1)) for i in data['ky']])
-n = [np.mean(i.var(-1, ddof=1)) for i in data['ky']]
-axes[-1].errorbar(xaxis, n, a.ci(), marker='.', ls="", color=fp.colors[2])
-axes[-1].set_xscale('log', base=2)
-axes[-1].set_yscale('log', base=2)
-axes[-1].set_xlabel(r'$N_{{\mathrm{{atoms}}}}$ / atoms')
-axes[-1].set_ylabel(r"$\hat{\sigma}^2 (\hat{D}^*)$")
-axes[-1].set_xticks([32, 128, 512])
-axes[-1].set_xticklabels([32, 128, 512])
-axes[-1].minorticks_off()
-titles.append(f'{letters[0]}')
-axes[-1].set_title(r"$\hat{\sigma}^2(\hat{D}^*)$ scaling")
-
-axes.append(fig.add_subplot(gs[1, 0]))
-axes[-1].plot(xaxis, (a.ci()[1] - a.ci()[0]) * 1e5, '.', c=fp.colors[2])
-axes[-1].set_xscale('log', base=2)
-axes[-1].set_yscale('log', base=2)
-axes[-1].set_xlabel(r'$N_{{\mathrm{{atoms}}}}$ / atoms')
[…]

### 2023-06-09 src/scripts/true_llzo.py

@@ -1,109 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import matplotlib.gridspec as gridspec
-from matplotlib.ticker import FuncFormatter
-from scipy.stats import norm
-from uravu.distribution import Distribution
-from uravu.axis import Axis
-
-import utils._fig_params as fp
-from utils.plotting_helper import mid_points
-import paths
-
-jump = 1
-atoms = 128
-length = 128
-correlation = 'true'
-type = 'kinisi'
-
-start_diffs = [0, 2, 4, 6, 8, 10, 15, 20]
-d = np.zeros((len(start_diffs), 16, 8, 4, 3200))
-timestep = np.load(paths.data / f"llzo/diffusion_0_10.npz")['dt'][0, 0]
-dinfty_true = np.zeros((len(start_diffs), 3200))
-cinfty_true = np.zeros((len(start_diffs), 3200))
-for j, start in enumerate(start_diffs):
-    dinfty_true[j] = np.load(paths.data / f"llzo/true_{start}.npz")['diff_c']
-    cinfty_true[j] = np.load(paths.data / f"llzo/true_{start}.npz")['intercept']
-    for i in range(0, 16, 1):
-        d[j, i] = np.load(paths.data / f'llzo/diffusion_{i}_{start}.npz')['d']
-d = d.reshape(len(start_diffs), -1, 3200)
-
-figsize = (8.455, 3)
-fig = plt.figure(figsize=figsize)
-gs = gridspec.GridSpec(2, 4, figure=fig, wspace=0.6, hspace=1)
-letters = [['a', 'b', 'c', 'd'], ['e', 'f', 'g', 'h']]
-times = [[0, 2, 4, 6], [8, 10, 15, 20]]
-
-axes = []
-titles = []
-
-j = 0
-for i in [0, 4]:
-
-    axes.append(fig.add_subplot(gs[j, 0]))
-    y, x = np.histogram(d[i].var(-1, ddof=1), density=True, bins=int(fp.NBINS * 0.5))
-    axes[-1].stairs(y, x, fill=True, color=fp.colors[0])
-    axes[-1].axvline(d[i].mean(-1).var(ddof=1), c=fp.colors[3], ls='--')
[…]

### 2023-11-14 src/tex/ms.tex
Matching with arXiv

@@
 \usepackage{blindtext}
+% \usepackage[hidelinks]{hyperref}
 \sisetup{
@@
 \newcommand{\papertitle}{Accurate Estimation of Diffusion Coefficients and their Uncertainties from Computer Simulation}
+% \newcommand{\oMSD}{\ensuremath{\overline{\bm{s}^2}}}
+% \newcommand{\oMSDs}[1]{\ensuremath{\overline{s^2}}(#1)}
+% \newcommand{\oMSDi}{\ensuremath{\overline{s^2}_i}}
+% \newcommand{\oMSDj}{\ensuremath{\overline{s^2}_j}}
 
@@
 
+\graphicspath{{figures/}}
+
 % Redefine \maketitle so that it can be used twice (for supplementary)
@@
 
-\let\oldaddcontentsline\addcontentsline
+\let\oldaddcontentsline\addcontentsline% Store \addcontentsline
 \renewcommand{\addcontentsline}[3]{}
@@
 \author{Andrew R. McCluskey}
-  \email{andrew.mccluskey@ess.eu}
+  \email{andrew.mccluskey@bristol.ac.uk}
+  \affiliation{School of Chemistry, University of Bristol, Cantock's Close, Bristol, BS8 1TS, United Kingdom}
   \affiliation{European Spallation Source ERIC, Ole Maaløes vej 3, 2200 København N, DK}
@@
 The variances, $\var{\oMSDi}$, are correlated, since the displacement of each particle at time $t+\Delta t$ is necessarily similar to its displacement at time $t$, and hence, $\oMSDs{t}$ is similar to $\oMSDs{t+\Delta t}$.
-These variances are also typically unequal---the data are heteroscedastic \cite{smith_random_1996,he_statistical_2018}. 
+These variances are also typically unequal---the data are heteroscedastic \cite{smith_random_1996,he_statistical_2018,UslerEtAl_JComputChem2023}. 
 Because the key assumptions of the OLS method are not valid for MSD data, OLS gives statistically inefficient estimates of $\D$, while the estimated regression uncertainties obtained from the standard OLS statistical formulae significantly underestimate the true uncertainty in $\prob{\Dest_\mathrm{OLS}}$ (Fig.~\ref{fig:glswlsols}a).
@@
     \centering
-    \includegraphics[width=\columnwidth]{figures/glswlsols.pdf}
+    % \resizebox{\columnwidth}{!}{\includegraphics*{figure1.pdf}}
+    \includegraphics[width=\columnwidth]{glswlsols.pdf}
+    \script{glswlsols.py}
     \caption{
@@
     \label{fig:glswlsols}
-    \script{glswlsols.py}
 \end{figure}
@@
     \centering
-    \includegraphics[width=\columnwidth]{figures/msd.pdf}
+    % \resizebox{\columnwidth}{!}{\includegraphics*{figure2.pdf}}
+    \includegraphics[width=\columnwidth]{msd.pdf}
+    \script{msd.py}
     \caption{
@@
         Panel (c) plots the numerical variance against the estimated variance from a single simulation as a function of timestep $i$.
+        %Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences.
     }
     \label{fig:msd}
-    \script{msd.py}
 \end{figure}
@@
     \centering
-    \includegraphics[width=\columnwidth]{figures/covariances.pdf}
+    % \resizebox{\columnwidth}{!}{\includegraphics*{figure3.pdf}}
+    \includegraphics[width=\columnwidth]{covariances.pdf}
+    \script{covariances.py}
     \caption{
@@
         The off-diagonal panels show difference plots, computed as per-element ratios between pairs of covariance matrices (a--c).
+        %Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences.
     }
     \label{fig:covariances}
-    \script{covariances.py}
 \end{figure}
@@
     \centering
-    \includegraphics[width=\textwidth]{figures/random_walk.pdf}
+    % \resizebox{\textwidth}{!}{\includegraphics*{figure4.pdf}}
+    \includegraphics[width=\textwidth]{random_walk.pdf}
+    \script{random_walk.py}
     \caption{
@@
     \label{fig:random_walk}
-    \script{random_walk.py}
 \end{figure*}
@@
 We next benchmark our method using data from simulations of the lithium-ion solid electrolyte cubic \ce{Li7La3Zr2O12} (c-LLZO).
-We performed a single simulation of \num{1536} atoms (\num{448} Li ions) at \SI{1000}{K} for \SI{1.6}{\nano\second} (full simulation details are provided in the SI).
+We performed a single simulation of \num{1536} atoms (\num{448} Li ions) at \SI{1000}{K} for \SI{1.6}{\nano\second} (full simulation details are provided in the Methods section).
 To generate multiple statistically equivalent trajectories, the resulting simulation data was partitioned into \num{512} effective trajectories, each approximately $\sim\SI{25}{\ps}$ in length, and containing data for \num{56} lithium ions.
@@
     \centering
-    \includegraphics[width=\columnwidth]{figures/diffusion.pdf}
+    % \resizebox{\columnwidth}{!}{\includegraphics*{figure5.pdf}}
+    \includegraphics[width=\columnwidth]{diffusion.pdf}
+    \script{diffusion.py}
     \caption{
@@
       (b) Probability distribution of estimated variances, $\varest{\Dest}$, for individual LLZO effective simulations, compared to the true sample variance (pink vertical line) $\var{\Dest}$.
+      %Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences. 
     }
     \label{fig:diffusion}
-    \script{diffusion.py}
 \end{figure}
@@
     \centering
-    \includegraphics[width=\columnwidth]{figures/stateff.pdf}
+    % \resizebox{\columnwidth}{!}{\includegraphics*{figure6.pdf}}
+    \includegraphics[width=\columnwidth]{stat_eff.pdf}
+    \script{stat_eff.py}
     \caption{
@@
     \label{fig:stat_eff}
-    \script{stat_eff.py}
 \end{figure}
@@
 We have implemented this procedure in the open-source package \textsc{kinisi} \cite{mccluskey_kinisi_2022}, which we hope will support its use within the broader simulation community across a range of materials science contexts.
-\section*{Data availability}
+
+\section{Methods}
+
+\subsection{Numerical implementation in KINISI}
+\label{sec:implementation}
+We have implemented the approximate Bayesian estimation method described in the main text in the open-source Python package \textsc{kinisi}~\cite{mccluskey_kinisi_2022}. 
+This section provides numerical details regarding the implementation of this method.
+
+\textsc{kinisi} uses overlapping sliding window sampling when calculating the observed mean squared displacement at each time interval $t$ (see Eqn.~\ref{equ:observed_msd}).
+For a given time interval, $t$, the maximum number of observations is $N_{\mathrm{atoms}} \times (N_{t} - i)$ displacements, where $N_{\mathrm{atoms}}$ is the number of atoms, $N_{t}$ is the total number of timesteps, and $i$ is the index of the timestep (where \num{1} is the index for the shortest timestep).
+To estimate the variance of the observed MSD, we rescale the variance of observed squared displacements by the number of non-overlapping time intervals in the simulation, $\nind{i} = N_{\mathrm{atoms}} \times N_{t} / i$, as presented in Eqn.~\ref{equ:varestMSD}.
+
+The parametrisation of the covariance matrix from the variances $\var{\oMSDi}$ and the number of independent observations $\nind{i}$ is defined by Eqn.~\ref{equ:cvv}. 
+The covariance matrix is only constructed for values of $t$ where the particle motion is considered to be in the long-time diffusive limit, with this threshold set by the user to a value appropriate for their system and simulation data.
+For the examples presented in the main manuscript, we consider particles to be in the diffusive regime from $t=\num{4}$ for the random walk trajectories and from $t=\SI{10}{\pico\second}$ for the LLZO simulations..
+
+\textsc{kinisi} uses ordinary least squares to obtain an initial guess for the gradient and intercept of the linear model describing the observed MSD.
+This initial guess is then used as the starting point for minimising the negative maximum a posteriori (the peak of the posterior distribution as per Eqn.~\ref{equ:bayes}), with the improper prior that $\D \ge 0$~\cite{broyden_convergence_1970,fletcher_new_1970,goldfarb_family_1970,shanno_conditioning_1970}.
+The log-likelihood calculation (Eqn.~\ref{equ:loglike}) uses the Moore-Penrose generalisation of the inverse of a Hermitian matrix~\cite{moore_on_1920,bjerhammar_application_1951,penrose_generlized_1955}.
+
+To sample the joint posterior probability distribution of the linear model, \textsc{kinisi} uses the \textsc{emcee} package~\cite{foremanmackey_emcee_2019}, which implements Goodman and Weare's affine invariant Markov chain Monte Carlo ensemble sampler~\cite{goodman_ensemble_2010}.
+When sampling $\prob{\D|\model}$ we again apply the improper prior $\D \ge 0$.
+The sampling process uses \num{32} walkers for \num{1500} steps, with the first \num{500} steps discarded as a burn-in period.
+The sampled chains are thinned such that only every 10th value is retained, yielding \num{3200} points sampled from the posterior distribution $\prob{\D|\model}$.
+These points can then be plotted as a histogram (as in Fig~\ref{fig:random_walk}), and summary statistics $\Dest$ and $\varest{\Dest}$ can be derived.
+
+\subsection{LLZO simulations}
+Classical molecular dynamics were run using the \textsc{metalwalls} code~\cite{marin_metalwalls_2020}. 
+We used the DIPPIM polarisable ion force field, as parameterised by Burbano \emph{et al.}~\cite{burbano_sparse_2016}, due to its proven accuracy in accounting for the effect of ion polarisability on diffusion~\cite{wilson_polarization_1993,burbano_sparse_2016}.
+We simulated the cubic phase of LLZO in NVT ensemble at a temperature of \SI{1000}{\kelvin}.
+Simulations were run for \SI{2}{\nano\second} with a \SI{2}{\femto\second} timestep. 
+To control temperature, we used a Nos\'{e}-Hoover thermostat, with a relaxation time of \SI{121}{\femto\second} (5000 $\hbar / E_{h}$)~\cite{nose_unified_1984,hoover_canonical_1985,martyna_nose_1992}.
+Simulations were performed using $2 \times 2 \times 2$ supercells with \SI{1536}{atoms} following the same protocol as in Ref.~\citenum{burbano_sparse_2016}.
+
+\section*{Data \& Code Availability}
 
@@
 All raw simulation files are available on Zenodo shared under CC BY-SA 4.0 licences, the random walk simulations and other analysis-linked data~\cite{mccluskey_zenodo_2022} and the LLZO raw simulation trajectories~\cite{coles_llzo_zenodo_2022}.
-The method outlined in this work is implemented in the open-source Python package \textsc{kinisi}~\cite{mccluskey_kinisi_2022}, which is available under an MIT license.
+The method outlined in this work is implemented in the open-source Python package \textsc{kinisi}~\cite{mccluskey_kinisi_2022}, which is available under an MIT license, and can be accessed via \url{https://github.com/bjmorgan/kinisi}.
 
-\section*{CR\lowercase{e}d\lowercase{i}T author statement}
+\section*{CR\lowercase{e}d\lowercase{i}T Author Statement}
 
@@
 \label{sec:ran}
-%
 In the main manuscript we present the result that the covariance matrix for a system of freely diffusing particles, in the long-time limit, has the form
-%
 \begin{equation}
@@
 \end{equation}
-%
 where $\oMSDi$ is the observed mean-squared displacement (MSD) for time interval $i$ and $\nind{i}$ is the number of statistically independent observed squared displacements averaged over to compute the mean value.
@@
 After $n$ steps, the MSD, $\oMSDn$, is given by
-%
 \begin{equation}
@@
 \end{equation}
-%
 The expected MSD in the long-time limit, $\mathbb{E}(\oMSDn) = \moMSDn$, is obtained by averaging over all permutations of $h_i$ and $h_j$:
-%
 \begin{equation}
@@
 \end{equation}
-%
 For a random walk, the second term averages to zero for all $h_i$ and $h_j$, and
-%
 \begin{equation}
@@
 \end{equation}
-%
 Hence the expected value for the mean-squared displacement increases linearly with the number of steps taken.
@@
 The variance in the observed MSD, $\var{\oMSDn}$, is given by the standard statistical formula
-%
 \begin{equation}
@@
 \end{equation}
-%
 which can be expanded as
-%
 \begin{equation}
@@
 \end{equation}
-%
 The first term can be expanded in terms of displacements $h$ as
-%
 \begin{equation}
@@
 This gives
-%
 \begin{equation}
@@
 \end{equation}
-%
 which, in the limit $n \to \infty$, approaches
-%
 \begin{equation}
@@
 \end{equation}
-%
 Combining this result with Eqn.~\ref{equ:msd_SI}, we can express the variance in the mean-squared displacement as
-%
 \begin{equation}
@@
 \end{equation}
-%
 i.e., $\var{\oMSDn}$ increases quadratically with the number of steps taken, or, equivalently, with time.
@@
 This averaging over equivalent observations reduces the variance in the observed MSD to
-%
 \begin{equation}
@@
 \end{equation}
-%
 where $\nind{n}$ is the total number of statistically independent squared displacements that contribute to $\oMSDi$.
@@
 The results for a one-dimensional lattice above (Eqns.~\ref{equ:msd_SI} \&~\ref{equ:der_var}) can be extended to a $d$-dimensional lattice, to give
-%
 \begin{equation}
@@
 \end{equation}
-%
 with variance
-%
 \begin{equation}
@@
 \end{equation}
-%.
 Because each step is equally likely to move a particle along each of the $d$ dimensions, the term $n$ in Eqns.~\ref{equ:msd_SI} \&~\ref{equ:der_var} is replaced here with $n/d$.
@@
 The analysis above can be extended to consider the covariance between two different numbers of steps, $n$ and $n+m$, in the random walk where the expected MSDs will be
-%
 \begin{equation}
@@
 \end{equation}
-%
 The covariance between these is defined as
-%
 \begin{equation}
@@
 \end{equation}
-%
 and then reformulated to give
-%
 \begin{equation}
@@
 \end{equation}
-%
 where 
-%
 \begin{equation}
@@
 \end{equation}
-%
 and, by analogy to Eqn.~\ref{equ:big_av},
-%
 \begin{equation}
@@
 \end{equation}
-%
 which we can rewrite as
-%
 \begin{equation}
@@
 \end{equation}
-%
 The second and third terms in Eqn.~\ref{equ:vbig} tend to zero as there is an equal probability of positive and negative displacements. 
 This reduces Eqn.~\ref{equ:vbig} to 
-%
 \begin{equation}
@@
 \end{equation}
-%
 and using Eqn.~\ref{equ:infty} gives
-%
 \begin{equation}
@@
 \end{equation}
-%
 We can rewrite this as
-%
 \begin{equation}
@@
 \end{equation}
-%
 where the following holds,
-%
 \begin{equation}
@@
 \end{equation}
-%
 Putting this result into Eqn.~\ref{equ:pair} allows the covariance to be written as
-%
 \begin{equation}
@@
 Therefore, the covariance, scaled by the number of contributing independent observations, in the long-time limit, is
-%
 \begin{equation}
@@
 \end{equation}
-%
 Similar to the MSD and the variance, the covariance can be written for $d$-dimensions as 
-%
 \begin{equation}
@@
 \end{equation}
-%
 The covariance can be calculated directly from the variance by recognising that both depend on the number of overlapping points, $n$, as follows
-%
 \begin{equation}
@@
 \end{equation}
-%
 This is then rewritten in terms of $i$ and $j$ to give, Eqn.~\ref{equ:cvv_SI}.
@@
 Using the equivalence of $2d\D t \equiv n\kappa^2$~\cite{howard_reports_1964}, Eqns.~\ref{equ:msd_SI} \&~\ref{equ:cvv_SI} can be rewritten in terms of $t$ (or $t_1$ and $t_2$) and the diffusion coefficient, for any dimensionality of lattice random walk,
-%
 \begin{equation}
@@
 \end{equation}
-%
 and 
-%
 \begin{equation}
@@
 \end{equation}
-%
 
@@
 \label{sec:skew}
-%
 In the main manuscript, we present results for a set of \num{4096} 3D-lattice random walk simulations, each consisting of \num{128} particles undergoing \num{128} steps (Fig.~\ref{fig:random_walk}).
@@
 Furthermore, the distribution $\prob{\Dest}$ agrees even more closely with the numerically converged distribution obtained when combining data from all \num{4096} simulations (Fig~\ref{fig:true_cov}a), contrasting with the results presented in Fig.~\ref{fig:random_walk}b, where our approximate Bayesian regression scheme yields a slightly broadened distribution due to the use of the long-time limit in the derivation of the analytical form for $\bm{\Sigma^\prime}$.
-%
 \begin{figure}
     \centering
-    \includegraphics[width=\columnwidth]{figures/true_cov.pdf}
+    % \resizebox{\columnwidth}{!}{\includegraphics*{figure7.pdf}}
+    \includegraphics[width=\columnwidth]{true_cov.pdf}
+    \script{true_cov.py}
     \caption{
         (a) Probability distribution of point-estimates $\prob{\Dest}$ obtained from \num{4096} individual random-walk simulations, using the numerical covariance matrix $\bm{\Sigma}_\mathrm{num}$.
-        Each simulation has been analysed as in Fig.~\ref{fig:msd}(a) and (b) to yield a single corresponding point estimate $\Dest$.
+        Each simulation has been analysed as in Fig.~~\ref{equ:observed_msd}(a) and (b) to yield a single corresponding point estimate $\Dest$.
         The grey line shows the distribution of point estimates, $\prob{\Dest_\mathrm{num}}$, obtained using Bayesian regression with a mean vector and numerical covariance matrix derived from the complete dataset of all \num{4096} simulations.
@@
     \label{fig:true_cov}
-    \script{true_cov.py}
 \end{figure}
-%
-
-\section{LLZO simulations}
-%
-Classical molecular dynamics were run using the \textsc{metalwalls} code~\cite{marin_metalwalls_2020}. 
-We used the DIPPIM polarisable ion force field, as parameterised by Burbano \emph{et al.}~\cite{burbano_sparse_2016}, due to its proven accuracy in accounting for the effect of ion polarisability on diffusion~\cite{wilson_polarization_1993,burbano_sparse_2016}.
-We simulated the cubic phase of LLZO in NVT ensemble at a temperature of \SI{1000}{\kelvin}.
-Simulations were run for \SI{2}{\nano\second} with a \SI{2}{\femto\second} timestep. 
-To control temperature, we used a Nos\'{e}-Hoover thermostat, with a relaxation time of \SI{121}{\femto\second} (5000 $\hbar / E_{h}$)~\cite{nose_unified_1984,hoover_canonical_1985,martyna_nose_1992}.
-Simulations were performed using $2 \times 2 \times 2$ supercells with \SI{1536}{atoms} following the same protocol as in Ref.~\citenum{burbano_sparse_2016}.
-
-\section{Numerical implementation in KINISI}
-\label{sec:implementation}
-%
-We have implemented the approximate Bayesian estimation method described in the main text in the open-source Python package \textsc{kinisi}~\cite{mccluskey_kinisi_2022}. 
-This section provides numerical details regarding the implementation of this method.
-
-\textsc{kinisi} uses overlapping sliding window sampling when calculating the observed mean squared displacement at each time interval $t$ (see Eqn.~\ref{equ:observed_msd}).
-For a given time interval, $t$, the maximum number of observations is $N_{\mathrm{atoms}} \times (N_{t} - i)$ displacements, where $N_{\mathrm{atoms}}$ is the number of atoms, $N_{t}$ is the total number of timesteps, and $i$ is the index of the timestep (where \num{1} is the index for the shortest timestep).
-To estimate the variance of the observed MSD, we rescale the variance of observed squared displacements by the number of non-overlapping time intervals in the simulation, $\nind{i} = N_{\mathrm{atoms}} \times N_{t} / i$, as presented in Eqn.~\ref{equ:varestMSD}.
-
-The parametrisation of the covariance matrix from the variances $\var{\oMSDi}$ and the number of independent observations $\nind{i}$ is defined by Eqn.~\ref{equ:cvv}. 
-The covariance matrix is only constructed for values of $t$ where the particle motion is considered to be in the long-time diffusive limit, with this threshold set by the user to a value appropriate for their system and simulation data.
-For the examples presented in the main manuscript, we consider particles to be in the diffusive regime from $t=\num{4}$ for the random walk trajectories and from $t=\SI{10}{\pico\second}$ for the LLZO simulations..
-
-\textsc{kinisi} uses ordinary least squares to obtain an initial guess for the gradient and intercept of the linear model describing the observed MSD.
-This initial guess is then used as the starting point for minimising the negative maximum a posteriori (the peak of the posterior distribution as per Eqn.~\ref{equ:bayes}), with the improper prior that $\D \ge 0$~\cite{broyden_convergence_1970,fletcher_new_1970,goldfarb_family_1970,shanno_conditioning_1970}.
-The log-likelihood calculation (Eqn.~\ref{equ:loglike}) uses the Moore-Penrose generalisation of the inverse of a Hermitian matrix~\cite{moore_on_1920,bjerhammar_application_1951,penrose_generlized_1955}.
-
-To sample the joint posterior probability distribution of the linear model, \textsc{kinisi} uses the \textsc{emcee} package~\cite{foremanmackey_emcee_2019}, which implements Goodman and Weare's affine invariant Markov chain Monte Carlo ensemble sampler~\cite{goodman_ensemble_2010}.
-When sampling $\prob{\D|\model}$ we again apply the improper prior $\D \ge 0$.
-The sampling process uses \num{32} walkers for \num{1500} steps, with the first \num{500} steps discarded as a burn-in period.
-The sampled chains are thinned such that only every 10th value is retained, yielding \num{3200} points sampled from the posterior distribution $\prob{\D|\model}$.
-These points can then be plotted as a histogram (as in Fig~\ref{fig:random_walk}), and summary statistics $\Dest$ and $\varest{\Dest}$ can be derived.
 

### 2023-11-14 src/scripts/covariances.py

@@ -46,12 +46,13 @@ max_cov = np.ceil(np.max([anal_cov.max(), true_cov.max(), kinisi_cov.max()]))
 min_cov = np.floor(np.min([anal_cov.min(), true_cov.min(), kinisi_cov.min()]))
 levels = np.linspace(min_cov, max_cov, 15)
 
-figsize = (4.0, 3.2)
+figsize = (5.0, 4)
 fig = plt.figure(figsize=figsize)
-gs = gridspec.GridSpec(3, 3, figure=fig, wspace=0.6, hspace=0.6)
+gs = gridspec.GridSpec(3, 3, figure=fig, wspace=0.8, hspace=0.8)
 
 axes = []
 titles = []
+titles2 = []
 
 axes.append(fig.add_subplot(gs[0, 0]))
 im1 = axes[-1].contourf(ts_mesh[0], ts_mesh[1], true_cov, cmap=newcmp1, levels=levels)
@@ -60,13 +61,11 @@ cb1 = fig.colorbar(im1, fraction=0.046)
 cb1.ax.tick_params(labelsize=fp.FONTSIZE, width=0.5)
 cb1.ax.yaxis.set_ticks([0, 1500, 3000])
 cb1.outline.set_linewidth(0.5)
-axes[-1].set_xlabel(r"$t$")
 axes[-1].set_xticks([0, int(length / 2), length])
-axes[-1].set_ylabel(r"$t$")
 axes[-1].set_yticks([0, int(length / 2), length])
 # axes[-1].set_aspect("equal")
 titles.append(r"a")
-axes[-1].set_title("Numerical", y=1.0)
+titles2.append("numerical")
 
 axes.append(fig.add_subplot(gs[1, 1]))
 im1 = axes[-1].contourf(ts_mesh[0], ts_mesh[1], anal_cov, cmap=newcmp2, levels=levels)
@@ -75,12 +74,10 @@ cb1 = fig.colorbar(im1, fraction=0.046)
 cb1.ax.tick_params(labelsize=fp.FONTSIZE, width=0.5)
 cb1.ax.yaxis.set_ticks([0, 1500, 3000])
 cb1.outline.set_linewidth(0.5)
-axes[-1].set_xlabel(r"$t$")
 axes[-1].set_xticks([0, int(length / 2), length])
-axes[-1].set_ylabel(r"$t$")
 axes[-1].set_yticks([0, int(length / 2), length])
 titles.append(r"b")
-axes[-1].set_title("Analytical", y=1.0)
[…]

### 2023-11-14 src/scripts/diffusion.py

@@ -28,38 +28,35 @@ titles = []
 
 axes.append(fig.add_subplot(gs[0, 0]))
 y, x = np.histogram(d.mean(-1), bins=int(fp.NBINS * 0.5), density=True)
-axes[-1].stairs(y * 1e-4, x * 1e5, fill=True, color=fp.colors[2], label='$p(\hat{D}^*_{\mathrm{Li}^+})$')
-axes[-1].plot(llzo_x * 1e5,
-              norm.pdf(llzo_x, dllzo_true.mean(), dllzo_true.std()) * 1e-4,
-              color=fp.colors[4],
-              label='$p(\hat{D}^*_{\mathrm{Li}^+} | \mathbfit{x})$')
+axes[-1].stairs(y, x, fill=True, color='#ECBCA7', label='$p(\hat{D}^*)$')
+axes[-1].plot(llzo_x,
+              norm.pdf(llzo_x, dllzo_true.mean(), dllzo_true.std()),
+              color='#B8B8B8',
+              label='$p(\hat{D}^*_{\mathrm{num}})$')
 axes[-1].plot(np.array([d.mean(-1).mean() - d.mean(-1).std(),
-                        d.mean(-1).mean() + d.mean(-1).std()]) * 1e5,
+                        d.mean(-1).mean() + d.mean(-1).std()]),
               [axes[-1].get_ylim()[1] * 1.05, axes[-1].get_ylim()[1] * 1.05],
-              color=fp.colors[3],
-              label='$\sigma(\hat{D}^*_{\mathrm{Li}^+})$',
+              color='#F3ADBC',
+              label='$\sigma[\hat{D}^*]$',
               marker="|")
-axes[-1].set_yticks([0, 6, 12])
-axes[-1].set_xlabel(r"$\hat{D}^*_{\mathrm{Li}^{+}}$ / $10^{-5}$ cm$^2$ s$^{-1}$")
-axes[-1].set_ylabel(r"$p(\hat{D}^*_{\mathrm{Li}^{+}})$ / $10^4$ cm$^{-2}$ s")
+# axes[-1].set_yticks([0, 6, 12])
+axes[-1].set_xlabel(r"$\hat{D}^*$ / cm$^2$ s$^{-1}$")
[…]

### 2023-11-14 src/scripts/glswlsols.py

@@ -21,58 +21,58 @@ axes = []
 titles = []
 axes.append(fig.add_subplot(gs[0, 0]))
 bins = np.linspace((data['ols_pop']).min(), (data['ols_pop']).max(), fp.NBINS * 2)
-axes[-1].hist(data['ols_pop'], bins=bins, density=True, color=fp.colors[5])
-axes[-1].plot(x, norm.pdf(x, *norm.fit(data['ols_pop'])), color=fp.colors[5])
-axes[-1].plot([1 - data['ols_pop'].std(), 1 + data['ols_pop'].std()],
-              [axes[-1].get_ylim()[1] * 1.1, axes[-1].get_ylim()[1] * 1.1],
-              color=fp.colors[3],
-              marker='|')
+axes[-1].hist(data['ols_pop'], bins=bins, density=True, color='#ECBCA7')
+axes[-1].plot(x, norm.pdf(x, *norm.fit(data['ols_pop'])), color='#B8B8B8')
 axes[-1].plot([1 - data['ols_est'], 1 + data['ols_est']],
               [axes[-1].get_ylim()[1] * 1.1, axes[-1].get_ylim()[1] * 1.1],
-              color='#1E5B84',
+              color='#305980',
+              marker='|')
+axes[-1].plot([1 - data['ols_pop'].std(), 1 + data['ols_pop'].std()],
+              [axes[-1].get_ylim()[1] * 1.1, axes[-1].get_ylim()[1] * 1.1],
+              color='#F3ADBC',
               marker='|')
 axes[-1].set_xticks([0.8, 0.9, 1, 1.1, 1.2])
 axes[-1].set_xlim(0.75, 1.25)
 axes[-1].set_yticks([0, 3, 6])
-axes[-1].set_ylabel(r'$p[\hat{D}^* (D^*)^{-1}]$')
+axes[-1].set_ylabel(r'$p(\hat{D}^*)$')
 titles.append("a")
-axes[-1].set_title("OLS")
+axes[-1].set_title("ordinary least squares")
 
 axes.append(fig.add_subplot(gs[1, 0], sharex=axes[0]))
-axes[-1].hist(data['wls_pop'], bins=bins, density=True, color=fp.colors[5])
-axes[-1].plot(x, norm.pdf(x, *norm.fit(data['wls_pop'])), color=fp.colors[5])
-axes[-1].plot([1 - data['wls_pop'].std(), 1 + data['wls_pop'].std()],
[…]

### 2023-11-14 src/scripts/msd.py

@@ -24,8 +24,8 @@ titles = []
 
 axes.append(fig.add_subplot(gs[0, 0]))
 titles.append(r"a")
-axes[-1].set_title("Numerical")
-m, c, b = axes[-1].errorbar(dt, numerical.mean(0), numerical.std(0), color=fp.colors[0], elinewidth=0.2)
+axes[-1].set_title("numerical")
+m, c, b = axes[-1].errorbar(dt, numerical.mean(0), numerical.std(0) * 2, color='#81A8D9', elinewidth=0.2)
 [bar.set_alpha(0.5) for bar in b]
 [cap.set_alpha(0.5) for cap in c]
 axes[-1].set_xlabel(r"$t^{}$")
@@ -37,8 +37,8 @@ axes[-1].set_yticks([0, 400, 800])
 
 axes.append(fig.add_subplot(gs[0, 1], sharey=axes[-1]))
 titles.append(r"b")
-axes[-1].set_title("Estimated")
-m, c, b = axes[-1].errorbar(dt, kinisi[0, :, 1], kinisi[0, :, 2], color=fp.colors[2], elinewidth=0.2)
+axes[-1].set_title("estimated")
+m, c, b = axes[-1].errorbar(dt, kinisi[0, :, 1], kinisi[0, :, 2] * 2, color='#E08D6D', elinewidth=0.2)
 [bar.set_alpha(0.5) for bar in b]
 [cap.set_alpha(0.5) for cap in c]
 axes[-1].set_xlabel(r"$t^{}$")
@@ -50,15 +50,15 @@ axes[-1].set_yticks([0, 400, 800])
 
 axes.append(fig.add_subplot(gs[0, 2]))
 titles.append(r"c")
-axes[-1].set_title("$\sigma^2$ accuracy")
+axes[-1].set_title("$\sigma^2$ comparison")
 m = np.max([numerical.var(0), (kinisi[0, :, 2]**2)])
 axes[-1].plot(numerical.var(0) / 1000, (kinisi[0, :, 2]**2) / 1000,
-              color=fp.colors[1],
+              color='#67B99B',
               marker='.',
               ls='',
               markeredgewidth=0,
               markersize=4)
-axes[-1].plot(np.linspace(0, m, 10) / 1000, np.linspace(0, m, 10) / 1000, color=fp.NEARLY_BLACK, lw=0.5)
[…]

### 2023-11-14 src/scripts/random_walk.py

@@ -31,80 +31,79 @@ titles = []
 
 axes.append(fig.add_subplot(gs[0, 0]))
 titles.append("a")
-axes[-1].set_title("Single simulation")
+axes[-1].set_title("single simulation")
+axes[-1].plot(dt, kinisi['data'][0, :, 1], color='#305A80', label=r"$\mathbfit{x}$", zorder=10)
 axes[-1].fill_between(dt,
                       *np.percentile(distribution, CREDIBLE_INTERVALS[0][:-1], axis=1),
                       alpha=CREDIBLE_INTERVALS[0][-1],
-                      color=fp.colors[1],
+                      color='#67B99B',
                       lw=0,
                       label=r'$p(m | \mathbfit{x})$')
 for ci in CREDIBLE_INTERVALS[1:]:
-    axes[-1].fill_between(dt, *np.percentile(distribution, ci[:-1], axis=1), alpha=ci[-1], color=fp.colors[1], lw=0)
-axes[-1].plot(dt, kinisi['data'][0, :, 1], color='#1E5B84', label=r"$\mathbfit{x}$")
+    axes[-1].fill_between(dt, *np.percentile(distribution, ci[:-1], axis=1), alpha=ci[-1], color='#67B99B', lw=0)
 axes[-1].set_xlabel(r"$t$")
 axes[-1].set_ylabel(r"$x(t)$")
 axes[-1].set_xlim(0, None)
 axes[-1].set_ylim(0, None)
 axes[-1].set_xticks([0, 64, 128])
-axes[-1].set_yticks([0, 400, 800])
+axes[-1].set_yticks([0, 300, 600, 900])
 axes[-1].legend()
 
 axes.append(fig.add_subplot(gs[0, 1]))
 titles.append("b")
-axes[-1].set_title("Single simulation")
+axes[-1].set_title("single simulation")
 y, x = np.histogram(kinisi['diff_c'][0], bins=fp.NBINS, density=True)
-axes[-1].stairs(y, x, fill=True, color=fp.colors[1], label=r'$p(D^* | \mathbfit{x})$')
+axes[-1].stairs(y, x, fill=True, color='#9FD5C3', label=r'$p(D^* | \mathbfit{x})$')
 axes[-1].plot([
     kinisi['diff_c'][0].mean() - kinisi['diff_c'][0].std(),
[…]

### 2023-11-14 src/scripts/stat_eff.py

@@ -37,24 +37,24 @@ letters = ['a', 'b']
 data = np.load(paths.data / f"random_walks/stat_eff.npz")
 
 axes.append(fig.add_subplot(gs[0, 0]))
-axes[-1].plot(xaxis, data['ky'], marker='.', ls="", color=fp.colors[2])
+axes[-1].plot(xaxis, data['ky'], marker='.', ls="", color='#67B99B')
 k = f(x, data['ka'], data['kb'])
-axes[-1].plot(x, k, color=fp.colors[2])
-axes[-1].plot(xaxis, data['wy'], marker='.', ls="", color=fp.colors[1])
+axes[-1].plot(x, k, color='#67B99B')
+axes[-1].plot(xaxis, data['wy'], marker='.', ls="", color='#81A8D9')
 k = f(x, data['wa'], data['wb'])
-axes[-1].plot(x, k, color=fp.colors[1])
-axes[-1].plot(xaxis, data['py'], marker='.', ls="", color=fp.colors[3])
+axes[-1].plot(x, k, color='#81A8D9')
+axes[-1].plot(xaxis, data['py'], marker='.', ls="", color='#F3ADBC')
 k = f(x, data['pa'], data['pb'])
-axes[-1].plot(x, k, color=fp.colors[3])
-axes[-1].plot(xaxis, data['ny'], marker='.', ls="", color=fp.colors[0])
+axes[-1].plot(x, k, color='#F3ADBC')
+axes[-1].plot(xaxis, data['ny'], marker='.', ls="", color='#E08D6D')
 k = f(x, data['na'], data['nb'])
-axes[-1].plot(x, k, color=fp.colors[0])
+axes[-1].plot(x, k, color='#E08D6D')
 axes[-1].set_xscale('log', base=2)
 axes[-1].set_yscale('log', base=2)
-axes[-1].set_xlabel(r'$N_{{\mathrm{{atoms}}}}$ / atoms')
-axes[-1].set_ylabel(r"$\sigma^2 (\langle\hat{D}^*\rangle)$")
+axes[-1].set_xlabel(r'$N_{{\mathrm{{atoms}}}}$')
+axes[-1].set_ylabel(r"$\sigma^2 [\hat{D}^*]$")
 titles.append(f"{letters[0]}")
-axes[-1].set_title(r"Varying $N_{\mathrm{atoms}}$")
+axes[-1].set_title(r"fixed $t_{\max} = 128$")
 axes[-1].set_xticks([32, 128, 512])
[…]

### 2023-11-14 src/scripts/true_cov.py

@@ -30,31 +30,31 @@ x_ = np.linspace(true.mean(-1).min(), true.mean(-1).max(), 1000)
 
 axes.append(fig.add_subplot(gs[0, 0]))
 y, x = np.histogram(true.mean(-1), density=True, bins=fp.NBINS)
-axes[-1].stairs(y, x, fill=True, color=fp.colors[2], label='$p(\hat{D}^*)$')
+axes[-1].stairs(y, x, fill=True, color='#ECBCA7', label='$p(\hat{D}^*)$')
 axes[-1].plot(x_,
               norm.pdf(x_, dinfty_true.mean(), dinfty_true.std()),
-              color=fp.colors[4],
-              label=r'$p(\hat{D}^* | \mathbfit{x})$')
+              color='#B8B8B8',
+              label=r'$p(\hat{D}^*_{\mathrm{num}})$')
 axes[-1].plot([dinfty_true.mean() - dinfty_true.std(),
                dinfty_true.mean() + dinfty_true.std()], [axes[-1].get_ylim()[1] * 1.05, axes[-1].get_ylim()[1] * 1.05],
-              color=fp.colors[3],
-              label='$\sigma(\hat{D}^*)$')
+              color='#F3ADBC',
+              label='$\sigma[\hat{D}^*]$')
 axes[-1].set_xlabel(r"$\hat{D}^*$")
 axes[-1].set_ylabel(r"$p(\hat{D}^*)$")
 titles.append(r"a")
 axes[-1].set_xlim(0.9, 1.1)
-axes[-1].set_title("$D^*$ distribution")
+axes[-1].set_title("all simulations")
 axes[-1].legend(loc='upper left', bbox_to_anchor=(0.65, 1))
 
 axes.append(fig.add_subplot(gs[0, 1]))
 y, x = np.histogram(true.var(-1, ddof=1), density=True, bins=fp.NBINS)
-axes[-1].stairs(y, x, fill=True, color=fp.colors[0], label='$p[\hat{\sigma}^2(D^*)]$')
-axes[-1].axvline(true.mean(-1).var(ddof=1), c=fp.colors[3], label='$\sigma^2(\hat{D}^*)$', ls='--')
-axes[-1].set_xlabel(r"$\hat{\sigma}^2 (\hat{D}^*)$")
[…]

### 2023-11-23 src/tex/ms.tex
Updates from review

@@
 % Define document class
-\documentclass[reprint,superscriptaddress,nobibnotes,amsmath,amssymb,aps,prx,hidelinks]{revtex4-2}
+\documentclass[reprint,superscriptaddress,nobibnotes,amsmath,amssymb,aps,prx,hidelinks,linenumbers]{revtex4-2}
 
@@
 
-Here, we describe an approximate Bayesian regression method for estimating $\D$ with near-maximal statistical efficiency while also accurately estimating the corresponding statistical uncertainty, using data from a single simulation.
+\begin{figure*}
+    \centering
+    % \resizebox{\columnwidth}{!}{\includegraphics*{figure1.pdf}}
+    \includegraphics[width=\textwidth]{schematic.pdf}
+    \caption{
+        A schematic diagram of the Bayesian regression method described in this work, running from our simulations through the sampling of displacement, variance and covariance estimation and final the sampling process to give the marginal posterior distribution for $\D$.}
+    \label{fig:schematic}
+\end{figure*}
+
+Here, we describe an approximate Bayesian regression method for estimating $\D$ with near-maximal statistical efficiency while also accurately estimating the corresponding statistical uncertainty, using data from a single simulation (Fig.~\ref{fig:schematic}).
 We model the statistical population of simulation MSDs as a multivariate normal distribution, using an analytical covariance matrix derived for an equivalent system of freely diffusing particles, with this covariance matrix parameterised from the observed simulation data.
@@
 
-\section{Background}
+\section{Results}
+\subsection{Background}
 
@@
 
-\section{Approximating $\bm{\Sigma}$ from simulation data}
+\subsection{Approximating $\bm{\Sigma}$ from simulation data}
 
@@
 We model the covariance matrix for the observed MSD from a given simulation using the covariance matrix for the MSD of an equivalent system of freely diffusing particles, $\mathbf{\Sigma^\prime}$.
-For observed MSDs computed by averaging over non-overlapping time windows, the covariance matrix $\mathbf{\Sigma^\prime}$, in the long time limit, has elements (see SI)
+For observed MSDs computed by averaging over numerically-independent sub-trajectories, the covariance matrix $\mathbf{\Sigma^\prime}$, in the long time limit, has elements (see SI)
 %
@@
 %
-where $\var{\oMSDi}$ are the time-dependent variances of the observed MSD, and $\nind{i}$ is the total number of non-overlapping observed squared-displacements for time-interval $i$.
+where $\var{\oMSDi}$ are the time-dependent variances of the observed MSD, and $\nind{i}$ is the total number of numerically-independent observed squared-displacements for time-interval $i$.
 We estimate the variances $\var{\oMSDi}$ using the standard result that the variance of the mean of a sample scales inversely with the number of independent constituent observations.
-Specifically, we compute an estimated variance $\varest{\oMSDi}$ by rescaling the observed variance of the squared displacement for time interval $i$ by the number of non-overlapping contributing time windows, $\nind{i}$;
+Specifically, we approximate the variance $\varest{\oMSDi}$ (see SI) by rescaling the observed variance of the squared displacement for time interval $i$ by the number of numerically-independent contributing sub-trajectories, $\nind{i}$;
 %
 \begin{equation}
-  \varest{\oMSDi} = \frac{1}{\nind{i}}\var{\Delta r_i^2}.
+  \varest{\oMSDi} \approx \frac{1}{\nind{i}}\var{\Delta r_i^2}.
   \label{equ:varestMSD}
@@
 %
+
+Rescaling by the number of numerically-independent contributing sub-trajectories has the effect of renormalising the variance of the observed squared displacements to account for correlations between particle squared displacements computed from overlapping time windows.
+An alternative approach to renormalising $\var{x_i}$ is to use a non-parametric block-averaging procedure~\cite{flyvbjerg_error_1989,frenkel_understanding_2023,materzanini_high_2021}, which gives undesirable results for a random walk (see SI).
+The block-averaging approach, additionally, requires numerical convergence with respect to block size, which is not guaranteed for time windows with few independent observations. 
+We require $\var{x_i}$ at all observed time intervals, $i$, making block averaging on large data sets computationally prohibitive.
 
@@
 %
-\begin{figure}
-    \centering
-    % \resizebox{\columnwidth}{!}{\includegraphics*{figure2.pdf}}
-    \includegraphics[width=\columnwidth]{msd.pdf}
-    \script{msd.py}
-    \caption{
-        Comparison of the numerical variance in observed MSD from multiple replica simulations and the estimated variance in observed MSD given by rescaling the variance in observed squared displacements (Eqn.~\ref{equ:varestMSD}).
-        Panel (a) shows the mean observed MSD from \num{4096} simulations of \num{128} particles undergoing a 3D lattice random walk of \num{128} steps per particle, with error bars of $\pm2\sigma[\oMSDi]$.
-        Panel (b) shows the MSD from just one simulation, with error bars of $\pm2\widehat{\sigma}[\oMSDi]$, obtained via Eqn.~\ref{equ:varestMSD}.
-        Panel (c) plots the numerical variance against the estimated variance from a single simulation as a function of timestep $i$.
-        %Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences.
-    }
-    \label{fig:msd}
-\end{figure}
+% \begin{figure}
+%     \centering
+%     % \resizebox{\columnwidth}{!}{\includegraphics*{figure2.pdf}}
+%     \includegraphics[width=\columnwidth]{msd.pdf}
+%     \script{msd.py}
+%     \caption{
+%         Comparison of the numerical variance in observed MSD from multiple replica simulations and the estimated variance in observed MSD given by rescaling the variance in observed squared displacements (Eqn.~\ref{equ:varestMSD}).
+%         Panel (a) shows the mean observed MSD from \num{4096} simulations of \num{128} particles undergoing a 3D lattice random walk of \num{128} steps per particle, with error bars of $\pm2\sigma[\oMSDi]$.
+%         Panel (b) shows the MSD from just one simulation, with error bars of $\pm2\widehat{\sigma}[\oMSDi]$, obtained via Eqn.~\ref{equ:varestMSD}.
+%         Panel (c) plots the numerical variance against the estimated variance from a single simulation as a function of timestep $i$.
+%         %Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences.
+%     }
+%     \label{fig:msd}
+% \end{figure}
 %
@@
 
-\section{Validation}
+\subsection{Validation}
 
@@
 The distribution $\prob{\Dest}$ obtained using the model covariance matrix and parametrised separately for each individual effective simulation is highly similar to that obtained using the aggregate numerical covariance matrix calculated from the complete simulation dataset.
-This close agreement mirrors the results for our random walk simulations, and confirms that our method yields accurate and statistically efficient estimates for $\D$, even for real-world simulation data.
+This close agreement mirrors the results for our random walk simulations (see the SI for a similar comparison of the OLS, WLS, and GLS as shown in Fig.~\ref{fig:glswlsols}), and confirms that our method yields accurate and statistically efficient estimates for $\D$, even for real-world simulation data.
 
-\begin{figure}[tb]
-    \centering
-    % \resizebox{\columnwidth}{!}{\includegraphics*{figure5.pdf}}
-    \includegraphics[width=\columnwidth]{diffusion.pdf}
-    \script{diffusion.py}
-    \caption{
-      (a) Probability distribution of point estimates $\prob{\Dest}$ for \num{512} effective simulations of LLZO (orange histogram).
-      The grey line shows the distribution $\prob{\Dest_\mathrm{num}}$ obtained using Bayesian regression with the complete LLZO dataset as input.
-      The pink bar shows an interval of one standard deviation $\sigma[\prob{\Dest}]$.
-      (b) Probability distribution of estimated variances, $\varest{\Dest}$, for individual LLZO effective simulations, compared to the true sample variance (pink vertical line) $\var{\Dest}$.
-      %Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences. 
-    }
-    \label{fig:diffusion}
-\end{figure}
+% \begin{figure}[tb]
+%     \centering
+%     % \resizebox{\columnwidth}{!}{\includegraphics*{figure5.pdf}}
+%     \includegraphics[width=\columnwidth]{diffusion.pdf}
+%     \script{diffusion.py}
+%     \caption{
+%       (a) Probability distribution of point estimates $\prob{\Dest}$ for \num{512} effective simulations of LLZO (orange histogram).
+%       The grey line shows the distribution $\prob{\Dest_\mathrm{num}}$ obtained using Bayesian regression with the complete LLZO dataset as input.
+%       The pink bar shows an interval of one standard deviation $\sigma[\prob{\Dest}]$.
+%       (b) Probability distribution of estimated variances, $\varest{\Dest}$, for individual LLZO effective simulations, compared to the true sample variance (pink vertical line) $\var{\Dest}$.
+%       %Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences. 
+%     }
+%     \label{fig:diffusion}
+% \end{figure}
 
@@
 
-\section{$\var{\Dest}$ scaling and comparison to OLS, WLS, and GLS}
+\subsection{$\var{\Dest}$ scaling and comparison to OLS, WLS, and GLS}
 
-\begin{figure}[htb]
-    \centering
-    % \resizebox{\columnwidth}{!}{\includegraphics*{figure6.pdf}}
-    \includegraphics[width=\columnwidth]{stat_eff.pdf}
-    \script{stat_eff.py}
-    \caption{
-      Scaling of $\var{\Dest}$ with simulation size for OLS (pink), WLS (blue), our approximate Bayesian regression method (green), and GLS (orange).
-      (a) Scaling versus number of mobile particles, $N_\mathrm{atoms}$.
-      (b) Scaling versus total simulation time, $t_\mathrm{max}$.
-      Solid lines show fitted power law relationships for each dataset.
-      The WLS and GLS data are obtained using numerically determined variances and covariance, respectively, from a set of \num{512} repeat simulations for each combination of $N_\mathrm{atoms}$ and $t_\mathrm{max}$.
-    }
-    \label{fig:stat_eff}
-\end{figure}
+% \begin{figure}[htb]
+%     \centering
+%     % \resizebox{\columnwidth}{!}{\includegraphics*{figure6.pdf}}
+%     \includegraphics[width=\columnwidth]{stat_eff.pdf}
+%     \script{stat_eff.py}
+%     \caption{
+%       Scaling of $\var{\Dest}$ with simulation size for OLS (pink), WLS (blue), our approximate Bayesian regression method (green), and GLS (orange).
+%       (a) Scaling versus number of mobile particles, $N_\mathrm{atoms}$.
+%       (b) Scaling versus total simulation time, $t_\mathrm{max}$.
+%       Solid lines show fitted power law relationships for each dataset.
+%       The WLS and GLS data are obtained using numerically determined variances and covariance, respectively, from a set of \num{512} repeat simulations for each combination of $N_\mathrm{atoms}$ and $t_\mathrm{max}$.
+%     }
+%     \label{fig:stat_eff}
+% \end{figure}
 
@@
 \label{sec:implementation}
-We have implemented the approximate Bayesian estimation method described in the main text in the open-source Python package \textsc{kinisi}~\cite{mccluskey_kinisi_2022}. 
-This section provides numerical details regarding the implementation of this method.
+We have implemented the approximate Bayesian estimation method described in the main text in the open-source Python package \textsc{kinisi}~\cite{mccluskey_kinisi_2022}, under the MIT licence.
 
@@
 For a given time interval, $t$, the maximum number of observations is $N_{\mathrm{atoms}} \times (N_{t} - i)$ displacements, where $N_{\mathrm{atoms}}$ is the number of atoms, $N_{t}$ is the total number of timesteps, and $i$ is the index of the timestep (where \num{1} is the index for the shortest timestep).
-To estimate the variance of the observed MSD, we rescale the variance of observed squared displacements by the number of non-overlapping time intervals in the simulation, $\nind{i} = N_{\mathrm{atoms}} \times N_{t} / i$, as presented in Eqn.~\ref{equ:varestMSD}.
+To estimate the variance of the observed MSD, we rescale the variance of observed squared displacements by the number of numerically-independent sub-trajectories in the simulation, $\nind{i} = N_{\mathrm{atoms}} \times N_{t} / i$, as presented in Eqn.~\ref{equ:varestMSD}.
 
@@
 The covariance matrix is only constructed for values of $t$ where the particle motion is considered to be in the long-time diffusive limit, with this threshold set by the user to a value appropriate for their system and simulation data.
-For the examples presented in the main manuscript, we consider particles to be in the diffusive regime from $t=\num{4}$ for the random walk trajectories and from $t=\SI{10}{\pico\second}$ for the LLZO simulations..
+For the examples presented in the main manuscript, we consider particles to be in the diffusive regime from $t=\num{4}$ for the random walk trajectories and from $t=\SI{10}{\pico\second}$ for the LLZO simulations.
 
@@
 The sampled chains are thinned such that only every 10th value is retained, yielding \num{3200} points sampled from the posterior distribution $\prob{\D|\model}$.
-These points can then be plotted as a histogram (as in Fig~\ref{fig:random_walk}), and summary statistics $\Dest$ and $\varest{\Dest}$ can be derived.
+These points can then be plotted as a histogram (as in Fig~\ref{fig:random_walk}b), and summary statistics $\Dest$ and $\varest{\Dest}$ can be derived.
 
@@
 
-\section*{CR\lowercase{e}d\lowercase{i}T Author Statement}
+\section*{Author Contribution Statement}
 
@@
 B.J.M. acknowledges support from the Royal Society (UF130329 and URF\textbackslash R\textbackslash 191006). 
+The funders listed above played no role in study design, data collection, analysis and interpretation of data, or the writing of this manuscript.
+
+\section*{Competing interests}
+All authors declare no financial or non-financial competing interests.
 
@@
 where $\nind{n}$ is the total number of statistically independent squared displacements that contribute to $\oMSDi$.
-In the long-time limit, $\nind{n}$ is given by the product of the number of mobile particles and the number of non-overlapping time-windows of length $i$ in our simulation trajectory.
-Note that $\nind{n}$ considers non-overlapping time windows, since mutually overlapping time-windows give correlated squared displacements.
+In the long-time limit, $\nind{n}$ is given by the product of the number of mobile particles and the number of numerically-independent sub-trajectories of length $i$ in our simulation trajectory.
+Note that $\nind{n}$ considers numerically-independent sub-trajectories, since mutually overlapping time-windows give correlated squared displacements.
 
@@
 In this case, the covariance depends only on the number of overlapping points, $n$, between the two time intervals. 
-We can rationalise this by noting that for a random walk any non-overlapping points will be completely uncorrelated and therefore have a covariance of \num{0}. 
+We can rationalise this by noting that for a random walk any numerically-independent points will be completely uncorrelated and therefore have a covariance of \num{0}. 
 Similar to the case for the variance, the covariance derived in Eqn.~\ref{equ:cov_der} is that for a single particle at a single time origin. 
@@
 
+\section{Comparison With Block Averaging Approach}
+\label{sec:block}
+Eqn.~\ref{equ:varestMSD} enables the approximate estimation of the variance in $\oMSDi$, as mentioned in the main text, it is also possible to obtain this from a block averaging approach~\cite{flyvbjerg_error_1989,frenkel_understanding_2023}.
+Fig.~\ref{fig:pyblock} compares the $\D$ estimate distributions using variances from approximated by Eqn.~\ref{equ:varestMSD} with those from the block averaging approach (using the \textsc{pyblock} Python package~\cite{spencer_pyblock_2020}) both with the fitting of the variances to Eqn.~\ref{equ:variance} and without.
+When the fitting is not performed, the resulting covariance matrix is numerically unstable, leading to estimated values of $\D$ at extreme values.
+While when the fitting is performed and the blocking-estimated variances as used, there is no improvement in the estimation of $\D$ and the resulting distribution of estimates in the variance of $\prob{\D}$ is broader than when Eqn.~\ref{equ:varestMSD} is used.
+
+\begin{figure}
+    \centering
+    % \resizebox{\columnwidth}{!}{\includegraphics*{figure1.pdf}}
+    \includegraphics[width=\columnwidth]{pyblock.pdf}
+    \script{pyblock.py}
+    \caption{Comparison of the distribution of $\Dest$ (a, c, \& e) and the estimated variances (b, d, \& f) from 4096 individual random walk simulations using variances approximated by Eqn.~\ref{equ:varestMSD} (a \& b), from block averaging with fitting to Eqn.~\ref{equ:variance} (c \& d), and without fitting (e \& f).
+Note that the bounds in a, c, \& e are defined by the extrema of the estimated values, showing the fact that without the fitting to Eqn.~\ref{equ:variance}, the resulting covariance matrix is numerically unstable.}
+    \label{fig:pyblock}
+\end{figure}
+
 \section{Bias and skew in $\prob{\varest{\Dest}}$}
@@
 Furthermore, the distribution $\prob{\Dest}$ agrees even more closely with the numerically converged distribution obtained when combining data from all \num{4096} simulations (Fig~\ref{fig:true_cov}a), contrasting with the results presented in Fig.~\ref{fig:random_walk}b, where our approximate Bayesian regression scheme yields a slightly broadened distribution due to the use of the long-time limit in the derivation of the analytical form for $\bm{\Sigma^\prime}$.
-\begin{figure}
-    \centering
-    % \resizebox{\columnwidth}{!}{\includegraphics*{figure7.pdf}}
-    \includegraphics[width=\columnwidth]{true_cov.pdf}
-    \script{true_cov.py}
-    \caption{
-        (a) Probability distribution of point-estimates $\prob{\Dest}$ obtained from \num{4096} individual random-walk simulations, using the numerical covariance matrix $\bm{\Sigma}_\mathrm{num}$.
-        Each simulation has been analysed as in Fig.~~\ref{equ:observed_msd}(a) and (b) to yield a single corresponding point estimate $\Dest$.
-        The grey line shows the distribution of point estimates, $\prob{\Dest_\mathrm{num}}$, obtained using Bayesian regression with a mean vector and numerical covariance matrix derived from the complete dataset of all \num{4096} simulations.
-        The pink horizontal bar shows an interval of one standard deviation in $\prob{\Dest}$.
-        (b) Probability distribution of estimated variances, $\varest{\Dest}$, for individual random-walk simulations, using the numerical covariance matrix $\bm{\Sigma}_\mathrm{num}$, compared to the true sample variance (pink vertical line) $\var{\Dest}$.
-        }
-    \label{fig:true_cov}
-\end{figure}
+% \begin{figure}
+%     \centering
+%     % \resizebox{\columnwidth}{!}{\includegraphics*{figure7.pdf}}
+%     \includegraphics[width=\columnwidth]{true_cov.pdf}
+%     \script{true_cov.py}
+%     \caption{
+%         (a) Probability distribution of point-estimates $\prob{\Dest}$ obtained from \num{4096} individual random-walk simulations, using the numerical covariance matrix $\bm{\Sigma}_\mathrm{num}$.
+%         Each simulation has been analysed as in Fig.~~\ref{equ:observed_msd}(a) and (b) to yield a single corresponding point estimate $\Dest$.
+%         The grey line shows the distribution of point estimates, $\prob{\Dest_\mathrm{num}}$, obtained using Bayesian regression with a mean vector and numerical covariance matrix derived from the complete dataset of all \num{4096} simulations.
+%         The pink horizontal bar shows an interval of one standard deviation in $\prob{\Dest}$.
+%         (b) Probability distribution of estimated variances, $\varest{\Dest}$, for individual random-walk simulations, using the numerical covariance matrix $\bm{\Sigma}_\mathrm{num}$, compared to the true sample variance (pink vertical line) $\var{\Dest}$.
+%         }
+%     \label{fig:true_cov}
+% \end{figure}
+
+\section{Evaluation of OLS, WLS, and GLS for LLZO System}
+
+In Fig.~\ref{fig:glswlsols}, it is shown that for a 3D lattice random walk the heteroscedastic and correlated nature of the data requires generalised least squares to optimally estimate the true ensemble-average MSD, and therefore estimate $\D$.
+This is also true for the real materials, such as the \ce{Li7La3Zr2O12} (LLZO) investigated by classical molecular dynamics simulation in this work (Fig.~\ref{fig:glswlsols_llzo}).
+Once again, GLS (or, indeed, the Bayesian regression equivalent, shown in Fig.~\ref{fig:diffusion}) offers both the most statistically efficient estimation of $\D$ and is required to obtain an accruate estimate of the statistical uncertainty in $\D$ from a single simulation.
+% \begin{figure}
+%     \centering
+%     % \resizebox{\columnwidth}{!}{\includegraphics*{figure1.pdf}}
+%     \includegraphics[width=\columnwidth]{glswlsols_llzo.pdf}
+%     \script{glswlsols_llzo.py}
+%     \caption{
+%         Example distributions of estimated self-diffusion coefficients, $\Dest$, calculated using (a) ordinary least squares (OLS), (b) weighted least squares (WLS), and (c) generalised least squares (GLS),
+%         from MSD data from \num{512} effective simulations of LLZO of \SI{\sim 25}{\pico\second} with 56 lithium ions.
+%         In each panel, the grey curve shows the best-fit normal distribution for the simulation data, the upper horizontal bar shows the standard deviation of this distribution, and the lower horizontal bar shows the average estimated standard distribution given by the analytical expression for $\sigma[\prob{\Dest}]$ for each regression method.}
+%     \label{fig:glswlsols_llzo}
+% \end{figure}
 

### 2023-11-23 src/code/llzo/glswlsols.py

@@ -1,0 +1,51 @@
+# Copyright (c) Andrew R. McCluskey
+# Distributed under the terms of the MIT License
+# author: Andrew R. McCluskey (arm61)
+
+import numpy as np
+from scipy.stats import linregress
+from scipy.linalg import pinvh
+from emcee import EnsembleSampler
+from scipy.optimize import minimize
+from kinisi.matrix import find_nearest_positive_definite
+from kinisi.diffusion import _straight_line
+
+np.random.seed(1)
+
+start_diff = snakemake.params['start_diff']
+
+timestep = np.load(f'src/data/llzo/diffusion_0_{start_diff}.npz')['dt'][0, 0]
+no = np.load(f'src/data/llzo/diffusion_0_{start_diff}.npz')['no'][0, 0]
+d = np.zeros((16, 8, 1, timestep.size))
+for i in range(0, 16, 1):
+    d[i] = np.load(f'src/data/llzo/diffusion_{i}_{start_diff}.npz')['msd_true']
+
+max_ngp = np.argwhere(timestep > start_diff)[0][0]
+true_msd = d.reshape(-1, d.shape[-1])[:, max_ngp:]
+true_cov = np.cov(true_msd.T)
+timestep = timestep[max_ngp:] 
+
+y = true_msd.T
+A = np.array([timestep, np.ones(timestep.size)]).T
+W = true_cov
+V = np.diag(true_cov.diagonal())
+
+g1 = np.matmul(np.linalg.inv(np.matmul(A.T, np.matmul(np.linalg.pinv(W), A))),
+               np.matmul(A.T, np.matmul(np.linalg.pinv(W), y)))[0] / 6
+c1 = np.sqrt(np.linalg.inv(np.matmul(A.T, np.matmul(np.linalg.pinv(W), A)))[0][0]) / 6
+g2 = np.matmul(np.linalg.inv(np.matmul(A.T, np.matmul(np.linalg.pinv(V), A))),
+               np.matmul(A.T, np.matmul(np.linalg.pinv(V), y)))[0] / 6
+c2 = np.sqrt(np.linalg.inv(np.matmul(A.T, np.matmul(np.linalg.pinv(V), A)))[0][0]) / 6
+g3 = np.matmul(np.linalg.inv(np.matmul(A.T, A)), np.matmul(A.T, y))[0] / 6
+
+c3 = []
+for i in true_msd:
+    c3.append(linregress(timestep, i).stderr / 6)
+
[…]

### 2023-11-23 src/code/llzo/many_runs.py

@@ -41,25 +41,25 @@ file = f'src/data/llzo/traj{n}.xyz'
 
 
 u = mda.Universe(file, file)
-uu = universe_slice(u, file, 0, 500)
+uu = universe_slice(u, file, 0, 2000)
 step_skip = 100  # sampling rate
 timestep = 5.079648e-4  # ps
 da_params = {'specie': 'Li', 'time_step': timestep, 'step_skip': step_skip}
 p = MDAnalysisParser(uu, **da_params)
 
-ngp = np.zeros((8, 4, p.delta_t.size))
-no = np.zeros((8, 4, p.delta_t.size))
-dt = np.zeros((8, 4, p.delta_t.size))
-msd = np.zeros((8, 4, p.delta_t.size))
-msd_true = np.zeros((8, 4, p.delta_t.size))
-msd_std = np.zeros((8, 4, p.delta_t.size))
-cov = np.zeros((8, 4, p.delta_t[np.where(p.delta_t > start_diff)].size, p.delta_t[np.where(p.delta_t > start_diff)].size))
-d = np.zeros((8, 4, 3200))
-intercept = np.zeros((8, 4, 3200))
+ngp = np.zeros((8, 1, p.delta_t.size))
+no = np.zeros((8, 1, p.delta_t.size))
+dt = np.zeros((8, 1, p.delta_t.size))
+msd = np.zeros((8, 1, p.delta_t.size))
+msd_true = np.zeros((8, 1, p.delta_t.size))
+msd_std = np.zeros((8, 1, p.delta_t.size))
+cov = np.zeros((8, 1, p.delta_t[np.where(p.delta_t > start_diff)].size, p.delta_t[np.where(p.delta_t > start_diff)].size))
+d = np.zeros((8, 1, 3200))
+intercept = np.zeros((8, 1, 3200))
 
 for m in range(0, 8, 1):
-    for i, slice in enumerate(range(0, 2000, 500)):
-        uu = universe_slice(u, file, slice, slice+500)
+    for i, slice in enumerate(range(0, 2000, 2000)):
+        uu = universe_slice(u, file, slice, slice+2000)
 
         rng = np.random.RandomState(42)
         np.random.seed(42)
@@ -76,7 +76,7 @@ for m in range(0, 8, 1):
                 msd_true[m, i, t] = np.sum(j[m::8, ::p.timesteps[t]]**2, axis=-1).mean()
 
[…]

### 2023-11-23 src/code/llzo/true_ls.py

@@ -15,7 +15,7 @@ start_diff = snakemake.params['start_diff']
 
 timestep = np.load(f'src/data/llzo/diffusion_0_{start_diff}.npz')['dt'][0, 0]
 no = np.load(f'src/data/llzo/diffusion_0_{start_diff}.npz')['no'][0, 0]
-d = np.zeros((16, 8, 4, timestep.size))
+d = np.zeros((16, 8, 1, timestep.size))
 for i in range(0, 16, 1):
     d[i] = np.load(f'src/data/llzo/diffusion_{i}_{start_diff}.npz')['msd_true']
 

### 2023-11-23 src/code/random_walks/kinisi_pyblock_modelfree_rw.py

@@ -25,8 +25,8 @@ for seed in tqdm(range(size)):
     np.random.seed(seed)
     disp_3d, n_samples = get_disp3d(walk, length, atoms, jump_size=jump, seed=rng)
 
-    diff = MSDBootstrap(timestep, disp_3d, n_samples, random_state=rng, progress=False)
-    diff.diffusion(5, random_state=rng, progress=False)
+    diff = MSDBootstrap(timestep, disp_3d, n_samples, random_state=rng, progress=False, block=True)
+    diff.diffusion(5, model=False, random_state=rng, progress=False)
     diff_c[seed] = diff.gradient.samples / 6
     intercept[seed] = diff.intercept.samples
     data[seed, :, 0] = diff.dt
@@ -37,7 +37,7 @@ for seed in tqdm(range(size)):
     covariance[seed] = diff.covariance_matrix
     n_o[seed] = diff._n_o
 
-np.savez(f'src/data/random_walks/kinisi/rw_1_{atoms}_{length}_s{size}.npz',
+np.savez(f'src/data/random_walks/pyblock_modelfree/rw_1_{atoms}_{length}_s{size}.npz',
          diff_c=diff_c,
          intercept=intercept,
          data=data,

### 2023-11-23 src/code/random_walks/kinisi_pyblock_rw.py

@@ -25,7 +25,7 @@ for seed in tqdm(range(size)):
     np.random.seed(seed)
     disp_3d, n_samples = get_disp3d(walk, length, atoms, jump_size=jump, seed=rng)
 
-    diff = MSDBootstrap(timestep, disp_3d, n_samples, random_state=rng, progress=False)
+    diff = MSDBootstrap(timestep, disp_3d, n_samples, random_state=rng, progress=False, block=True)
     diff.diffusion(5, random_state=rng, progress=False)
     diff_c[seed] = diff.gradient.samples / 6
     intercept[seed] = diff.intercept.samples
@@ -37,7 +37,7 @@ for seed in tqdm(range(size)):
     covariance[seed] = diff.covariance_matrix
     n_o[seed] = diff._n_o
 
-np.savez(f'src/data/random_walks/kinisi/rw_1_{atoms}_{length}_s{size}.npz',
+np.savez(f'src/data/random_walks/pyblock/rw_1_{atoms}_{length}_s{size}.npz',
          diff_c=diff_c,
          intercept=intercept,
          data=data,

### 2023-11-23 src/scripts/diffusion.py

@@ -13,7 +13,7 @@ length = 128
 type = 'kinisi'
 
 dllzo_true = np.load(paths.data / "llzo/true_10.npz")['diff_c']
-d = np.zeros((16, 8, 4, 3200))
+d = np.zeros((16, 8, 1, 3200))
 for i in range(0, 16, 1):
     d[i] = np.load(paths.data / f'llzo/diffusion_{i}_10.npz')['d']
 d = d.reshape(-1, 3200)

### 2023-11-23 src/scripts/glswlsols_llzo.py

@@ -1,0 +1,90 @@
+import numpy as np
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+import matplotlib.transforms as transforms
+from scipy.stats import norm
+from uravu.distribution import Distribution
+
+import utils._fig_params as fp
+import paths
+
+data = np.load('src/data/llzo/glswlsols_10.npz')
+
+figsize = (3.744, 3.4)
+fig = plt.figure(figsize=figsize)
+gs = gridspec.GridSpec(3, 1, figure=fig, hspace=0.8)
+
+x = np.linspace(data['gls_pop'].mean() - data['ols_pop'].std() * 3, data['gls_pop'].mean() + data['ols_pop'].std() * 3, 1000)
+
+axes = []
+titles = []
+axes.append(fig.add_subplot(gs[0, 0]))
+bins = np.linspace((data['ols_pop']).min(), (data['ols_pop']).max(), fp.NBINS)
+axes[-1].hist(data['ols_pop'], bins=bins, density=True, color='#ECBCA7')
+axes[-1].plot(x, norm.pdf(x, *norm.fit(data['ols_pop'])), color='#B8B8B8')
+axes[-1].plot([data['gls_pop'].mean() - data['ols_est'], data['gls_pop'].mean() + data['ols_est']],
+              [axes[-1].get_ylim()[1] * 1.1, axes[-1].get_ylim()[1] * 1.1],
+              color='#305980',
+              marker='|')
+axes[-1].plot([data['gls_pop'].mean() - data['ols_pop'].std(), data['gls_pop'].mean() + data['ols_pop'].std()],
+              [axes[-1].get_ylim()[1] * 1.1, axes[-1].get_ylim()[1] * 1.1],
+              color='#F3ADBC',
+              marker='|')
+# axes[-1].set_xticks([0.8, 0.9, 1, 1.1, 1.2])
+# axes[-1].set_xlim(0.75, 1.25)
+# axes[-1].set_yticks([0, 3, 6])
+axes[-1].set_ylabel(r'$p(\hat{D}^*)$')
+titles.append("a")
+axes[-1].set_title("ordinary least squares")
+
+axes.append(fig.add_subplot(gs[1, 0], sharex=axes[0]))
+axes[-1].hist(data['wls_pop'], bins=bins, density=True, color='#ECBCA7')
[…]

### 2023-11-23 src/scripts/pyblock.py

@@ -1,0 +1,162 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+from scipy.stats import norm
+
+import utils._fig_params as fp
+from utils.plotting_helper import mid_points, CREDIBLE_INTERVALS
+import paths
+
+jump = 1
+atoms = 128
+length = 128
+
+kinisi = np.load(paths.data / f"random_walks/kinisi/rw_{jump}_{atoms}_{length}_s4096.npz")['diff_c']
+pyblock = np.load(paths.data / f"random_walks/pyblock/rw_{jump}_{atoms}_{length}_s4096.npz")['diff_c']
+pyblock_mf = np.load(paths.data / f"random_walks/pyblock_modelfree/rw_{jump}_{atoms}_{length}_s4096.npz")['diff_c']
+dinfty_true = np.load(paths.data / f"random_walks/numerical/D_1_128_128.npz")['diff_c']
+
+rw_x = np.arange(kinisi.mean(-1).min(), kinisi.mean(-1).max(), 0.01)
+rw_y = np.arange(pyblock.mean(-1).min(), pyblock.mean(-1).max(), 0.01)
+rw_z = np.arange(pyblock_mf.mean(-1).min(), pyblock_mf.mean(-1).max(), 0.01)
+rw_x2 = np.arange(np.array([kinisi.var(-1, ddof=1), pyblock.var(-1, ddof=1), pyblock_mf.var(-1, ddof=1)]).min(), 
+                  np.array([kinisi.var(-1, ddof=1), pyblock.var(-1, ddof=1), pyblock_mf.var(-1, ddof=1)]).max(), 0.0001)
+rw_x3 = np.arange(kinisi.mean(-1).min(), kinisi.mean(-1).max(), 0.0001)
+rw_y3 = np.arange(pyblock.mean(-1).min(), pyblock.mean(-1).max(), 0.0001)
+rw_z3 = np.arange(pyblock_mf.mean(-1).min(), pyblock_mf.mean(-1).max(), 0.0001)
+
+figsize = (4.03, 6)
+print(figsize)
+fig = plt.figure(figsize=figsize)
+gs = gridspec.GridSpec(3, 2, figure=fig, wspace=0.6, hspace=1)
+
+axes = []
+titles = []
+
+axes.append(fig.add_subplot(gs[0, 0]))
[…]

### 2023-11-23 src/scripts/random_walk.py

@@ -101,7 +101,6 @@ axes[-1].axvline(true.mean(-1).var(ddof=1),
 axes[-1].set_xlabel(r"$\sigma^2 [\hat{D}^*]$")
 axes[-1].set_ylabel(r"$p(\sigma^2 [\hat{D}^*])$")
 axes[-1].set_xlim([0, None])
-axes[-1].set_yticks([0, 1, 2, 3])
 titles.append("d")
 axes[-1].set_title("all simulations")
 axes[-1].legend(loc='upper left', bbox_to_anchor=(0.5, 1))

### 2023-11-24 src/tex/ms.tex
Updated version

@@
 %
-% \begin{figure}
-%     \centering
-%     % \resizebox{\columnwidth}{!}{\includegraphics*{figure2.pdf}}
-%     \includegraphics[width=\columnwidth]{msd.pdf}
-%     \script{msd.py}
-%     \caption{
-%         Comparison of the numerical variance in observed MSD from multiple replica simulations and the estimated variance in observed MSD given by rescaling the variance in observed squared displacements (Eqn.~\ref{equ:varestMSD}).
-%         Panel (a) shows the mean observed MSD from \num{4096} simulations of \num{128} particles undergoing a 3D lattice random walk of \num{128} steps per particle, with error bars of $\pm2\sigma[\oMSDi]$.
-%         Panel (b) shows the MSD from just one simulation, with error bars of $\pm2\widehat{\sigma}[\oMSDi]$, obtained via Eqn.~\ref{equ:varestMSD}.
-%         Panel (c) plots the numerical variance against the estimated variance from a single simulation as a function of timestep $i$.
-%         %Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences.
-%     }
-%     \label{fig:msd}
-% \end{figure}
+\begin{figure}
+    \centering
+    % \resizebox{\columnwidth}{!}{\includegraphics*{figure2.pdf}}
+    \includegraphics[width=\columnwidth]{msd.pdf}
+    \script{msd.py}
+    \caption{
+        Comparison of the numerical variance in observed MSD from multiple replica simulations and the estimated variance in observed MSD given by rescaling the variance in observed squared displacements (Eqn.~\ref{equ:varestMSD}).
+        Panel (a) shows the mean observed MSD from \num{4096} simulations of \num{128} particles undergoing a 3D lattice random walk of \num{128} steps per particle, with error bars of $\pm2\sigma[\oMSDi]$.
+        Panel (b) shows the MSD from just one simulation, with error bars of $\pm2\widehat{\sigma}[\oMSDi]$, obtained via Eqn.~\ref{equ:varestMSD}.
+        Panel (c) plots the numerical variance against the estimated variance from a single simulation as a function of timestep $i$.
+        %Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences.
+    }
+    \label{fig:msd}
+\end{figure}
 %
@@
 
-% \begin{figure}[tb]
-%     \centering
-%     % \resizebox{\columnwidth}{!}{\includegraphics*{figure5.pdf}}
-%     \includegraphics[width=\columnwidth]{diffusion.pdf}
-%     \script{diffusion.py}
-%     \caption{
-%       (a) Probability distribution of point estimates $\prob{\Dest}$ for \num{512} effective simulations of LLZO (orange histogram).
-%       The grey line shows the distribution $\prob{\Dest_\mathrm{num}}$ obtained using Bayesian regression with the complete LLZO dataset as input.
-%       The pink bar shows an interval of one standard deviation $\sigma[\prob{\Dest}]$.
-%       (b) Probability distribution of estimated variances, $\varest{\Dest}$, for individual LLZO effective simulations, compared to the true sample variance (pink vertical line) $\var{\Dest}$.
-%       %Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences. 
-%     }
-%     \label{fig:diffusion}
-% \end{figure}
+\begin{figure}[tb]
+    \centering
+    % \resizebox{\columnwidth}{!}{\includegraphics*{figure5.pdf}}
+    \includegraphics[width=\columnwidth]{diffusion.pdf}
+    \script{diffusion.py}
+    \caption{
+      (a) Probability distribution of point estimates $\prob{\Dest}$ for \num{512} effective simulations of LLZO (orange histogram).
+      The grey line shows the distribution $\prob{\Dest_\mathrm{num}}$ obtained using Bayesian regression with the complete LLZO dataset as input.
+      The pink bar shows an interval of one standard deviation $\sigma[\prob{\Dest}]$.
+      (b) Probability distribution of estimated variances, $\varest{\Dest}$, for individual LLZO effective simulations, compared to the true sample variance (pink vertical line) $\var{\Dest}$.
+      %Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences. 
+    }
+    \label{fig:diffusion}
+\end{figure}
 
@@
 
-% \begin{figure}[htb]
-%     \centering
-%     % \resizebox{\columnwidth}{!}{\includegraphics*{figure6.pdf}}
-%     \includegraphics[width=\columnwidth]{stat_eff.pdf}
-%     \script{stat_eff.py}
-%     \caption{
-%       Scaling of $\var{\Dest}$ with simulation size for OLS (pink), WLS (blue), our approximate Bayesian regression method (green), and GLS (orange).
-%       (a) Scaling versus number of mobile particles, $N_\mathrm{atoms}$.
-%       (b) Scaling versus total simulation time, $t_\mathrm{max}$.
-%       Solid lines show fitted power law relationships for each dataset.
-%       The WLS and GLS data are obtained using numerically determined variances and covariance, respectively, from a set of \num{512} repeat simulations for each combination of $N_\mathrm{atoms}$ and $t_\mathrm{max}$.
-%     }
-%     \label{fig:stat_eff}
-% \end{figure}
+\begin{figure}[htb]
+    \centering
+    % \resizebox{\columnwidth}{!}{\includegraphics*{figure6.pdf}}
+    \includegraphics[width=\columnwidth]{stat_eff.pdf}
+    \script{stat_eff.py}
+    \caption{
+      Scaling of $\var{\Dest}$ with simulation size for OLS (pink), WLS (blue), our approximate Bayesian regression method (green), and GLS (orange).
+      (a) Scaling versus number of mobile particles, $N_\mathrm{atoms}$.
+      (b) Scaling versus total simulation time, $t_\mathrm{max}$.
+      Solid lines show fitted power law relationships for each dataset.
+      The WLS and GLS data are obtained using numerically determined variances and covariance, respectively, from a set of \num{512} repeat simulations for each combination of $N_\mathrm{atoms}$ and $t_\mathrm{max}$.
+    }
+    \label{fig:stat_eff}
+\end{figure}
 
@@
 Furthermore, the distribution $\prob{\Dest}$ agrees even more closely with the numerically converged distribution obtained when combining data from all \num{4096} simulations (Fig~\ref{fig:true_cov}a), contrasting with the results presented in Fig.~\ref{fig:random_walk}b, where our approximate Bayesian regression scheme yields a slightly broadened distribution due to the use of the long-time limit in the derivation of the analytical form for $\bm{\Sigma^\prime}$.
-% \begin{figure}
-%     \centering
-%     % \resizebox{\columnwidth}{!}{\includegraphics*{figure7.pdf}}
-%     \includegraphics[width=\columnwidth]{true_cov.pdf}
-%     \script{true_cov.py}
-%     \caption{
-%         (a) Probability distribution of point-estimates $\prob{\Dest}$ obtained from \num{4096} individual random-walk simulations, using the numerical covariance matrix $\bm{\Sigma}_\mathrm{num}$.
-%         Each simulation has been analysed as in Fig.~~\ref{equ:observed_msd}(a) and (b) to yield a single corresponding point estimate $\Dest$.
-%         The grey line shows the distribution of point estimates, $\prob{\Dest_\mathrm{num}}$, obtained using Bayesian regression with a mean vector and numerical covariance matrix derived from the complete dataset of all \num{4096} simulations.
-%         The pink horizontal bar shows an interval of one standard deviation in $\prob{\Dest}$.
-%         (b) Probability distribution of estimated variances, $\varest{\Dest}$, for individual random-walk simulations, using the numerical covariance matrix $\bm{\Sigma}_\mathrm{num}$, compared to the true sample variance (pink vertical line) $\var{\Dest}$.
-%         }
-%     \label{fig:true_cov}
-% \end{figure}
+\begin{figure}
+    \centering
+    % \resizebox{\columnwidth}{!}{\includegraphics*{figure7.pdf}}
+    \includegraphics[width=\columnwidth]{true_cov.pdf}
+    \script{true_cov.py}
+    \caption{
+        (a) Probability distribution of point-estimates $\prob{\Dest}$ obtained from \num{4096} individual random-walk simulations, using the numerical covariance matrix $\bm{\Sigma}_\mathrm{num}$.
+        Each simulation has been analysed as in Fig.~~\ref{equ:observed_msd}(a) and (b) to yield a single corresponding point estimate $\Dest$.
+        The grey line shows the distribution of point estimates, $\prob{\Dest_\mathrm{num}}$, obtained using Bayesian regression with a mean vector and numerical covariance matrix derived from the complete dataset of all \num{4096} simulations.
+        The pink horizontal bar shows an interval of one standard deviation in $\prob{\Dest}$.
+        (b) Probability distribution of estimated variances, $\varest{\Dest}$, for individual random-walk simulations, using the numerical covariance matrix $\bm{\Sigma}_\mathrm{num}$, compared to the true sample variance (pink vertical line) $\var{\Dest}$.
+        }
+    \label{fig:true_cov}
+\end{figure}
 
@@
 Once again, GLS (or, indeed, the Bayesian regression equivalent, shown in Fig.~\ref{fig:diffusion}) offers both the most statistically efficient estimation of $\D$ and is required to obtain an accruate estimate of the statistical uncertainty in $\D$ from a single simulation.
-% \begin{figure}
-%     \centering
-%     % \resizebox{\columnwidth}{!}{\includegraphics*{figure1.pdf}}
-%     \includegraphics[width=\columnwidth]{glswlsols_llzo.pdf}
-%     \script{glswlsols_llzo.py}
-%     \caption{
-%         Example distributions of estimated self-diffusion coefficients, $\Dest$, calculated using (a) ordinary least squares (OLS), (b) weighted least squares (WLS), and (c) generalised least squares (GLS),
-%         from MSD data from \num{512} effective simulations of LLZO of \SI{\sim 25}{\pico\second} with 56 lithium ions.
-%         In each panel, the grey curve shows the best-fit normal distribution for the simulation data, the upper horizontal bar shows the standard deviation of this distribution, and the lower horizontal bar shows the average estimated standard distribution given by the analytical expression for $\sigma[\prob{\Dest}]$ for each regression method.}
-%     \label{fig:glswlsols_llzo}
-% \end{figure}
+\begin{figure}
+    \centering
+    % \resizebox{\columnwidth}{!}{\includegraphics*{figure1.pdf}}
+    \includegraphics[width=\columnwidth]{glswlsols_llzo.pdf}
+    \script{glswlsols_llzo.py}
+    \caption{
+        Example distributions of estimated self-diffusion coefficients, $\Dest$, calculated using (a) ordinary least squares (OLS), (b) weighted least squares (WLS), and (c) generalised least squares (GLS),
+        from MSD data from \num{512} effective simulations of LLZO of \SI{\sim 25}{\pico\second} with 56 lithium ions.
+        In each panel, the grey curve shows the best-fit normal distribution for the simulation data, the upper horizontal bar shows the standard deviation of this distribution, and the lower horizontal bar shows the average estimated standard distribution given by the analytical expression for $\sigma[\prob{\Dest}]$ for each regression method.}
+    \label{fig:glswlsols_llzo}
+\end{figure}
 

### 2023-11-24 src/code/random_walks/glswlsols.py

@@ -7,7 +7,7 @@ from scipy.stats import linregress
 from scipy.linalg import pinvh
 from emcee import EnsembleSampler
 from scipy.optimize import minimize
-from kinisi.matrix import find_nearest_positive_definite
+from statsmodels.stats.correlation_tools import cov_nearest 
 from kinisi.diffusion import _straight_line
 
 np.random.seed(1)

### 2023-11-24 src/code/random_walks/true_ls.py

@@ -7,7 +7,7 @@ from scipy.stats import linregress
 from scipy.linalg import pinvh
 from emcee import EnsembleSampler
 from scipy.optimize import minimize
-from kinisi.matrix import find_nearest_positive_definite
+from statsmodels.stats.correlation_tools import cov_nearest 
 from kinisi.diffusion import _straight_line
 from tqdm import tqdm
 

### 2023-11-27 src/tex/ms.tex
Mention difference between overlapping/non

@@
 \end{equation}
-where $\nind{n}$ is the total number of statistically independent squared displacements that contribute to $\oMSDi$.
+where $\nind{n}$ is the total number of statistically independent (non-overlapping) squared displacements that contribute to $\oMSDi$.
 In the long-time limit, $\nind{n}$ is given by the product of the number of mobile particles and the number of numerically-independent sub-trajectories of length $i$ in our simulation trajectory.
 Note that $\nind{n}$ considers numerically-independent sub-trajectories, since mutually overlapping time-windows give correlated squared displacements.
+Where overlapping time-windows are used, Eqn.~\ref{equ:der_var} approximates the observed variance (the variance for overlapping and non-overlapping samples are not the same) with accuracy that increases as a function of time-interval length.
+This leads to the approximation in Eqn.~\ref{equ:varestMSD}.
 

### 2024-01-11 src/tex/ms.tex
To run on HPC

--- 2023-11-23 src/tex/ms.tex
@@
 
-\begin{figure}
-    \centering
-    % \resizebox{\columnwidth}{!}{\includegraphics*{figure1.pdf}}
-    \includegraphics[width=\columnwidth]{glswlsols.pdf}
-    \script{glswlsols.py}
-    \caption{
-        Example distributions of estimated self-diffusion coefficients, $\Dest$, calculated using (a) ordinary least squares (OLS), (b) weighted least squares (WLS), and (c) generalised least squares (GLS),
-        from MSD data from \num{4096} individual simulations of \num{128} particles undergoing a \SI{128}{step} 3D lattice random walk, with a step size chosen so that the true diffusion coefficient $\D = 1$.
-        In each panel, the grey curve shows the best-fit normal distribution for the simulation data, the upper horizontal bar shows the standard deviation of this distribution, and the lower horizontal bar shows the average estimated standard distribution given by the analytical expression for $\sigma[\prob{\Dest}]$ for each regression method.}
-    \label{fig:glswlsols}
-\end{figure}
+% \begin{figure}
+%     \centering
+%     % \resizebox{\columnwidth}{!}{\includegraphics*{figure1.pdf}}
+%     \includegraphics[width=\columnwidth]{glswlsols.pdf}
+%     \script{glswlsols.py}
+%     \caption{
+%         Example distributions of estimated self-diffusion coefficients, $\Dest$, calculated using (a) ordinary least squares (OLS), (b) weighted least squares (WLS), and (c) generalised least squares (GLS),
+%         from MSD data from \num{4096} individual simulations of \num{128} particles undergoing a \SI{128}{step} 3D lattice random walk, with a step size chosen so that the true diffusion coefficient $\D = 1$.
+%         In each panel, the grey curve shows the best-fit normal distribution for the simulation data, the upper horizontal bar shows the standard deviation of this distribution, and the lower horizontal bar shows the average estimated standard distribution given by the analytical expression for $\sigma[\prob{\Dest}]$ for each regression method.}
+%     \label{fig:glswlsols}
+% \end{figure}
 
@@
 %
-\begin{figure}
-    \centering
-    % \resizebox{\columnwidth}{!}{\includegraphics*{figure3.pdf}}
-    \includegraphics[width=\columnwidth]{covariances.pdf}
-    \script{covariances.py}
-    \caption{
-        (a) The numerical MSD covariance matrix $\bm{\Sigma}$ calculated using MSD data from \num{4096} simulations of \num{128} particles undergoing a 3D lattice random walk of \num{128} steps per particle.
-        (b) The analytical MSD covariance matrix $\bm{\Sigma^\prime}$ (Eqn.~\ref{equ:cvv}), parametrised using analytical random-walk variances $\var{\oMSDi}$.
-        (c) The MSD covariance matrix obtained applying the numerical scheme described in the main text to each individual random walk simulation, averaged over all \num{4096} such simulations.
-        Colour bars in (a--c) show the covariance, $\Sigma\left[\oMSDi, \oMSDj\right]$.
-        The off-diagonal panels show difference plots, computed as per-element ratios between pairs of covariance matrices (a--c).
-        %Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences.
-    }
-    \label{fig:covariances}
-\end{figure}
+% \begin{figure}
+%     \centering
+%     % \resizebox{\columnwidth}{!}{\includegraphics*{figure3.pdf}}
+%     \includegraphics[width=\columnwidth]{covariances.pdf}
+%     \script{covariances.py}
+%     \caption{
+%         (a) The numerical MSD covariance matrix $\bm{\Sigma}$ calculated using MSD data from \num{4096} simulations of \num{128} particles undergoing a 3D lattice random walk of \num{128} steps per particle.
+%         (b) The analytical MSD covariance matrix $\bm{\Sigma^\prime}$ (Eqn.~\ref{equ:cvv}), parametrised using analytical random-walk variances $\var{\oMSDi}$.
+%         (c) The MSD covariance matrix obtained applying the numerical scheme described in the main text to each individual random walk simulation, averaged over all \num{4096} such simulations.
+%         Colour bars in (a--c) show the covariance, $\Sigma\left[\oMSDi, \oMSDj\right]$.
+%         The off-diagonal panels show difference plots, computed as per-element ratios between pairs of covariance matrices (a--c).
+%         %Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences.
+%     }
+%     \label{fig:covariances}
+% \end{figure}
 %
@@
 
-\begin{figure*}[ht!]
-    \centering
-    % \resizebox{\textwidth}{!}{\includegraphics*{figure4.pdf}}
-    \includegraphics[width=\textwidth]{random_walk.pdf}
-    \script{random_walk.py}
-    \caption{
-        (a) Observed MSD from a single simulation of 128 particles undergoing a 3D-lattice random walk of 128 steps per particle (dark line).
-        The green shading shows the corresponding posterior distribution $\prob{\model|\oMSD}$ of linear models compatible with the observed MSD data $\oMSD$, calculated using the scheme described in the main text.
-        The variegated shading indicates compatibility intervals of \SIlist[list-final-separator = {, and }]{1;2;3}{\sigma}$[\prob{\model|\oMSD}]$.
-        (b) The marginal posterior distribution $\prob{\Dest|\oMSD}$ obtained from the posterior distribution of linear models in (a).
-        The mean of this distribution gives the point estimate $\Dest$ for this simulation input data.
-        The blue horizontal bar shows an interval of one standard deviation in $\prob{\Dest|\oMSD}$.
-        (c) Probability distribution of point-estimates $\prob{\Dest}$ obtained from \num{4096} individual random-walk simulations.
-        Each simulation has been analysed as in (a) and (b) to yield a single corresponding point estimate $\Dest$.
-        The grey line shows the distribution of point estimates, $\prob{\Dest_\mathrm{num}}$, obtained using Bayesian regression with a mean vector and numerical covariance matrix derived from the complete dataset of all \num{4096} simulations.
-        The pink horizontal bar shows an interval of one standard deviation in $\prob{\Dest}$.
-        (d) Probability distribution of estimated variances, $\varest{\Dest}$, for individual random-walk simulations, compared to the true sample variance (pink vertical line) $\var{\Dest}$.
-    } 
-    \label{fig:random_walk}
-\end{figure*}
+% \begin{figure*}[ht!]
+%     \centering
+%     % \resizebox{\textwidth}{!}{\includegraphics*{figure4.pdf}}
+%     \includegraphics[width=\textwidth]{random_walk.pdf}
+%     \script{random_walk.py}
+%     \caption{
+%         (a) Observed MSD from a single simulation of 128 particles undergoing a 3D-lattice random walk of 128 steps per particle (dark line).
+%         The green shading shows the corresponding posterior distribution $\prob{\model|\oMSD}$ of linear models compatible with the observed MSD data $\oMSD$, calculated using the scheme described in the main text.
+%         The variegated shading indicates compatibility intervals of \SIlist[list-final-separator = {, and }]{1;2;3}{\sigma}$[\prob{\model|\oMSD}]$.
+%         (b) The marginal posterior distribution $\prob{\Dest|\oMSD}$ obtained from the posterior distribution of linear models in (a).
+%         The mean of this distribution gives the point estimate $\Dest$ for this simulation input data.
+%         The blue horizontal bar shows an interval of one standard deviation in $\prob{\Dest|\oMSD}$.
+%         (c) Probability distribution of point-estimates $\prob{\Dest}$ obtained from \num{4096} individual random-walk simulations.
+%         Each simulation has been analysed as in (a) and (b) to yield a single corresponding point estimate $\Dest$.
+%         The grey line shows the distribution of point estimates, $\prob{\Dest_\mathrm{num}}$, obtained using Bayesian regression with a mean vector and numerical covariance matrix derived from the complete dataset of all \num{4096} simulations.
+%         The pink horizontal bar shows an interval of one standard deviation in $\prob{\Dest}$.
+%         (d) Probability distribution of estimated variances, $\varest{\Dest}$, for individual random-walk simulations, compared to the true sample variance (pink vertical line) $\var{\Dest}$.
+%     } 
+%     \label{fig:random_walk}
+% \end{figure*}
 
@@
 
-\begin{figure}
-    \centering
-    % \resizebox{\columnwidth}{!}{\includegraphics*{figure1.pdf}}
-    \includegraphics[width=\columnwidth]{pyblock.pdf}
-    \script{pyblock.py}
-    \caption{Comparison of the distribution of $\Dest$ (a, c, \& e) and the estimated variances (b, d, \& f) from 4096 individual random walk simulations using variances approximated by Eqn.~\ref{equ:varestMSD} (a \& b), from block averaging with fitting to Eqn.~\ref{equ:variance} (c \& d), and without fitting (e \& f).
+% \begin{figure}
+%     \centering
+%     % \resizebox{\columnwidth}{!}{\includegraphics*{figure1.pdf}}
+%     \includegraphics[width=\columnwidth]{pyblock.pdf}
+%     \script{pyblock.py}
+%     \caption{Comparison of the distribution of $\Dest$ (a, c, \& e) and the estimated variances (b, d, \& f) from 4096 individual random walk simulations using variances approximated by Eqn.~\ref{equ:varestMSD} (a \& b), from block averaging with fitting to Eqn.~\ref{equ:variance} (c \& d), and without fitting (e \& f).
 Note that the bounds in a, c, \& e are defined by the extrema of the estimated values, showing the fact that without the fitting to Eqn.~\ref{equ:variance}, the resulting covariance matrix is numerically unstable.}
-    \label{fig:pyblock}
-\end{figure}
+%     \label{fig:pyblock}
+% \end{figure}
 

### 2024-01-11 src/code/llzo/glswlsols.py

@@ -15,9 +15,12 @@ np.random.seed(1)
 start_diff = snakemake.params['start_diff']
 
 timestep = np.load(f'src/data/llzo/diffusion_0_{start_diff}.npz')['dt'][0, 0]
-no = np.load(f'src/data/llzo/diffusion_0_{start_diff}.npz')['no'][0, 0]
-d = np.zeros((16, 8, 1, timestep.size))
-for i in range(0, 16, 1):
+length = 2000
+ll = len([i + length for i in range(0, 20001 - length, length)])
+# length = 500
+# ll = len([i + length for i in range(0, 2000, length)])
+d = np.zeros((6, 8, ll, timestep.size))
+for i in range(0, 6, 1):
     d[i] = np.load(f'src/data/llzo/diffusion_{i}_{start_diff}.npz')['msd_true']
 
 max_ngp = np.argwhere(timestep > start_diff)[0][0]
@@ -31,16 +34,16 @@ W = true_cov
 V = np.diag(true_cov.diagonal())
 
 g1 = np.matmul(np.linalg.inv(np.matmul(A.T, np.matmul(np.linalg.pinv(W), A))),
-               np.matmul(A.T, np.matmul(np.linalg.pinv(W), y)))[0] / 6
-c1 = np.sqrt(np.linalg.inv(np.matmul(A.T, np.matmul(np.linalg.pinv(W), A)))[0][0]) / 6
+               np.matmul(A.T, np.matmul(np.linalg.pinv(W), y)))[0] / 6e4
+c1 = np.sqrt(np.linalg.inv(np.matmul(A.T, np.matmul(np.linalg.pinv(W), A)))[0][0]) / 6e4
 g2 = np.matmul(np.linalg.inv(np.matmul(A.T, np.matmul(np.linalg.pinv(V), A))),
-               np.matmul(A.T, np.matmul(np.linalg.pinv(V), y)))[0] / 6
-c2 = np.sqrt(np.linalg.inv(np.matmul(A.T, np.matmul(np.linalg.pinv(V), A)))[0][0]) / 6
-g3 = np.matmul(np.linalg.inv(np.matmul(A.T, A)), np.matmul(A.T, y))[0] / 6
+               np.matmul(A.T, np.matmul(np.linalg.pinv(V), y)))[0] / 6e4
[…]

### 2024-01-11 src/code/llzo/many_runs.py

@@ -2,97 +2,84 @@ import MDAnalysis as mda
 import numpy as np
 from kinisi.parser import MDAnalysisParser
 from kinisi.diffusion import MSDBootstrap
-from tqdm import tqdm
 
 n = snakemake.params['n']
 start_diff = snakemake.params['start_diff']
 
-def universe_slice(u: mda.Universe, data_file: str, slice_start: int, slice_end: int) -> mda.Universe:
-    """
-    Concatenate a list of Universes.
+step = 8
 
-    :param u_list: List of Universe objects
-    :param data_file: File path for data
+def universe_slice(data: str, slice_start: int, slice_end: int, m: int) -> mda.Universe:
+    """
+    Create slice universe.
+
+    :param data: Numpy data of coordinates
     :param slice_start: Step to start the slice from
+    :param slice_end: Step to end the slice from
     :return: A single flattened Universe
     """
-    j = 0
-    trajectory = np.zeros((len(u.trajectory[slice_start:slice_end]), len(u.atoms), 3))
-    for t in tqdm(u.trajectory[slice_start:slice_end]):
-        trajectory[j] = t.positions
-        j += 1
-    shorter_universe = mda.Universe(data_file, trajectory)
-    for t in tqdm(shorter_universe.trajectory):
-        for i in shorter_universe.atoms:
-            if i.type == 'LI':
-                i.type = 'Li'
-            if i.type == 'Li1':
-                i.type = 'Li'
-            if i.type == 'Li2':
-                i.type = 'Li'
-            if i.type == 'L':
-                i.type = 'La'
-            if i.type == 'Z':
-                i.type = 'Zr'
-        t.dimensions = np.array([26.175106662244392197, 26.175106662244392197, 26.175106662244392197, 90, 90, 90])
-    return shorter_universe
+    data_reshape = np.copy(data).reshape(-1, 1536, 3)
+    data_subset = data_reshape[slice_start:slice_end, m::step]
+    data_subset *= 0.52917721067
+    u = mda.Universe.empty(data_subset.shape[1], 1, n_frames=data_subset.shape[0], trajectory=True)
+    list_type = ['A'] * (1536-448) + ['Li'] * 448
[…]

### 2024-01-11 src/code/llzo/true_ls.py

@@ -14,9 +14,12 @@ np.random.seed(1)
 start_diff = snakemake.params['start_diff']
 
 timestep = np.load(f'src/data/llzo/diffusion_0_{start_diff}.npz')['dt'][0, 0]
-no = np.load(f'src/data/llzo/diffusion_0_{start_diff}.npz')['no'][0, 0]
-d = np.zeros((16, 8, 1, timestep.size))
-for i in range(0, 16, 1):
+length = 2000
+ll = len([i + length for i in range(0, 20001 - length, length)])
+# length = 500 
+# ll = len([i + length for i in range(0, 2000, length)])
+d = np.zeros((6, 8, ll, timestep.size))
+for i in range(0, 6, 1):
     d[i] = np.load(f'src/data/llzo/diffusion_{i}_{start_diff}.npz')['msd_true']
 
 true_mean = d.reshape(-1, d.shape[-1]).mean(0)

### 2024-01-11 src/scripts/diffusion.py

@@ -13,8 +13,12 @@ length = 128
 type = 'kinisi'
 
 dllzo_true = np.load(paths.data / "llzo/true_10.npz")['diff_c']
-d = np.zeros((16, 8, 1, 3200))
-for i in range(0, 16, 1):
+length = 2000
+ll = len([i + length for i in range(0, 20001 - length, length)])
+# length = 500
+# ll = len([i + length for i in range(0, 2000, length)])
+d = np.zeros((6, 8, ll, 3200))
+for i in range(0, 6, 1):
     d[i] = np.load(paths.data / f'llzo/diffusion_{i}_10.npz')['d']
 d = d.reshape(-1, 3200)
 llzo_x = np.linspace(dllzo_true.min(), dllzo_true.max(), 1000)
@@ -25,6 +29,8 @@ gs = gridspec.GridSpec(1, 2, figure=fig, wspace=1)
 
 axes = []
 titles = []
+
+print(d.shape)
 
 axes.append(fig.add_subplot(gs[0, 0]))
 y, x = np.histogram(d.mean(-1), bins=int(fp.NBINS * 0.5), density=True)

### 2024-01-11 src/scripts/random_walk.py

@@ -75,11 +75,11 @@ axes[-1].plot(rw_x,
               norm.pdf(rw_x, dinfty_true.mean(), dinfty_true.std()),
               color='#B8B8B8',
               label='$p(\hat{D}^*_{\mathrm{num}})$')
-axes[-1].plot([true.mean(-1).mean() - true.mean(-1).std(),
-               true.mean(-1).mean() + true.mean(-1).std()],
+axes[-1].plot([dinfty_true.mean() - dinfty_true.std(),
+               dinfty_true.mean() + dinfty_true.std()],
               [axes[-1].get_ylim()[1] * 1.1, axes[-1].get_ylim()[1] * 1.1],
               color='#F3ADBC',
-              label='$\sigma[\hat{D}^*]$',
+              label='$\sigma[\hat{D}^*_{\mathrm{num}}]$',
               marker='|')
 axes[-1].set_yticks([0, 10, 20])
 axes[-1].set_xlabel(r"$\hat{D}^*$")
@@ -93,14 +93,14 @@ axes[-1].legend(loc='upper left', bbox_to_anchor=(0.6, 1))
 axes.append(fig.add_subplot(gs[0, 3]))
 y, x = np.histogram(true.var(-1, ddof=1), bins=fp.NBINS, density=True)
 axes[-1].stairs(y, x, fill=True, color='#B3CBE8', label='$p(\hat{\sigma}^2[D^*])$')
-axes[-1].axvline(true.mean(-1).var(ddof=1),
+axes[-1].axvline(dinfty_true.var(ddof=1),
                  ymax=0.85,
                  c='#F3ADBC',
-                 label='$\sigma^2[\hat{D}^*]$',
+                 label=r'$\sigma^2[\hat{D}^*_{\mathrm{num}}]$',
                  ls='-')
 axes[-1].set_xlabel(r"$\sigma^2 [\hat{D}^*]$")
 axes[-1].set_ylabel(r"$p(\sigma^2 [\hat{D}^*])$")
-axes[-1].set_xlim([0, None])
+axes[-1].set_xlim([0, 3.5e-3])
 titles.append("d")
 axes[-1].set_title("all simulations")
 axes[-1].legend(loc='upper left', bbox_to_anchor=(0.5, 1))

### 2024-01-19 src/tex/ms.tex
Complete using HR

--- 2023-11-27 src/tex/ms.tex
@@
 %
-where $\var{\oMSDi}$ are the time-dependent variances of the observed MSD, and $\nind{i}$ is the total number of numerically-independent observed squared-displacements for time-interval $i$.
+where $\var{\oMSDi}$ are the time-dependent variances of the observed MSD, and $\nind{i}$ is the total number of numerically-independent observed squared-displacements for time-interval $i$, accounting for both overlapping sampling and concerted motion (see Eqn.~\ref{equ:der_var}).
 We estimate the variances $\var{\oMSDi}$ using the standard result that the variance of the mean of a sample scales inversely with the number of independent constituent observations.
@@
 A quantitative comparison between the true MSD variance and the single-trajectory estimated MSD variance is made in Fig.~\ref{fig:msd}c: the close numerical agreement confirms that Eqn~\ref{equ:varestMSD} can be used to estimate $\var{\oMSD}$, which can then be used to parameterise the model covariance matrix $\mathbf{\Sigma^\prime}$ via Eqn.~\ref{equ:cvv}.
-%
+
 \begin{figure}
@@
 \end{figure}
-%
 
@@
 the corresponding analytical model covariance matrix, $\mathbf{\Sigma^\prime}$, defined by Eqn.~\ref{equ:cvv} and parametrised using analytical variances $\var{\oMSDi}$ (Fig.~\ref{fig:covariances}b); and the average model covariance matrix obtained by parametrising Eqn.~\ref{equ:cvv} using smoothed variances estimated from individual simulation trajectories, and averaging over the resulting set of all \num{4096} matrices (Fig.~\ref{fig:covariances}c).
-%
+
 \begin{figure}
@@
 
-We next benchmark our method using data from simulations of the lithium-ion solid electrolyte cubic \ce{Li7La3Zr2O12} (c-LLZO).
-We performed a single simulation of \num{1536} atoms (\num{448} Li ions) at \SI{1000}{K} for \SI{1.6}{\nano\second} (full simulation details are provided in the Methods section).
-To generate multiple statistically equivalent trajectories, the resulting simulation data was partitioned into \num{512} effective trajectories, each approximately $\sim\SI{25}{\ps}$ in length, and containing data for \num{56} lithium ions.
+We next benchmark our method using data from simulations of the cubic lithium-ion solid electrolyte \ce{Li7La3Zr2O12} (c-LLZO).
+We performed a single simulation of \num{1536} atoms (\num{448} Li ions) at \SI{700}{K} for \SI{6}{\nano\second} (full simulation details are provided in the Methods section).
+To generate multiple statistically equivalent trajectories, the resulting simulation data was partitioned into \num{480} effective trajectories, each approximately $\sim\SI{100}{\ps}$ in length, and containing data for \num{56} lithium ions.
 We then perform the same approximate Bayesian regression analysis as above on each effective trajectory, excluding the first \SI{10}{ps} of MSD data in each case to remove short-time data corresponding to ballistic and sub-diffusive regimes \cite{he_statistical_2018}.
@@
 We used the DIPPIM polarisable ion force field, as parameterised by Burbano \emph{et al.}~\cite{burbano_sparse_2016}, due to its proven accuracy in accounting for the effect of ion polarisability on diffusion~\cite{wilson_polarization_1993,burbano_sparse_2016}.
-We simulated the cubic phase of LLZO in NVT ensemble at a temperature of \SI{1000}{\kelvin}.
-Simulations were run for \SI{2}{\nano\second} with a \SI{2}{\femto\second} timestep. 
+We simulated the cubic phase of LLZO in NVT ensemble at a temperature of \SI{700}{\kelvin}.
+Simulations were run for \SI{6}{\nano\second} with a \SI{2}{\femto\second} timestep. 
 To control temperature, we used a Nos\'{e}-Hoover thermostat, with a relaxation time of \SI{121}{\femto\second} (5000 $\hbar / E_{h}$)~\cite{nose_unified_1984,hoover_canonical_1985,martyna_nose_1992}.
@@
 \end{equation}
-For a random walk, the second term averages to zero for all $h_i$ and $h_j$, and
+For a random walk, the second term averages to zero for all $h_i$ and $h_j$ (and where the walk has some temporal correlation, this introduces a scaling factor \cite{friauf_correlation_1962} to the total MSD), and
 \begin{equation}
@@
 Hence the expected value for the mean-squared displacement increases linearly with the number of steps taken.
+In the event that the walk has some temporal correlation, as is the case for vacancy mediated diffusion, the second term does not average to zero, making Eqn.~\ref{equ:msd_SI} a proportional relationship~\cite{friauf_correlation_1962}.
+This proportionality would be carried through this derivation but would have no impact on the overall Bayesian regression method as the variances that are estimated from the simulation would include this scaling.
 
@@
 \end{equation}
-where $\nind{n}$ is the total number of statistically independent (non-overlapping) squared displacements that contribute to $\oMSDi$.
-In the long-time limit, $\nind{n}$ is given by the product of the number of mobile particles and the number of numerically-independent sub-trajectories of length $i$ in our simulation trajectory.
+where $\nind{n}$ is the total number of statistically independent (non-overlapping and accounting for concerted motion) squared displacements that contribute to $\oMSDi$.
+In the long-time limit, $\nind{n}$ is given by the product of the number of mobile particles, $N_{\textrm{atoms}}$, the number of numerically-independent sub-trajectories of lenth $i$ in our simulation trajectory, $N_t(t)$, and the ratio of the mean-squared displacement and the mean-square of the displacement of the centre of mass for all mobile particles,
+\begin{equation}
+    H_R(t_{\textrm{diff}}) = \frac{\sum^{N(t_{\textrm{diff}})}_{j=1}{\left[\Delta\mathbf{r}_j(t_{\textrm{diff}})\right]}^2}{\sum^{N_{t}(t_{\textrm{diff}})}_{l=1}\left|\left[\sum^{N_{\textrm{atoms}}}_{k=1}\Delta r_{k,l}(t_{\textrm{diff}})\right]^2\right|^2},
+\end{equation}
+at the start of the diffusive regime, $t_{\textrm{diff}}$.
 Note that $\nind{n}$ considers numerically-independent sub-trajectories, since mutually overlapping time-windows give correlated squared displacements.
 Where overlapping time-windows are used, Eqn.~\ref{equ:der_var} approximates the observed variance (the variance for overlapping and non-overlapping samples are not the same) with accuracy that increases as a function of time-interval length.
+The inclusion of $H_R(t_{\textrm{diff}})$ account for concerted motion between individual particles in the simulation~\cite{vargas_dynamic_2019,canepa_pushing_2022}, which is particularly important in solid and liquid electrolyte systems.
+The selection of $t_{\textrm{diff}}$ as the point at which to use $H_R$ is taken to maximise the number of observations of the centre-of-mass displacement. 
 This leads to the approximation in Eqn.~\ref{equ:varestMSD}.

### 2024-01-19 src/code/llzo/many_runs.py

@@ -1,7 +1,7 @@
 import MDAnalysis as mda
 import numpy as np
 from kinisi.parser import MDAnalysisParser
-from kinisi.diffusion import MSDBootstrap
+from kinisi.diffusion import MSDDiffusion
 
 n = snakemake.params['n']
 start_diff = snakemake.params['start_diff']
@@ -57,7 +57,7 @@ for m in range(0, step, 1):
 
         da_params = {'specie': 'Li', 'time_step': timestep, 'step_skip': step_skip}
         p = MDAnalysisParser(uu, **da_params)
-        b = MSDBootstrap(p.delta_t, p.disp_3d, p._n_o)
+        b = MSDDiffusion(p.delta_t, p.disp_3d, p._n_o)
         b.diffusion(start_diff, random_state=rng)
         # print(b._popt)
         dt[m, i] = b.dt
@@ -68,8 +68,7 @@ for m in range(0, step, 1):
         cov[m, i] = b.covariance_matrix
         d[m, i] = b.D.samples
         g[m, i] = b.gradient.samples
-        f[m, i] = b._f
-        print(b._f)
+        f[m, i] = b._hr
         intercept[m, i] = b.intercept.samples
         
 np.savez(f'src/data/llzo/diffusion_{n}_{start_diff}.npz',

### 2024-01-19 src/code/random_walks/kinisi_pyblock_modelfree_rw.py

@@ -3,7 +3,7 @@
 # author: Andrew R. McCluskey (arm61)
 
 import numpy as np
-from kinisi.diffusion import MSDBootstrap
+from kinisi.diffusion import MSDDiffusion
 from tqdm import tqdm
 from random_walk import get_disp3d, walk
 
@@ -14,8 +14,8 @@ size = int(snakemake.params['n'])
 
 timestep = np.arange(1, length + 1, 1, dtype=int)
 data = np.zeros((size, timestep.size, 4))
-covariance = np.zeros((size, timestep.size - 4, timestep.size - 4))
-npd_covariance = np.zeros((size, timestep.size - 4, timestep.size - 4))
+covariance = np.zeros((size, timestep.size, timestep.size))
+npd_covariance = np.zeros((size, timestep.size, timestep.size))
 n_o = np.zeros((size, timestep.size))
 diff_c = np.zeros((size, 3200))
 intercept = np.zeros((size, 3200))
@@ -25,14 +25,14 @@ for seed in tqdm(range(size)):
     np.random.seed(seed)
     disp_3d, n_samples = get_disp3d(walk, length, atoms, jump_size=jump, seed=rng)
 
-    diff = MSDBootstrap(timestep, disp_3d, n_samples, random_state=rng, progress=False, block=True)
-    diff.diffusion(5, model=False, random_state=rng, progress=False)
+    diff = MSDDiffusion(timestep, disp_3d, n_samples, random_state=rng, progress=False, block=True)
+    diff.diffusion(0, model=False, random_state=rng, progress=False)
     diff_c[seed] = diff.gradient.samples / 6
     intercept[seed] = diff.intercept.samples
     data[seed, :, 0] = diff.dt
     data[seed, :, 1] = diff.n
     data[seed, :, 2] = diff.s
-    data[seed, 4:, 3] = diff._model_v
+    data[seed, :, 3] = diff._model_v
     npd_covariance[seed] = diff._npd_covariance_matrix
     covariance[seed] = diff.covariance_matrix
     n_o[seed] = diff._n_o

### 2024-01-19 src/code/random_walks/kinisi_pyblock_rw.py

@@ -3,7 +3,7 @@
 # author: Andrew R. McCluskey (arm61)
 
 import numpy as np
-from kinisi.diffusion import MSDBootstrap
+from kinisi.diffusion import MSDDiffusion
 from tqdm import tqdm
 from random_walk import get_disp3d, walk
 
@@ -14,8 +14,8 @@ size = int(snakemake.params['n'])
 
 timestep = np.arange(1, length + 1, 1, dtype=int)
 data = np.zeros((size, timestep.size, 4))
-covariance = np.zeros((size, timestep.size - 4, timestep.size - 4))
-npd_covariance = np.zeros((size, timestep.size - 4, timestep.size - 4))
+covariance = np.zeros((size, timestep.size, timestep.size))
+npd_covariance = np.zeros((size, timestep.size, timestep.size))
 n_o = np.zeros((size, timestep.size))
 diff_c = np.zeros((size, 3200))
 intercept = np.zeros((size, 3200))
@@ -25,14 +25,14 @@ for seed in tqdm(range(size)):
     np.random.seed(seed)
     disp_3d, n_samples = get_disp3d(walk, length, atoms, jump_size=jump, seed=rng)
 
-    diff = MSDBootstrap(timestep, disp_3d, n_samples, random_state=rng, progress=False, block=True)
-    diff.diffusion(5, random_state=rng, progress=False)
+    diff = MSDDiffusion(timestep, disp_3d, n_samples, random_state=rng, progress=False, block=True)
+    diff.diffusion(0, random_state=rng, progress=False)
     diff_c[seed] = diff.gradient.samples / 6
     intercept[seed] = diff.intercept.samples
     data[seed, :, 0] = diff.dt
     data[seed, :, 1] = diff.n
     data[seed, :, 2] = diff.s
-    data[seed, 4:, 3] = diff._model_v
+    data[seed, :, 3] = diff._model_v
     npd_covariance[seed] = diff._npd_covariance_matrix
     covariance[seed] = diff.covariance_matrix
     n_o[seed] = diff._n_o

### 2024-01-19 src/code/random_walks/kinisi_rw.py

@@ -3,7 +3,7 @@
 # author: Andrew R. McCluskey (arm61)
 
 import numpy as np
-from kinisi.diffusion import MSDBootstrap
+from kinisi.diffusion import MSDDiffusion
 from tqdm import tqdm
 from random_walk import get_disp3d, walk
 
@@ -16,6 +16,7 @@ timestep = np.arange(1, length + 1, 1, dtype=int)
 data = np.zeros((size, timestep.size, 4))
 covariance = np.zeros((size, timestep.size - 4, timestep.size - 4))
 npd_covariance = np.zeros((size, timestep.size - 4, timestep.size - 4))
+f = np.zeros((size, timestep.size))
 n_o = np.zeros((size, timestep.size))
 diff_c = np.zeros((size, 3200))
 intercept = np.zeros((size, 3200))
@@ -25,7 +26,7 @@ for seed in tqdm(range(size)):
     np.random.seed(seed)
     disp_3d, n_samples = get_disp3d(walk, length, atoms, jump_size=jump, seed=rng)
 
-    diff = MSDBootstrap(timestep, disp_3d, n_samples, random_state=rng, progress=False)
+    diff = MSDDiffusion(timestep, disp_3d, n_samples, random_state=rng, progress=False)
     diff.diffusion(5, random_state=rng, progress=False)
     diff_c[seed] = diff.gradient.samples / 6
     intercept[seed] = diff.intercept.samples
@@ -36,6 +37,8 @@ for seed in tqdm(range(size)):
     npd_covariance[seed] = diff._npd_covariance_matrix
     covariance[seed] = diff.covariance_matrix
     n_o[seed] = diff._n_o
+    f[seed] = diff._hr
+
 
 np.savez(f'src/data/random_walks/kinisi/rw_1_{atoms}_{length}_s{size}.npz',
          diff_c=diff_c,
@@ -43,4 +46,5 @@ np.savez(f'src/data/random_walks/kinisi/rw_1_{atoms}_{length}_s{size}.npz',
          data=data,
          covariance=covariance,
          npd_covariance=npd_covariance,
-         n_o=n_o)
+         n_o=n_o,
+         f=f)

### 2024-01-19 src/scripts/covariances.py

@@ -29,7 +29,7 @@ true_cov = np.cov(true_msd[:, 4:length].T)
 kinisi_data = np.load(paths.data / f"random_walks/kinisi/rw_{jump}_{atoms}_{length}_s4096.npz")
 
 kinisi_cov = kinisi_data["covariance"].mean(0)
-no = kinisi_data["n_o"][0]
+no = (kinisi_data["n_o"] * kinisi_data['f'][:, 4, np.newaxis]).mean(0)
 timestep = np.arange(1, length + 1, 1)
 
 ts_mesh = np.meshgrid(timestep[4:], timestep[4:])
@@ -93,7 +93,7 @@ titles2.append("estimated")
 
 axes.append(fig.add_subplot(gs[0, 1]))
 ratio = (true_cov / anal_cov)
-mod = np.max([1 - np.min(ratio), np.abs(1 - np.max(ratio))])
+mod = 0.5
 im1 = axes[-1].contourf(ts_mesh[0], ts_mesh[1], ratio, cmap=newcmp4, levels=np.linspace(1 - mod, 1 + mod, 15))
 fig.subplots_adjust(right=0.8)
 cb1 = fig.colorbar(im1, fraction=0.046)
@@ -106,7 +106,6 @@ titles.append(r"$\mathrm{a}\div\mathrm{b}$")
 
 axes.append(fig.add_subplot(gs[0, 2]))
 ratio = (true_cov / kinisi_cov)
-mod = np.max([1 - np.min(ratio), np.abs(1 - np.max(ratio))])
 im1 = axes[-1].contourf(ts_mesh[0], ts_mesh[1], ratio, cmap=newcmp4, levels=np.linspace(1 - mod, 1 + mod, 15))
 fig.subplots_adjust(right=0.8)
 cb1 = fig.colorbar(im1, fraction=0.046)
@@ -137,7 +136,7 @@ for i, ax in enumerate(axes):
     print(ax.get_window_extent().x1 - ax.get_window_extent().x0)
     print(ax.get_window_extent().y1 - ax.get_window_extent().y0)
     x = ax.get_window_extent().x0 - x_correction[i]
-    y = ax.get_window_extent().y1 + 11.2
+    y = ax.get_window_extent().y1 + 11.2 
     x, y = fig.transFigure.inverted().transform([x, y])
     fig.text(x, y, titles[i], ha='left')
 for i, ax in enumerate(axes[:3]):

### 2024-01-19 src/scripts/diffusion.py

@@ -56,7 +56,7 @@ axes[-1].legend(loc='upper left', bbox_to_anchor=(0.6, 1))
 axes.append(fig.add_subplot(gs[0, 1]))
 y, x = np.histogram(d.var(-1, ddof=1), bins=int(fp.NBINS * 0.5), density=True)
 axes[-1].stairs(y, x, fill=True, color='#B3CBE8', label=r'$p(\hat{\sigma}^2[D^*])$')
-axes[-1].axvline(d.mean(-1).var(ddof=1), c='#F3ADBC', ls='-', label=r'$\sigma^2(\hat{D}^*)$')
+axes[-1].axvline(d.mean(-1).var(ddof=1), c='#F3ADBC', ls='-', ymax=0.95, label=r'$\sigma^2(\hat{D}^*)$')
 axes[-1].set_xlabel(r"$\hat{\sigma}^2(\hat{D}^*)$ / cm$^4$ s$^{-2}$")
 axes[-1].set_ylabel(r"$p(\hat{\sigma}^2[\hat{D}^*])$ / cm$^{-4}$ s$^2$")
 axes[-1].set_xlim([0, None])

### 2024-01-19 src/scripts/pyblock.py

@@ -59,7 +59,7 @@ axes.append(fig.add_subplot(gs[0, 1]))
 y, x = np.histogram(kinisi.var(-1, ddof=1), bins=rw_x2, density=True)
 axes[-1].stairs(y, x, fill=True, color='#B3CBE8', label='$p(\hat{\sigma}^2[D^*])$')
 axes[-1].axvline(kinisi.mean(-1).var(ddof=1),
-                 ymax=0.85,
+                 ymax=0.95,
                  c='#F3ADBC',
                  label='$\sigma^2[\hat{D}^*]$',
                  ls='-')
@@ -97,7 +97,7 @@ axes.append(fig.add_subplot(gs[1, 1], sharex=axes[-2]))
 y, x = np.histogram(pyblock.var(-1, ddof=1), bins=rw_x2, density=True)
 axes[-1].stairs(y, x, fill=True, color='#B3CBE8', label='$p(\hat{\sigma}^2[D^*])$')
 axes[-1].axvline(pyblock.mean(-1).var(ddof=1),
-                 ymax=0.85,
+                 ymax=0.95,
                  c='#F3ADBC',
                  label='$\sigma^2[\hat{D}^*]$',
                  ls='-')
@@ -135,7 +135,7 @@ axes.append(fig.add_subplot(gs[2, 1], sharex=axes[-2]))
 y, x = np.histogram(pyblock_mf.var(-1, ddof=1), bins=rw_x2, density=True)
 axes[-1].stairs(y, x, fill=True, color='#B3CBE8', label='$p(\hat{\sigma}^2[D^*])$')
 axes[-1].axvline(pyblock_mf.mean(-1).var(ddof=1),
-                 ymax=0.85,
+                 ymax=0.95,
                  c='#F3ADBC',
                  label='$\sigma^2[\hat{D}^*]$',
                  ls='-')

### 2024-01-19 src/scripts/random_walk.py

@@ -94,13 +94,13 @@ axes.append(fig.add_subplot(gs[0, 3]))
 y, x = np.histogram(true.var(-1, ddof=1), bins=fp.NBINS, density=True)
 axes[-1].stairs(y, x, fill=True, color='#B3CBE8', label='$p(\hat{\sigma}^2[D^*])$')
 axes[-1].axvline(dinfty_true.var(ddof=1),
-                 ymax=0.85,
+                 ymax=0.95,
                  c='#F3ADBC',
                  label=r'$\sigma^2[\hat{D}^*_{\mathrm{num}}]$',
                  ls='-')
 axes[-1].set_xlabel(r"$\sigma^2 [\hat{D}^*]$")
 axes[-1].set_ylabel(r"$p(\sigma^2 [\hat{D}^*])$")
-axes[-1].set_xlim([0, 3.5e-3])
+# axes[-1].set_xlim([0, 3.5e-3])
 titles.append("d")
 axes[-1].set_title("all simulations")
 axes[-1].legend(loc='upper left', bbox_to_anchor=(0.5, 1))

### 2024-07-30 src/tex/ms.tex
Latest version

--- 2023-11-27 src/tex/ms.tex
@@
   \email{andrew.mccluskey@bristol.ac.uk}
-  \affiliation{School of Chemistry, University of Bristol, Cantock's Close, Bristol, BS8 1TS, United Kingdom}
+  \affiliation{Centre for Computational Chemistry, School of Chemistry, University of Bristol, Cantock's Close, Bristol, BS8 1TS, UK}
   \affiliation{European Spallation Source ERIC, Ole Maaløes vej 3, 2200 København N, DK}
+  \affiliation{Diamond Light Source, Rutherford Appleton Laboratory, Harwell Science and Innovation Campus, Didcot, OX11 0DE, UK}
 \author{Samuel W. Coles}
@@
     We present a scheme for estimating $\D$ from a single simulation trajectory with high statistical efficiency and accurately estimating the uncertainty in the predicted value.
-    The statistical distribution of MSDs observable from a given simulation is modelled as a multivariate normal distribution using an analytical covariance matrix for an equivalent system of freely diffusing particles, which we parameterise from the available simulation data.
+    The statistical distribution of MSDs observable from a given simulation is modelled as a multivariate normal distribution using an eestimated covariance matrix, which we parameterise from the available simulation data.
     We then perform Bayesian regression to sample the distribution of linear models that are compatible with this model multivariate normal distribution, to obtain a statistically efficient estimate of $D^*$ and an accurate estimate of the associated statistical uncertainty.
@@
 Here, we describe an approximate Bayesian regression method for estimating $\D$ with near-maximal statistical efficiency while also accurately estimating the corresponding statistical uncertainty, using data from a single simulation (Fig.~\ref{fig:schematic}).
-We model the statistical population of simulation MSDs as a multivariate normal distribution, using an analytical covariance matrix derived for an equivalent system of freely diffusing particles, with this covariance matrix parameterised from the observed simulation data.
+We model the statistical population of simulation MSDs as a multivariate normal distribution, using an estimated covariance matrix, parameterised from the observed simulation data.
 We then use Markov-chain Monte Carlo to sample the posterior distribution of linear models compatible with this multivariate normal model.
@@
 
-\subsection{Approximating $\bm{\Sigma}$ from simulation data}
+\subsection{Estimating $\mathbf{\Sigma}$ from simulation data}
 
 For Bayesian regression and GLS, we require the covariance matrix for the observed MSD, $\mathbf{\Sigma}$, which is generally unknown.
-To proceed, we replace $\mathbf{\Sigma}$ with a model covariance matrix, $\mathbf{\Sigma^\prime}$, with a known analytical form, that we parameterise from the available simulation data.
-Providing the correlation structure of $\mathbf{\Sigma^\prime}$ is similar to that of $\mathbf{\Sigma}$, this model correlation matrix can be used in approximate Bayesian or GLS schemes to estimate the ensemble-average MSD, and hence $\D$, with high efficiency and accurate estimated uncertainties.
+To proceed, we replace $\mathbf{\Sigma}$ with an estimated covariance matrix, $\widehat{\mathbf{\Sigma}}$, that is estimated directly from the simulation data.
+Providing the correlation structure of $\widehat{\mathbf{\Sigma}}$ is similar to that of $\mathbf{\Sigma}$, this estimated correlation matrix can be used in approximate Bayesian or GLS schemes to estimate the ensemble-average MSD, and hence $\D$, with high efficiency and accurate estimated uncertainties.
 
-We model the covariance matrix for the observed MSD from a given simulation using the covariance matrix for the MSD of an equivalent system of freely diffusing particles, $\mathbf{\Sigma^\prime}$.
-For observed MSDs computed by averaging over numerically-independent sub-trajectories, the covariance matrix $\mathbf{\Sigma^\prime}$, in the long time limit, has elements (see SI)
-%
-\begin{equation}
-  \Sigma^\prime\left[\oMSDi, \oMSDj\right]= \Sigma^\prime\left[\oMSDj, \oMSDi\right] =
-  \var{\oMSDi} \frac{\nind{i}}{\nind{j}},\hspace{1em} \forall\,i \leq j,
-  \label{equ:cvv}
-\end{equation} 
-%
-where $\var{\oMSDi}$ are the time-dependent variances of the observed MSD, and $\nind{i}$ is the total number of numerically-independent observed squared-displacements for time-interval $i$.
-We estimate the variances $\var{\oMSDi}$ using the standard result that the variance of the mean of a sample scales inversely with the number of independent constituent observations.
-Specifically, we approximate the variance $\varest{\oMSDi}$ (see SI) by rescaling the observed variance of the squared displacement for time interval $i$ by the number of numerically-independent contributing sub-trajectories, $\nind{i}$;
+We estimate the covariance matrix for the observed MSD from a given simulation using the estimates variances. 
+These estimated variances, $\varest{\oMSDi}$, are found using the standard result that ths variance of the mean of a sample scales inversely with the number of independent constituent observations.
+Specifically, we estimate the variance $\varest{\oMSDi}$ by rescaling the observed variance of the squared displacement for time interval $i$ by the number of numerically-independent contributing sub-trajectories, $\nind{i}$;
 %
@@
 %
+From this, the full estimated covariance matrix, can be populated, in the long time limit (see SI) as follows, 
+%
+\begin{equation}
+  \widehat{\Sigma}\left[\oMSDi, \oMSDj\right]= \widehat{\Sigma}\left[\oMSDj, \oMSDi\right] =
+  \varest{\oMSDi} \frac{\nind{i}}{\nind{j}},\hspace{1em} \forall\,i \leq j,
+  \label{equ:cvv}
+\end{equation} 
+%
+where $\nind{i}$ is the total number of numerically-independent observed squared-displacements for a random walk of independent particles in time-interval $i$.
 
-Rescaling by the number of numerically-independent contributing sub-trajectories has the effect of renormalising the variance of the observed squared displacements to account for correlations between particle squared displacements computed from overlapping time windows.
+Rescaling the variance by the number of numerically-independent contributing sub-trajectories has the effect of renormalising the variance of the observed squared displacements to account for correlations between particle squared displacements computed from overlapping time windows.
 An alternative approach to renormalising $\var{x_i}$ is to use a non-parametric block-averaging procedure~\cite{flyvbjerg_error_1989,frenkel_understanding_2023,materzanini_high_2021}, which gives undesirable results for a random walk (see SI).
@@
 
-The estimated variance $\varest{\oMSD}$ can be calculated from a single simulation trajectory, and provides an accurate estimate of the true variance $\var{\oMSD}$.
+The estimated variance $\varest{\oMSD}$ can be determined from a single simulation trajectory, and provides an accurate estimate of the true variance $\var{\oMSD}$.
 To demonstrate this, we performed \num{4096} independent simulations of \num{128} particles undergoing a three-dimensional cubic-lattice random walk of \num{128} steps per particle.
@@
 We also compute the MSD and estimated variance using data from a single simulation trajectory (Fig.~\ref{fig:msd}b), using the scheme described above.
-A quantitative comparison between the true MSD variance and the single-trajectory estimated MSD variance is made in Fig.~\ref{fig:msd}c: the close numerical agreement confirms that Eqn~\ref{equ:varestMSD} can be used to estimate $\var{\oMSD}$, which can then be used to parameterise the model covariance matrix $\mathbf{\Sigma^\prime}$ via Eqn.~\ref{equ:cvv}.
-%
+A quantitative comparison between the true MSD variance and the single-trajectory estimated MSD variance is made in Fig.~\ref{fig:msd}c: the close numerical agreement confirms that Eqn~\ref{equ:varestMSD} can be used to estimate $\var{\oMSD}$, which can then be used to parameterise the estimated covariance matrix $\widehat{\mathbf{\Sigma}}$ via Eqn.~\ref{equ:cvv}.
+
 \begin{figure}
@@
 \end{figure}
-%
 
-The practical implementation of both GLS and Bayesian regression requires that the covariance matrix $\mathbf{\Sigma^\prime}$ is invertible (positive definite); see Eqns.~\ref{equ:gls} and \ref{equ:loglike}.
-The estimated MSD variances derived from simulation data via Eqn.~\ref{equ:varestMSD} are statistically noisy and using these to directly parameterise $\mathbf{\Sigma^\prime}$ can yield non-invertible singular matrices.
-To make our scheme numerically tractable, we therefore fit our estimated MSD variances to the analytical variance for an analogous system of particles undergoing random walks~\cite{smith_random_1996};
-%
-\begin{equation}
-    \var{\oMSDi} = a\frac{t^2_i}{\nind{i}},
-    \label{equ:variance}
-\end{equation}
-%
-where $a$ is a scaling parameter determined by fitting Eqn.~\ref{equ:variance} to the directly estimated MSD variances.
-This smoothing of $\var{\oMSDi}$ guarantees that the resulting model covariance matrix $\mathbf{\Sigma^\prime}$ is invertible and thus suitable for GLS or Bayesian regression.
+The practical implementation of both GLS and Bayesian regression requires that the covariance matrix $\widehat{\mathbf{\Sigma}}$ is invertible (positive definite); see Eqns.~\ref{equ:gls} and \ref{equ:loglike}.
+The estimated covariance matrix however in constructed from a limited number of samples, and therefore may have a high condition number~\cite{higham_restoring_2016}. 
+Statistical noise and high condition numbers result in numerical sensitivity in linear equations, such as inversion. 
+Therefore, to make our scheme numerically tractable, the estimated covariance matrix is reconditioned using the minimum eigenvalue method~\cite{tabeart_improving_2020}. 
+This approach ensures that the condition number for the estimated covariance matrix is equal to a user defined parameter, that can be set to produce an invertible matrix that is suitable for GLS or Bayesian regression.
 
-To illustrate the complete numerical procedure for deriving the model covariance matrix, $\bm{\Sigma^\prime}$, we present in Fig.~\ref{fig:covariances} the MSD covariance matrix for \num{4096} random-walk simulations, as described above, at three differing levels of approximation:
+To illustrate the complete numerical procedure for deriving the estimated covariance matrix, $\widehat{\mathbf{\Sigma}}$, we present in Fig.~\ref{fig:covariances} the MSD covariance matrix for \num{4096} random-walk simulations, as described above, at three differing levels of approximation:
 the numerically converged covariance matrix, $\bm{\Sigma}$, computed using the data from all \num{4096} simulations (Fig.~\ref{fig:covariances}a);
-the corresponding analytical model covariance matrix, $\mathbf{\Sigma^\prime}$, defined by Eqn.~\ref{equ:cvv} and parametrised using analytical variances $\var{\oMSDi}$ (Fig.~\ref{fig:covariances}b); and the average model covariance matrix obtained by parametrising Eqn.~\ref{equ:cvv} using smoothed variances estimated from individual simulation trajectories, and averaging over the resulting set of all \num{4096} matrices (Fig.~\ref{fig:covariances}c).
-%
-\begin{figure}
-    \centering
-    % \resizebox{\columnwidth}{!}{\includegraphics*{figure3.pdf}}
-    \includegraphics[width=\columnwidth]{covariances.pdf}
-    \script{covariances.py}
-    \caption{
-        (a) The numerical MSD covariance matrix $\bm{\Sigma}$ calculated using MSD data from \num{4096} simulations of \num{128} particles undergoing a 3D lattice random walk of \num{128} steps per particle.
-        (b) The analytical MSD covariance matrix $\bm{\Sigma^\prime}$ (Eqn.~\ref{equ:cvv}), parametrised using analytical random-walk variances $\var{\oMSDi}$.
-        (c) The MSD covariance matrix obtained applying the numerical scheme described in the main text to each individual random walk simulation, averaged over all \num{4096} such simulations.
-        Colour bars in (a--c) show the covariance, $\Sigma\left[\oMSDi, \oMSDj\right]$.
-        The off-diagonal panels show difference plots, computed as per-element ratios between pairs of covariance matrices (a--c).
-        %Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences.
-    }
-    \label{fig:covariances}
-\end{figure}
-%
+the corresponding analytical covariance matrix, $\mathbf{\Sigma^\prime}$, derived in the SI (Fig.~\ref{fig:covariances}b); and the average estimated covariance matrix obtained by parametrising Eqn.~\ref{equ:cvv} using smoothed variances estimated from individual simulation trajectories, and averaging over the resulting set of all \num{4096} matrices (Fig.~\ref{fig:covariances}c).
+
+% \begin{figure}
+%     \centering
+%     % \resizebox{\columnwidth}{!}{\includegraphics*{figure3.pdf}}
+%     \includegraphics[width=\columnwidth]{covariances.pdf}
+%     \script{covariances.py}
+%     \caption{
+%         (a) The numerical MSD covariance matrix $\bm{\Sigma}$ calculated using MSD data from \num{4096} simulations of \num{128} particles undergoing a 3D lattice random walk of \num{128} steps per particle.
+%         (b) The analytical MSD covariance matrix $\bm{\Sigma^\prime}$ (Eqn.~\ref{equ:cvv_SI}).
+%         (c) The MSD covariance matrix obtained applying the numerical scheme described in the main text to each individual random walk simulation, averaged over all \num{4096} such simulations.
+%         Colour bars in (a--c) show the covariance, $\Sigma\left[\oMSDi, \oMSDj\right]$.
+%         The off-diagonal panels show scale plots, computed as per-element ratios between pairs of covariance matrices (a--c).
+%         %Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences.
+%     }
+%     \label{fig:covariances}
+% \end{figure}
 
 While the analytical and average estimated covariance matrices show some systematic deviation from the numerically converged covariance matrix, the general correlation structure is preserved.
-The discrepancy between the model and numerical covariance matrices largely stems from the approximation made in deriving the analytical form that $t$ is large, which leads to an overestimation of the variance at low $t$.
+The discrepancy between the estimated and numerical covariance matrices largely stems from the assumption that the analytical covariance structure is equivalent for the estimated covariance matrix. 
 Despite this, the average estimated covariance matrix reproduces well the correlation structure of the true numerical covariance matrix, indicating that the covariance matrices estimated from individual simulation trajectories may be used within approximate GLS or Bayesian regression schemes to estimate $\D$ and $\var{\Dest}$.
@@
 Fig.~\ref{fig:random_walk}a shows the observed MSD from a single 3D-lattice random-walk simulation, along with the estimated posterior distribution of linear models compatible with the observed MSD data, $\prob{\model|\oMSD}$, calculated via Eqns.~\ref{equ:bayes} and \ref{equ:loglike}.
-The corresponding marginal posterior distribution of estimated diffusion coefficients $\prob{\D|\oMSD}$ is shown in Fig.~\ref{fig:random_walk}b; this distribution is approximately Gaussian and is centred on the true self-diffusion coefficient $\D = \num{1}$, demonstrating that for this example trajectory we obtain a good point-estimate of $\D$.
+The corresponding marginal posterior distribution of estimated diffusion coefficients $\prob{\D|\oMSD}$ is shown in Fig.~\ref{fig:random_walk}b; this distribution is approximately Gaussian and close to the true self-diffusion coefficient $\D = \num{1}$, demonstrating that for this example trajectory we obtain a good point-estimate of $\D$.
 
@@
 The distribution of estimated diffusion coefficients from single simulations is slightly broader than the exact numerical results.
-This minor deviation is a consequence of the overestimation of $\varest{\oMSDi}$ at short times, noted above, which results from our use of the long-time limit in the derivation of the analytical model covariance matrix.
+This minor deviation is a consequence of the noise in data collected from a single simulation trajectory.
 
@@
 We also show the true variance of individual point estimates, $\var{\Dest}$, which characterises the spread of the histogram in Fig.~\ref{fig:random_walk}c.
-The distribution of estimated variances is biased relative to the true variance and skewed, due to numerical differences between the true covariance matrix $\mathbf{\Sigma}$ and the model covariance matrix $\mathbf{\Sigma^\prime}$ (further details are provided in the SI).
+The distribution of estimated variances is biased relative to the true variance, due to numerical differences between the true covariance matrix $\mathbf{\Sigma}$ and the estimated covariance matrix $\widehat{\mathbf{\Sigma}}$ (further details are provided in the SI).
 In general, however, the distribution of the estimated variance shows good agreement with the true sample variance.
@@
 
-We next benchmark our method using data from simulations of the lithium-ion solid electrolyte cubic \ce{Li7La3Zr2O12} (c-LLZO).
-We performed a single simulation of \num{1536} atoms (\num{448} Li ions) at \SI{1000}{K} for \SI{1.6}{\nano\second} (full simulation details are provided in the Methods section).
-To generate multiple statistically equivalent trajectories, the resulting simulation data was partitioned into \num{512} effective trajectories, each approximately $\sim\SI{25}{\ps}$ in length, and containing data for \num{56} lithium ions.
+We next benchmark our method using data from simulations of the cubic lithium-ion solid electrolyte \ce{Li7La3Zr2O12} (c-LLZO).
+We performed a single simulation of \num{1536} atoms (\num{448} Li ions) at \SI{700}{K} for \SI{6}{\nano\second} (full simulation details are provided in the Methods section).
+To generate multiple statistically equivalent trajectories, the resulting simulation data was partitioned into \num{192} effective trajectories, each approximately $\sim\SI{500}{\ps}$ in length, and containing data for \num{28} lithium ions, which were selected randomly without replacement.
 We then perform the same approximate Bayesian regression analysis as above on each effective trajectory, excluding the first \SI{10}{ps} of MSD data in each case to remove short-time data corresponding to ballistic and sub-diffusive regimes \cite{he_statistical_2018}.
 
-The resulting distribution of the point estimates, $\Dest$, from analysis of all \num{512} effective trajectories is shown in Fig.~\ref{fig:diffusion}a.
+The resulting distribution of the point estimates, $\Dest$, from analysis of all \num{420} effective trajectories is shown in Fig.~\ref{fig:diffusion}a.
 Again, the corresponding distribution of $\Dest$ estimates derived using Bayesian regression and a well-converged numerical covariance matrix calculated from the full LLZO dataset is also shown for comparison.
-The distribution $\prob{\Dest}$ obtained using the model covariance matrix and parametrised separately for each individual effective simulation is highly similar to that obtained using the aggregate numerical covariance matrix calculated from the complete simulation dataset.
+The distribution $\prob{\Dest}$ obtained using the estimated covariance matrix and parametrised separately for each individual effective simulation is highly similar to that obtained using the aggregate numerical covariance matrix calculated from the complete simulation dataset.
 This close agreement mirrors the results for our random walk simulations (see the SI for a similar comparison of the OLS, WLS, and GLS as shown in Fig.~\ref{fig:glswlsols}), and confirms that our method yields accurate and statistically efficient estimates for $\D$, even for real-world simulation data.
 
-\begin{figure}[tb]
-    \centering
-    % \resizebox{\columnwidth}{!}{\includegraphics*{figure5.pdf}}
-    \includegraphics[width=\columnwidth]{diffusion.pdf}
-    \script{diffusion.py}
-    \caption{
-      (a) Probability distribution of point estimates $\prob{\Dest}$ for \num{512} effective simulations of LLZO (orange histogram).
-      The grey line shows the distribution $\prob{\Dest_\mathrm{num}}$ obtained using Bayesian regression with the complete LLZO dataset as input.
-      The pink bar shows an interval of one standard deviation $\sigma[\prob{\Dest}]$.
-      (b) Probability distribution of estimated variances, $\varest{\Dest}$, for individual LLZO effective simulations, compared to the true sample variance (pink vertical line) $\var{\Dest}$.
-      %Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences. 
-    }
-    \label{fig:diffusion}
-\end{figure}
+% \begin{figure}[tb]
+%     \centering
+%     % \resizebox{\columnwidth}{!}{\includegraphics*{figure5.pdf}}
+%     \includegraphics[width=\columnwidth]{diffusion.pdf}
+%     \script{diffusion.py}
+%     \caption{
+%       (a) Probability distribution of point estimates $\prob{\Dest}$ for \num{480} effective simulations of LLZO (orange histogram).
+%       The grey line shows the distribution $\prob{\Dest_\mathrm{num}}$ obtained using Bayesian regression with the complete LLZO dataset as input.
+%       The pink bar shows an interval of one standard deviation $\sigma[\prob{\Dest}]$.
+%       (b) Probability distribution of estimated variances, $\varest{\Dest}$, for individual LLZO effective simulations, compared to the true sample variance (pink vertical line) $\var{\Dest}$.
+%       %Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences. 
+%     }
+%     \label{fig:diffusion}
+% \end{figure}
 
@@
 While the estimated variances deviate somewhat from the true distribution $\prob{\var{\Dest}}$, the agreement is reasonable and mirrors our results for the random walk simulations.
-Hence, our method provides reasonably accurate estimates of the uncertainty in $\Dest$ for our c-LLZO dataset, even when applied to single effective trajectories with limited displacement data (only 56 mobile ions, and \SI{25}{ps} simulation length).
+Hence, our method provides reasonably accurate estimates of the uncertainty in $\Dest$ for our c-LLZO dataset, even when applied to single effective trajectories with limited displacement data (only 32 mobile ions, and \SI{2}{ns} simulation length).
 
@@
 
-To model the covariance matrix, we use an analytical solution derived for an equivalent system of freely diffusing particles.
-To parameterise this model covariance matrix, we rescale the variance of the observed squared displacements from the input simulation trajectory, followed by a smoothing step to ensure a positive-definite matrix.
-The resulting model covariance matrix preserves the correlation structure of the true simulation MSD covariance matrix, and gives a multivariate normal model for the population of observable simulation MSDs that depends solely on the model parameters, $\D$ and $c$.
+To estimate the covariance matrix, we rescale the variance of the observed squared displacements from the input simulation trajectory and use these to populate then full covariance matrix based on the relationship between the variances and covariance matrix for an equivalent system of freely diffusing particles.
+The resulting estimated covariance matrix preserves the correlation structure of the true simulation MSD covariance matrix, and gives a multivariate normal model for the population of observable simulation MSDs that depends solely on the model parameters, $\D$ and $c$.
 
@@
 In both cases, we obtain a distribution of estimates for $\D$ that closely matches the theoretically optimal distribution obtained using a numerical covariance matrix derived from a large number of replica simulation trajectories.
+The assumption of freely diffusing particles means may overlook short-time interval effects that are present in real systems, such as correlations between different atoms.
+Such correlations would reduce the value of $nind{i}$, however, the impact of this is anticipated to be minimal and potentially limited to the ballistic regime.
 
@@
 We have implemented the approximate Bayesian estimation method described in the main text in the open-source Python package \textsc{kinisi}~\cite{mccluskey_kinisi_2022}, under the MIT licence.
+Throughout this work, kinisi-[INSERT VERSION NUMBER] was used for all analyses.
 
@@
 The covariance matrix is only constructed for values of $t$ where the particle motion is considered to be in the long-time diffusive limit, with this threshold set by the user to a value appropriate for their system and simulation data.
-For the examples presented in the main manuscript, we consider particles to be in the diffusive regime from $t=\num{4}$ for the random walk trajectories and from $t=\SI{10}{\pico\second}$ for the LLZO simulations.
+For the examples presented in the main manuscript, we consider particles to be in the diffusive regime from $t=\num{2}$ for the random walk trajectories and from $t=\SI{10}{\pico\second}$ for the LLZO simulations.
+The covariance matrix is reconditioned using the minimum eigenvalue method~\cite{tabeart_improving_2020} for all simulations, the maximum condition number was \num{1e16}.
 
@@
 We used the DIPPIM polarisable ion force field, as parameterised by Burbano \emph{et al.}~\cite{burbano_sparse_2016}, due to its proven accuracy in accounting for the effect of ion polarisability on diffusion~\cite{wilson_polarization_1993,burbano_sparse_2016}.
-We simulated the cubic phase of LLZO in NVT ensemble at a temperature of \SI{1000}{\kelvin}.
-Simulations were run for \SI{2}{\nano\second} with a \SI{2}{\femto\second} timestep. 
+We simulated the cubic phase of LLZO in NVT ensemble at a temperature of \SI{700}{\kelvin}.
+Simulations were run for \SI{6}{\nano\second} with a \SI{0.5}{\femto\second} timestep. 
 To control temperature, we used a Nos\'{e}-Hoover thermostat, with a relaxation time of \SI{121}{\femto\second} (5000 $\hbar / E_{h}$)~\cite{nose_unified_1984,hoover_canonical_1985,martyna_nose_1992}.
@@
 \begin{equation}
-  \Sigma^\prime\left[\oMSDi, \oMSDj\right]= \Sigma^\prime\left[\oMSDj, \oMSDi\right] =
-  \var{\oMSDi} \frac{\nind{i}}{\nind{j}},\hspace{1em} \forall\,i \leq j,
+  \widehat{\Sigma}\left[\oMSDi, \oMSDj\right]= \widehat{\Sigma}\left[\oMSDj, \oMSDi\right] =
+  \varest{\oMSDi} \frac{\nind{i}}{\nind{j}},\hspace{1em} \forall\,i \leq j,
   \label{equ:cvv_SI}
@@
 To derive this result, we first present a derivation of the expected variance for the MSD at timestep $i$, $\var{\oMSD}$, following the approach of Smith and Gillan~\cite{smith_random_1996}.
-We then derive an expression for the covariance $\Sigma^\prime\left[\oMSDi, \oMSDj\right]$ to obtain the result above.
+We then derive an expression for the covariance $\widehat{\Sigma}\left[\oMSDi, \oMSDj\right]$ to obtain the result above.
 
@@
     \var{\oMSDn}_d = \sum^d{\frac{2n^2\kappa^4}{d^2\nind{n}}} = \frac{2n^2\kappa^4}{d\nind{n}},
+    \label{equ:analvar}
 \end{equation}
@@
 Eqn.~\ref{equ:varestMSD} enables the approximate estimation of the variance in $\oMSDi$, as mentioned in the main text, it is also possible to obtain this from a block averaging approach~\cite{flyvbjerg_error_1989,frenkel_understanding_2023}.
-Fig.~\ref{fig:pyblock} compares the $\D$ estimate distributions using variances from approximated by Eqn.~\ref{equ:varestMSD} with those from the block averaging approach (using the \textsc{pyblock} Python package~\cite{spencer_pyblock_2020}) both with the fitting of the variances to Eqn.~\ref{equ:variance} and without.
-When the fitting is not performed, the resulting covariance matrix is numerically unstable, leading to estimated values of $\D$ at extreme values.
-While when the fitting is performed and the blocking-estimated variances as used, there is no improvement in the estimation of $\D$ and the resulting distribution of estimates in the variance of $\prob{\D}$ is broader than when Eqn.~\ref{equ:varestMSD} is used.
+Fig.~\ref{fig:pyblock} compares the $\D$ estimate distributions using variances from approximated by Eqn.~\ref{equ:varestMSD} with those from the block averaging approach (using the \textsc{pyblock} Python package~\cite{spencer_pyblock_2020}).
+The estimates of $\D$ are similar between the two approaches, however with pyblocking approach typically produces covariance matrices that cannot be conditioned easily, leading to numerical instabilities -- leading to a large standard deviation.
+The more significant difference is in the estimated variances, where the block averaging approach consistently underestimates the variance in $\Dest$. 
+Generally speaking, any approach to estimate the variance of a quantity should prefer to overestimate the variance, such that undue confidence is not placed in the results.
 
@@
     \script{pyblock.py}
-    \caption{Comparison of the distribution of $\Dest$ (a, c, \& e) and the estimated variances (b, d, \& f) from 4096 individual random walk simulations using variances approximated by Eqn.~\ref{equ:varestMSD} (a \& b), from block averaging with fitting to Eqn.~\ref{equ:variance} (c \& d), and without fitting (e \& f).
-Note that the bounds in a, c, \& e are defined by the extrema of the estimated values, showing the fact that without the fitting to Eqn.~\ref{equ:variance}, the resulting covariance matrix is numerically unstable.}
+    \caption{Comparison of the distribution of $\Dest$ (a \& c) and the estimated variances (b \& d) from 4096 individual random walk simulations using variances approximated by Eqn.~\ref{equ:varestMSD} (a \& b), from block averaging (c \& d).}
     \label{fig:pyblock}
@@
 Once again, GLS (or, indeed, the Bayesian regression equivalent, shown in Fig.~\ref{fig:diffusion}) offers both the most statistically efficient estimation of $\D$ and is required to obtain an accruate estimate of the statistical uncertainty in $\D$ from a single simulation.
-\begin{figure}
-    \centering
-    % \resizebox{\columnwidth}{!}{\includegraphics*{figure1.pdf}}
-    \includegraphics[width=\columnwidth]{glswlsols_llzo.pdf}
-    \script{glswlsols_llzo.py}
-    \caption{
-        Example distributions of estimated self-diffusion coefficients, $\Dest$, calculated using (a) ordinary least squares (OLS), (b) weighted least squares (WLS), and (c) generalised least squares (GLS),
-        from MSD data from \num{512} effective simulations of LLZO of \SI{\sim 25}{\pico\second} with 56 lithium ions.
-        In each panel, the grey curve shows the best-fit normal distribution for the simulation data, the upper horizontal bar shows the standard deviation of this distribution, and the lower horizontal bar shows the average estimated standard distribution given by the analytical expression for $\sigma[\prob{\Dest}]$ for each regression method.}
-    \label{fig:glswlsols_llzo}
-\end{figure}
+% \begin{figure}
+%     \centering
+%     % \resizebox{\columnwidth}{!}{\includegraphics*{figure1.pdf}}
+%     \includegraphics[width=\columnwidth]{glswlsols_llzo.pdf}
+%     \script{glswlsols_llzo.py}
+%     \caption{
+%         Example distributions of estimated self-diffusion coefficients, $\Dest$, calculated using (a) ordinary least squares (OLS), (b) weighted least squares (WLS), and (c) generalised least squares (GLS),
+%         from MSD data from \num{512} effective simulations of LLZO of \SI{\sim 25}{\pico\second} with 56 lithium ions.
+%         In each panel, the grey curve shows the best-fit normal distribution for the simulation data, the upper horizontal bar shows the standard deviation of this distribution, and the lower horizontal bar shows the average estimated standard distribution given by the analytical expression for $\sigma[\prob{\Dest}]$ for each regression method.}
+%     \label{fig:glswlsols_llzo}
+% \end{figure}
 

### 2024-07-30 src/code/llzo/glswlsols.py

@@ -13,15 +13,15 @@ from kinisi.diffusion import _straight_line
 np.random.seed(1)
 
 start_diff = snakemake.params['start_diff']
+length = snakemake.params['length']
 
-timestep = np.load(f'src/data/llzo/diffusion_0_{start_diff}.npz')['dt'][0, 0]
-length = 2000
+timestep = np.load(f'src/data/llzo/diffusion_0_{start_diff}_{length}.npz')['dt'][0, 0]
 ll = len([i + length for i in range(0, 20001 - length, length)])
 # length = 500
 # ll = len([i + length for i in range(0, 2000, length)])
-d = np.zeros((6, 8, ll, timestep.size))
-for i in range(0, 6, 1):
-    d[i] = np.load(f'src/data/llzo/diffusion_{i}_{start_diff}.npz')['msd_true']
+d = np.zeros((5, 16, ll, timestep.size))
+for i in range(1, 6, 1):
+    d[i-1] = np.load(f'src/data/llzo/diffusion_{i}_{start_diff}_{length}.npz')['msd_true']
 
 max_ngp = np.argwhere(timestep > start_diff)[0][0]
 true_msd = d.reshape(-1, d.shape[-1])[:, max_ngp:]
@@ -45,7 +45,7 @@ c3 = []
 for i in true_msd:
     c3.append(linregress(timestep, i).stderr / 6e4)
 
-np.savez(f'src/data/llzo/glswlsols_{start_diff}.npz',
+np.savez(f'src/data/llzo/glswlsols_{start_diff}_{length}.npz',
          gls_pop=g1,
          wls_pop=g2,
          ols_pop=g3,

### 2024-07-30 src/code/llzo/many_runs.py

@@ -1,12 +1,20 @@
 import MDAnalysis as mda
 import numpy as np
 from kinisi.parser import MDAnalysisParser
-from kinisi.diffusion import MSDDiffusion
+from kinisi.diffusion import MSDBootstrap
 
 n = snakemake.params['n']
 start_diff = snakemake.params['start_diff']
+length = snakemake.params['length']
 
-step = 8
+step = 16
+n_atoms_drift = int((1536-448) / step)
+n_atoms_li = int(448 / step)
+rng = np.random.RandomState(n*42)
+np.random.seed(n*42)
+choice_drift = rng.choice(np.arange(0, 1536-448, dtype=int), size=(step, n_atoms_drift), replace=False)
+choice_li = rng.choice(np.arange(1536-448, 1536, dtype=int), size=(step, n_atoms_li), replace=False)
+choice = np.concatenate([choice_drift, choice_li], axis=-1)
 
 def universe_slice(data: str, slice_start: int, slice_end: int, m: int) -> mda.Universe:
     """
@@ -18,18 +26,17 @@ def universe_slice(data: str, slice_start: int, slice_end: int, m: int) -> mda.U
     :return: A single flattened Universe
     """
     data_reshape = np.copy(data).reshape(-1, 1536, 3)
-    data_subset = data_reshape[slice_start:slice_end, m::step]
+    data_subset = data_reshape[slice_start:slice_end, choice_li[m]]
     data_subset *= 0.52917721067
     u = mda.Universe.empty(data_subset.shape[1], 1, n_frames=data_subset.shape[0], trajectory=True)
-    list_type = ['A'] * (1536-448) + ['Li'] * 448
-    u.add_TopologyAttr('type', list_type[m::step])
+    list_type = np.array(['Li'] * n_atoms_li)
+    u.add_TopologyAttr('type', list(list_type))
     for i, t in enumerate(u.trajectory):
         t.positions = data_subset[i]
         t.dimensions = np.array([26.30938114, 26.30938114, 26.30938114, 90, 90, 90])
     return u
 
 data = np.loadtxt(f'src/data/llzo/traj{n}.out', usecols=[0, 1, 2])
-length = 2000
 uu = universe_slice(data, 0, length, 0)
 step_skip = 100  # sampling rate
 timestep = 5.079648e-4  # ps
@@ -45,21 +52,16 @@ n_o = np.zeros((step, ll, p.delta_t.size))
[…]

### 2024-07-30 src/code/random_walks/efficiency.py

@@ -5,7 +5,7 @@ from uravu.relationship import Relationship
 
 jump = 1
 
-xaxis = np.array([16, 32, 64, 128, 256, 512, 1024])
+xaxis = np.array([16, 32])#, 64, 128, 256, 512, 1024])
 
 kinisi_atoms = np.zeros((xaxis.size, 512, 3200))
 weighted_atoms = np.zeros((xaxis.size, 512, 3200))

### 2024-07-30 src/code/random_walks/glswlsols.py

@@ -16,9 +16,9 @@ atoms = int(snakemake.params['atoms'])
 length = int(snakemake.params['length'])
 jump = int(snakemake.params['jump'])
 
-true_msd = np.load(f'src/data/random_walks/numerical/rw_1_{atoms}_{length}_s4096.npz')['data'][:, 4:]
+true_msd = np.load(f'src/data/random_walks/numerical/rw_1_{atoms}_{length}_s4096.npz')['data'][:, 1:]
 true_cov = np.cov(true_msd.T)
-timestep = np.arange(4, length, 1)
+timestep = np.arange(1, length, 1)
 
 y = true_msd.T
 A = np.array([timestep, np.ones(timestep.size)]).T

### 2024-07-30 src/code/random_walks/kinisi_pyblock_rw.py

@@ -3,7 +3,7 @@
 # author: Andrew R. McCluskey (arm61)
 
 import numpy as np
-from kinisi.diffusion import MSDDiffusion
+from kinisi.diffusion import MSDBootstrap
 from tqdm import tqdm
 from random_walk import get_disp3d, walk
 
@@ -14,8 +14,7 @@ size = int(snakemake.params['n'])
 
 timestep = np.arange(1, length + 1, 1, dtype=int)
 data = np.zeros((size, timestep.size, 4))
-covariance = np.zeros((size, timestep.size, timestep.size))
-npd_covariance = np.zeros((size, timestep.size, timestep.size))
+covariance = np.zeros((size, timestep.size-1, timestep.size-1))
 n_o = np.zeros((size, timestep.size))
 diff_c = np.zeros((size, 3200))
 intercept = np.zeros((size, 3200))
@@ -25,15 +24,14 @@ for seed in tqdm(range(size)):
     np.random.seed(seed)
     disp_3d, n_samples = get_disp3d(walk, length, atoms, jump_size=jump, seed=rng)
 
-    diff = MSDDiffusion(timestep, disp_3d, n_samples, random_state=rng, progress=False, block=True)
-    diff.diffusion(0, random_state=rng, progress=False)
+    diff = MSDBootstrap(timestep, disp_3d, n_samples, random_state=rng, progress=False, block=True)
+    diff.diffusion(2, random_state=rng, progress=False, cond_max=1e16)
     diff_c[seed] = diff.gradient.samples / 6
     intercept[seed] = diff.intercept.samples
     data[seed, :, 0] = diff.dt
     data[seed, :, 1] = diff.n
     data[seed, :, 2] = diff.s
-    data[seed, :, 3] = diff._model_v
-    npd_covariance[seed] = diff._npd_covariance_matrix
+    data[seed, 1:, 3] = diff._model_v
     covariance[seed] = diff.covariance_matrix
     n_o[seed] = diff._n_o
 
@@ -42,5 +40,4 @@ np.savez(f'src/data/random_walks/pyblock/rw_1_{atoms}_{length}_s{size}.npz',
          intercept=intercept,
          data=data,
          covariance=covariance,
-         npd_covariance=npd_covariance,
          n_o=n_o)

### 2024-07-30 src/scripts/covariances.py

@@ -24,15 +24,15 @@ newcmp3 = sns.light_palette(colors[2], as_cmap=True)
 newcmp4 = sns.diverging_palette(311.4, 161.2, as_cmap=True)
 
 true_msd = np.load(paths.data / f"random_walks/numerical/rw_{jump}_{atoms}_{length}_s4096.npz")["data"]
-true_cov = np.cov(true_msd[:, 4:length].T)
+true_cov = np.cov(true_msd[:, 1:length].T)
 
-kinisi_data = np.load(paths.data / f"random_walks/kinisi/rw_{jump}_{atoms}_{length}_s4096.npz")
+kinisi_data = np.load(paths.data / f"random_walks/kinisi/rw_{jump}_{atoms}_{length}_s512.npz")
 
 kinisi_cov = kinisi_data["covariance"].mean(0)
-no = (kinisi_data["n_o"] * kinisi_data['f'][:, 4, np.newaxis]).mean(0)
+no = (kinisi_data["n_o"]).mean(0)# * kinisi_data['f'][:, 4, np.newaxis]).mean(0)
 timestep = np.arange(1, length + 1, 1)
 
-ts_mesh = np.meshgrid(timestep[4:], timestep[4:])
+ts_mesh = np.meshgrid(timestep[1:], timestep[1:])
 
 anal_cov = np.zeros((length, length))
 for i in range(0, anal_cov.shape[0]):
@@ -40,7 +40,7 @@ for i in range(0, anal_cov.shape[0]):
         value = 8 * dimensionality**2 * D**2 * timestep[i]**2 / (dimensionality * no[j])
         anal_cov[i, j] = value
         anal_cov[j, i] = np.copy(anal_cov[i, j])
-anal_cov = anal_cov[4:, 4:]
+anal_cov = anal_cov[1:, 1:]
 
 max_cov = np.ceil(np.max([anal_cov.max(), true_cov.max(), kinisi_cov.max()]))
 min_cov = np.floor(np.min([anal_cov.min(), true_cov.min(), kinisi_cov.min()]))

### 2024-07-30 src/scripts/pyblock.py

@@ -12,32 +12,22 @@ atoms = 128
 length = 128
 
 kinisi = np.load(paths.data / f"random_walks/kinisi/rw_{jump}_{atoms}_{length}_s4096.npz")['diff_c']
-pyblock = np.load(paths.data / f"random_walks/pyblock/rw_{jump}_{atoms}_{length}_s4096.npz")['diff_c']
-pyblock_mf = np.load(paths.data / f"random_walks/pyblock_modelfree/rw_{jump}_{atoms}_{length}_s4096.npz")['diff_c']
+pyblock_mf = np.load(paths.data / f"random_walks/pyblock/rw_{jump}_{atoms}_{length}_s4096.npz")['diff_c']
 dinfty_true = np.load(paths.data / f"random_walks/numerical/D_1_128_128.npz")['diff_c']
 
-rw_x = np.arange(kinisi.mean(-1).min(), kinisi.mean(-1).max(), 0.01)
-rw_y = np.arange(pyblock.mean(-1).min(), pyblock.mean(-1).max(), 0.01)
-rw_z = np.arange(pyblock_mf.mean(-1).min(), pyblock_mf.mean(-1).max(), 0.01)
-rw_x2 = np.arange(np.array([kinisi.var(-1, ddof=1), pyblock.var(-1, ddof=1), pyblock_mf.var(-1, ddof=1)]).min(), 
-                  np.array([kinisi.var(-1, ddof=1), pyblock.var(-1, ddof=1), pyblock_mf.var(-1, ddof=1)]).max(), 0.0001)
-rw_x3 = np.arange(kinisi.mean(-1).min(), kinisi.mean(-1).max(), 0.0001)
-rw_y3 = np.arange(pyblock.mean(-1).min(), pyblock.mean(-1).max(), 0.0001)
-rw_z3 = np.arange(pyblock_mf.mean(-1).min(), pyblock_mf.mean(-1).max(), 0.0001)
-
-figsize = (4.03, 6)
+figsize = (4.03, 4)
 print(figsize)
 fig = plt.figure(figsize=figsize)
-gs = gridspec.GridSpec(3, 2, figure=fig, wspace=0.6, hspace=1)
+gs = gridspec.GridSpec(2, 2, figure=fig, wspace=0.6, hspace=1)
 
 axes = []
 titles = []
 
 axes.append(fig.add_subplot(gs[0, 0]))
-y, x = np.histogram(kinisi.mean(-1), bins=rw_x, density=True)
[…]

### 2024-07-30 src/scripts/random_walk.py

@@ -57,18 +57,18 @@ axes[-1].stairs(y, x, fill=True, color='#9FD5C3', label=r'$p(D^* | \mathbfit{x})
 axes[-1].plot([
     kinisi['diff_c'][0].mean() - kinisi['diff_c'][0].std(),
     kinisi['diff_c'][0].mean() + kinisi['diff_c'][0].std()
-], [axes[-1].get_ylim()[1] * 1.1, axes[-1].get_ylim()[1] * 1.1],
+], [33, 33],
               color='#81A8D9',
               label='$\hat{\sigma}[\hat{D}^*]$',
               marker='|')
-axes[-1].axvline(kinisi['diff_c'][0].mean(), ymax=0.85, color='#E08D6D', label=r'$\hat{D}^*$', ls='--')
-axes[-1].set_yticks([0, 5, 10, 15])
+axes[-1].axvline(kinisi['diff_c'][0].mean(), ymax=0.95, color='#E08D6D', label=r'$\hat{D}^*$', ls='--')
+axes[-1].set_yticks([0, 15, 30])
 axes[-1].set_xlabel(r"$D^* | \mathbfit{x}$")
 axes[-1].set_ylabel(r"$p(D^* | \mathbfit{x})$")
-axes[-1].set_xlim(0.87, 1.13)
+axes[-1].set_xlim(0.89, 1.11)
 axes[-1].legend(loc='upper left', bbox_to_anchor=(0.65, 1))
 
-axes.append(fig.add_subplot(gs[0, 2]))
+axes.append(fig.add_subplot(gs[0, 2], sharex=axes[-1], sharey=axes[-1]))
 y, x = np.histogram(true.mean(-1), bins=fp.NBINS, density=True)
 axes[-1].stairs(y, x, fill=True, color='#ECBCA7', label='$p(\hat{D}^*)$')
 axes[-1].plot(rw_x,
@@ -77,30 +77,31 @@ axes[-1].plot(rw_x,
               label='$p(\hat{D}^*_{\mathrm{num}})$')
 axes[-1].plot([dinfty_true.mean() - dinfty_true.std(),
                dinfty_true.mean() + dinfty_true.std()],
-              [axes[-1].get_ylim()[1] * 1.1, axes[-1].get_ylim()[1] * 1.1],
+              [33, 33],
               color='#F3ADBC',
               label='$\sigma[\hat{D}^*_{\mathrm{num}}]$',
[…]

### 2024-09-12 src/tex/ms.tex
Latest version

@@
     \caption{
-        A schematic diagram of the Bayesian regression method described in this work, running from our simulations through the sampling of displacement, variance and covariance estimation and final the sampling process to give the marginal posterior distribution for $\D$.}
+        A schematic diagram of the Bayesian regression method described in this work, from running simulations through the sampling of displacement, variance and covariance estimation and final the sampling process to give the marginal posterior distribution for $\D$.}
     \label{fig:schematic}
@@
     \centering
-    % \resizebox{\columnwidth}{!}{\includegraphics*{figure1.pdf}}
+    % \resizebox{\columnwidth}{!}{\includegraphics*{figure2.pdf}}
     \includegraphics[width=\columnwidth]{glswlsols.pdf}
@@
     \centering
-    % \resizebox{\columnwidth}{!}{\includegraphics*{figure2.pdf}}
+    % \resizebox{\columnwidth}{!}{\includegraphics*{figure3.pdf}}
     \includegraphics[width=\columnwidth]{msd.pdf}
@@
 the numerically converged covariance matrix, $\bm{\Sigma}$, computed using the data from all \num{4096} simulations (Fig.~\ref{fig:covariances}a);
-the corresponding analytical covariance matrix, $\mathbf{\Sigma^\prime}$, derived in the SI (Fig.~\ref{fig:covariances}b); and the average estimated covariance matrix obtained by parametrising Eqn.~\ref{equ:cvv} using smoothed variances estimated from individual simulation trajectories, and averaging over the resulting set of all \num{4096} matrices (Fig.~\ref{fig:covariances}c).
+the corresponding analytical covariance matrix, $\mathbf{\Sigma^\prime}$, derived in the SI (Fig.~\ref{fig:covariances}b); and the average estimated covariance matrix obtained by parametrising Eqn.~\ref{equ:cvv} using calculated variances estimated from individual simulation trajectories, and averaging over the resulting set of all \num{4096} matrices (Fig.~\ref{fig:covariances}c).
 
-% \begin{figure}
-%     \centering
-%     % \resizebox{\columnwidth}{!}{\includegraphics*{figure3.pdf}}
-%     \includegraphics[width=\columnwidth]{covariances.pdf}
-%     \script{covariances.py}
-%     \caption{
-%         (a) The numerical MSD covariance matrix $\bm{\Sigma}$ calculated using MSD data from \num{4096} simulations of \num{128} particles undergoing a 3D lattice random walk of \num{128} steps per particle.
-%         (b) The analytical MSD covariance matrix $\bm{\Sigma^\prime}$ (Eqn.~\ref{equ:cvv_SI}).
-%         (c) The MSD covariance matrix obtained applying the numerical scheme described in the main text to each individual random walk simulation, averaged over all \num{4096} such simulations.
-%         Colour bars in (a--c) show the covariance, $\Sigma\left[\oMSDi, \oMSDj\right]$.
-%         The off-diagonal panels show scale plots, computed as per-element ratios between pairs of covariance matrices (a--c).
-%         %Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences.
-%     }
-%     \label{fig:covariances}
-% \end{figure}
+\begin{figure}
+    \centering
+    % \resizebox{\columnwidth}{!}{\includegraphics*{figure4.pdf}}
+    \includegraphics[width=\columnwidth]{covariances.pdf}
+    \script{covariances.py}
+    \caption{
+        (a) The numerical MSD covariance matrix $\bm{\Sigma}$ calculated using MSD data from \num{4096} simulations of \num{128} particles undergoing a 3D lattice random walk of \num{128} steps per particle.
+        (b) The analytical MSD covariance matrix $\bm{\Sigma^\prime}$ (Eqn.~\ref{equ:cvv_SI}).
+        (c) The MSD covariance matrix obtained applying the numerical scheme described in the main text to each individual random walk simulation, averaged over all \num{4096} such simulations.
+        Colour bars in (a--c) show the covariance, $\Sigma\left[\oMSDi, \oMSDj\right]$.
+        The off-diagonal panels show scale plots, computed as per-element ratios between pairs of covariance matrices (a--c).
+        %Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences.
+    }
+    \label{fig:covariances}
+\end{figure}
 
@@
     \centering
-    % \resizebox{\textwidth}{!}{\includegraphics*{figure4.pdf}}
+    % \resizebox{\textwidth}{!}{\includegraphics*{figure5.pdf}}
     \includegraphics[width=\textwidth]{random_walk.pdf}
@@
 
-% \begin{figure}[tb]
-%     \centering
-%     % \resizebox{\columnwidth}{!}{\includegraphics*{figure5.pdf}}
-%     \includegraphics[width=\columnwidth]{diffusion.pdf}
-%     \script{diffusion.py}
-%     \caption{
-%       (a) Probability distribution of point estimates $\prob{\Dest}$ for \num{480} effective simulations of LLZO (orange histogram).
-%       The grey line shows the distribution $\prob{\Dest_\mathrm{num}}$ obtained using Bayesian regression with the complete LLZO dataset as input.
-%       The pink bar shows an interval of one standard deviation $\sigma[\prob{\Dest}]$.
-%       (b) Probability distribution of estimated variances, $\varest{\Dest}$, for individual LLZO effective simulations, compared to the true sample variance (pink vertical line) $\var{\Dest}$.
-%       %Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences. 
-%     }
-%     \label{fig:diffusion}
-% \end{figure}
+\begin{figure}[tb]
+    \centering
+    % \resizebox{\columnwidth}{!}{\includegraphics*{figure6.pdf}}
+    \includegraphics[width=\columnwidth]{diffusion.pdf}
+    \script{diffusion.py}
+    \caption{
+      (a) Probability distribution of point estimates $\prob{\Dest}$ for \num{480} effective simulations of LLZO (orange histogram).
+      The grey line shows the distribution $\prob{\Dest_\mathrm{num}}$ obtained using Bayesian regression with the complete LLZO dataset as input.
+      The pink bar shows an interval of one standard deviation $\sigma[\prob{\Dest}]$.
+      (b) Probability distribution of estimated variances, $\varest{\Dest}$, for individual LLZO effective simulations, compared to the true sample variance (pink vertical line) $\var{\Dest}$.
+      %Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences. 
+    }
+    \label{fig:diffusion}
+\end{figure}
 
@@
     \centering
-    % \resizebox{\columnwidth}{!}{\includegraphics*{figure6.pdf}}
+    % \resizebox{\columnwidth}{!}{\includegraphics*{figure7.pdf}}
     \includegraphics[width=\columnwidth]{stat_eff.pdf}
@@
 We have implemented the approximate Bayesian estimation method described in the main text in the open-source Python package \textsc{kinisi}~\cite{mccluskey_kinisi_2022}, under the MIT licence.
-Throughout this work, kinisi-[INSERT VERSION NUMBER] was used for all analyses.
+Throughout this work, \textsc{kinisi}-1.1.0 was used for all analyses.
 
@@
     \centering
-    % \resizebox{\columnwidth}{!}{\includegraphics*{figure1.pdf}}
+    % \resizebox{\columnwidth}{!}{\includegraphics*{figure8.pdf}}
     \includegraphics[width=\columnwidth]{pyblock.pdf}
@@
     \label{fig:pyblock}
+\end{figure}
+
+\begin{figure}
+    \centering
+    % \resizebox{\columnwidth}{!}{\includegraphics*{figure8.pdf}}
+    \includegraphics[width=\columnwidth]{msd_blocking.pdf}
+    \script{msd_blocking.py}
+    \caption{Comparison of the distribution of $\Dest$ (a \& c) and the estimated variances (b \& d) from 4096 individual random walk simulations using variances approximated by Eqn.~\ref{equ:varestMSD} (a \& b), from block averaging (c \& d).}
+    \label{fig:msd_blocking}
 \end{figure}
@@
     \centering
-    % \resizebox{\columnwidth}{!}{\includegraphics*{figure7.pdf}}
+    % \resizebox{\columnwidth}{!}{\includegraphics*{figure9.pdf}}
     \includegraphics[width=\columnwidth]{true_cov.pdf}
@@
 Once again, GLS (or, indeed, the Bayesian regression equivalent, shown in Fig.~\ref{fig:diffusion}) offers both the most statistically efficient estimation of $\D$ and is required to obtain an accruate estimate of the statistical uncertainty in $\D$ from a single simulation.
-% \begin{figure}
-%     \centering
-%     % \resizebox{\columnwidth}{!}{\includegraphics*{figure1.pdf}}
-%     \includegraphics[width=\columnwidth]{glswlsols_llzo.pdf}
-%     \script{glswlsols_llzo.py}
-%     \caption{
-%         Example distributions of estimated self-diffusion coefficients, $\Dest$, calculated using (a) ordinary least squares (OLS), (b) weighted least squares (WLS), and (c) generalised least squares (GLS),
-%         from MSD data from \num{512} effective simulations of LLZO of \SI{\sim 25}{\pico\second} with 56 lithium ions.
-%         In each panel, the grey curve shows the best-fit normal distribution for the simulation data, the upper horizontal bar shows the standard deviation of this distribution, and the lower horizontal bar shows the average estimated standard distribution given by the analytical expression for $\sigma[\prob{\Dest}]$ for each regression method.}
-%     \label{fig:glswlsols_llzo}
-% \end{figure}
+\begin{figure}
+    \centering
+    % \resizebox{\columnwidth}{!}{\includegraphics*{figure10.pdf}}
+    \includegraphics[width=\columnwidth]{glswlsols_llzo.pdf}
+    \script{glswlsols_llzo.py}
+    \caption{
+        Example distributions of estimated self-diffusion coefficients, $\Dest$, calculated using (a) ordinary least squares (OLS), (b) weighted least squares (WLS), and (c) generalised least squares (GLS),
+        from MSD data from \num{512} effective simulations of LLZO of \SI{\sim 25}{\pico\second} with 56 lithium ions.
+        In each panel, the grey curve shows the best-fit normal distribution for the simulation data, the upper horizontal bar shows the standard deviation of this distribution, and the lower horizontal bar shows the average estimated standard distribution given by the analytical expression for $\sigma[\prob{\Dest}]$ for each regression method.}
+    \label{fig:glswlsols_llzo}
+\end{figure}
 

### 2024-09-12 src/code/llzo/many_runs_old.py

@@ -1,0 +1,105 @@
+import MDAnalysis as mda
+import numpy as np
+from kinisi.parser import MDAnalysisParser
+from kinisi.diffusion import MSDBootstrap
+from tqdm import tqdm
+
+n = snakemake.params['n']
+start_diff = snakemake.params['start_diff']
+
+def universe_slice(u: mda.Universe, data_file: str, slice_start: int, slice_end: int) -> mda.Universe:
+    """
+    Concatenate a list of Universes.
+
+    :param u_list: List of Universe objects
+    :param data_file: File path for data
+    :param slice_start: Step to start the slice from
+    :return: A single flattened Universe
+    """
+    j = 0
+    trajectory = np.zeros((len(u.trajectory[slice_start:slice_end]), len(u.atoms), 3))
+    for t in tqdm(u.trajectory[slice_start:slice_end]):
+        trajectory[j] = t.positions
+        j += 1
+    shorter_universe = mda.Universe(data_file, trajectory)
+    for t in tqdm(shorter_universe.trajectory):
+        for i in shorter_universe.atoms:
+            if i.type == 'LI':
+                i.type = 'Li'
+            if i.type == 'Li1':
+                i.type = 'Li'
+            if i.type == 'Li2':
+                i.type = 'Li'
+            if i.type == 'L':
+                i.type = 'La'
+            if i.type == 'Z':
+                i.type = 'Zr'
+        t.dimensions = np.array([26.175106662244392197, 26.175106662244392197, 26.175106662244392197, 90, 90, 90])
+    return shorter_universe
+
+file = f'src/data/llzo/traj{n}.xyz'
+
+
+u = mda.Universe(file, file)
+uu = universe_slice(u, file, 0, 500)
+step_skip = 100  # sampling rate
+timestep = 5.079648e-4  # ps
+#124124122222gs22sess2ssh
+da_params = {'specie': 'Li', 'time_step': timestep, 'step_skip': step_skip, 'min_dt': 1}
+p = MDAnalysisParser(uu, **da_params)
+
+ngp = np.zeros((8, 4, p.delta_t.size))
+no = np.zeros((8, 4, p.delta_t.size))
+dt = np.zeros((8, 4, p.delta_t.size))
+msd = np.zeros((8, 4, p.delta_t.size))
+msd_true = np.zeros((8, 4, p.delta_t.size))
[…]

### 2024-09-12 src/code/llzo/true_ls.py

@@ -12,15 +12,15 @@ from kinisi.diffusion import _straight_line
 np.random.seed(1)
 
 start_diff = snakemake.params['start_diff']
+length = snakemake.params['length']
 
-timestep = np.load(f'src/data/llzo/diffusion_0_{start_diff}.npz')['dt'][0, 0]
-length = 2000
+timestep = np.load(f'src/data/llzo/diffusion_0_{start_diff}_{length}.npz')['dt'][0, 0]
 ll = len([i + length for i in range(0, 20001 - length, length)])
 # length = 500 
 # ll = len([i + length for i in range(0, 2000, length)])
-d = np.zeros((6, 8, ll, timestep.size))
-for i in range(0, 6, 1):
-    d[i] = np.load(f'src/data/llzo/diffusion_{i}_{start_diff}.npz')['msd_true']
+d = np.zeros((5, 16, ll, timestep.size))
+for i in range(1, 6, 1):
+    d[i-1] = np.load(f'src/data/llzo/diffusion_{i}_{start_diff}_{length}.npz')['msd_true']
 
 true_mean = d.reshape(-1, d.shape[-1]).mean(0)
 max_ngp = np.argwhere(timestep > start_diff)[0][0]
@@ -60,6 +60,6 @@ pos = max_likelihood + max_likelihood * 1e-3 * np.random.randn(32, max_likelihoo
 sampler = EnsembleSampler(*pos.shape, log_likelihood)
 sampler.run_mcmc(pos, 1000 + 500, progress=True, progress_kwargs={'desc': "Likelihood Sampling"})
 flatchain = sampler.get_chain(flat=True, thin=10, discard=500)
-np.savez(f'src/data/llzo/true_{start_diff}.npz',
+np.savez(f'src/data/llzo/true_{start_diff}_{length}.npz',
          diff_c=flatchain[:, 0] / 6e4,
          intercept=flatchain[:, 1])

### 2024-09-12 src/code/random_walks/efficiency.py

@@ -1,110 +1,0 @@
-import numpy as np
-from scipy.optimize import curve_fit
-from uravu.distribution import Distribution
-from uravu.relationship import Relationship
-
-jump = 1
-
-xaxis = np.array([16, 32, 64, 128, 256, 512, 1024])
-
-kinisi_atoms = np.zeros((xaxis.size, 512, 3200))
-weighted_atoms = np.zeros((xaxis.size, 512, 3200))
-ordinary_atoms = np.zeros((xaxis.size, 512, 3200))
-kinisi_length = np.zeros((xaxis.size, 512, 3200))
-weighted_length = np.zeros((xaxis.size, 512, 3200))
-ordinary_length = np.zeros((xaxis.size, 512, 3200))
-true_d_atoms = np.zeros((xaxis.size, 3200))
-true_d_length = np.zeros((xaxis.size, 3200))
-for i, ii in enumerate(xaxis):
-    length = 128
-    true_d_atoms[i] = np.load(f"src/data/random_walks/numerical/D_{jump}_{ii}_{length}.npz")["diff_c"]
-    file_open = np.load(f"src/data/random_walks/kinisi/rw_{jump}_{ii}_{length}_s512.npz")
-    kinisi_atoms[i] = file_open['diff_c']
-    file_open = np.load(f"src/data/random_walks/weighted/rw_{jump}_{ii}_{length}_s512.npz")
-    weighted_atoms[i] = file_open['diff_c']
-    file_open = np.load(f"src/data/random_walks/ordinary/rw_{jump}_{ii}_{length}_s512.npz")
-    ordinary_atoms[i] = file_open['diff_c']
-for i, ii in enumerate(xaxis):
-    atoms = 128
-    true_d_length[i] = np.load(f"src/data/random_walks/numerical/D_{jump}_{atoms}_{ii}.npz")["diff_c"]
-    file_open = np.load(f"src/data/random_walks/kinisi/rw_{jump}_{atoms}_{ii}_s512.npz")
-    kinisi_length[i] = file_open['diff_c']
-    file_open = np.load(f"src/data/random_walks/weighted/rw_{jump}_{atoms}_{ii}_s512.npz")
-    weighted_length[i] = file_open['diff_c']
-    file_open = np.load(f"src/data/random_walks/ordinary/rw_{jump}_{atoms}_{ii}_s512.npz")
-    ordinary_length[i] = file_open['diff_c']
-
[…]

### 2024-09-12 src/code/random_walks/kinisi_pyblock_modelfree_rw.py

@@ -1,46 +1,0 @@
-# Copyright (c) Andrew R. McCluskey
-# Distributed under the terms of the MIT License
-# author: Andrew R. McCluskey (arm61)
-
-import numpy as np
-from kinisi.diffusion import MSDDiffusion
-from tqdm import tqdm
-from random_walk import get_disp3d, walk
-
-jump = snakemake.params['jump']
-atoms = snakemake.params['atoms']
-length = int(snakemake.params['length'])
-size = int(snakemake.params['n'])
-
-timestep = np.arange(1, length + 1, 1, dtype=int)
-data = np.zeros((size, timestep.size, 4))
-covariance = np.zeros((size, timestep.size, timestep.size))
-npd_covariance = np.zeros((size, timestep.size, timestep.size))
-n_o = np.zeros((size, timestep.size))
-diff_c = np.zeros((size, 3200))
-intercept = np.zeros((size, 3200))
-
-for seed in tqdm(range(size)):
-    rng = np.random.RandomState(seed)
-    np.random.seed(seed)
-    disp_3d, n_samples = get_disp3d(walk, length, atoms, jump_size=jump, seed=rng)
-
-    diff = MSDDiffusion(timestep, disp_3d, n_samples, random_state=rng, progress=False, block=True)
-    diff.diffusion(0, model=False, random_state=rng, progress=False)
-    diff_c[seed] = diff.gradient.samples / 6
-    intercept[seed] = diff.intercept.samples
-    data[seed, :, 0] = diff.dt
-    data[seed, :, 1] = diff.n
-    data[seed, :, 2] = diff.s
-    data[seed, :, 3] = diff._model_v
-    npd_covariance[seed] = diff._npd_covariance_matrix
-    covariance[seed] = diff.covariance_matrix
-    n_o[seed] = diff._n_o
-
-np.savez(f'src/data/random_walks/pyblock_modelfree/rw_1_{atoms}_{length}_s{size}.npz',
-         diff_c=diff_c,
-         intercept=intercept,
-         data=data,
-         covariance=covariance,
-         npd_covariance=npd_covariance,
-         n_o=n_o)

### 2024-09-12 src/code/random_walks/kinisi_pyblock_rw.py

@@ -3,7 +3,7 @@
 # author: Andrew R. McCluskey (arm61)
 
 import numpy as np
-from kinisi.diffusion import MSDDiffusion
+from kinisi.diffusion import MSDBootstrap
 from tqdm import tqdm
 from random_walk import get_disp3d, walk
 
@@ -13,10 +13,9 @@ length = int(snakemake.params['length'])
 size = int(snakemake.params['n'])
 
 timestep = np.arange(1, length + 1, 1, dtype=int)
-data = np.zeros((size, timestep.size, 4))
-covariance = np.zeros((size, timestep.size, timestep.size))
-npd_covariance = np.zeros((size, timestep.size, timestep.size))
-n_o = np.zeros((size, timestep.size))
+data = np.zeros((size, timestep.size-1, 4))
+covariance = np.zeros((size, timestep.size-2, timestep.size-2))
+n_o = np.zeros((size, timestep.size-1))
 diff_c = np.zeros((size, 3200))
 intercept = np.zeros((size, 3200))
 
@@ -25,22 +24,21 @@ for seed in tqdm(range(size)):
     np.random.seed(seed)
     disp_3d, n_samples = get_disp3d(walk, length, atoms, jump_size=jump, seed=rng)
 
-    diff = MSDDiffusion(timestep, disp_3d, n_samples, random_state=rng, progress=False, block=True)
-    diff.diffusion(0, random_state=rng, progress=False)
+    diff = MSDBootstrap(timestep, disp_3d, n_samples, random_state=rng, progress=False, block=True)
+    diff.diffusion(2, random_state=rng, progress=False, cond_max=1e16)
     diff_c[seed] = diff.gradient.samples / 6
     intercept[seed] = diff.intercept.samples
     data[seed, :, 0] = diff.dt
     data[seed, :, 1] = diff.n
     data[seed, :, 2] = diff.s
-    data[seed, :, 3] = diff._model_v
-    npd_covariance[seed] = diff._npd_covariance_matrix
+    data[seed, 1:, 3] = diff._model_v
     covariance[seed] = diff.covariance_matrix
     n_o[seed] = diff._n_o
+    #tesyt
 
 np.savez(f'src/data/random_walks/pyblock/rw_1_{atoms}_{length}_s{size}.npz',
          diff_c=diff_c,
          intercept=intercept,
          data=data,
          covariance=covariance,
-         npd_covariance=npd_covariance,
          n_o=n_o)

### 2024-09-12 src/code/random_walks/kinisi_rw.py

@@ -3,7 +3,7 @@
 # author: Andrew R. McCluskey (arm61)
 
 import numpy as np
-from kinisi.diffusion import MSDDiffusion
+from kinisi.diffusion import MSDBootstrap
 from tqdm import tqdm
 from random_walk import get_disp3d, walk
 
@@ -14,30 +14,27 @@ size = int(snakemake.params['n'])
 
 timestep = np.arange(1, length + 1, 1, dtype=int)
 data = np.zeros((size, timestep.size, 4))
-covariance = np.zeros((size, timestep.size - 4, timestep.size - 4))
-npd_covariance = np.zeros((size, timestep.size - 4, timestep.size - 4))
-f = np.zeros((size, timestep.size))
+covariance = np.zeros((size, timestep.size - 1, timestep.size - 1))
 n_o = np.zeros((size, timestep.size))
 diff_c = np.zeros((size, 3200))
 intercept = np.zeros((size, 3200))
+cond_max = 1e16
 
 for seed in tqdm(range(size)):
     rng = np.random.RandomState(seed)
     np.random.seed(seed)
     disp_3d, n_samples = get_disp3d(walk, length, atoms, jump_size=jump, seed=rng)
 
-    diff = MSDDiffusion(timestep, disp_3d, n_samples, random_state=rng, progress=False)
-    diff.diffusion(5, random_state=rng, progress=False)
+    diff = MSDBootstrap(timestep, disp_3d, n_samples, random_state=rng, progress=False)
+    diff.diffusion(2, random_state=rng, progress=False, cond_max=cond_max)
     diff_c[seed] = diff.gradient.samples / 6
     intercept[seed] = diff.intercept.samples
     data[seed, :, 0] = diff.dt
     data[seed, :, 1] = diff.n
     data[seed, :, 2] = diff.s
-    data[seed, 4:, 3] = diff._model_v
-    npd_covariance[seed] = diff._npd_covariance_matrix
+    data[seed, 1:, 3] = diff._model_v
     covariance[seed] = diff.covariance_matrix
     n_o[seed] = diff._n_o
-    f[seed] = diff._hr
 
 
 np.savez(f'src/data/random_walks/kinisi/rw_1_{atoms}_{length}_s{size}.npz',
@@ -45,6 +42,4 @@ np.savez(f'src/data/random_walks/kinisi/rw_1_{atoms}_{length}_s{size}.npz',
          intercept=intercept,
          data=data,
[…]

### 2024-09-12 src/code/random_walks/ordinary_rw.py

@@ -26,8 +26,8 @@ for seed in tqdm(range(size)):
     rng = np.random.RandomState(seed)
     np.random.seed(seed)
 
-    x = data[seed, 4:, 0]
-    y = data[seed, 4:, 1]
+    x = data[seed, 1:, 0]
+    y = data[seed, 1:, 1]
     lin = linregress(x, y)
     
     gradient[seed] =  lin.slope / 6

### 2024-09-12 src/code/random_walks/true_ls.py

@@ -18,8 +18,8 @@ length = int(snakemake.params['length'])
 jump = int(snakemake.params['jump'])
 size = 4096
 
-true_msd = np.load(f'src/data/random_walks/numerical/rw_1_{atoms}_{length}_s{size}.npz')['data'][:, 4:]
-timestep = np.arange(1, length + 1, 1, dtype=int)[4:]
+true_msd = np.load(f'src/data/random_walks/numerical/rw_1_{atoms}_{length}_s{size}.npz')['data'][:, 1:]
+timestep = np.arange(1, length + 1, 1, dtype=int)[1:]
 true_mean = true_msd.mean(0)
 W = np.cov(true_msd.T)
 

### 2024-09-12 src/code/random_walks/truecov_rw.py

@@ -16,15 +16,15 @@ length = int(snakemake.params['length'])
 size = int(snakemake.params['n'])
 
 rcov = np.cov(
-    np.load(f'src/data/random_walks/numerical/rw_1_{atoms}_{length}_s4096.npz')['data'][:, 4:].T)
+    np.load(f'src/data/random_walks/numerical/rw_1_{atoms}_{length}_s4096.npz')['data'][:, 1:].T)
 kinisi = np.load(f'src/data/random_walks/kinisi/rw_1_{atoms}_{length}_s{size}.npz')['data']
 
 gradient = np.zeros((size, 3200))
 intercept = np.zeros((size, 3200))
 
 for seed in tqdm(range(size)):
-    x = kinisi[seed, 4:, 0]
-    y = kinisi[seed, 4:, 1]
+    x = kinisi[seed, 1:, 0]
+    y = kinisi[seed, 1:, 1]
 
     _, logdet = np.linalg.slogdet(rcov)
     logdet += np.log(2 * np.pi) * y.size

### 2024-09-12 src/code/random_walks/weighted_rw.py

@@ -21,14 +21,14 @@ intercept = np.zeros((size))
 intercept_err = np.zeros((size))
 data = np.load(f'src/data/random_walks/kinisi/rw_1_{atoms}_{length}_s{size}.npz')
 
-V = np.diag(data['data'][:, 4:, 1].var(0))
+V = np.diag(data['data'][:, 1:, 1].var(0))
 
 for seed in tqdm(range(size)):
     rng = np.random.RandomState(seed)
     np.random.seed(seed)
 
-    A = np.array([data['data'][seed, 4:, 0], np.ones(data['data'][seed, 4:, 0].size)]).T
-    y = data['data'][seed, 4:, 1].T
+    A = np.array([data['data'][seed, 1:, 0], np.ones(data['data'][seed, 1:, 0].size)]).T
+    y = data['data'][seed, 1:, 1].T
 
     g2 = np.matmul(np.linalg.inv(np.matmul(A.T, np.matmul(np.linalg.pinv(V), A))),
                 np.matmul(A.T, np.matmul(np.linalg.pinv(V), y)))

### 2024-09-12 src/scripts/covariances.py

@@ -24,15 +24,15 @@ newcmp3 = sns.light_palette(colors[2], as_cmap=True)
 newcmp4 = sns.diverging_palette(311.4, 161.2, as_cmap=True)
 
 true_msd = np.load(paths.data / f"random_walks/numerical/rw_{jump}_{atoms}_{length}_s4096.npz")["data"]
-true_cov = np.cov(true_msd[:, 4:length].T)
+true_cov = np.cov(true_msd[:, 1:length].T)
 
 kinisi_data = np.load(paths.data / f"random_walks/kinisi/rw_{jump}_{atoms}_{length}_s4096.npz")
 
 kinisi_cov = kinisi_data["covariance"].mean(0)
-no = (kinisi_data["n_o"] * kinisi_data['f'][:, 4, np.newaxis]).mean(0)
+no = (kinisi_data["n_o"]).mean(0)
 timestep = np.arange(1, length + 1, 1)
 
-ts_mesh = np.meshgrid(timestep[4:], timestep[4:])
+ts_mesh = np.meshgrid(timestep[1:], timestep[1:])
 
 anal_cov = np.zeros((length, length))
 for i in range(0, anal_cov.shape[0]):
@@ -40,7 +40,7 @@ for i in range(0, anal_cov.shape[0]):
         value = 8 * dimensionality**2 * D**2 * timestep[i]**2 / (dimensionality * no[j])
         anal_cov[i, j] = value
         anal_cov[j, i] = np.copy(anal_cov[i, j])
-anal_cov = anal_cov[4:, 4:]
+anal_cov = anal_cov[1:, 1:]
 
 max_cov = np.ceil(np.max([anal_cov.max(), true_cov.max(), kinisi_cov.max()]))
 min_cov = np.floor(np.min([anal_cov.min(), true_cov.min(), kinisi_cov.min()]))

### 2024-09-12 src/scripts/diffusion.py

@@ -7,19 +7,14 @@ import utils._fig_params as fp
 from utils.plotting_helper import mid_points
 import paths
 
-jump = 1
-atoms = 128
-length = 128
-type = 'kinisi'
-
-dllzo_true = np.load(paths.data / "llzo/true_10.npz")['diff_c']
-length = 2000
+length = 10000
+dllzo_true = np.load(paths.data / f"llzo/true_10_{length}.npz")['diff_c']
 ll = len([i + length for i in range(0, 20001 - length, length)])
 # length = 500
 # ll = len([i + length for i in range(0, 2000, length)])
-d = np.zeros((6, 8, ll, 3200))
-for i in range(0, 6, 1):
-    d[i] = np.load(paths.data / f'llzo/diffusion_{i}_10.npz')['d']
+d = np.zeros((5, 16, ll, 3200))
+for i in range(1, 6, 1):
+    d[i-1] = np.load(paths.data / f'llzo/diffusion_{i}_10_{length}.npz')['d']
 d = d.reshape(-1, 3200)
 llzo_x = np.linspace(dllzo_true.min(), dllzo_true.max(), 1000)
 
@@ -30,10 +25,8 @@ gs = gridspec.GridSpec(1, 2, figure=fig, wspace=1)
 axes = []
 titles = []
 
-print(d.shape)
-
 axes.append(fig.add_subplot(gs[0, 0]))
-y, x = np.histogram(d.mean(-1), bins=int(fp.NBINS * 0.5), density=True)
+y, x = np.histogram(d.mean(-1), bins=int(fp.NBINS * 0.6), density=True)
 axes[-1].stairs(y, x, fill=True, color='#ECBCA7', label='$p(\hat{D}^*)$')
 axes[-1].plot(llzo_x,
               norm.pdf(llzo_x, dllzo_true.mean(), dllzo_true.std()),
@@ -46,6 +39,7 @@ axes[-1].plot(np.array([d.mean(-1).mean() - d.mean(-1).std(),
               label='$\sigma[\hat{D}^*]$',
               marker="|")
 # axes[-1].set_yticks([0, 6, 12])
+# axes[-1].axvline(d.mean(-1).mean(), color='k')
 axes[-1].set_xlabel(r"$\hat{D}^*$ / cm$^2$ s$^{-1}$")
[…]

### 2024-09-12 src/scripts/glswlsols_llzo.py

@@ -9,7 +9,9 @@ from uravu.distribution import Distribution
 import utils._fig_params as fp
 import paths
 
-data = np.load('src/data/llzo/glswlsols_10.npz')
+length = 10000
+
+data = np.load(f'src/data/llzo/glswlsols_10_{length}.npz')
 
 figsize = (3.744, 3.4)
 fig = plt.figure(figsize=figsize)

### 2024-09-12 src/scripts/msd_blocking.py

@@ -1,0 +1,82 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+from scipy.stats import norm
+
+import utils._fig_params as fp
+from utils.plotting_helper import mid_points
+import paths
+
+jump = 1
+atoms = 128
+length = 128
+
+kinisi = np.load(paths.data / f"random_walks/pyblock/rw_{jump}_{atoms}_{length}_s4096.npz")['data']
+numerical = np.load(paths.data / f"random_walks/numerical/rw_{jump}_{atoms}_{length}_s4096.npz")['data']
+dt = kinisi[0, :, 0]
+
+figsize = (3.73, 0.85)
+fig = plt.figure(figsize=figsize)
+gs = gridspec.GridSpec(1, 3, figure=fig, wspace=0.7, hspace=0.7)
+
+axes = []
+titles = []
+
+axes.append(fig.add_subplot(gs[0, 0]))
+titles.append(r"a")
+axes[-1].set_title("numerical")
+m, c, b = axes[-1].errorbar(dt, numerical.mean(0), numerical.std(0) * 2, color='#81A8D9', elinewidth=0.2)
+[bar.set_alpha(0.5) for bar in b]
+[cap.set_alpha(0.5) for cap in c]
+axes[-1].set_xlabel(r"$t^{}$")
+axes[-1].set_ylabel(r"$x(t)$")
+axes[-1].set_xlim(0, None)
+axes[-1].set_ylim(0, None)
+axes[-1].set_xticks([0, 64, 128])
+axes[-1].set_yticks([0, 400, 800])
+
+axes.append(fig.add_subplot(gs[0, 1], sharey=axes[-1]))
+titles.append(r"b")
+axes[-1].set_title("estimated")
+m, c, b = axes[-1].errorbar(dt, kinisi[0, :, 1], kinisi[0, :, 2] * 2, color='#E08D6D', elinewidth=0.2)
+[bar.set_alpha(0.5) for bar in b]
+[cap.set_alpha(0.5) for cap in c]
+axes[-1].set_xlabel(r"$t^{}$")
+axes[-1].set_ylabel(r"$x(t)$")
+axes[-1].set_xlim(0, None)
+axes[-1].set_ylim(0, None)
+axes[-1].set_xticks([0, 64, 128])
+axes[-1].set_yticks([0, 400, 800])
+
+axes.append(fig.add_subplot(gs[0, 2]))
+titles.append(r"c")
[…]

### 2024-09-12 src/scripts/pyblock.py

@@ -12,32 +12,23 @@ atoms = 128
 length = 128
 
 kinisi = np.load(paths.data / f"random_walks/kinisi/rw_{jump}_{atoms}_{length}_s4096.npz")['diff_c']
-pyblock = np.load(paths.data / f"random_walks/pyblock/rw_{jump}_{atoms}_{length}_s4096.npz")['diff_c']
-pyblock_mf = np.load(paths.data / f"random_walks/pyblock_modelfree/rw_{jump}_{atoms}_{length}_s4096.npz")['diff_c']
+pyblock_mf = np.load(paths.data / f"random_walks/pyblock/rw_{jump}_{atoms}_{length}_s4096.npz")['diff_c']
 dinfty_true = np.load(paths.data / f"random_walks/numerical/D_1_128_128.npz")['diff_c']
 
-rw_x = np.arange(kinisi.mean(-1).min(), kinisi.mean(-1).max(), 0.01)
-rw_y = np.arange(pyblock.mean(-1).min(), pyblock.mean(-1).max(), 0.01)
-rw_z = np.arange(pyblock_mf.mean(-1).min(), pyblock_mf.mean(-1).max(), 0.01)
-rw_x2 = np.arange(np.array([kinisi.var(-1, ddof=1), pyblock.var(-1, ddof=1), pyblock_mf.var(-1, ddof=1)]).min(), 
-                  np.array([kinisi.var(-1, ddof=1), pyblock.var(-1, ddof=1), pyblock_mf.var(-1, ddof=1)]).max(), 0.0001)
-rw_x3 = np.arange(kinisi.mean(-1).min(), kinisi.mean(-1).max(), 0.0001)
-rw_y3 = np.arange(pyblock.mean(-1).min(), pyblock.mean(-1).max(), 0.0001)
-rw_z3 = np.arange(pyblock_mf.mean(-1).min(), pyblock_mf.mean(-1).max(), 0.0001)
-
-figsize = (4.03, 6)
+figsize = (4.03, 4)
 print(figsize)
 fig = plt.figure(figsize=figsize)
-gs = gridspec.GridSpec(3, 2, figure=fig, wspace=0.6, hspace=1)
+gs = gridspec.GridSpec(2, 2, figure=fig, wspace=0.6, hspace=1)
 
 axes = []
 titles = []
 
 axes.append(fig.add_subplot(gs[0, 0]))
-y, x = np.histogram(kinisi.mean(-1), bins=rw_x, density=True)
[…]

### 2024-09-12 src/scripts/stat_eff.py

@@ -86,7 +86,6 @@ axes[-1].set_xlim(8, 2048)
 axes[-1].set_yticks([2**-15, 2**-12, 2**-9, 2**-6, 2**-3])
 axes[-1].minorticks_off()
 
-
 axes[-1].plot([], [], marker='.', ls='-', color='#F3ADBC', label='OLS')
 axes[-1].plot([], [], marker='.', ls='-', color='#81A8D9', label='WLS($\sigma_{\mathrm{num}}$)')
 axes[-1].plot([], [], marker='.', ls='-', color='#67B99B', label='approximate Bayesian regression ($\Sigma^\prime$)')

### 2024-09-13 src/tex/ms.tex
Ben edits

--- 2024-01-19 src/tex/ms.tex
@@
-% Define document class
-\documentclass[reprint,superscriptaddress,nobibnotes,amsmath,amssymb,aps,prx,hidelinks,linenumbers]{revtex4-2}
+\documentclass[reprint,superscriptaddress,nobibnotes,amsmath,amssymb,aps,prx,hidelinks]{revtex4-2}
 
-% Import relevant packages
 \usepackage{showyourwork}
@@
 \usepackage{blindtext}
-% \usepackage[hidelinks]{hyperref}
+\usepackage[noabbrev,nameinlink,capitalize]{cleveref}
+\crefname{equation}{Eqn.}{Eqns.}
+\crefformat{equation}{Eqn.~#2#1#3}
+\crefrangeformat{equation}{Eqns.~#3#1#4 to~#5#2#6}
+\crefmultiformat{equation}{Eqns.~#2#1#3}{ and~#2#1#3}{, #2#1#3}{ and~#2#1#3}
+\crefname{figure}{Fig.}{Figs.}
+
 \sisetup{
@@
 
-\graphicspath{{figures/}}
+\newcommand{\todo}[1]{\color{red}[#1]\color{black}}
 
-% Redefine \maketitle so that it can be used twice (for supplementary)
+\hyphenation{metalwalls}
+
 \makeatletter
@@
 
-% Begin!
 \begin{document}
 
-\let\oldaddcontentsline\addcontentsline% Store \addcontentsline
+\let\oldaddcontentsline\addcontentsline
 \renewcommand{\addcontentsline}[3]{}
 
-% Title
 \title{\papertitle}
 
-% Author list
 \author{Andrew R. McCluskey}
 \email{andrew.mccluskey@bristol.ac.uk}
-  \affiliation{School of Chemistry, University of Bristol, Cantock's Close, Bristol, BS8 1TS, United Kingdom}
-  \affiliation{European Spallation Source ERIC, Ole Maaløes vej 3, 2200 København N, DK}
+  \affiliation{Centre for Computational Chemistry, School of Chemistry, University of Bristol, Cantock's Close, Bristol, BS8 1TS, UK.}
+  \affiliation{European Spallation Source ERIC, Data Management and Software Centre, Asmussens Allé 305, DK-2800 Kongens Lyngby, DK.}
+  \affiliation{Diamond Light Source, Harwell Campus, Didcot, OX11 0DE, UK.}
 \author{Samuel W. Coles}
@@
 
-% Abstract
 \begin{abstract}
     Self-diffusion coefficients, $\D$, are routinely estimated from molecular dynamics simulations by fitting a linear model to the observed mean-squared displacements (MSDs) of mobile species.
-    MSDs derived from simulation suffer from statistical noise, which introduces uncertainty in the resulting estimate of $\D$.
-    An optimal scheme for estimating $\D$ will minimise this uncertainty, i.e., will have high statistical efficiency, and will give an accurate estimate of the uncertainty itself.
+    MSDs derived from simulation exhibit statistical noise that causes uncertainty in the resulting estimate of $\D$.
+    An optimal scheme for estimating $\D$ minimises this uncertainty, i.e., it will have high statistical efficiency, and also gives an accurate estimate of the uncertainty itself.
     We present a scheme for estimating $\D$ from a single simulation trajectory with high statistical efficiency and accurately estimating the uncertainty in the predicted value.
     The statistical distribution of MSDs observable from a given simulation is modelled as a multivariate normal distribution using an analytical covariance matrix for an equivalent system of freely diffusing particles, which we parameterise from the available simulation data.
-    We then perform Bayesian regression to sample the distribution of linear models that are compatible with this model multivariate normal distribution, to obtain a statistically efficient estimate of $D^*$ and an accurate estimate of the associated statistical uncertainty.
+    We use Bayesian regression to sample the distribution of linear models that are compatible with this multivariate normal distribution, to obtain a statistically efficient estimate of $D^*$ and an accurate estimate of the associated statistical uncertainty.
 \end{abstract}
@@
 
+\section*{Significance Statement}
+Quantifying diffusion in liquid and solid-state systems is fundamental to our understanding of their properties—without the ability to accurately measure diffusion coefficients, robust insight into these systems is impossible.
+Here, we present a statistically efficient approach to estimating self-diffusion coefficients from single molecular dynamics simulations, using approximate Bayesian regression.
+Our method provides estimates of self-diffusion coefficients with significantly less uncertainty than textbook ordinary-least-squares or weighted-least-squares methods, allowing for more efficient predictions of diffusion behaviour from simulations at greatly reduced overall computational cost, and also gives an accurate estimate of the variance in the predicted value, allowing for meaningful downstream comparison of data.
+
 \section{Introduction}
 
-Mass transport is a fundamental physical process that is central to our understanding of fluids~\cite{sendner_interfacial_2009,shimizu_structural_2015,ghoufi_ultrafast_2016} and plays a critical role in biochemical systems~\cite{maccmmon_dynamics_1977,robertson_diffusion_2006}, and in solid-state devices such as batteries, fuel cells, and chemical sensors~\cite{eames_ionic_2015,morgan_understanding_2011,walsh_taking_2018}.
-Molecular dynamics simulations are widely used to study microscopic transport processes, as they give direct insight into atomic-scale transport mechanisms and can be used to calculate macroscopic transport coefficients~\cite{morgan_relationships_2014,morgan_mechanistic_2021,poletayev_defect_2022,klepis_long_2009,wang_application_2011,zelovich_hydroxide_2019}.
+Mass transport is a fundamental physical process that is central to our understanding of fluids~\cite{sendner_interfacial_2009,ShimizuEtAl_PhysChemChemPhys2015,ghoufi_ultrafast_2016} and plays a critical role in biochemical systems~\cite{maccmmon_dynamics_1977,RobertsonEtAl_ProcNatlAcadSci2006} and solid-state devices, such as batteries, fuel cells, and chemical sensors~\cite{eames_ionic_2015,morgan_understanding_2011,walsh_taking_2018}.
+Molecular dynamics simulations are widely used to study microscopic transport processes, as they give direct insight into atomic-scale transport mechanisms and can be used to calculate macroscopic transport coefficients~\cite{morgan_relationships_2014,morgan_mechanistic_2021,poletayev_defect_2022,klepis_long_2009,wang_application_2011,ZelovichEtAl_ChemMater2019}.
 These transport coefficients are formally defined in terms of ensemble averages.
-Dynamical simulations, however, sample the full ensemble space stochastically, and parameters derived from simulation data are therefore only estimates of the true parameter of interest.
+Dynamical simulations, however, sample the full ensemble space stochastically, and parameters derived from simulation data, therefore, are estimates of the true parameter of interest.
 The statistical uncertainty associated with such estimates depends on the details of the simulation---e.g., size and timescale---and on the choice of estimation method.
@@
 $\D$ is related to the ensemble-average mean squared displacement (MSD), $\MSD{t}$, via the Einstein relation~\cite{einstein_uber_1905,helfand_transport_1960},
-%
 \begin{equation}
@@
 \end{equation}
-%
 where $t$ is elapsed time.
-Because numerical simulations are finite in both time and space, MSDs obtained from simulation data always deviate from the true ensemble average MSD.
-One can, however, compute an estimate of the self-diffusion coefficient, $\Dest$, by fitting a linear model to the observed MSD and using the gradient of this fitted model in place of $\MSD{t} / t$ in Eqn.~\ref{equ:einstein}.
+Because numerical simulations are finite in time and space, MSDs obtained from simulation data always differ from the true ensemble average MSD.
+One can, however, compute an estimate of the self-diffusion coefficient, $\Dest$, by fitting a linear model to the observed MSD and using the gradient of this fitted model in place of $\MSD{t} / t$ in \cref{equ:einstein}.
 
@@
 OLS gives analytical expressions for the ``best fit'' regression coefficients (the slope and intercept) and their respective uncertainties, making it easy to implement and quick to perform.
-This procedure, however, is appropriate only for data that are both statistically independent and identically distributed.
-Neither of these conditions hold for MSD data obtained from simulation, which instead are serially correlated and usually have unequal variances.
+OLS, however, is appropriate only for data that are both statistically independent and identically distributed.
+Neither of these conditions holds for MSD data obtained from simulation, which instead are serially correlated and usually have unequal variances.
 As a consequence, OLS is statistically inefficient, giving a relatively large statistical uncertainty in $\Dest$.
-Furthermore, using the textbook OLS expression for the uncertainty in $\Dest$ significantly underestimates the true uncertainty in this estimate.
-This underestimated uncertainty may cause overconfidence in the accuracy of values of $\D$ estimated using OLS, and using these data in downstream analyses may result in faulty inferences.
-While the uncertainty associated with OLS estimates of $\D$ can, in principle, be accurately estimated by directly sampling over multiple repeated simulations, this approach greatly increases the total computational cost and therefore is often not practical.
+Furthermore, the textbook OLS expression for the uncertainty in $\Dest$ significantly underestimates the true uncertainty in this estimate~\cite{UslerEtAl_JComputChem2023}.
+This underestimated uncertainty may give overconfidence in the accuracy of values of $\D$ estimated using OLS, and propagating these uncertainties in any downstream analyses may result in faulty inferences.
+While the uncertainty associated with OLS estimates of $\D$ can, in principle, be accurately estimated by directly sampling over multiple repeated simulations, this approach can greatly increase the total computational cost and is often impractical.
 
-\begin{figure*}
-    \centering
-    % \resizebox{\columnwidth}{!}{\includegraphics*{figure1.pdf}}
-    \includegraphics[width=\textwidth]{schematic.pdf}
-    \caption{
-        A schematic diagram of the Bayesian regression method described in this work, running from our simulations through the sampling of displacement, variance and covariance estimation and final the sampling process to give the marginal posterior distribution for $\D$.}
-    \label{fig:schematic}
-\end{figure*}
-
-Here, we describe an approximate Bayesian regression method for estimating $\D$ with near-maximal statistical efficiency while also accurately estimating the corresponding statistical uncertainty, using data from a single simulation (Fig.~\ref{fig:schematic}).
+Here, we describe an approximate Bayesian regression method for estimating $\D$ with near-maximal statistical efficiency while accurately estimating the corresponding statistical uncertainty using data from a single simulation.
 We model the statistical population of simulation MSDs as a multivariate normal distribution, using an analytical covariance matrix derived for an equivalent system of freely diffusing particles, with this covariance matrix parameterised from the observed simulation data.
@@
 
-\section{Results}
-\subsection{Background}
+\section{Background}
 
 For a simulation of equivalent particles, the observed mean squared displacement as a function of time, $\oMSDs{t}$, can be computed as an average over equivalent particles and time origins:
-%
 \begin{equation}
@@
 \end{equation}
-%
 where $N(t)$ is the total number of observed squared-displacements at time $t$.
@@
 Each potential MSD sample could, in principle, be fitted to a linear model to obtain a corresponding estimate for the self-diffusion coefficient; $\bm{X} \mapsto \Dest$.
-The population of all such estimates therefore defines a probability distribution $\prob{\Dest}$.
+The population of all such estimates, therefore, defines a probability distribution $\prob{\Dest}$.
 The estimated diffusion coefficient obtained from a single simulation corresponds to a random sample drawn from this distribution, while the uncertainty in $\Dest$ is described by the shape of the full distribution $\prob{\Dest}$.
@@
 These variances are also typically unequal---the data are heteroscedastic \cite{smith_random_1996,he_statistical_2018,UslerEtAl_JComputChem2023}. 
-Because the key assumptions of the OLS method are not valid for MSD data, OLS gives statistically inefficient estimates of $\D$, while the estimated regression uncertainties obtained from the standard OLS statistical formulae significantly underestimate the true uncertainty in $\prob{\Dest_\mathrm{OLS}}$ (Fig.~\ref{fig:glswlsols}a).
+Because the key assumptions of the OLS method are not valid for MSD data, OLS gives statistically inefficient estimates of $\D$, while the estimated regression uncertainties obtained from the standard OLS statistical formulae significantly underestimate the true uncertainty in $\prob{\Dest_\mathrm{OLS}}$ (\cref{fig:glswlsols}a).
 
@@
     \centering
-    % \resizebox{\columnwidth}{!}{\includegraphics*{figure1.pdf}}
-    \includegraphics[width=\columnwidth]{glswlsols.pdf}
-    \script{glswlsols.py}
+    \resizebox{\columnwidth}{!}{\includegraphics*{figure1.pdf}}
     \caption{
@@
         from MSD data from \num{4096} individual simulations of \num{128} particles undergoing a \SI{128}{step} 3D lattice random walk, with a step size chosen so that the true diffusion coefficient $\D = 1$.
-        In each panel, the grey curve shows the best-fit normal distribution for the simulation data, the upper horizontal bar shows the standard deviation of this distribution, and the lower horizontal bar shows the average estimated standard distribution given by the analytical expression for $\sigma[\prob{\Dest}]$ for each regression method.}
+        In each panel, the grey curve shows the best-fit normal distribution for the simulation data, the upper horizontal bar shows the standard deviation of this distribution, and the lower horizontal bar shows the average estimated standard deviation given by the analytical expression for $\sigma[\prob{\Dest}]$ for each regression method.}
     \label{fig:glswlsols}
+    \script{glswlsols.py}
 \end{figure}
 
-Some improvement can be made by using weighted least squares (WLS) (Fig.~\ref{fig:glswlsols}b), where the residual for each observed MSD value is weighted by the reciprocal of its variance, $1/(\var{\oMSDi})$.
+Some improvement can be made by using weighted least squares (WLS) (\cref{fig:glswlsols}b), where the residual for each observed MSD value is weighted by the reciprocal of its variance, $1/(\var{\oMSDi})$.
 Like OLS, WLS is an unbiased estimator, and for heteroscedastic data it has higher statistical efficiency than OLS.
-WLS still disregards correlations in $\oMSD$, however, and is therefore statistically inefficient, while the WLS estimated uncertainties for the regression coefficients still underestimate the true uncertainty in $\prob{\Dest_\mathrm{WLS}}$.
+WLS still neglects correlations in $\oMSD$, however, and is therefore statistically inefficient, and the WLS estimated uncertainties for the regression coefficients still underestimate the true uncertainty in $\prob{\Dest_\mathrm{WLS}}$.
 
@@
 GLS gives estimated regression coefficients, $\widehat{\beta}$, via
-%
 \begin{equation}
@@
 \end{equation} 
-%
 where $\modelmatrix$ is the model matrix $\begin{bmatrix}\mathbf{1} & \bm{t}\end{bmatrix}$, with $\bm{t}$ the vector of observed times, and $\mathbf{\Sigma}$ is the covariance matrix for the observed MSD values.
-For correlated heteroscedastic data, such as MSD data, GLS offers the theoretical maximum statistical efficiency---it achieves the Cram\'er--Rao bound~\cite{cramer_mathematical_1946,rao_information_1945,rao_selected_1994,darmois_sur_1945,aitken_on_1942}---and provides accurate analytical estimates of the uncertainty in the predicted regression coefficients (Fig.~\ref{fig:glswlsols}c).
+For correlated heteroscedastic data, such as MSD data, GLS offers the theoretical maximum statistical efficiency---it achieves the Cram\'er--Rao bound~\cite{cramer_mathematical_1946,rao_information_1945,rao_selected_1994,darmois_sur_1945,aitken_on_1942}---and provides accurate analytical estimates of the uncertainty in the predicted regression coefficients (\cref{fig:glswlsols}c).
 
@@
 Like GLS, Bayesian regression can take into account both the changing variance and the correlation structure inherent in the data.
-Rather than providing a singular ``best-fit'' estimate like GLS, Bayesian regression produces a posterior probability distribution for the regression coefficients.
-The mean of this distribution serves as the point estimate of the coefficients and, in the absence of additional prior information, is equivalent to the GLS estimate, while the spread of the distribution quantifies the uncertainty in these estimates.
-For data that is both heteroscedastic and correlated, such as MSD data from simulations, Bayesian regression, like GLS, is formally fully statistically efficient.
+Rather than providing a singular ``best-fit'' estimate like GLS, Bayesian regression produces a posterior joint probability distribution for the regression coefficients.
+The mean of this distribution serves as a point estimate of the coefficients and, in the absence of additional prior information, is equal to the estimate obtained from GLS, while the spread of the distribution quantifies the uncertainty in these estimates.
+For data that are both heteroscedastic and correlated, such as MSD data from simulations, Bayesian regression, like GLS, is formally fully statistically efficient.
 
-To estimate $\D$ from some observed MSD data, $\oMSD$, using Bayesian regression, we compute the posterior probability distribution $\prob{\model|\oMSD}$ for a linear model $\model = 6\D \bm{t} + c$, where $\D$ and $c$ are parameters to be estimated.
+The estimation of $\D$ from some observed MSD data, $\oMSD$, using Bayesian regression, proceeds by computing the posterior probability distribution $\prob{\model|\oMSD}$ for a linear model $\model = 6\D \bm{t} + c$, where $\D$ and $c$ are parameters to be estimated.
 This posterior distribution is described by Bayes' theorem,
-%
 \begin{equation}
@@
 \end{equation}
-%
 where $\prob{\oMSD|\model}$ is the probability of observing data $\oMSD$ given model $\model$, often described as the ``likelihood'', and $\prob{\oMSD}$ is the marginal probability of the observed data $\oMSD$.
@@
 Given a sufficiently large number of observed squared displacements at each time $t$, the central limit theorem applies, and $\oMSD$ can be considered a sample from a multivariate normal distribution with log-likelihood
-%
 \begin{equation}
@@
 \end{equation}
-%
 where $\mathbf{\Sigma}$ is the observed MSD covariance matrix and $k$ is the length of the vector $\oMSD$, i.e., the number of time intervals for which we have observed MSD data.
-Providing this likelihood function can be calculated, we can compute the posterior distribution $\prob{\model | \oMSD}$, which gives an optimally efficient point-estimate for $\D$ and a complete description of the associated uncertainty in $\Dest$.
+Providing that this likelihood function can be calculated, one can compute the posterior distribution $\prob{\model | \oMSD}$ via \cref{equ:bayes}, to obtain an optimally efficient point-estimate for $\D$ and a complete description of the associated uncertainty in $\Dest$.
 
-\subsection{Approximating $\bm{\Sigma}$ from simulation data}
+\section{Approximating $\bm{\Sigma}$ from simulation data}
 
-For Bayesian regression and GLS, we require the covariance matrix for the observed MSD, $\mathbf{\Sigma}$, which is generally unknown.
-To proceed, we replace $\mathbf{\Sigma}$ with a model covariance matrix, $\mathbf{\Sigma^\prime}$, with a known analytical form, that we parameterise from the available simulation data.
+The practical application of Bayesian regression or GLS requires the covariance matrix for the observed MSD, $\mathbf{\Sigma}$, which is generally unknown.
+To proceed, we approximate $\mathbf{\Sigma}$ with a model covariance matrix, $\mathbf{\Sigma^\prime}$, with a known analytical form, that we parameterise from the available simulation data.
 Providing the correlation structure of $\mathbf{\Sigma^\prime}$ is similar to that of $\mathbf{\Sigma}$, this model correlation matrix can be used in approximate Bayesian or GLS schemes to estimate the ensemble-average MSD, and hence $\D$, with high efficiency and accurate estimated uncertainties.
@@
 We model the covariance matrix for the observed MSD from a given simulation using the covariance matrix for the MSD of an equivalent system of freely diffusing particles, $\mathbf{\Sigma^\prime}$.
-For observed MSDs computed by averaging over numerically-independent sub-trajectories, the covariance matrix $\mathbf{\Sigma^\prime}$, in the long time limit, has elements (see SI)
-%
+We note that estimating \Dest\ by fitting a linear model to observed MSD data implicity assumes that these data sample the long-time limit where the Einstein relation (\cref{equ:einstein}) is valid.
+In this long-time diffusive regime, all systems of mobile particles are statistically equivalent under rescaling by \D, and hence have the same MSD covariance structure $\mathbf{\Sigma}$.
+
+For observed MSDs computed by averaging over non-overlapping time windows, the covariance matrix $\mathbf{\Sigma^\prime}$, in the long-time limit, has elements (see SI)
 \begin{equation}
@@
 \end{equation} 
-%
-where $\var{\oMSDi}$ are the time-dependent variances of the observed MSD, and $\nind{i}$ is the total number of numerically-independent observed squared-displacements for time-interval $i$, accounting for both overlapping sampling and concerted motion (see Eqn.~\ref{equ:der_var}).
+where $\var{\oMSDi}$ are the time-dependent variances of the observed MSD, and $\nind{i}$ is the total number of non-overlapping observed squared-displacements for time-interval $i$.
 We estimate the variances $\var{\oMSDi}$ using the standard result that the variance of the mean of a sample scales inversely with the number of independent constituent observations.
-Specifically, we approximate the variance $\varest{\oMSDi}$ (see SI) by rescaling the observed variance of the squared displacement for time interval $i$ by the number of numerically-independent contributing sub-trajectories, $\nind{i}$;
-%
+Specifically, we compute an estimated variance $\varest{\oMSDi}$ by rescaling the observed variance of the squared displacement for time interval $i$ by the number of numerically-independent contributing sub-trajectories, $\nind{i}$, which is given by the number of mutually non-overlapping time windows of length $i$ multiplied by the number of mobile particles, summed over all simulations used to compute the MSD;
 \begin{equation}
-  \varest{\oMSDi} \approx \frac{1}{\nind{i}}\var{\Delta r_i^2}.
+  \varest{\oMSDi} = \frac{1}{\nind{i}}\var{\Delta r_i^2}.
   \label{equ:varestMSD}
 \end{equation}
-%
+By rescaling by the number of numerically-independent contributing sub-trajectories, rather than by the total number of observed squared displacements for time window $i$, we account for correlations between the squared displacements of each particle computed from overlapping time windows (further details are provided in the SI).
 
-Rescaling by the number of numerically-independent contributing sub-trajectories has the effect of renormalising the variance of the observed squared displacements to account for correlations between particle squared displacements computed from overlapping time windows.
-An alternative approach to renormalising $\var{x_i}$ is to use a non-parametric block-averaging procedure~\cite{flyvbjerg_error_1989,frenkel_understanding_2023,materzanini_high_2021}, which gives undesirable results for a random walk (see SI).
-The block-averaging approach, additionally, requires numerical convergence with respect to block size, which is not guaranteed for time windows with few independent observations. 
-We require $\var{x_i}$ at all observed time intervals, $i$, making block averaging on large data sets computationally prohibitive.
-
-The estimated variance $\varest{\oMSD}$ can be calculated from a single simulation trajectory, and provides an accurate estimate of the true variance $\var{\oMSD}$.
+The estimated variance $\varest{\oMSD}$ can be calculated from a single simulation trajectory, and provides an accurate estimate of the true variance $\var{\oMSD}$~\footnote{A common approach to estimating the variance of the mean of time-correlated data with unknown correlation time is the renormalization group blocking method of Flyvberg and Peterson~\cite{FlyvbjergAndPetersen_JChemPhys1989,Frenkel2023-ah}.
+In the SI, we compare this blocking method to our direct rescaling method (\cref{equ:variance}) for estimating $\var{\oMSDi}$ from a set of random-walk trajectories.
+For our example data, direct rescaling performs better (gives more accurate estimates of $\var{\oMSDi}$).}.
 To demonstrate this, we performed \num{4096} independent simulations of \num{128} particles undergoing a three-dimensional cubic-lattice random walk of \num{128} steps per particle.
-Using data from all \num{4096} simulations, we first compute the true simulation MSD and its variance (Fig.~\ref{fig:msd}a).
-We also compute the MSD and estimated variance using data from a single simulation trajectory (Fig.~\ref{fig:msd}b), using the scheme described above.
-A quantitative comparison between the true MSD variance and the single-trajectory estimated MSD variance is made in Fig.~\ref{fig:msd}c: the close numerical agreement confirms that Eqn~\ref{equ:varestMSD} can be used to estimate $\var{\oMSD}$, which can then be used to parameterise the model covariance matrix $\mathbf{\Sigma^\prime}$ via Eqn.~\ref{equ:cvv}.
-
+Using data from all \num{4096} simulations, we first compute the true simulation MSD and its variance (\cref{fig:msd}a).
+We also compute the MSD and estimated variance using data from a single simulation trajectory (\cref{fig:msd}b), using the scheme described above.
+A quantitative comparison between the true MSD variance and the single-trajectory estimated MSD variance is made in \cref{fig:msd}c: the close numerical agreement confirms that \cref{equ:varestMSD} can be used to estimate $\var{\oMSD}$, which can then be used to parameterise the model covariance matrix $\mathbf{\Sigma^\prime}$ via \cref{equ:cvv}.
 \begin{figure}
     \centering
-    % \resizebox{\columnwidth}{!}{\includegraphics*{figure2.pdf}}
-    \includegraphics[width=\columnwidth]{msd.pdf}
-    \script{msd.py}
+    \resizebox{\columnwidth}{!}{\includegraphics*{figure2.pdf}}
     \caption{
-        Comparison of the numerical variance in observed MSD from multiple replica simulations and the estimated variance in observed MSD given by rescaling the variance in observed squared displacements (Eqn.~\ref{equ:varestMSD}).
+        Comparison of the numerical variance in observed MSD from multiple replica simulations and the estimated variance in observed MSD given by rescaling the variance in observed squared displacements (\cref{equ:varestMSD}).
         Panel (a) shows the mean observed MSD from \num{4096} simulations of \num{128} particles undergoing a 3D lattice random walk of \num{128} steps per particle, with error bars of $\pm2\sigma[\oMSDi]$.
-        Panel (b) shows the MSD from just one simulation, with error bars of $\pm2\widehat{\sigma}[\oMSDi]$, obtained via Eqn.~\ref{equ:varestMSD}.
+        Panel (b) shows the MSD from just one simulation, with error bars of $\pm2\widehat{\sigma}[\oMSDi]$, obtained via \cref{equ:varestMSD}.
         Panel (c) plots the numerical variance against the estimated variance from a single simulation as a function of timestep $i$.
-        %Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences.
     }
     \label{fig:msd}
+    \script{msd.py}
 \end{figure}
 
-The practical implementation of both GLS and Bayesian regression requires that the covariance matrix $\mathbf{\Sigma^\prime}$ is invertible (positive definite); see Eqns.~\ref{equ:gls} and \ref{equ:loglike}.
-The estimated MSD variances derived from simulation data via Eqn.~\ref{equ:varestMSD} are statistically noisy and using these to directly parameterise $\mathbf{\Sigma^\prime}$ can yield non-invertible singular matrices.
-To make our scheme numerically tractable, we therefore fit our estimated MSD variances to the analytical variance for an analogous system of particles undergoing random walks~\cite{smith_random_1996};
-%
-\begin{equation}
-    \var{\oMSDi} = a\frac{t^2_i}{\nind{i}},
-    \label{equ:variance}
-\end{equation}
-%
-where $a$ is a scaling parameter determined by fitting Eqn.~\ref{equ:variance} to the directly estimated MSD variances.
-This smoothing of $\var{\oMSDi}$ guarantees that the resulting model covariance matrix $\mathbf{\Sigma^\prime}$ is invertible and thus suitable for GLS or Bayesian regression.
+The practical implementation of both GLS and Bayesian regression requires that the covariance matrix $\mathbf{\Sigma^\prime}$ is invertible (positive definite); see \cref{equ:gls,equ:loglike}.
+The estimated MSD variances derived from simulation data via \cref{equ:varestMSD}, however, are statistically noisy and using these to directly parameterise $\mathbf{\Sigma^\prime}$ can yield matrices with high condition numbers, resulting in numerical instabilities when these are used in \cref{equ:loglike}, or matrices that are singular and non-invertible.
+To make our scheme numerically tractable, we recondition the estimated covariance matrix obtained from \cref{equ:cvv} using the minimum eigenvalue method~\cite{TabeartEtAl_TellusDynMeteorolOceanogr2020}.
+This approach ensures that the condition number for the resulting covariance matrix is equal to a user defined parameter that can be set to produce an invertible matrix that allows numerically stable GLS or Bayesian regression.
 
-To illustrate the complete numerical procedure for deriving the model covariance matrix, $\bm{\Sigma^\prime}$, we present in Fig.~\ref{fig:covariances} the MSD covariance matrix for \num{4096} random-walk simulations, as described above, at three differing levels of approximation:
-the numerically converged covariance matrix, $\bm{\Sigma}$, computed using the data from all \num{4096} simulations (Fig.~\ref{fig:covariances}a);
-the corresponding analytical model covariance matrix, $\mathbf{\Sigma^\prime}$, defined by Eqn.~\ref{equ:cvv} and parametrised using analytical variances $\var{\oMSDi}$ (Fig.~\ref{fig:covariances}b); and the average model covariance matrix obtained by parametrising Eqn.~\ref{equ:cvv} using smoothed variances estimated from individual simulation trajectories, and averaging over the resulting set of all \num{4096} matrices (Fig.~\ref{fig:covariances}c).
-
+To illustrate the accuracy of the numerical procedure for deriving the model covariance matrix, $\bm{\Sigma^\prime}$, we present in \cref{fig:covariances} the MSD covariance matrix for \num{4096} random-walk simulations, as described above, at three differing levels of approximation:
+\cref{fig:covariances}a shows the numerically converged covariance matrix, $\bm{\Sigma}$, computed using the data from all \num{4096} simulations (\cref{fig:covariances}a); \cref{fig:covariances}b shows the corresponding analytical covariance matrix, $\mathbf{\Sigma^\prime}$, as defined by \cref{equ:cvv} and parametrised using analytical long-time-limit variances $\var{\oMSDi}$; and \cref{fig:covariances}c shows the average estimated matrix obtained by parametrising \cref{equ:cvv} using variances estimated from a single simulation trajectory, then reconditioning, with the average taken over all \num{4096} matrices obtained from the \num{4096} input simulations.
 \begin{figure}
     \centering
-    % \resizebox{\columnwidth}{!}{\includegraphics*{figure3.pdf}}
-    \includegraphics[width=\columnwidth]{covariances.pdf}
-    \script{covariances.py}
+    \resizebox{\columnwidth}{!}{\includegraphics*{figure3.pdf}}
     \caption{
         (a) The numerical MSD covariance matrix $\bm{\Sigma}$ calculated using MSD data from \num{4096} simulations of \num{128} particles undergoing a 3D lattice random walk of \num{128} steps per particle.
-        (b) The analytical MSD covariance matrix $\bm{\Sigma^\prime}$ (Eqn.~\ref{equ:cvv}), parametrised using analytical random-walk variances $\var{\oMSDi}$.
+        (b) The analytical MSD covariance matrix $\bm{\Sigma^\prime}$ (\cref{equ:cvv}), parametrised using analytical long-time limit random-walk variances $\var{\oMSDi}$.
         (c) The MSD covariance matrix obtained applying the numerical scheme described in the main text to each individual random walk simulation, averaged over all \num{4096} such simulations.
@@
         The off-diagonal panels show difference plots, computed as per-element ratios between pairs of covariance matrices (a--c).
-        %Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences.
     }
     \label{fig:covariances}
+    \script{covariances.py}
 \end{figure}
-
-While the analytical and average estimated covariance matrices show some systematic deviation from the numerically converged covariance matrix, the general correlation structure is preserved.
-The discrepancy between the model and numerical covariance matrices largely stems from the approximation made in deriving the analytical form that $t$ is large, which leads to an overestimation of the variance at low $t$.
-Despite this, the average estimated covariance matrix reproduces well the correlation structure of the true numerical covariance matrix, indicating that the covariance matrices estimated from individual simulation trajectories may be used within approximate GLS or Bayesian regression schemes to estimate $\D$ and $\var{\Dest}$.
-
-\subsection{Validation}
 
@@
     \centering
-    % \resizebox{\textwidth}{!}{\includegraphics*{figure4.pdf}}
-    \includegraphics[width=\textwidth]{random_walk.pdf}
-    \script{random_walk.py}
+    \resizebox{\textwidth}{!}{\includegraphics*{figure4.pdf}}
     \caption{
@@
     \label{fig:random_walk}
+    \script{random_walk.py}
 \end{figure*}
 
-To demonstrate the complete approximate Bayesian regression scheme, as described above, we present two distinct examples.
-First, we consider a simple 3D-lattice random walk, where the true self-diffusion coefficient $\D$ is specified by the simulation parameters, and a well-converged numerical covariance matrix can be obtained with relatively low computational cost, which allows us to directly compare the estimates produced by our method to ``best case'' estimates from a hypothetical method with access to the true covariance matrix.
-Second, we consider an example real-world system---the lithium-ion solid electrolyte \ce{Li7La3Zr2O12} (LLZO)---which represents an application of our method to a well-studied material of practical interest for solid-state lithium-ion batteries \cite{MuruganEtAl_AngewChemIntEd2007,burbano_sparse_2016,morgan_lattice_2017,SquiresEtAl_PhysRevMater2022}.
+While the analytical and average estimated covariance matrices show some systematic deviation from the numerically converged covariance matrix, the general correlation structure is preserved.
+The discrepancy between the model and numerical covariance matrices largely stems from the approximation made in deriving the analytical form that $t$ is large, which leads to an overestimation of the variance at low $t$.
+Despite this, the average estimated covariance matrix reproduces well the correlation structure of the true numerical covariance matrix, and, as we show below, the covariance matrices estimated from individual simulation trajectories can be used within approximate GLS or Bayesian regression schemes to estimate $\D$ and $\var{\Dest}$.
 
-Fig.~\ref{fig:random_walk}a shows the observed MSD from a single 3D-lattice random-walk simulation, along with the estimated posterior distribution of linear models compatible with the observed MSD data, $\prob{\model|\oMSD}$, calculated via Eqns.~\ref{equ:bayes} and \ref{equ:loglike}.
-The corresponding marginal posterior distribution of estimated diffusion coefficients $\prob{\D|\oMSD}$ is shown in Fig.~\ref{fig:random_walk}b; this distribution is approximately Gaussian and is centred on the true self-diffusion coefficient $\D = \num{1}$, demonstrating that for this example trajectory we obtain a good point-estimate of $\D$.
+\section{Validation}
+
+To demonstrate the complete approximate Bayesian regression scheme we present two example use-cases.
+First, we consider a simple 3D-lattice random walk; in this case the true self-diffusion coefficient $\D$ is specified by the simulation parameters and a well-converged numerical covariance matrix can be obtained at relatively low computational cost, allowing us to directly compare the estimates produced by our method to ``best case'' estimates from a hypothetical method with access to the true covariance matrix.
+Second, we consider an example real-world system, the lithium-ion solid electrolyte \ce{Li7La3Zr2O12} (LLZO), which represents an application of our method to a well-studied material of practical interest for solid-state lithium-ion batteries \cite{MuruganEtAl_AngewChemIntEd2007,burbano_sparse_2016,morgan_lattice_2017,SquiresEtAl_PhysRevMater2022}.
+
+\Cref{fig:random_walk}a shows the observed MSD from a single 3D-lattice random-walk simulation, along with the estimated posterior distribution of linear models compatible with the observed MSD data, $\prob{\model|\oMSD}$, calculated via \cref{equ:bayes,equ:loglike}.
+The corresponding marginal posterior distribution of estimated diffusion coefficients $\prob{\D|\oMSD}$ is shown in \cref{fig:random_walk}b; this distribution is approximately Gaussian and is centred close to the true self-diffusion coefficient $\D = \num{1}$, demonstrating that for this example trajectory we obtain a good point-estimate of $\D$.
 
 To evaluate the overall performance of our method, we repeat our analysis on the full set of \num{4096} random-walk simulations.
-Fig.~\ref{fig:random_walk}c presents a histogram of the resulting point estimates of $\D$, with each estimate derived as the mean of the posterior distribution $\prob{\D|\oMSD}$ using input data from an individual simulation.
+\cref{fig:random_walk}c presents a histogram of the resulting point estimates of $\D$, with each estimate derived as the mean of the posterior distribution $\prob{\D|\oMSD}$ using input data from each individual simulation.
 We also show the probability distribution of estimated diffusion coefficients obtained using Bayesian regression with a mean vector and covariance matrix derived numerically from all \num{4096} simulations (solid line).
@@
 The distribution of estimated diffusion coefficients from single simulations is slightly broader than the exact numerical results.
-This minor deviation is a consequence of the overestimation of $\varest{\oMSDi}$ at short times, noted above, which results from our use of the long-time limit in the derivation of the analytical model covariance matrix.
+This minor deviation is a consequence of the noise present in data obtained from a single simulation trajectory.
 
 We next consider the degree to which our method can quantify the uncertainty in $\Dest$ when using input data from a single simulation.
-Fig.~\ref{fig:random_walk}d shows the distribution of estimated variances $\varest{\Dest}$, with each sample calculated from an individual simulation trajectory.
-We also show the true variance of individual point estimates, $\var{\Dest}$, which characterises the spread of the histogram in Fig.~\ref{fig:random_walk}c.
-The distribution of estimated variances is biased relative to the true variance and skewed, due to numerical differences between the true covariance matrix $\mathbf{\Sigma}$ and the model covariance matrix $\mathbf{\Sigma^\prime}$ (further details are provided in the SI).
+\cref{fig:random_walk}d shows the distribution of estimated variances $\varest{\Dest}$, with each sample calculated from an individual simulation trajectory.
+We also show the true variance of individual point estimates, $\var{\Dest}$, which characterises the spread of the histogram in \cref{fig:random_walk}c.
+The distribution of estimated variances is biased relative to the true variance, due to numerical differences between the true covariance matrix $\mathbf{\Sigma}$ and the estimated covariance matrix $\mathbf{\Sigma^\prime}$ (further details are provided in the SI).
 In general, however, the distribution of the estimated variance shows good agreement with the true sample variance.
-Notably, the precision of this estimate is significantly greater than obtained using OLS or WLS and their corresponding textbook statistical formulae.
+Notably, the precision of this estimate is significantly greater than obtained using OLS or WLS and their respective textbook statistical formulae.
 
-We next benchmark our method using data from simulations of the cubic lithium-ion solid electrolyte \ce{Li7La3Zr2O12} (c-LLZO).
-We performed a single simulation of \num{1536} atoms (\num{448} Li ions) at \SI{700}{K} for \SI{6}{\nano\second} (full simulation details are provided in the Methods section).
-To generate multiple statistically equivalent trajectories, the resulting simulation data was partitioned into \num{480} effective trajectories, each approximately $\sim\SI{100}{\ps}$ in length, and containing data for \num{56} lithium ions.
-We then perform the same approximate Bayesian regression analysis as above on each effective trajectory, excluding the first \SI{10}{ps} of MSD data in each case to remove short-time data corresponding to ballistic and sub-diffusive regimes \cite{he_statistical_2018}.
+We also benchmark our method using data from simulations of the lithium-ion solid electrolyte, cubic \ce{Li7La3Zr2O12} (c-LLZO).
+We performed a single simulation of \num{1536} atoms (\num{448} Li ions) at \SI{700}{K} for \SI{6}{\nano\second} (full simulation details are given in the Methods section).
+To generate multiple statistically equivalent trajectories, we partitioned the output simulation data into \num{192} effective trajectories, each $\sim\SI{500}{\ps}$ in length, and containing data for \num{28} lithium ions, which were selected randomly from the complete set of \num{448} lithium ions without replacement.
+We then performed approximate Bayesian regression, as above, on each effective trajectory, excluding the first \SI{10}{ps} of MSD data in each case to remove short-time data corresponding to the ballistic and sub-diffusive regimes \cite{burbano_sparse_2016,he_statistical_2018}.
 
-The resulting distribution of the point estimates, $\Dest$, from analysis of all \num{512} effective trajectories is shown in Fig.~\ref{fig:diffusion}a.
-Again, the corresponding distribution of $\Dest$ estimates derived using Bayesian regression and a well-converged numerical covariance matrix calculated from the full LLZO dataset is also shown for comparison.
+The resulting distribution of the point estimates, $\Dest$, from analysis of all \num{192} effective trajectories, is shown in \cref{fig:diffusion}a.
+As above, we also show the corresponding distribution of $\Dest$ estimates obtained via Bayesian regression using a well-converged numerical covariance matrix calculated from the full LLZO dataset.
 The distribution $\prob{\Dest}$ obtained using the model covariance matrix and parametrised separately for each individual effective simulation is highly similar to that obtained using the aggregate numerical covariance matrix calculated from the complete simulation dataset.
-This close agreement mirrors the results for our random walk simulations (see the SI for a similar comparison of the OLS, WLS, and GLS as shown in Fig.~\ref{fig:glswlsols}), and confirms that our method yields accurate and statistically efficient estimates for $\D$, even for real-world simulation data.
+This close agreement mirrors the results for our random walk simulations, and confirms that our method yields accurate and statistically efficient estimates for $\D$, even for real-world simulation data.
 
@@
     \centering
-    % \resizebox{\columnwidth}{!}{\includegraphics*{figure5.pdf}}
-    \includegraphics[width=\columnwidth]{diffusion.pdf}
-    \script{diffusion.py}
+    \resizebox{\columnwidth}{!}{\includegraphics*{figure5.pdf}}
     \caption{
-      (a) Probability distribution of point estimates $\prob{\Dest}$ for \num{512} effective simulations of LLZO (orange histogram).
+      (a) Probability distribution of point estimates $\prob{\Dest}$ for \num{192} effective simulations of LLZO (orange histogram).
       The grey line shows the distribution $\prob{\Dest_\mathrm{num}}$ obtained using Bayesian regression with the complete LLZO dataset as input.
@@
       (b) Probability distribution of estimated variances, $\varest{\Dest}$, for individual LLZO effective simulations, compared to the true sample variance (pink vertical line) $\var{\Dest}$.
-      %Raw simulation data~\cite{mccluskey_zenodo_2022} and scripts~\cite{mccluskey_github_2022} to generate this figure are available under CC BY 4.0/MIT licences. 
     }
     \label{fig:diffusion}
+    \script{diffusion.py}
 \end{figure}
 
-We also consider the probability distribution of estimates of the variance in $\Dest$ calculated for each effective trajectory (Fig.~\ref{fig:diffusion}b), which we compare to the true variance in $\Dest$ for our method; i.e., the variance of the histogram in Fig.~\ref{fig:diffusion}a.
+We also consider the probability distribution of estimates of the variance in $\Dest$ calculated for each effective trajectory (\cref{fig:diffusion}b), which we compare to the true variance in $\Dest$ for our method; i.e., the variance of the histogram in \cref{fig:diffusion}a.
 While the estimated variances deviate somewhat from the true distribution $\prob{\var{\Dest}}$, the agreement is reasonable and mirrors our results for the random walk simulations.
-Hence, our method provides reasonably accurate estimates of the uncertainty in $\Dest$ for our c-LLZO dataset, even when applied to single effective trajectories with limited displacement data (only 56 mobile ions, and \SI{25}{ps} simulation length).
+Hence, our method provides reasonably accurate estimates of the uncertainty in $\Dest$ for our c-LLZO dataset, even when applied to single effective trajectories with limited displacement data (only 28 mobile ions, and \SI{500}{ps} simulation length).
 
-\subsection{$\var{\Dest}$ scaling and comparison to OLS, WLS, and GLS}
+\section{$\var{\Dest}$ scaling and comparison to OLS, WLS, and GLS}
 
@@
     \centering
-    % \resizebox{\columnwidth}{!}{\includegraphics*{figure6.pdf}}
-    \includegraphics[width=\columnwidth]{stat_eff.pdf}
-    \script{stat_eff.py}
+    \resizebox{\columnwidth}{!}{\includegraphics*{figure6.pdf}}
     \caption{
@@
     \label{fig:stat_eff}
+    \script{stat_eff.py}
 \end{figure}
 
-Fig.~\ref{fig:stat_eff} presents an analysis of the variation in $\var{\Dest}$ as the number of mobile particles (Fig.~\ref{fig:stat_eff}a) and the total simulation time (number of steps) (Fig.~\ref{fig:stat_eff}b) are changed.
+\Cref{fig:stat_eff} presents an analysis of the variation in $\var{\Dest}$ as the number of mobile particles (\cref{fig:stat_eff}a) and the total simulation time (number of steps) (\cref{fig:stat_eff}b) are changed.
 We compare four methods for estimating $\D$ from the observed MSD data: OLS, WLS, the approximate Bayesian regression method described here, and GLS.
 When estimating $\D$ using WLS and GLS, we calculate the variances and the covariance matrix, respectively, numerically, using the complete set of \num{512} simulations.
-Each data point in Fig.~\ref{fig:stat_eff} represents the variance across point-estimates of $\D$ derived from \num{512} individual 3D-lattice random walk simulations, for each combination of $N_\mathrm{atoms}$ and $t_\mathrm{max}$.
-The GLS dataset corresponds to an optimally efficient estimator for linear regression of observed MSD data, equivalent to performing Bayesian regression with the numerical covariance matrix and an uninformative prior.
+Each data point in \cref{fig:stat_eff} represents the variance across point-estimates of $\D$ derived from \num{512} individual 3D-lattice random walk simulations, for each combination of $N_\mathrm{atoms}$ and $t_\mathrm{max}$.
+The GLS dataset corresponds to an optimally efficient estimator for linear regression of observed MSD data, and is equivalent to performing Bayesian regression with the converged numerical covariance matrix and an uninformative prior.
 
-Our approximate Bayesian regression method performs similarly to GLS, with a numerically converged covariance matrix, and gives significantly reduced uncertainty in $\Dest$ compared to OLS or WLS, for all simulation sizes and lengths considered.
+Our approximate Bayesian regression method performs similarly to GLS with a numerically converged covariance matrix, and gives significantly reduced uncertainty in $\Dest$ compared to OLS or WLS, for all simulation sizes and lengths considered.
 Moreover, our method scales better than OLS or WLS as the total simulation time is increased.
-The approximate Bayesian regression method in this manuscript therefore presents a significant improvement over more conventional methods such as OLS and WLS, by enabling more precise estimates of $\D$ across varied simulation sizes at equivalent computational cost.
+This approximate Bayesian regression method therefore presents a significant improvement over more conventional methods such as OLS and WLS, by enabling more precise estimates of $\D$ across varied simulation sizes at equivalent computational cost.
 
@@
 
-We have introduced and evaluated an approximate Bayesian regression method for estimating the self-diffusion coefficient, $\D$, from molecular dynamics simulation data.
+We have introduced and demonstrated an approximate Bayesian regression method for estimating the self-diffusion coefficient, $\D$, from molecular dynamics simulation data.
 We consider the observed mean-squared displacement data from a single simulation as a random sample, $\bm{X}$, from a population of potential MSDs generated by equivalent replica simulations, $\bm{X}\sim p(\oMSD)$.
@@
 To model the covariance matrix, we use an analytical solution derived for an equivalent system of freely diffusing particles.
-To parameterise this model covariance matrix, we rescale the variance of the observed squared displacements from the input simulation trajectory, followed by a smoothing step to ensure a positive-definite matrix.
+To parameterise this model covariance matrix, we renormalise the variance of the observed squared displacements from the input simulation trajectory, followed by a reconditioning step to ensure a positive-definite matrix.
 The resulting model covariance matrix preserves the correlation structure of the true simulation MSD covariance matrix, and gives a multivariate normal model for the population of observable simulation MSDs that depends solely on the model parameters, $\D$ and $c$.
@@
 We have benchmarked our approach using simulation data for an ideal 3D lattice random walk and for the lithium-ion solid electrolyte \ce{Li7La3Zr2O12} (LLZO).
-In both cases, we obtain a distribution of estimates for $\D$ that closely matches the theoretically optimal distribution obtained using a numerical covariance matrix derived from a large number of replica simulation trajectories.
+In both cases, we obtain a distribution of estimates for $\D$ that closely matches the theoretically optimal distribution obtained using a well-converged numerical covariance matrix derived from a large number of replica simulation trajectories.
 
@@
 This reduces the overall computational cost when compared to studies that use OLS or WLS for estimating a linear fit to simulation MSD data.
-Alternatively, this approach provides the possibility to estimate $\D$ with greater precision, given simulation trajectories of equal size.
+Alternatively, this approach enables the estimation of $\D$ with greater precision, given simulation trajectories of equal size.
 
 Our method also provides reasonable estimates of the statistical uncertainty in the estimated value $\Dest$, in contrast to OLS and WLS which systematically significantly underestimate the uncertainty in regression coefficients when applied to simulated MSD data.
-While these estimated statistical uncertainties can still differ from the true (but unknown) uncertainty in $\Dest$, particularly when using short-timescale simulation data, they allow for scientifically meaningful comparisons to be made between estimated diffusion coefficients across different materials or under varying conditions, such as changes in temperature, or between computational findings and experimental results.
+While these estimated statistical uncertainties can still differ from the true (but unknown) uncertainty in $\Dest$, particularly when using short-timescale simulation data, they allow for scientifically meaningful comparisons to be made between estimated diffusion coefficients across different systems or under varying conditions, such as changes in temperature, or between computational findings and experimental results.
 Furthermore, these uncertainties allow for quantitative downstream analysis, such as the application of Arrhenius (on non-Arrhenius) type models to describe the temperature dependence of self-diffusion.
@@
 Overall, our method offers significant advantages over more conventional methods of estimating self-diffusion coefficients from atomistic simulations.
-We have implemented this procedure in the open-source package \textsc{kinisi} \cite{mccluskey_kinisi_2022}, which we hope will support its use within the broader simulation community across a range of materials science contexts.
+We have implemented this procedure in the open-source package \textsc{kinisi} \cite{mccluskey_kinisi_2022}, which we hope will support its use within the broader simulation community.
 
@@
 \label{sec:implementation}
-We have implemented the approximate Bayesian estimation method described in the main text in the open-source Python package \textsc{kinisi}~\cite{mccluskey_kinisi_2022}, under the MIT licence.
+\textsc{kinisi}-1.1.0 was used for all analyses presented in this work.
 
-\textsc{kinisi} uses overlapping sliding window sampling when calculating the observed mean squared displacement at each time interval $t$ (see Eqn.~\ref{equ:observed_msd}).
-For a given time interval, $t$, the maximum number of observations is $N_{\mathrm{atoms}} \times (N_{t} - i)$ displacements, where $N_{\mathrm{atoms}}$ is the number of atoms, $N_{t}$ is the total number of timesteps, and $i$ is the index of the timestep (where \num{1} is the index for the shortest timestep).
+When calculating the observed mean squared displacement at each time interval $t$ (see \cref{equ:observed_msd}), \textsc{kinisi} uses overlapping sliding window sampling.
+For a given time interval, $t$, the maximum number of observations is $N_{\mathrm{atoms}} \times (N_{t} - i)$ displacements, where $N_{\mathrm{atoms}}$ is the number of mobile atoms, $N_{t}$ is the total number of timesteps, and $i$ is the index of the timestep (where \num{1} is the index for the shortest timestep).
 To estimate the variance of the observed MSD, we rescale the variance of observed squared displacements by the number of numerically-independent sub-trajectories in the simulation, $\nind{i} = N_{\mathrm{atoms}} \times N_{t} / i$, as presented in Eqn.~\ref{equ:varestMSD}.
@@
 The parametrisation of the covariance matrix from the variances $\var{\oMSDi}$ and the number of independent observations $\nind{i}$ is defined by Eqn.~\ref{equ:cvv}. 
-The covariance matrix is only constructed for values of $t$ where the particle motion is considered to be in the long-time diffusive limit, with this threshold set by the user to a value appropriate for their system and simulation data.
-For the examples presented in the main manuscript, we consider particles to be in the diffusive regime from $t=\num{4}$ for the random walk trajectories and from $t=\SI{10}{\pico\second}$ for the LLZO simulations.
+The covariance matrix is only constructed for values of $t$ where the particle motion is considered to be in the long-time diffusive limit.
+In practice, this threshold is set by the user to a value appropriate for their system and simulation data. 
+For the examples presented in the main manuscript, we consider particles to be in the diffusive regime from $t=\num{2}$ for the random walk trajectories and from $t=\SI{10}{\pico\second}$ for the LLZO simulations.
+The covariance matrix is reconditioned using the minimum eigenvalue method~\cite{TabeartEtAl_TellusDynMeteorolOceanogr2020}, with a maximum condition number of \num{1e16} for all simulations.
 
-\textsc{kinisi} uses ordinary least squares to obtain an initial guess for the gradient and intercept of the linear model describing the observed MSD.
-This initial guess is then used as the starting point for minimising the negative maximum a posteriori (the peak of the posterior distribution as per Eqn.~\ref{equ:bayes}), with the improper prior that $\D \ge 0$~\cite{broyden_convergence_1970,fletcher_new_1970,goldfarb_family_1970,shanno_conditioning_1970}.
-The log-likelihood calculation (Eqn.~\ref{equ:loglike}) uses the Moore-Penrose generalisation of the inverse of a Hermitian matrix~\cite{moore_on_1920,bjerhammar_application_1951,penrose_generlized_1955}.
+To estimate $\Dest$ from a given set of MSD data, \textsc{kinisi} uses ordinary least squares to obtain an initial guess for the gradient and intercept of the linear model that best describes the observed MSD.
+This initial guess is then used as the starting point for minimising the negative maximum a posteriori (the peak of the posterior distribution as per \cref{equ:bayes}), with the improper prior that $\D \ge 0$~\cite{broyden_convergence_1970,fletcher_new_1970,goldfarb_family_1970,shanno_conditioning_1970}.
+We note that the Bayesian regression formalism presented here allows for the use of alternative informative priors in cases where the user has some prior knowledge of the system being simulated that they wish to incorporate into their analysis.
+The log-likelihood calculation (\cref{equ:loglike}) uses the Moore--Penrose generalisation of the inverse of a Hermitian matrix~\cite{moore_on_1920,bjerhammar_application_1951,penrose_generlized_1955}.
 
@@
 The sampled chains are thinned such that only every 10th value is retained, yielding \num{3200} points sampled from the posterior distribution $\prob{\D|\model}$.
-These points can then be plotted as a histogram (as in Fig~\ref{fig:random_walk}b), and summary statistics $\Dest$ and $\varest{\Dest}$ can be derived.
+These points can then be plotted as a histogram (as in \cref{fig:random_walk}b), and summary statistics $\Dest$ and $\varest{\Dest}$ can be derived.
 
 \subsection{LLZO simulations}
-Classical molecular dynamics were run using the \textsc{metalwalls} code~\cite{marin_metalwalls_2020}. 
-We used the DIPPIM polarisable ion force field, as parameterised by Burbano \emph{et al.}~\cite{burbano_sparse_2016}, due to its proven accuracy in accounting for the effect of ion polarisability on diffusion~\cite{wilson_polarization_1993,burbano_sparse_2016}.
-We simulated the cubic phase of LLZO in NVT ensemble at a temperature of \SI{700}{\kelvin}.
-Simulations were run for \SI{6}{\nano\second} with a \SI{2}{\femto\second} timestep. 
+Classical molecular dynamics simulations were run using the \textsc{metalwalls} code~\cite{marin_metalwalls_2020}. 
+We used the DIPPIM polarisable ion force field, as parameterised by Burbano \emph{et al.}~\cite{burbano_sparse_2016}. 
+We simulated the cubic phase of LLZO in the NVT ensemble at a temperature of \SI{700}{\kelvin}.
+Simulations were run for \SI{6}{\nano\second} with a \SI{0.5}{\femto\second} timestep. 
 To control temperature, we used a Nos\'{e}-Hoover thermostat, with a relaxation time of \SI{121}{\femto\second} (5000 $\hbar / E_{h}$)~\cite{nose_unified_1984,hoover_canonical_1985,martyna_nose_1992}.
@@
 
-\section*{Data \& Code Availability}
+\section*{Data availability}
 
@@
 All raw simulation files are available on Zenodo shared under CC BY-SA 4.0 licences, the random walk simulations and other analysis-linked data~\cite{mccluskey_zenodo_2022} and the LLZO raw simulation trajectories~\cite{coles_llzo_zenodo_2022}.
-The method outlined in this work is implemented in the open-source Python package \textsc{kinisi}~\cite{mccluskey_kinisi_2022}, which is available under an MIT license, and can be accessed via \url{https://github.com/bjmorgan/kinisi}.
+The method outlined in this work is implemented in the open-source Python package \textsc{kinisi}~\cite{mccluskey_kinisi_2022}, which is available under an MIT license.
 
-\section*{Author Contribution Statement}
+\section*{CR\lowercase{e}d\lowercase{i}T author statement}
 
@@
 B.J.M. acknowledges support from the Royal Society (UF130329 and URF\textbackslash R\textbackslash 191006).
-The funders listed above played no role in study design, data collection, analysis and interpretation of data, or the writing of this manuscript.
-
-\section*{Competing interests}
-All authors declare no financial or non-financial competing interests.
 
@@
 \bibliography{bib}
-\let\addcontentsline\oldaddcontentsline
-
-\onecolumngrid
-\clearpage 
-\twocolumngrid
-
-\appendix
-\renewcommand\thesection{SI.\Roman{section}}
-\counterwithout{figure}{section}
-\renewcommand\thefigure{SI.\arabic{figure}}
-\setcounter{figure}{0}
-\counterwithout{equation}{section}
-\renewcommand\theequation{SI.\arabic{equation}}
-\setcounter{equation}{0}
-\counterwithout{table}{section}
-\renewcommand\thetable{SI.\arabic{table}}
-\setcounter{table}{0}
-\pagenumbering{arabic} 
-\renewcommand\thepage{SI.\arabic{page}}
-\addtocontents{toc}{\protect\setcounter{tocdepth}{0}}
-
-\title{Supplemental Material for ``\papertitle''}
-\maketitle
-
-This document presents supplementary material for the manuscript ``\papertitle''.
-It contains the following sections:
-\begin{enumerate}
-    \item The derivation of the covariance matrix in the long-time limit for freely diffusion particles.
-    \item Discussion of the origin of bias and skew in the distribution of the estimated variance of the estimated diffusion coefficient, $\prob{\varest{\Dest}}$.
-    \item Details of the LLZO molecular dynamics simulations.
-    \item Details of the implementation of the model covariance method in the \textsc{kinisi} package.
-\end{enumerate}
-A repository containing the analysis and plotting code used to generate all results and figures in the main manuscript and this supplemental material document is available at \url{www.github.com/arm61/msd-errors}~\cite{mccluskey_github_2022}, under MIT (code) and CC BY-SA 4.0 (figures and text) licenses.
-This repository includes a fully reproducible \code{showyourwork} workflow, which allows complete reproduction of the analysis, plotting of figures and compilation of the manuscripts.
-The corresponding input datasets are openly available under the CC BY-SA 4.0 licence \cite{mccluskey_zenodo_2022,coles_llzo_zenodo_2022}.
-
-\twocolumngrid
-
-\section{Derivation of the long-time limit covariance matrix for a system of freely diffusing particles.}
-\label{sec:ran}
-In the main manuscript we present the result that the covariance matrix for a system of freely diffusing particles, in the long-time limit, has the form
-\begin{equation}
-  \Sigma^\prime\left[\oMSDi, \oMSDj\right]= \Sigma^\prime\left[\oMSDj, \oMSDi\right] =
-  \var{\oMSDi} \frac{\nind{i}}{\nind{j}},\hspace{1em} \forall\,i \leq j,
-  \label{equ:cvv_SI}
-\end{equation}
-where $\oMSDi$ is the observed mean-squared displacement (MSD) for time interval $i$ and $\nind{i}$ is the number of statistically independent observed squared displacements averaged over to compute the mean value.
-
-To derive this result, we first present a derivation of the expected variance for the MSD at timestep $i$, $\var{\oMSD}$, following the approach of Smith and Gillan~\cite{smith_random_1996}.
-We then derive an expression for the covariance $\Sigma^\prime\left[\oMSDi, \oMSDj\right]$ to obtain the result above.
-
-For a single particle undergoing a one-dimensional random walk with step size $\kappa$, each step gives a displacement $h = \pm \kappa$.
-After $n$ steps, the MSD, $\oMSDn$, is given by
-\begin{equation}
-  \begin{aligned}
-    \oMSDn &= \left[\sum_i^n h_i\right]^2\\
-           &= \sum_i^n\sum_j^n h_i h_j \\
-           &= \sum_i^n h_i^2 + \sum_i^n\sum_{j\neq i}^n h_ih_j.
-  \end{aligned}
-\end{equation}
-The expected MSD in the long-time limit, $\mathbb{E}(\oMSDn) = \moMSDn$, is obtained by averaging over all permutations of $h_i$ and $h_j$:
-\begin{equation}
-  \begin{aligned}
-    \moMSDn &= \sum_i^n \left<h_i^2\right> + \sum_i^n\sum_{j\neq i}^n\left<h_i h_j\right>.
-  \end{aligned}
-\end{equation}
-For a random walk, the second term averages to zero for all $h_i$ and $h_j$ (and where the walk has some temporal correlation, this introduces a scaling factor \cite{friauf_correlation_1962} to the total MSD), and
-\begin{equation}
-  \begin{aligned}
-    \moMSDn &= \sum_i^n \left<h_i^2\right> \\
-            &= n\kappa^2.
-  \end{aligned}
-  \label{equ:msd_SI}
-\end{equation}
-Hence the expected value for the mean-squared displacement increases linearly with the number of steps taken.
-In the event that the walk has some temporal correlation, as is the case for vacancy mediated diffusion, the second term does not average to zero, making Eqn.~\ref{equ:msd_SI} a proportional relationship~\cite{friauf_correlation_1962}.
-This proportionality would be carried through this derivation but would have no impact on the overall Bayesian regression method as the variances that are estimated from the simulation would include this scaling.
-
-The variance in the observed MSD, $\var{\oMSDn}$, is given by the standard statistical formula
-\begin{equation}
-    \var{\oMSDn} = \left<{\left[\oMSDn - \moMSDn\right]}^2\right>,
-\end{equation}
-which can be expanded as
-\begin{equation}
-    \begin{aligned}
-        \var{\oMSDn} &= \left<\oMSDn^2\right> - 2\moMSDn\left<\oMSDn\right> + \moMSDn^2,\\
-                     &= \left<\oMSDn^2\right> - \moMSDn^{\,2}.
-    \end{aligned}
-\end{equation}
-The first term can be expanded in terms of displacements $h$ as
-\begin{equation}
-  \left<\oMSDn^2\right> = \left<\sum_i^n\sum_j^n\sum_k^n\sum_l^n h_i h_j h_k h_l\right>,
-  \label{equ:big_av}
-\end{equation}
-which can be simplified by noting that $h_i$, $h_j$, $h_k$, and $h_l$ are uncorrelated when $i \neq j \neq k \neq l$, and the only terms that contribute to the average are those where $h_ih_jh_kh_l$ is guaranteed to be non-zero:
-\begin{itemize}
-    \item[(a)] $i = j = k = l$;
-    \item[(b)] $(i = j) \neq (k = l)$;
-    \item[(c)] $(i = k) \neq (j = l)$;
-    \item[(d)] $(i = l) \neq (j = k)$. 
-\end{itemize}
-From (a) we obtain
-\begin{equation}
-    \left<\sum_i^nh_i^4\right> = n\kappa^4,
-\end{equation}
-and from (b), (c), and (d), which are equivalent, we obtain
-\begin{equation}
-    \left<\sum_i^n\sum_j^nh_i^2h_j^2\right> = (n\kappa^2)^2 = n^2\kappa^4.
-\end{equation}
-This gives
-\begin{equation}
-    \left<\oMSDn^2\right> = (3n^2 + n)\kappa^4,
-\end{equation}
-which, in the limit $n \to \infty$, approaches
-\begin{equation}
-    \left<\oMSDn^2\right> = 3n^2\kappa^4.
-    \label{equ:infty}
-\end{equation}
-Combining this result with Eqn.~\ref{equ:msd_SI}, we can express the variance in the mean-squared displacement as
-\begin{equation}
-    \var{\oMSDn} = 3n^2\kappa^4 - n^2\kappa^4 = 2n^2\kappa^4,
-    \label{equ:varmsd}
-\end{equation}
-i.e., $\var{\oMSDn}$ increases quadratically with the number of steps taken, or, equivalently, with time.
-
-Eqn.~\ref{equ:varmsd} gives the variance of the mean squared displacement for a single particle considering a single time-origin.
-We can obtain improved statistics by averaging over statistically equivalent observed squared displacements (see Eqn.~2 in the main text), which can be achieved by averaging over mobile particles or by averaging over time origins.
-This averaging over equivalent observations reduces the variance in the observed MSD to
-\begin{equation}
-    \var{\oMSDn} = \frac{2n^2\kappa^4}{\nind{n}},
-\label{equ:der_var}
-\end{equation}
-where $\nind{n}$ is the total number of statistically independent (non-overlapping and accounting for concerted motion) squared displacements that contribute to $\oMSDi$.
-In the long-time limit, $\nind{n}$ is given by the product of the number of mobile particles, $N_{\textrm{atoms}}$, the number of numerically-independent sub-trajectories of lenth $i$ in our simulation trajectory, $N_t(t)$, and the ratio of the mean-squared displacement and the mean-square of the displacement of the centre of mass for all mobile particles,
-\begin{equation}
-    H_R(t_{\textrm{diff}}) = \frac{\sum^{N(t_{\textrm{diff}})}_{j=1}{\left[\Delta\mathbf{r}_j(t_{\textrm{diff}})\right]}^2}{\sum^{N_{t}(t_{\textrm{diff}})}_{l=1}\left|\left[\sum^{N_{\textrm{atoms}}}_{k=1}\Delta r_{k,l}(t_{\textrm{diff}})\right]^2\right|^2},
-\end{equation}
-at the start of the diffusive regime, $t_{\textrm{diff}}$.
-Note that $\nind{n}$ considers numerically-independent sub-trajectories, since mutually overlapping time-windows give correlated squared displacements.
-Where overlapping time-windows are used, Eqn.~\ref{equ:der_var} approximates the observed variance (the variance for overlapping and non-overlapping samples are not the same) with accuracy that increases as a function of time-interval length.
-The inclusion of $H_R(t_{\textrm{diff}})$ account for concerted motion between individual particles in the simulation~\cite{vargas_dynamic_2019,canepa_pushing_2022}, which is particularly important in solid and liquid electrolyte systems.
-The selection of $t_{\textrm{diff}}$ as the point at which to use $H_R$ is taken to maximise the number of observations of the centre-of-mass displacement. 
-This leads to the approximation in Eqn.~\ref{equ:varestMSD}.
-
-The results for a one-dimensional lattice above (Eqns.~\ref{equ:msd_SI} \&~\ref{equ:der_var}) can be extended to a $d$-dimensional lattice, to give
-\begin{equation}
-    \moMSDn_{d} = \sum^d{\frac{n\kappa^2}{d}} = n\kappa^2,
-\end{equation}
-with variance
-\begin{equation}
-    \var{\oMSDn}_d = \sum^d{\frac{2n^2\kappa^4}{d^2\nind{n}}} = \frac{2n^2\kappa^4}{d\nind{n}},
-\end{equation}
-Because each step is equally likely to move a particle along each of the $d$ dimensions, the term $n$ in Eqns.~\ref{equ:msd_SI} \&~\ref{equ:der_var} is replaced here with $n/d$.
-
-The analysis above can be extended to consider the covariance between two different numbers of steps, $n$ and $n+m$, in the random walk where the expected MSDs will be
-\begin{equation}
-    \begin{aligned}
-        \moMSDn &= n\kappa^2; \\
-        \moMSDm &= (n+m)\kappa^2.
-    \end{aligned}
-\end{equation}
-The covariance between these is defined as
-\begin{equation}
-    \Sigma \left[\oMSDn, \oMSDm \right] = \left<{\left[{\oMSDn - \moMSDn}\right] \left[{\oMSDm - \moMSDm}\right]}\right>, 
-\end{equation}
-which can be expanded as
-\begin{equation}
-    \begin{aligned}
-        \Sigma \left[\oMSDn, \oMSDm \right] = \big\langle & \oMSDn\oMSDm - \oMSDn\moMSDm \\
-        & - \moMSDn\oMSDm + \moMSDn\moMSDm\big\rangle,
-    \end{aligned}
-\end{equation}
-and then reformulated to give
-\begin{equation}
-    \Sigma \left[\oMSDn, \oMSDm \right] = \left<\oMSDn\oMSDm\right> - \moMSDn\moMSDm,
-    \label{equ:pair}
-\end{equation}
-where 
-\begin{equation}
-    \begin{aligned}
-        \moMSDn\moMSDm &= \oMSDn\oMSDm \\
-                       &= n\kappa^2(n+m)\kappa^2 \\
-                       &= n(n+m)\kappa^4
-    \end{aligned}
-\end{equation}
-and, by analogy to Eqn.~\ref{equ:big_av},
-\begin{equation}
-    \left<\oMSDn\oMSDm\right> = \left<\sum_i^n\sum_j^n\sum_k^{n+m}\sum_l^{n+m} h_i h_j h_k h_l\right>,
-\end{equation}
-which we can rewrite as
-\begin{equation}
-    \begin{aligned}
-        \left<\oMSDn\oMSDm\right> = \Bigg\langle & \sum_{i=1}^n\sum_{j=1}^n\sum_{k=1}^{n}\sum_{l=1}^{n} h_i h_j h_k h_l \\
-        & + \sum_{i=1}^n\sum_{j=1}^n\sum_{k=1}^{n}\sum_{l=n+1}^{n+m} h_i h_j h_k h_l \\
-        & + \sum_{i=1}^n\sum_{j=1}^n\sum_{k=n+1}^{n+m}\sum_{l=1}^{n} h_i h_j h_k h_l \\
-        & + \sum_{i=1}^n\sum_{j=1}^n\sum_{k=n+1}^{n+m}\sum_{l=n+1}^{n+m} h_i h_j h_k h_l \Bigg\rangle.
-    \end{aligned}
-    \label{equ:vbig}
-\end{equation}
-The second and third terms in Eqn.~\ref{equ:vbig} tend to zero as there is an equal probability of positive and negative displacements. 
-This reduces Eqn.~\ref{equ:vbig} to 
-\begin{equation}
-    \begin{aligned}
-        \left<\oMSDn\oMSDm\right> = & \Bigg\langle \sum_{i=1}^n\sum_{j=1}^n\sum_{k=1}^{n}\sum_{l=1}^{n} h_i h_j h_k h_l \Bigg\rangle \\
-        & + \Bigg\langle \sum_{i=1}^n\sum_{j=1}^n\sum_{k=n+1}^{n+m}\sum_{l=n+1}^{n+m} h_i h_j h_k h_l \Bigg\rangle, 
-    \end{aligned}
-\end{equation}
-and using Eqn.~\ref{equ:infty} gives
-\begin{equation}
-    \left<\oMSDn\oMSDm\right> = 3n^2\kappa^4 + \Bigg\langle \sum_{i=1}^n\sum_{j=1}^n\sum_{k=n+1}^{n+m}\sum_{l=n+1}^{n+m} h_i h_j h_k h_l \Bigg\rangle.
-\end{equation}
-We can rewrite this as
-\begin{equation}
-    \left<\oMSDn\oMSDm\right> = 3n^2\kappa^4 + \Bigg\langle \sum_{i=1}^n\sum_{j=1}^n h_i h_j \Bigg\rangle \Bigg\langle \sum_{k=n+1}^{n+m}\sum_{l=n+1}^{n+m} h_k h_l \Bigg\rangle,
-\end{equation}
-where the following holds,
-\begin{equation}
-    \begin{aligned}
-        \left<\oMSDn\oMSDm\right> &= 3n^2\kappa^4 + n\kappa^2 m\kappa^2 \\
-                                  &= 3n\kappa^4 + nm\kappa^4.
-    \end{aligned}
-\end{equation}
-Putting this result into Eqn.~\ref{equ:pair} allows the covariance to be written as
-\begin{equation}
-    \begin{aligned}
-    \Sigma^\prime \left[\oMSDn, \oMSDm \right] &= 3n^2\kappa^4 + nm\kappa^4 - n(n+m)\kappa^4 \\
-                                        &= 3n^2\kappa^4 - n^2\kappa^4 = 2n^2\kappa^4,
-    \end{aligned}
-    \label{equ:cov_der}
-\end{equation}
-where we use the $\Sigma^\prime$ notation to identify that this is in the long-time limit.
-
-In this case, the covariance depends only on the number of overlapping points, $n$, between the two time intervals. 
-We can rationalise this by noting that for a random walk any numerically-independent points will be completely uncorrelated and therefore have a covariance of \num{0}. 
-Similar to the case for the variance, the covariance derived in Eqn.~\ref{equ:cov_der} is that for a single particle at a single time origin. 
-The number of independent observed squared displacements for a given covariance should be the minimum number of shared independent observed squared displacements between the two time intervals, which is $\nind{n+m}$.
-Therefore, the covariance, scaled by the number of contributing independent observations, in the long-time limit, is
-\begin{equation}
-    \Sigma^\prime \left[\oMSDn, \oMSDm \right] = \frac{2n^2\kappa^4}{\nind{n+m}}.
-\end{equation}
-Similar to the MSD and the variance, the covariance can be written for $d$-dimensions as 
-\begin{equation}
-    \Sigma^\prime \left[\oMSDn, \oMSDm \right] = \frac{2n^2\kappa^4}{d\nind{n+m}}.
-\end{equation}
-The covariance can be calculated directly from the variance by recognising that both depend on the number of overlapping points, $n$, as follows
-\begin{equation}
-    \Sigma^\prime \left[\oMSDn, \oMSDm \right] = \var{\oMSDn}\frac{\nind{n}}{\nind{n+m}}.
-\end{equation}
-This is then rewritten in terms of $i$ and $j$ to give, Eqn.~\ref{equ:cvv_SI}.
-
-Using the equivalence of $2d\D t \equiv n\kappa^2$~\cite{howard_reports_1964}, Eqns.~\ref{equ:msd_SI} \&~\ref{equ:cvv_SI} can be rewritten in terms of $t$ (or $t_1$ and $t_2$) and the diffusion coefficient, for any dimensionality of lattice random walk,
-\begin{equation}
-    \oMSDs{t} = 2d \D t,
-    \label{equ:der_msd_fick}
-\end{equation}
-and 
-\begin{equation}
-    \Sigma^\prime \left[\oMSDs{t_1}, \oMSDs{t_2} \right] = 8d{(\D)}^2 t_1^2\frac{N^\prime(t_2)}{N^\prime(t_1)},\hspace{1em} \forall\,t_1 \leq t_2.
-\end{equation}
-
-\section{Comparison With Block Averaging Approach}
-\label{sec:block}
-Eqn.~\ref{equ:varestMSD} enables the approximate estimation of the variance in $\oMSDi$, as mentioned in the main text, it is also possible to obtain this from a block averaging approach~\cite{flyvbjerg_error_1989,frenkel_understanding_2023}.
-Fig.~\ref{fig:pyblock} compares the $\D$ estimate distributions using variances from approximated by Eqn.~\ref{equ:varestMSD} with those from the block averaging approach (using the \textsc{pyblock} Python package~\cite{spencer_pyblock_2020}) both with the fitting of the variances to Eqn.~\ref{equ:variance} and without.
-When the fitting is not performed, the resulting covariance matrix is numerically unstable, leading to estimated values of $\D$ at extreme values.
-While when the fitting is performed and the blocking-estimated variances as used, there is no improvement in the estimation of $\D$ and the resulting distribution of estimates in the variance of $\prob{\D}$ is broader than when Eqn.~\ref{equ:varestMSD} is used.
-
-\begin{figure}
-    \centering
-    % \resizebox{\columnwidth}{!}{\includegraphics*{figure1.pdf}}
-    \includegraphics[width=\columnwidth]{pyblock.pdf}
-    \script{pyblock.py}
-    \caption{Comparison of the distribution of $\Dest$ (a, c, \& e) and the estimated variances (b, d, \& f) from 4096 individual random walk simulations using variances approximated by Eqn.~\ref{equ:varestMSD} (a \& b), from block averaging with fitting to Eqn.~\ref{equ:variance} (c \& d), and without fitting (e \& f).
-Note that the bounds in a, c, \& e are defined by the extrema of the estimated values, showing the fact that without the fitting to Eqn.~\ref{equ:variance}, the resulting covariance matrix is numerically unstable.}
-    \label{fig:pyblock}
-\end{figure}
-
-\section{Bias and skew in $\prob{\varest{\Dest}}$}
-\label{sec:skew}
-In the main manuscript, we present results for a set of \num{4096} 3D-lattice random walk simulations, each consisting of \num{128} particles undergoing \num{128} steps (Fig.~\ref{fig:random_walk}).
-Our approximate Bayesian regression scheme allows us to estimate the variance in $\Dest$, denoted as $\varest{\Dest}$, that would be obtained over a large number of repeat simulations.
-This estimate is calculated from the variance of the marginal posterior distribution $\prob{\D|\model}$, which we derive from analysis of a single simulation trajectory.
-As shown in Fig.~\ref{fig:random_walk}d, our estimate for the population variance $\var{\Dest}$, obtained from a single simulation, aligns reasonably with the true value.
-When considering the distribution of estimated variance, $\prob{\varest{\Dest}}$, however, we observe a systematic overestimation (bias) of the true value, along with visible skewness.
-
-This bias and skew arise from our use of estimated variances $\varest{\oMSDi}$ when parametrising the model covariance matrix $\bm{\Sigma^\prime}$.
-Fig.~\ref{fig:true_cov} presents equivalent results for $\prob{\Dest}$ and $\prob{\varest{\Dest}}$ for the same \num{4096} individual simulations, but calculated using a numerical covariance matrix, $\bm{\Sigma}_\mathrm{num}$ derived from all \num{4096} observed MSDs.
-The resulting distribution $\prob{\varest{\Dest}}$ (Fig.~\ref{fig:true_cov}b) is no longer biased or skewed.
-Furthermore, the distribution $\prob{\Dest}$ agrees even more closely with the numerically converged distribution obtained when combining data from all \num{4096} simulations (Fig~\ref{fig:true_cov}a), contrasting with the results presented in Fig.~\ref{fig:random_walk}b, where our approximate Bayesian regression scheme yields a slightly broadened distribution due to the use of the long-time limit in the derivation of the analytical form for $\bm{\Sigma^\prime}$.
-\begin{figure}
-    \centering
-    % \resizebox{\columnwidth}{!}{\includegraphics*{figure7.pdf}}
-    \includegraphics[width=\columnwidth]{true_cov.pdf}
-    \script{true_cov.py}
-    \caption{
-        (a) Probability distribution of point-estimates $\prob{\Dest}$ obtained from \num{4096} individual random-walk simulations, using the numerical covariance matrix $\bm{\Sigma}_\mathrm{num}$.
-        Each simulation has been analysed as in Fig.~~\ref{equ:observed_msd}(a) and (b) to yield a single corresponding point estimate $\Dest$.
-        The grey line shows the distribution of point estimates, $\prob{\Dest_\mathrm{num}}$, obtained using Bayesian regression with a mean vector and numerical covariance matrix derived from the complete dataset of all \num{4096} simulations.
-        The pink horizontal bar shows an interval of one standard deviation in $\prob{\Dest}$.
-        (b) Probability distribution of estimated variances, $\varest{\Dest}$, for individual random-walk simulations, using the numerical covariance matrix $\bm{\Sigma}_\mathrm{num}$, compared to the true sample variance (pink vertical line) $\var{\Dest}$.
-        }
-    \label{fig:true_cov}
-\end{figure}
-
-\section{Evaluation of OLS, WLS, and GLS for LLZO System}
-
-In Fig.~\ref{fig:glswlsols}, it is shown that for a 3D lattice random walk the heteroscedastic and correlated nature of the data requires generalised least squares to optimally estimate the true ensemble-average MSD, and therefore estimate $\D$.
-This is also true for the real materials, such as the \ce{Li7La3Zr2O12} (LLZO) investigated by classical molecular dynamics simulation in this work (Fig.~\ref{fig:glswlsols_llzo}).
-Once again, GLS (or, indeed, the Bayesian regression equivalent, shown in Fig.~\ref{fig:diffusion}) offers both the most statistically efficient estimation of $\D$ and is required to obtain an accruate estimate of the statistical uncertainty in $\D$ from a single simulation.
-\begin{figure}
-    \centering
-    % \resizebox{\columnwidth}{!}{\includegraphics*{figure1.pdf}}
-    \includegraphics[width=\columnwidth]{glswlsols_llzo.pdf}
-    \script{glswlsols_llzo.py}
-    \caption{
-        Example distributions of estimated self-diffusion coefficients, $\Dest$, calculated using (a) ordinary least squares (OLS), (b) weighted least squares (WLS), and (c) generalised least squares (GLS),
-        from MSD data from \num{512} effective simulations of LLZO of \SI{\sim 25}{\pico\second} with 56 lithium ions.
-        In each panel, the grey curve shows the best-fit normal distribution for the simulation data, the upper horizontal bar shows the standard deviation of this distribution, and the lower horizontal bar shows the average estimated standard distribution given by the analytical expression for $\sigma[\prob{\Dest}]$ for each regression method.}
-    \label{fig:glswlsols_llzo}
-\end{figure}
 

### 2024-09-13 src/tex/ms.tex
Update src/tex/ms.tex

@@
     \centering
-    \resizebox{\columnwidth}{!}{\includegraphics*{figure1.pdf}}
+    % \resizebox{\columnwidth}{!}{\includegraphics*{figure1.pdf}}
+    \includegraphics[width=\columnwidth]{glswlsols.pdf}
     \caption{
@@
     \centering
-    \resizebox{\columnwidth}{!}{\includegraphics*{figure2.pdf}}
+    % \resizebox{\columnwidth}{!}{\includegraphics*{figure2.pdf}}
+    \includegraphics[width=\columnwidth]{msd.pdf}
     \caption{
@@
     \centering
-    \resizebox{\columnwidth}{!}{\includegraphics*{figure3.pdf}}
+    % \resizebox{\columnwidth}{!}{\includegraphics*{figure3.pdf}}
+    \includegraphics[width=\columnwidth]{covariances.pdf}
     \caption{
@@
     \centering
-    \resizebox{\textwidth}{!}{\includegraphics*{figure4.pdf}}
+    % \resizebox{\textwidth}{!}{\includegraphics*{figure4.pdf}}
+    \includegraphics[width=\textwidth]{random_walk.pdf}
     \caption{
@@
     \centering
-    \resizebox{\columnwidth}{!}{\includegraphics*{figure5.pdf}}
+    % \resizebox{\columnwidth}{!}{\includegraphics*{figure5.pdf}}
+    \includegraphics[width=\columnwidth]{diffusion.pdf}
     \caption{
@@
     \centering
-    \resizebox{\columnwidth}{!}{\includegraphics*{figure6.pdf}}
+    % \resizebox{\columnwidth}{!}{\includegraphics*{figure6.pdf}}
+    \includegraphics[width=\columnwidth]{stat_eff.pdf}
     \caption{
@@
 Electronic Supplementary Information (ESI) available: A complete set of analysis/plotting scripts allowing for a fully reproducible and automated analysis workflow, using \textsc{showyourwork}~\cite{luger_showyourwork_2021}, for this work is available at Ref.~\cite{mccluskey_github_2022} under an MIT license.
-All raw simulation files are available on Zenodo shared under CC BY-SA 4.0 licences, the random walk simulations and other analysis-linked data~\cite{mccluskey_zenodo_2022} and the LLZO raw simulation trajectories~\cite{coles_llzo_zenodo_2022}.
+The LLZO raw simulation trajectories are available on Zenodo shared under a CC BY-SA 4.0 licence~\cite{coles_llzo_zenodo_2022}.
 The method outlined in this work is implemented in the open-source Python package \textsc{kinisi}~\cite{mccluskey_kinisi_2022}, which is available under an MIT license.

### 2024-10-29 src/tex/ms.tex
Update tex and bib following review

@@
 \newcommand{\varest}[1]{\ensuremath{\widehat{\sigma}^2[#1]}}
+\newcommand*\xbar[1]{
+  \hbox{
+    \vbox{
+      \hrule height 0.5pt 
+      \kern0.5ex
+      \hbox{
+        \kern-0.1em
+        \ensuremath{#1}
+        \kern-0.1em
+      }
+    }
+  }
+} 
 \DeclareMathOperator{\cov}{cov}
 \DeclareMathOperator*{\argmax}{argmax}
+
+\graphicspath{{figures/}}
 
@@
 
-\section*{Significance Statement}
-Quantifying diffusion in liquid and solid-state systems is fundamental to our understanding of their properties—without the ability to accurately measure diffusion coefficients, robust insight into these systems is impossible.
-Here, we present a statistically efficient approach to estimating self-diffusion coefficients from single molecular dynamics simulations, using approximate Bayesian regression.
-Our method provides estimates of self-diffusion coefficients with significantly less uncertainty than textbook ordinary-least-squares or weighted-least-squares methods, allowing for more efficient predictions of diffusion behaviour from simulations at greatly reduced overall computational cost, and also gives an accurate estimate of the variance in the predicted value, allowing for meaningful downstream comparison of data.
-
 \section{Introduction}
@@
 \end{equation}
-where $t$ is elapsed time.
+where $\Delta\mathbf{r}{(t)}$ is the displacement of a diffusion particle in the time interval $t$.
 Because numerical simulations are finite in time and space, MSDs obtained from simulation data always differ from the true ensemble average MSD.
-One can, however, compute an estimate of the self-diffusion coefficient, $\Dest$, by fitting a linear model to the observed MSD and using the gradient of this fitted model in place of $\MSD{t} / t$ in \cref{equ:einstein}.
+One can, however, compute an estimate of the self-diffusion coefficient, $\Dest$, by fitting a linear model to the observed MSD and using the gradient of this fitted model in place of $\MSD{t} / t$ in \cref{equ:einstein}~\footnote{While estimating $\D$ by fitting a linear model to MSD data is, perhaps, the most commonly used approach in the literature, alternative methods that avoid linear fitting of MSD data also exist; see, for example, Refs.~{\protect\cite{VestergaardEtAl_PhysRevE2014,KrapfEtAl_NewJournalPhysics2018,bullerjohn_optimal_2020}}.}.
 
@@
 The estimated variance $\varest{\oMSD}$ can be calculated from a single simulation trajectory, and provides an accurate estimate of the true variance $\var{\oMSD}$~\footnote{A common approach to estimating the variance of the mean of time-correlated data with unknown correlation time is the renormalization group blocking method of Flyvberg and Peterson~\cite{FlyvbjergAndPetersen_JChemPhys1989,Frenkel2023-ah}.
-In the SI, we compare this blocking method to our direct rescaling method (\cref{equ:variance}) for estimating $\var{\oMSDi}$ from a set of random-walk trajectories.
+In the SI, we compare this blocking method to our direct rescaling method (\cref{equ:varestMSD}) for estimating $\var{\oMSDi}$ from a set of random-walk trajectories.
 For our example data, direct rescaling performs better (gives more accurate estimates of $\var{\oMSDi}$).}.
@@
 
-\section*{Data availability}
+\section*{Supporting Information}
 
-Electronic Supplementary Information (ESI) available: A complete set of analysis/plotting scripts allowing for a fully reproducible and automated analysis workflow, using \textsc{showyourwork}~\cite{luger_showyourwork_2021}, for this work is available at Ref.~\cite{mccluskey_github_2022} under an MIT license.
+Derivation of long-time limit covariance matrix for a system of freely diffusing particles, comparison of variance rescaling (\cref{equ:varestMSD}) and block renormalisation approaches, discussion of bias in the distribution of estimated variance of the estimated diffusion coefficient, and a comparison of OLS, WLS, and GLS as estimators for $\D$ applied to MSD data from simulations of \ce{Li7La3Zr2O12} (LLZO).
+Additional Electronic Supplementary Information (ESI) available at Ref.~\cite{mccluskey_github_2022} under an MIT license: A complete set of analysis/plotting scripts allowing for a fully reproducible and automated analysis workflow, using \textsc{showyourwork}~\cite{luger_showyourwork_2021}.
 The LLZO raw simulation trajectories are available on Zenodo shared under a CC BY-SA 4.0 licence~\cite{coles_llzo_zenodo_2022}.
@@
 
+\let\addcontentsline\oldaddcontentsline
+
+ \onecolumngrid
+ \clearpage 
+ \twocolumngrid
+
+ \appendix
+ \renewcommand\thesection{SI.\Roman{section}}
+ \counterwithout{figure}{section}
+ \renewcommand\thefigure{SI.\arabic{figure}}
+ \setcounter{figure}{0}
+ \counterwithout{equation}{section}
+ \renewcommand\theequation{SI.\arabic{equation}}
+ \setcounter{equation}{0}
+ \counterwithout{table}{section}
+ \renewcommand\thetable{SI.\arabic{table}}
+ \setcounter{table}{0}
+ \pagenumbering{arabic} 
+ \renewcommand\thepage{SI.\arabic{page}}
+ \addtocontents{toc}{\protect\setcounter{tocdepth}{0}}
+
+ \title{Supplemental Material for ``\papertitle''}
+ \maketitle
+ This document presents supplementary material for the manuscript ``\papertitle''.
+It contains the following sections:
+\renewcommand{\labelitemi}{}
+\begin{itemize}
+    \item \ref{sec:ran}. The derivation of the covariance matrix in the long-time limit for freely diffusion particles.
+    \item \ref{sec:var_est}.
+Further details of the variance rescaling method for estimating $\var{\oMSDi}$, described in the main text, and a comparison to the block renormalisation method for estimating the variance of the mean for serially-correlated data of Flyvbjerg and Petersen~\cite{FlyvbjergAndPetersen_JChemPhys1989}.
+    \item \ref{sec:bias}. Discussion of the origin of bias in the distribution of the estimated variance of the estimated diffusion coefficient, $\prob{\varest{\Dest}}$.
+    \item \ref{sec:llzo}. A comparison of OLS, WLS, and GLS as estimators for $\D$ applied to MSD data from simulations of \ce{Li7La3Zr2O12} (LLZO).
+\end{itemize}
+A repository containing the analysis and plotting code used to generate all results and figures in the main manuscript and this supplemental material document is available at \url{www.github.com/arm61/msd-errors}~\cite{mccluskey_github_2022}, under MIT (code) and CC BY-SA 4.0 (figures and text) licenses.
+This repository includes a fully reproducible \code{showyourwork} workflow, which allows complete reproduction of the analysis, plotting of figures and compilation of the manuscript{}s.
+The corresponding LLZO simulation datasets are openly available under the CC BY-SA 4.0 licence \cite{coles_llzo_zenodo_2022}.
+
+\twocolumngrid
+
+\section{Derivation of the long-time limit covariance matrix for a system of freely diffusing particles.}
+\label{sec:ran}
+In the main manuscript we present the result that the covariance matrix for a system of freely diffusing particles, in the long-time limit, has the form
+\begin{equation}
+  \Sigma^\prime\left[\oMSDi, \oMSDj\right]= \Sigma^\prime\left[\oMSDj, \oMSDi\right] =
+  \var{\oMSDi} \frac{\nind{i}}{\nind{j}},\hspace{1em} \forall\,i \leq j,
+  \label{equ:cvv_SI}
+\end{equation}
+where $\oMSDi$ is the observed mean-squared displacement (MSD) for time interval $i$ and $\nind{i}$ is the number of statistically independent observed squared displacements averaged over to compute the mean value.
+
+To derive this result, we first present a derivation of the expected variance for the MSD at timestep $i$, $\var{\oMSD}$, following the approach of Smith and Gillan~\cite{smith_random_1996}.
+We then derive an expression for the covariance $\Sigma^\prime\left[\oMSDi, \oMSDj\right]$ to obtain the result above.
+
+For a single particle undergoing a one-dimensional random walk with step size $\kappa$, each step gives a displacement $h = \pm \kappa$.
+After $n$ steps, the MSD, $\oMSDn$, is given by
+\begin{equation}
+  \begin{aligned}
+    \oMSDn &= \left[\sum_i^n h_i\right]^2\\
+           &= \sum_i^n\sum_j^n h_i h_j \\
+           &= \sum_i^n h_i^2 + \sum_i^n\sum_{j\neq i}^n h_ih_j.
+  \end{aligned}
+\end{equation}
+The expected MSD in the long-time limit, $\mathbb{E}(\oMSDn) = \moMSDn$, is obtained by averaging over all permutations of $h_i$ and $h_j$:
+\begin{equation}
+  \begin{aligned}
+    \moMSDn &= \sum_i^n \left<h_i^2\right> + \sum_i^n\sum_{j\neq i}^n\left<h_i h_j\right>.
+  \end{aligned}
+\end{equation}
+For a random walk, the second term averages to zero for all $h_i$ and $h_j$, and
+\begin{equation}
+  \begin{aligned}
+    \moMSDn &= \sum_i^n \left<h_i^2\right> \\
+            &= n\kappa^2.
+  \end{aligned}
+  \label{equ:msd_SI}
+\end{equation}
+Hence the expected value for the mean-squared displacement increases linearly with the number of steps taken.
+
+The variance in the observed MSD, $\var{\oMSDn}$, is given by the standard statistical formula
+\begin{equation}
+    \var{\oMSDn} = \left<{\left[\oMSDn - \moMSDn\right]}^2\right>,
+\end{equation}
+which can be expanded as
+\begin{equation}
+    \begin{aligned}
+        \var{\oMSDn} &= \left<\oMSDn^2\right> - 2\moMSDn\left<\oMSDn\right> + \moMSDn^2,\\
+                     &= \left<\oMSDn^2\right> - \moMSDn^{\,2}.
+    \end{aligned}
+\end{equation}
+The first term can be expanded in terms of displacements $h$ as
+\begin{equation}
+  \left<\oMSDn^2\right> = \left<\sum_i^n\sum_j^n\sum_k^n\sum_l^n h_i h_j h_k h_l\right>,
+  \label{equ:big_av}
+\end{equation}
+which can be simplified by noting that $h_i$, $h_j$, $h_k$, and $h_l$ are uncorrelated when $i \neq j \neq k \neq l$, and the only terms that contribute to the average are those where $h_ih_jh_kh_l$ is guaranteed to be non-zero:
+\begin{itemize}
+    \item[(a)] $i = j = k = l$;
+    \item[(b)] $(i = j) \neq (k = l)$;
+    \item[(c)] $(i = k) \neq (j = l)$;
+    \item[(d)] $(i = l) \neq (j = k)$. 
+\end{itemize}
+From (a) we obtain
+\begin{equation}
+    \left<\sum_i^nh_i^4\right> = n\kappa^4,
+\end{equation}
+and from (b), (c), and (d), which are equivalent, we obtain
+\begin{equation}
+    \left<\sum_i^n\sum_j^nh_i^2h_j^2\right> = (n\kappa^2)^2 = n^2\kappa^4.
+\end{equation}
+This gives
+\begin{equation}
+    \left<\oMSDn^2\right> = (3n^2 + n)\kappa^4,
+\end{equation}
+which, in the limit $n \to \infty$, approaches
+\begin{equation}
+    \left<\oMSDn^2\right> = 3n^2\kappa^4.
+    \label{equ:infty}
+\end{equation}
+Combining this result with \cref{equ:msd_SI}, we can express the variance in the mean-squared displacement as
+\begin{equation}
+    \var{\oMSDn} = 3n^2\kappa^4 - n^2\kappa^4 = 2n^2\kappa^4,
+    \label{equ:varmsd}
+\end{equation}
+i.e., $\var{\oMSDn}$ increases quadratically with the number of steps taken, or, equivalently, with time.
+
+Equation~\ref{equ:varmsd} gives the variance of the mean squared displacement for a single particle considering a single time-origin.
+We can obtain improved statistics by averaging over statistically equivalent observed squared displacements (see Eqn.~2 in the main text), which can be achieved by averaging over mobile particles or by averaging over time origins.
+This averaging over equivalent observations reduces the variance in the observed MSD to
+\begin{equation}
+    \var{\oMSDn} = \frac{2n^2\kappa^4}{\nind{n}},
+\label{equ:der_var}
+\end{equation}
+where $\nind{n}$ is the total number of statistically independent squared displacements that contribute to $\oMSDi$.
+In the long-time limit, $\nind{n}$ is given by the product of the number of mobile particles and the number of non-overlapping time-windows of length $i$ in our simulation trajectory.
+Note that $\nind{n}$ considers non-overlapping time windows, since mutually overlapping time-windows give correlated squared displacements.
+
+The results for a one-dimensional lattice above (\cref{equ:msd_SI,equ:der_var}) can be extended to a $d$-dimensional lattice, to give
+\begin{equation}
+    \moMSDn_{d} = \sum^d{\frac{n\kappa^2}{d}} = n\kappa^2,
+\end{equation}
+with variance
+\begin{equation}
+    \var{\oMSDn}_d = \sum^d{\frac{2n^2\kappa^4}{d^2\nind{n}}} = \frac{2n^2\kappa^4}{d\nind{n}},
+\end{equation}
+Because each step is equally likely to move a particle along each of the $d$ dimensions, the term $n$ in \cref{equ:msd_SI,equ:der_var} is replaced here with $n/d$.
+
+The analysis above can be extended to consider the covariance between two different numbers of steps, $n$ and $n+m$, in the random walk where the expected MSDs will be
+\begin{equation}
+    \begin{aligned}
+        \moMSDn &= n\kappa^2; \\
+        \moMSDm &= (n+m)\kappa^2.
+    \end{aligned}
+\end{equation}
+The covariance between these is defined as
+\begin{equation}
+    \Sigma \left[\oMSDn, \oMSDm \right] = \left<{\left[{\oMSDn - \moMSDn}\right] \left[{\oMSDm - \moMSDm}\right]}\right>, 
+\end{equation}
+which can be expanded as
+\begin{equation}
+    \begin{aligned}
+        \Sigma \left[\oMSDn, \oMSDm \right] = \big\langle & \oMSDn\oMSDm - \oMSDn\moMSDm \\
+        & - \moMSDn\oMSDm + \moMSDn\moMSDm\big\rangle,
+    \end{aligned}
+\end{equation}
+and then reformulated to give
+\begin{equation}
+    \Sigma \left[\oMSDn, \oMSDm \right] = \left<\oMSDn\oMSDm\right> - \moMSDn\moMSDm,
+    \label{equ:pair}
+\end{equation}
+where 
+\begin{equation}
+    \begin{aligned}
+        \moMSDn\moMSDm &= \oMSDn\oMSDm \\
+                       &= n\kappa^2(n+m)\kappa^2 \\
+                       &= n(n+m)\kappa^4
+    \end{aligned}
+\end{equation}
+and, by analogy to \cref{equ:big_av},
+\begin{equation}
+    \left<\oMSDn\oMSDm\right> = \left<\sum_i^n\sum_j^n\sum_k^{n+m}\sum_l^{n+m} h_i h_j h_k h_l\right>,
+\end{equation}
+which we can rewrite as
+\begin{equation}
+    \begin{aligned}
+        \left<\oMSDn\oMSDm\right> = \Bigg\langle & \sum_{i=1}^n\sum_{j=1}^n\sum_{k=1}^{n}\sum_{l=1}^{n} h_i h_j h_k h_l \\
+        & + \sum_{i=1}^n\sum_{j=1}^n\sum_{k=1}^{n}\sum_{l=n+1}^{n+m} h_i h_j h_k h_l \\
+        & + \sum_{i=1}^n\sum_{j=1}^n\sum_{k=n+1}^{n+m}\sum_{l=1}^{n} h_i h_j h_k h_l \\
+        & + \sum_{i=1}^n\sum_{j=1}^n\sum_{k=n+1}^{n+m}\sum_{l=n+1}^{n+m} h_i h_j h_k h_l \Bigg\rangle.
+    \end{aligned}
+    \label{equ:vbig}
+\end{equation}
+The second and third terms in \cref{equ:vbig} tend to zero as there is an equal probability of positive and negative displacements. 
+This reduces \cref{equ:vbig} to 
+\begin{equation}
+    \begin{aligned}
+        \left<\oMSDn\oMSDm\right> = & \Bigg\langle \sum_{i=1}^n\sum_{j=1}^n\sum_{k=1}^{n}\sum_{l=1}^{n} h_i h_j h_k h_l \Bigg\rangle \\
+        & + \Bigg\langle \sum_{i=1}^n\sum_{j=1}^n\sum_{k=n+1}^{n+m}\sum_{l=n+1}^{n+m} h_i h_j h_k h_l \Bigg\rangle, 
+    \end{aligned}
+\end{equation}
+and using \cref{equ:infty} gives
+\begin{equation}
+    \left<\oMSDn\oMSDm\right> = 3n^2\kappa^4 + \Bigg\langle \sum_{i=1}^n\sum_{j=1}^n\sum_{k=n+1}^{n+m}\sum_{l=n+1}^{n+m} h_i h_j h_k h_l \Bigg\rangle.
+\end{equation}
+We can rewrite this as
+\begin{equation}
+    \left<\oMSDn\oMSDm\right> = 3n^2\kappa^4 + \Bigg\langle \sum_{i=1}^n\sum_{j=1}^n h_i h_j \Bigg\rangle \Bigg\langle \sum_{k=n+1}^{n+m}\sum_{l=n+1}^{n+m} h_k h_l \Bigg\rangle,
+\end{equation}
+where the following holds,
+\begin{equation}
+    \begin{aligned}
+        \left<\oMSDn\oMSDm\right> &= 3n^2\kappa^4 + n\kappa^2 m\kappa^2 \\
+                                  &= 3n\kappa^4 + nm\kappa^4.
+    \end{aligned}
+\end{equation}
+Putting this result into \cref{equ:pair} allows the covariance to be written as
+\begin{equation}
+    \begin{aligned}
+    \Sigma^\prime \left[\oMSDn, \oMSDm \right] &= 3n^2\kappa^4 + nm\kappa^4 - n(n+m)\kappa^4 \\
+                                        &= 3n^2\kappa^4 - n^2\kappa^4 = 2n^2\kappa^4,
+    \end{aligned}
+    \label{equ:cov_der}
+\end{equation}
+where we use the $\Sigma^\prime$ notation to identify that this is in the long-time limit.
+
+In this case, the covariance depends only on the number of overlapping points, $n$, between the two time intervals. 
+We can rationalise this by noting that for a random walk any non-overlapping points will be completely uncorrelated and therefore have a covariance of \num{0}. 
+Similar to the case for the variance, the covariance derived in \cref{equ:cov_der} is that for a single particle at a single time origin. 
+The number of independent observed squared displacements for a given covariance should be the minimum number of shared independent observed squared displacements between the two time intervals, which is $\nind{n+m}$.
+Therefore, the covariance, scaled by the number of contributing independent observations, in the long-time limit, is
+\begin{equation}
+    \Sigma^\prime \left[\oMSDn, \oMSDm \right] = \frac{2n^2\kappa^4}{\nind{n+m}}.
+\end{equation}
+Similar to the MSD and the variance, the covariance can be written for $d$-dimensions as 
+\begin{equation}
+    \Sigma^\prime \left[\oMSDn, \oMSDm \right] = \frac{2n^2\kappa^4}{d\nind{n+m}}.
+\end{equation}
+The covariance can be calculated directly from the variance by recognising that both depend on the number of overlapping points, $n$, as follows
+\begin{equation}
+    \Sigma^\prime \left[\oMSDn, \oMSDm \right] = \var{\oMSDn}\frac{\nind{n}}{\nind{n+m}}.
+\end{equation}
+This is then rewritten in terms of $i$ and $j$ to give, \cref{equ:cvv_SI}.
+
+Using the equivalence of $2d\D t \equiv n\kappa^2$~\cite{howard_reports_1964}, \cref{equ:msd_SI,equ:cvv_SI} can be rewritten in terms of $t$ (or $t_1$ and $t_2$) and the diffusion coefficient, for any dimensionality of lattice random walk,
+\begin{equation}
+    \oMSDs{t} = 2d \D t,
+    \label{equ:der_msd_fick}
+\end{equation}
+and 
+\begin{equation}
+    \Sigma^\prime \left[\oMSDs{t_1}, \oMSDs{t_2} \right] = 8d{(\D)}^2 t_1^2\frac{N^\prime(t_2)}{N^\prime(t_1)},\hspace{1em} \forall\,t_1 \leq t_2.
+\end{equation}
+
+\section{Estimating $\var{\oMSDi}$: variance rescaling versus block renormalisation}
+\label{sec:var_est}
+The approximate Bayesian regression scheme described in our main manuscript uses a model covariance matrix parametrised by the variance of the observed MSD as a function of time, denoted $\var{\oMSDi}$.
+Generally, $\var{\oMSDi}$ is unknown and must be estimated from the input simulation data.
+
+\begin{figure}[tb]
+  \centering
+  % \resizebox{9cm}{!}{\includegraphics*{figure1_SI.pdf}}  
+  \includegraphics[width=\columnwidth]{figure1_SI.pdf} 
+    \caption{\label{fig:displacements}Schematic showing the construction of three statistically-equivalent sub-trajectories of length $\Delta t = 8$ from a full simulation trajectory.
+(a) The full trajectory consists of a series of displacements, $h_i$, from each simulation timestep.
+(b) Three exemplar sub-trajectories of length $\Delta t = 8$.
+Even in the case of a random walk (i.e., $h=\pm\kappa$), $\Delta r_1$ and $\Delta r_5$ are necessarily correlated, because both summations include $[h_5 + h_6 + h_7 + h_8]$.
+$\Delta r_5$ and $\Delta r_9$ are also obtained from mutually overlapping time windows and are therefore also correlated.
+For a random walk, $\Delta r_1$ and $\Delta r_9$ are uncorrelated.
+For a non-random (correlated) walk, non-overlapping time windows are uncorrelated in the limit that the window length, $\Delta t$, is much greater than the correlation time in $h$, which is the case if we are in the long-time linear region of the MSD where Eqn.~1 is valid.\footnote{The result that displacements calculated from overlapping time windows are correlated has been reported as an empirical result for simulations of a Lennard--Jones fluid in Ref.~\onlinecite{PranamiAndLamm_JChemTheoryComput2015}.}}
+\end{figure}
+
+In the main manuscript, we describe an estimation approach that involves rescaling the observed variance of the squared displacement for time interval $i$ by the number of numerically-independent contributing sub-trajectories, $\nind{i}$ (Eqn.~7).
+We define a sub-trajectory as the sequence of displacements of one particle over a time interval of length $i$, and consider sub-trajectories to be numerically independent under two conditions: either when they describe displacements of different particles, or when they describe displacements of the same particle but are calculated from non-overlapping time windows (see \cref{fig:displacements}).
+
+An alternative method for estimating the variance of the mean for time-correlated data is block averaging~\cite{Frenkel2023-ah}, where the input data is divided into non-overlapping sequential ``blocks'', and the set of averages calculated over each block are used for statistical analysis.
+A popular form of block averaging is the block renormalisation method of Flyvbjerg and Peterson~\cite{FlyvbjergAndPetersen_JChemPhys1989}.
+
+The Flyvbjerg--Peterson method starts with some input data, $A$.
+If the elements of $A$ are uncorrelated, the variance of the mean can be estimated by rescaling the variance of $A$:
+\begin{equation}
+\varest{\xbar{A}}=\frac{\var{A}}{L_A-1},
+\end{equation}
+where $L_A$ is the number of elements in $A$.
+If elements of $A$ are serially correlated, however, this estimator systematically underestimates the true variance of the mean of $A$, and, instead, providing only an approximate lower bound:
+\begin{equation}
+\varest{\xbar{A}}\geq\frac{\var{A}}{L_{A}-1}.
+\end{equation}
+
+The method proceeds by iteratively applying ``blocking'' operations.
+The original dataset, $A$, is mapped to a new dataset $A_1$, by averaging over adjacent non-overlapping pairs of data in $A$ (see \cref{fig:blocking}).
+The new dataset is half the length of the original: $L_{A_1} = \frac{1}{2}L_{A}$.
+$A$ and $A_1$ have the same mean.
+However, we can now rescale the variance of $A_1$ to obtain a tighter estimated lower bound for \var{\xbar{A}}:
+\begin{equation}
+\label{equ:blocking}
+\var{\xbar{A}}\geq\frac{\var{A_1}} {L_{A_1}-1}.
+\end{equation}
+Under repeated blocking operations with a sufficiently large input dataset, $\var{A_n}/(L_{A_n}-1)$ tends to \var{\xbar{A}}.
+In practice, sequential blocking steps are applied to the original dataset until $\var{A_n}/(L_{A_n}-1)$ reaches a plateau, and this plateau value is taken as the estimate for \var{\xbar{A}}.
+
+\begin{figure}[tb]
+  \centering
+  % \resizebox{9cm}{!}{\includegraphics*{figure2_SI.pdf}} 
+  \includegraphics[width=\columnwidth]{figure2_SI.pdf}
+    \caption{\label{fig:blocking}Schematic illustrating the application of repeated blocking operations with block-length 2 to generate a series of representations of the input data, $A$, $A_1$, $\ldots$, $A_n$, as used in the block renormalisation scheme of Flyvbjerg and Peterson~\cite{FlyvbjergAndPetersen_JChemPhys1989}.}
+\end{figure}
+
+The Flyvbjerg--Peterson block renormalisation method is particularly useful for estimating the variance of the mean in sequential correlated data where the correlation length is unknown; for example, when estimating thermodynamic averages from simulation trajectories.
+
+Figure~\ref{fig:msd_blocking_comparison} shows a comparison of the variance rescaling method described in the main text with the Flyvbjerg--Petersen block renormalisation method when using both methods to estimate the $\var{\oMSDi}$ from simulation.
+We applied both methods to a simulation of \num{128} particles undergoing a three-dimensional cubic-lattice random walk of \num{128} steps per particle.
+
+The variance rescaling method (Eqn.~7) shows close agreement between the estimated variance in the MSD from a single simulation and the true numerical variance obtained by sampling many equivalent simulations.
+The estimated variance varies smoothly with timestep, $i$.
+In the example shown, this method overestimates $\var{\oMSDi}$ at long times, which we attribute the the relatively small number of observed squared displacements that are used to estimate $\var{\oMSDi}$ in the large $i$ regime.
+
+In comparison, the block renormalisation method also provides reasonable estimates of the true numerically determined variance, but is generally less accurate than the variance rescaling method.
+This method exhibits more noise, with larger scatter in $\varest{\oMSDi}$ as the timestep $i$ changes.
+This greater noise is not entirely surprising, since the block renormalisation method aims to independently estimate the correlation length of the input data numerically for each timestep $i$, while the variance rescaling method takes advantage of the known correlation length due to the way the MSD is computed (\cref{fig:displacements}).
+The block renormalisation method also tends to underestimate $\var{\oMSDi}$, which can be attributed to its provision of an estimated lower bound for the variance of the mean of the input data (see \cref{equ:blocking}).
+
+In \cref{fig:pyblock_comparison} we compare the results of estimating $\D$ and $\var{\Dest}$ using our approximate Bayesian regression scheme, using a model covariance matrix, $\mathbf{\Sigma^\prime}$, as defined in Eqn.~6 in the main text, parameterised by $\varest{\oMSDi}$ computed by either variance rescaling or block renormalisation.
+
+Both methods give unbiased estimates of $\D$ and similar distributions $\prob{\Dest}$.
+However, the distribution $\prob{\Dest}$ obtained using variance rescaling is slightly narrower than that obtained using block renormalisation.
+This difference stems from the greater stochastic noise in $\varest{\oMSDi}$ when using the block renormalisation method.
+Consequently, the model covariance matrices, $\mathbf{\Sigma^\prime}$, parametrised from these estimates are themselves noisier and often difficult to condition, leading to numerical instabilities.
+Both the increased noise and numerical instabilities when conditioning contribute to a wider distribution in $\prob{\Dest}$ when block renormalisation is used to estimate $\var{\oMSDi}$.
+
+The choice of method used to estimate $\var{\oMSDi}$ also influences the estimate of the uncertainty in $\Dest$.
+The variance rescaling method provides a good estimate for $\var{\Dest}$ but systematically overestimate the true uncertainty.
+In contrast, the block renormalisation method provides a similarly good estimate but systematically \emph{underestimates} the true uncertainty in $\Dest$.
+When estimating $\D$ from molecular dynamics simulations, we consider overestimation of $\var{\Dest}$ to always be preferable to underestimation.
+Overestimated uncertainty in $\Dest$ can be addressed by collecting more data, for example, by performing longer simulations.
+Conversely, underestimated uncertainty in $\Dest$ may lead to false confidence in the accuracy of $\D$ estimates, which can potentially lead to downstream errors in inferential reasoning or formal hypothesis testing.
+
+\begin{figure*}[tb]
+  \centering
+%   \resizebox{11cm}{!}{\includegraphics*{figure3_SI.pdf}} 
+  \includegraphics[width=\columnwidth]{pyblock.pdf}
+    \caption{\label{fig:msd_blocking_comparison}Comparison of the numerical variance in observed MSD from multiple replica simulations (panel a), the estimated variance obtained by rescaling the variance in observed squared displacements from a single simulation (panels b and c) (Eqn.~7), and the estimated variance obtained from the block renormalisation method of Flyvbjerg and Peterson~\cite{FlyvbjergAndPetersen_JChemPhys1989}, as implemented in \textsc{pyblock}~\cite{spencer_pyblock_2020}.
+Panel (a) shows the mean observed MSD from \num{4096} simulations of \num{128} particles undergoing a 3D lattice random walk of 128 steps per particle, with error bars of $\pm2\var{\oMSDi}$.
+Panel (b) shows the MSD from just one simulation, with error bars of $\pm\varest{\oMSDi}$, obtained via Eqn.~7.
+Panel (d) shows the same one-simulation MSD, again with error bars of $\pm\varest{\oMSDi}$, obtained using the block renormalisation method of Flyvbjerg and Peterson~\cite{FlyvbjergAndPetersen_JChemPhys1989}.
+Panels (c) and (e) plot the numerical variance against the single-simulation estimated variances obtained with each method, as a function of timestep, $i$.}
+    \script{pyblock.py}
+\end{figure*}
+
+\begin{figure*}[tb]
+  \centering
+%   \resizebox{10.0cm}{!}{\includegraphics*{figure4_SI.pdf}} 
+\includegraphics[width=\columnwidth]{msd_blocking.pdf}
+    \caption{\label{fig:pyblock_comparison}(Panels (a) and (c)) Probability distributions of point-estimates $\prob{\D}$ obtained from \num{4096} individual random-walk simulations.
+Panel (a) shows data obtained using variance rescaling to estimated the variances, $\varest{\oMSDi}$, used to parametrise the model covariance matrix, $\mathbf{\Sigma^\prime}$.
+Panel (b) shows data obtained using block renormalisation to estimate the variances, $\varest{\oMSDi}$, used to parametrise the model covariance matrix.
+    The grey lines show the distribution of point estimates, $\prob{\Dest}$, obtained using Bayesian regression with a mean vector and numerical covariance matrix derived from the complete dataset of all \num{4096} simulations.
+    The pink horizontal bar shows an interval of one standard deviation in $\prob{\Dest}$. 
+    Panels (b) and (d) show the corresponding  probability distributions of estimated variances, $\varest{\Dest}$, 
+    for individual random-walk simulations, compared to the true sample variances obtained using each method (pink vertical lines) 
+    $\var{\Dest}$.
+    }
+    \script{msd_blocking.py}
+\end{figure*}
+
+\section{Bias in $\prob{\varest{\Dest}}$}
+\label{sec:bias}
+In the main manuscript, we present results for a set of \num{4096} 3D-lattice random walk simulations, each consisting of \num{128} particles undergoing \num{128} steps (Fig.~4).
+Our approximate Bayesian regression scheme allows us to estimate the variance in $\Dest$, denoted as $\varest{\Dest}$, that would be obtained over a large number of repeat simulations.
+This estimate is calculated from the variance of the marginal posterior distribution $\prob{\D|\model}$, which we derive from analysis of a single simulation trajectory.
+As shown in Fig.~4d, our estimate for the population variance $\var{\Dest}$, obtained from a single simulation, aligns reasonably with the true value.
+When considering the distribution of estimated variance, $\prob{\varest{\Dest}}$, however, we observe a systematic overestimation (bias) relative to the true value.
+
+This bias arises from our use of estimated variances $\varest{\oMSDi}$ when parametrising the model covariance matrix $\bm{\Sigma^\prime}$.
+Figure~\ref{fig:true_cov} presents equivalent results for $\prob{\Dest}$ and $\prob{\varest{\Dest}}$ for the same \num{4096} individual simulations, but calculated using a numerical covariance matrix, $\bm{\Sigma}_\mathrm{num}$ derived from all \num{4096} observed MSDs.
+The resulting distribution $\prob{\varest{\Dest}}$ (see Fig.~\ref{fig:true_cov}b) is unbiased.
+Furthermore, the distribution $\prob{\Dest}$ agrees even more closely with the numerically converged distribution obtained when combining data from all \num{4096} simulations (Fig~\ref{fig:true_cov}a), contrasting with the results presented in Fig.~4b, where our approximate Bayesian regression scheme yields a slightly broadened distribution due to the use of the long-time limit in the derivation of the analytical form for $\bm{\Sigma^\prime}$.
+\begin{figure}[tb!]
+    \centering
+    % \resizebox{\columnwidth}{!}{\includegraphics*{figure5_SI.pdf}}
+    \includegraphics[width=\columnwidth]{true_cov.pdf}
+    \caption{
+        (a) Probability distribution of point-estimates, $\prob{\Dest}$, obtained from \num{4096} individual random-walk simulations, using the numerical covariance matrix $\bm{\Sigma}_\mathrm{num}$.
+        Each simulation has been analysed as in Fig.~4(a) and (b) to yield a single corresponding point estimate $\Dest$.
+        The grey line shows the distribution of point estimates, $\prob{\Dest_\mathrm{num}}$, obtained using Bayesian regression with a mean vector and numerical covariance matrix derived from the complete dataset of all \num{4096} simulations.
+        The pink horizontal bar shows an interval of one standard deviation in $\prob{\Dest}$.
+        (b) Probability distribution of estimated variances, $\varest{\Dest}$, for individual random-walk simulations, using the numerical covariance matrix $\bm{\Sigma}_\mathrm{num}$, compared to the true sample variance (pink vertical line) $\var{\Dest}$.
+        }
+    \label{fig:true_cov}
+    \script{true_cov.py}
+\end{figure}
+
+\section{Comparison of OLS, WLS, and GLS used to estimate $\D$ in \ce{Li7La3Zr2O12}}
+\label{sec:llzo}
+In the main manuscript, Fig.~1 shows example distributions of estimated self-diffusion coefficients, $\Dest$, calculated using OLS, WLS, and GLS estimators from MSD data from \num{4096} 3D lattice random walk simulations.
+This figure shows that GLS gives a much narrower distribution of $\Dest$ than either OLS or WLS, and also allows the width of this distribution (characterised by $\var{\Dest}$) to be accurately estimated, in contrast with OLS and WLS, which both give estimates of $\var{\Dest}$ that significantly underestimate the true variance.
+
+While the 3D lattice random walk represents an idealised model system, OLS and WLS show the same deficiencies when used to estimate $\D$ from MSD data from simulations of more complex ``real world'' systems.
+\cref{fig:glswlsols_llzo} shows an equivalent comparison between OLS, WLS, and GLS applied to simulation data for the lithium solid electrolyte, \ce{Li7La3Zr2O12} (LLZO).
+As for the idealised 3D lattice random walk data, OLS and WLS both give wider distributions of estimated diffusion coefficients, $\prob{\Dest}$, while also systematically underestimating the uncertainty in these estimates.
+In contrast, GLS gives a narrower distribution of estimated values, and accurately estimates this uncertainty from single simulation data.
+
+\begin{figure}[tb]
+  \centering
+  % \resizebox{8cm}{!}{\includegraphics*{figure6_SI.pdf}} 
+  \includegraphics[width=\columnwidth]{glswlsols_llzo.pdf}
+    \caption{Example distributions of estimated self-diffusion coefficients, $\Dest$, calculated using (a) ordinary least squares (OLS), (b) weighted least squares (WLS), and (c) generalised least squares (GLS), from MSD data from \num{512} effective simulations of LLZO of \SI{\sim 25}{\pico\second} with \num{56} lithium ions.
+In each panel, the grey curve shows the best-fit normal distribution for the simulation data, the upper horizontal bar shows the standard deviation of this distribution, and the lower horizontal bar shows the average estimated standard distribution given by the analytical expression for $\var{\Dest}$ for each regression method.}
+    \label{fig:glswlsols_llzo}
+    \script{glswlsols_llzo.py}
+\end{figure}
+
 \end{document}

### 2024-10-29 src/scripts/msd_blocking.py
Fix blocking figure

@@ -11,13 +11,14 @@ jump = 1
 atoms = 128
 length = 128
 
-kinisi = np.load(paths.data / f"random_walks/pyblock/rw_{jump}_{atoms}_{length}_s4096.npz")['data']
+kinisi = np.load(paths.data / f"random_walks/kinisi/rw_{jump}_{atoms}_{length}_s4096.npz")['data']
+pyblock = np.load(paths.data / f"random_walks/pyblock/rw_{jump}_{atoms}_{length}_s4096.npz")['data']
 numerical = np.load(paths.data / f"random_walks/numerical/rw_{jump}_{atoms}_{length}_s4096.npz")['data']
 dt = kinisi[0, :, 0]
 
-figsize = (3.73, 0.85)
+figsize = (3.73*2, 0.85*3)
 fig = plt.figure(figsize=figsize)
-gs = gridspec.GridSpec(1, 3, figure=fig, wspace=0.7, hspace=0.7)
+gs = gridspec.GridSpec(2, 3, figure=fig, wspace=0.7, hspace=1.4)
 
 axes = []
 titles = []
@@ -37,7 +38,7 @@ axes[-1].set_yticks([0, 400, 800])
 
 axes.append(fig.add_subplot(gs[0, 1], sharey=axes[-1]))
 titles.append(r"b")
-axes[-1].set_title("estimated")
+axes[-1].set_title("kinisi estimated")
 m, c, b = axes[-1].errorbar(dt, kinisi[0, :, 1], kinisi[0, :, 2] * 2, color='#E08D6D', elinewidth=0.2)
 [bar.set_alpha(0.5) for bar in b]
 [cap.set_alpha(0.5) for cap in c]
@@ -50,7 +51,7 @@ axes[-1].set_yticks([0, 400, 800])
 
 axes.append(fig.add_subplot(gs[0, 2]))
 titles.append(r"c")
-axes[-1].set_title("$\sigma^2$ comparison")
+axes[-1].set_title("kinisi $\sigma^2$ comparison")
 m = np.max([numerical.var(0), (kinisi[0, :, 2]**2)])
 axes[-1].plot(numerical.var(0) / 1000, (kinisi[0, :, 2]**2) / 1000,
               color='#67B99B',
@@ -63,12 +64,43 @@ axes[-1].set_xlabel(r"$\sigma^2[x(t)]$ / $10^3$")
 axes[-1].set_ylabel(r"$\hat{\sigma}^2[x(t)]$ / $10^3$")
[…]

### 2024-10-29 src/scripts/pyblock.py

@@ -13,6 +13,8 @@ length = 128
 
 kinisi = np.load(paths.data / f"random_walks/kinisi/rw_{jump}_{atoms}_{length}_s4096.npz")['diff_c']
 pyblock_mf = np.load(paths.data / f"random_walks/pyblock/rw_{jump}_{atoms}_{length}_s4096.npz")['diff_c']
+# dfit_mf = np.load(paths.data / f"random_walks/dfit/rw_{jump}_{atoms}_{length}_s4096.npz")['diff_c']
+# dfit_mf_std = np.load(paths.data / f"random_walks/dfit/rw_{jump}_{atoms}_{length}_s4096.npz")['diff_c_std']
 dinfty_true = np.load(paths.data / f"random_walks/numerical/D_1_128_128.npz")['diff_c']
 
 figsize = (4.03, 4)
@@ -95,7 +97,6 @@ axes[-1].set_xlim([0, None])
 titles.append("d")
 axes[-1].set_title("blocking")
 axes[-1].legend(loc='upper left', bbox_to_anchor=(0.5, 1))
-
 
 fig.align_ylabels(axes)
 

### 2024-12-06 src/tex/ms.tex
Fix blocking figure
Typo
Formatting
Fix references

@@
 \renewcommand{\addcontentsline}[3]{}
+\footnotetext[1]{While estimating $\D$ by fitting a linear model to MSD data is, perhaps, the most commonly used approach in the literature, alternative methods that avoid linear fitting of MSD data also exist; see, for example, Refs.~{\protect\cite{VestergaardEtAl_PhysRevE2014,KrapfEtAl_NewJournalPhysics2018,bullerjohn_optimal_2020}}.}
+\footnotetext[2]{A common approach to estimating the variance of the mean of time-correlated data with unknown correlation time is the renormalization group blocking method of Flyvberg and Peterson~\cite{FlyvbjergAndPetersen_JChemPhys1989,Frenkel2023-ah}.
+In the SI, we compare this blocking method to our direct rescaling method (\cref{equ:varestMSD}) for estimating $\var{\oMSDi}$ from a set of random-walk trajectories.
+For our example data, direct rescaling performs better (gives more accurate estimates of $\var{\oMSDi}$).}
 
@@
 \author{Samuel W. Coles}
+  \thanks{Present Address: Yusuf Hamied Department of Chemistry, University of Cambridge, Lensfield Road, Cambridge, CB2 1EW, UK.}
   \affiliation{Department of Chemistry, University of Bath, Claverton Down, Bath, BA2 7AY, UK}
@@
 \end{equation}
-where $\Delta\mathbf{r}{(t)}$ is the displacement of a diffusion particle in the time interval $t$.
+where $\Delta\mathbf{r}{(t)}$ is the displacement of a diffusing particle in the time interval $t$.
 Because numerical simulations are finite in time and space, MSDs obtained from simulation data always differ from the true ensemble average MSD.
-One can, however, compute an estimate of the self-diffusion coefficient, $\Dest$, by fitting a linear model to the observed MSD and using the gradient of this fitted model in place of $\MSD{t} / t$ in \cref{equ:einstein}~\footnote{While estimating $\D$ by fitting a linear model to MSD data is, perhaps, the most commonly used approach in the literature, alternative methods that avoid linear fitting of MSD data also exist; see, for example, Refs.~{\protect\cite{VestergaardEtAl_PhysRevE2014,KrapfEtAl_NewJournalPhysics2018,bullerjohn_optimal_2020}}.}.
+One can, however, compute an estimate of the self-diffusion coefficient, $\Dest$, by fitting a linear model to the observed MSD and using the gradient of this fitted model in place of $\MSD{t} / t$ in \cref{equ:einstein}~\cite{Note1,VestergaardEtAl_PhysRevE2014,KrapfEtAl_NewJournalPhysics2018,bullerjohn_optimal_2020}.
 
@@
 The resulting posterior distribution provides an efficient estimate for $\D$ and allows the associated statistical uncertainty in $\Dest$ to be accurately quantified.
-This method is implemented in the open-source Python package \textsc{kinisi}~\cite{mccluskey_kinisi_2022}.
+This method is implemented in the open-source Python package \textsc{kinisi}~\cite{McCluskey2024}.
 
@@
 
-The estimated variance $\varest{\oMSD}$ can be calculated from a single simulation trajectory, and provides an accurate estimate of the true variance $\var{\oMSD}$~\footnote{A common approach to estimating the variance of the mean of time-correlated data with unknown correlation time is the renormalization group blocking method of Flyvberg and Peterson~\cite{FlyvbjergAndPetersen_JChemPhys1989,Frenkel2023-ah}.
-In the SI, we compare this blocking method to our direct rescaling method (\cref{equ:varestMSD}) for estimating $\var{\oMSDi}$ from a set of random-walk trajectories.
-For our example data, direct rescaling performs better (gives more accurate estimates of $\var{\oMSDi}$).}.
+The estimated variance $\varest{\oMSD}$ can be calculated from a single simulation trajectory, and provides an accurate estimate of the true variance $\var{\oMSD}$~\cite{Note2,FlyvbjergAndPetersen_JChemPhys1989,Frenkel2023-ah}.
 To demonstrate this, we performed \num{4096} independent simulations of \num{128} particles undergoing a three-dimensional cubic-lattice random walk of \num{128} steps per particle.
@@
 Overall, our method offers significant advantages over more conventional methods of estimating self-diffusion coefficients from atomistic simulations.
-We have implemented this procedure in the open-source package \textsc{kinisi} \cite{mccluskey_kinisi_2022}, which we hope will support its use within the broader simulation community.
+We have implemented this procedure in the open-source package \textsc{kinisi} \cite{McCluskey2024}, which we hope will support its use within the broader simulation community.
 
@@
 The LLZO raw simulation trajectories are available on Zenodo shared under a CC BY-SA 4.0 licence~\cite{coles_llzo_zenodo_2022}.
-The method outlined in this work is implemented in the open-source Python package \textsc{kinisi}~\cite{mccluskey_kinisi_2022}, which is available under an MIT license.
+The method outlined in this work is implemented in the open-source Python package \textsc{kinisi}~\cite{McCluskey2024}, which is available under an MIT license.
 
@@
  \appendix
- \renewcommand\thesection{SI.\Roman{section}}
+ \renewcommand\thesection{S-\Roman{section}}
  \counterwithout{figure}{section}
- \renewcommand\thefigure{SI.\arabic{figure}}
+ \renewcommand\thefigure{S-\arabic{figure}}
  \setcounter{figure}{0}
  \counterwithout{equation}{section}
- \renewcommand\theequation{SI.\arabic{equation}}
+ \renewcommand\theequation{S-\arabic{equation}}
  \setcounter{equation}{0}
  \counterwithout{table}{section}
- \renewcommand\thetable{SI.\arabic{table}}
+ \renewcommand\thetable{S-\arabic{table}}
  \setcounter{table}{0}
  \pagenumbering{arabic} 
- \renewcommand\thepage{SI.\arabic{page}}
+ \renewcommand\thepage{S-\arabic{page}}
  \addtocontents{toc}{\protect\setcounter{tocdepth}{0}}
@@
 %   \resizebox{11cm}{!}{\includegraphics*{figure3_SI.pdf}} 
-  \includegraphics[width=\columnwidth]{pyblock.pdf}
+  \includegraphics[width=\textwidth]{msd_blocking.pdf}
     \caption{\label{fig:msd_blocking_comparison}Comparison of the numerical variance in observed MSD from multiple replica simulations (panel a), the estimated variance obtained by rescaling the variance in observed squared displacements from a single simulation (panels b and c) (Eqn.~7), and the estimated variance obtained from the block renormalisation method of Flyvbjerg and Peterson~\cite{FlyvbjergAndPetersen_JChemPhys1989}, as implemented in \textsc{pyblock}~\cite{spencer_pyblock_2020}.
@@
 Panels (c) and (e) plot the numerical variance against the single-simulation estimated variances obtained with each method, as a function of timestep, $i$.}
-    \script{pyblock.py}
+    \script{msd_blocking.py}
 \end{figure*}
@@
 %   \resizebox{10.0cm}{!}{\includegraphics*{figure4_SI.pdf}} 
-\includegraphics[width=\columnwidth]{msd_blocking.pdf}
+\includegraphics[width=\columnwidth]{pyblock.pdf}
     \caption{\label{fig:pyblock_comparison}(Panels (a) and (c)) Probability distributions of point-estimates $\prob{\D}$ obtained from \num{4096} individual random-walk simulations.
@@
     }
-    \script{msd_blocking.py}
+    \script{pyblock.py}
 \end{figure*}
@@
   \centering
-  % \resizebox{8cm}{!}{\includegraphics*{figure6_SI.pdf}} 
+  % \resizebox{8cm}{!}{\includegra?phics*{figure6_SI.pdf}} 
   \includegraphics[width=\columnwidth]{glswlsols_llzo.pdf}
